### 2024-08-22 doc/paper.tex
init paper

\documentclass[aps,prb,groupedaddress,twocolumn,showpacs,dvipdfmx, superscriptaddress,pdftex]{revtex4-2}
\usepackage[whole]{bxcjkjatype}
\usepackage{amsmath}
\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
\usepackage{amssymb}
\usepackage{bm}
\usepackage{dcolumn}
\usepackage{graphicx}
\usepackage{bm}
\usepackage{url}
\usepackage{here}
\usepackage{multirow}
\usepackage{ulem}
\usepackage{booktabs}
\usepackage[T1]{fontenc}
\usepackage[caption=false]{subfig}
\usepackage[usenames,dvipsnames]{color}
% \usepackage{pdfpages}
% \usepackage{epstopdf}
% \epstopdfDeclareGraphicsRule{.pdf}{png}{.png}{convert #1 \OutputFile}
% \AppendGraphicsExtensions{.pdf}
% \usepackage{hyperref}

% tiếng Việt
\usepackage[T5]{fontenc}
\usepackage[utf8]{inputenc}
\DeclareTextSymbolDefault{\DH}{T1}
%-------------------------------------
\newcommand{\tadd}[1]{\textcolor{red}{#1}}
\newcommand{\trem}[1]{{\color{red}\sout{#1}}}
\newcommand{\tcom}[1]{\textcolor{red}{\textbf{TI: #1}}}
\newcommand{\citehere}{\textcolor{red}{\textbf{(citation)}}}
\newcommand{\x}{\times}
%-------------------------------------
% \newcommand{\ghost}[1]{[#1]}
\newcommand{\ghost}[1]{}
%----------------------------------------
\def\red{\color{red}}
\def\blue{\color{blue}}
\def\green{\color{green}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\title{
  Meta-learning for multi foreign exchange movement prediction
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\author{Bao-Long Nguyen}
\email{mwklng2309@icloud.com}
\affiliation{School of Information Science, JAIST, 1-1 Asahidai,
  Nomi, Ishikawa, 923-1292, Japan}
%
\author{Tom Ichibha}
\email{ichibha@icloud.com}
\affiliation{School of Information Science, JAIST, 1-1 Asahidai,
  Nomi, Ishikawa, 923-1292, Japan}
%
\author{Ryo Maezono}
\email{rmaezono@mac.com}
\affiliation{School of Information Science, JAIST, 1-1 Asahidai,
  Nomi, Ishikawa, 923-1292, Japan}
%
\date{\today}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}

  \textbf{Abstract}. Xử lý bài toán dự đoán các chỉ số liên quan đến tài chính bằng machine learning gặp rất nhiều khó khăn vì loại dữ liệu này không có chu kỳ rõ ràng và không chỉ phụ thuộc vào lịch sử giao dịch, mà còn phụ thuộc rất nhiều vào tình hình kinh tế, chính trị, điều mà rất khó capture trong tập dữ liệu. Để vượt qua các thách thức này, chúng tôi sử dụng Meta-learning, kết hợp với \verb|LSTM| và \verb|CNN| để tổng hợp hiệu quả các đặc trưng ẩn của dữ liệu theo thời gian. Thực nghiệm trên bài toán dự đoán xu hướng tỷ giá ngoại hối của 15 cặp tiền tệ cho thấy phương pháp của chúng tôi đạt hội tụ ở mức cao khi so sánh với mô hình \verb|NHITS|. % đoạn này cần sửa lại xíu

\end{abstract}
%----------------------------------------
\keywords{Meta-learning, Machine leaning, \verb|LSTM|, \verb|CNN|, Time-series, Foreign Exchange}
\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Foreign exchange prediction nói riêng và dự đoán các chỉ số tài chính như giá chứng khoán nói chung từ lâu đã là vấn đề đáng quan tâm của nhiều nghiên cứu \citep{li2019multi, islam2021foreign, heryadi2021foreign}. Hai kỹ thuật chính được sử dụng trong financial time-series prediction là fundamental analysis and technical analysis \cite{ayitey2023forex}. Trong khi fundamental analysis thiên về phân tích các yếu tố như chính sách, chiến lược kinh tế của công ty, quốc gia để dự đoán tương lai; technical analysis dựa hoàn toàn vào lịch sử biến động giá để phân tích xu hướng thị trường.

\vspace{2mm}

Việc dự đoán các chỉ số tài chính này gặp một vài thách thức cố hữu. Trong đó có thể kể tới: (1) - Variance của loại dữ liệu này không hề cố định, từ đó giả thuyết rằng chúng tuân theo một phân phối để có thể xấp xỉ lỗi là không thể sử dụng được, dẫn đến việc các mô hình học máy khó có thể dự đoán chính xác giá cũng như xu hướng tăng giảm trong tương lai; (2) - Các chỉ số tài chính phần lớn không tuân theo bất kỳ quy luật rõ ràng nào nên việc học các đặc trưng ẩn của dữ liệu để tiến hành dự đoán gặp rất nhiều khó khăn; (3) - Một chỉ số tài chính (e.g. giá cổ phiếu của một công ty) không hoàn toàn phụ thuộc vào dữ liệu quá khứ mà còn phụ thuộc nhiều vào các tin tức, tình hình kinh tế, chính trị \cite{li2019multi}.

\vspace{2mm}

Đối với thách thức đầu tiên, các mô hình ensemble [cite something here] thường được sử dụng để hạn chế ảnh hưởng của sự biến đổi variance. Chúng tôi cũng tiếp cận bài toán theo hướng này nhưng ở mức cao hơn bằng cách sử dụng Meta-learning (ML) \cite{finn2017model}. Phương pháp này tổng hợp hiệu quả tham số của các mô hình cục bộ, giúp giảm thiểu đáng kể variance loss.

\vspace{2mm}

Để khắc phục thách thức thứ hai, các nghiên cứu phần lớn sử dụng các đặc trưng rút trích được từ mô hình Long short-term memory neural network (\verb|LSTM|) \cite{hochreiter1997long}, Artificial neural network (\verb|ANN|), và Convolution neural network (\verb|CNN|) \cite{lecun1989handwritten}. Cụ thể, trong năm 2022, 20\% tổng số các bài báo liên quan đến dự đoán chỉ số tài chính sử dụng \verb|LSTM|, 20\% sử dụng \verb|ANN| và 6\% sử dụng \verb|CNN| \cite{ayitey2023forex}. Để tận dụng hết được các đặc trưng được rút trích bởi các mô hình nêu trên, chúng tôi đề xuất phương pháp kết hợp các đặc trưng này.

\vspace{2mm}

Đối với thách thức thứ ba, nghiên cứu \cite{fama1970efficient} đưa ra giả thuyết rằng chỉ số tài chính của các công ty ngầm phản ánh tình hình kinh tế tài chính trên thị trường. Các nghiên cứu \citep{overreactioncontrarian, mech1993portfolio} cũng chỉ ra sự phụ thuộc giữa chỉ số tài chính của một công ty nhất định và các chỉ số của các công ty khác. Điều này càng làm tăng tính đúng đắn của giả thuyết trong nghiên cứu \cite{fama1970efficient}. Ngoài ra, chúng tôi cho rằng, các chỉ số tài chính còn có những phụ thuộc ngầm vào các thời điểm nhất định trong quá khứ (hidden-long-term dependency). Đối với cách tiếp cận truyền thống, người ta sử dụng một lượng dữ liệu quá khứ cố định (lookback window) để huấn luyện mô hình. Điều này gây một trở ngại lớn cho quá trình học vì các đặc trưng dài hạn theo thời gian sẽ bị quên. Mặt khác, ML chia nhỏ tập dữ liệu thành nhiều phần để học và tổng hợp hiệu quả các tham số học được nên có thể xử lý tốt thách thức này.

\vspace{2mm}

Cuối cùng, chúng tôi chứng minh tính ưu việt của thuật toán đề xuất bằng cách giải bài toán dự đoán xu hướng (tăng hoặc giảm) của tỉ giá ngoại hối và so sánh với mô hình SOTA hiện tại (\verb|NHITS| \cite{challu2023nhits}) trên hai loại dữ liệu: (1) - Dữ liệu tỉ giả của cặp tiền tệ USD/JPY; (2) - Dữ liệu tỷ giá của 60 cặp tiền tệ, cấu thành từ các nước Australia, Canada, Switzerland, Denmark, EU, United Kingdom, Hong Kong, Iceland, Japan, Norway, New Zealand, Singapore, Sweden, Turkey, United States, Mexico, China, South Africa. Các tập dữ liệu này được công khai trên Internet và có thể tải về dễ dàng. Bản cài đặt chính thức có thể xem tại [bỏ cái link github vào đây].

\vspace{2mm}

Tóm lại, đóng góp chính của chúng tôi như sau:

\begin{itemize}
  \item Tổng hợp hiệu quả tham số mô hình: Sử dụng ML thay cho các mô hình ensemble truyền thống trong việc tổng hợp kết quả từ các mô hình học máy.
  \item Kết hợp đặc trưng: Rút trích đặc trưng bằng cách kết hợp các đặc trưng của \verb|LSTM| và \verb|CNN|.
  \item Hidden-long-term dependency: Chứng minh thực nghiệm rằng một chỉ số tài chính không chỉ phụ thuộc vào các chỉ số tài chính khác mà còn có các phụ thuộc ẩn với chính nó.
  \item Experiment: Thực nghiệm trên các bộ dữ liệu về tỉ giá hối đoái và so sánh với mô hình SOTA NHITS để chứng minh tính hiệu quả của phương pháp đề xuất.
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Related work}
\label{sec.relatedWork}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{LSTM \& CNN in feature extraction}

Như đã đề cập, \verb|LSTM| là mạng neural rất phổ biến trong việc handle các bài toán liên quan đến dự đoán chỉ số tài chính nói riêng cũng như các bài toán trên dữ liệu time-series nói chung. \verb|LSTM| được sử dụng phổ biến như vậy bởi vì nó xử lý tốt vấn đề vanishing gradient và có thể khai thác các mối quan hệ phi tuyến trong dữ liệu. Thật vậy, bằng cách duy trì cell-state trong mỗi iteration, \verb|LSTM| có thể chống lại một cách hiệu quả vấn đề vanishing gradient, từ đó bảo toàn khả năng capture các phụ thuộc dài hạn \cite{cheng2018applied}. \verb|LSTM| cũng thực hiện rút trích đặc trưng với các hàm kích hoạt phi tuyến, giúp học được các tham số mô hình có thể capture tính phi tuyến của dữ liệu \cite{he2016exploiting}. Hai yếu tố nêu trên khiến cho \verb|LSTM| trở thành lựa chọn đầu tiên được nghĩ đến khi giải quyết các bài toán trên dữ liệu time-series.

\vspace{2mm}

\verb|CNN| được sử dụng rất nhiều trong các tác vụ xử lý hình ảnh \citep{naranjo2020review, sharma2018analysis} bởi khả năng tổng hợp các quan hệ cục bộ. Không chỉ vậy, \verb|CNN| còn được dùng rất nhiều trong các tác vụ xử lý dữ liệu time-series như speech recognition \cite{dua2022developing}, natural language processing \cite{varshitha2023natural}. Điều đó chứng tỏ được khả năng của \verb|CNN| trong việc khám phá mối quan hệ thời gian giữa các mẫu dữ liệu. Mặc dù vậy, \verb|CNN| lại rất ít được dùng trong các tác vụ dự đoán chỉ số tài chính. Trong nghiên cứu này, chúng tôi tận dụng khả năng rút trích đặc trưng cục bộ tuyệt vời của \verb|CNN| để tích hợp thêm thông tin ẩn vào quá trình huấn luyện của mô hình.

\subsection{Meta-learning}

Các thuật toán Meta-learning (ML), điển hình là Model-agnostic meta-learning (MAML) \cite{finn2017model} được biết đến với khả năng huấn luyện một mô hình có tính tổng quát cao, thích ứng nhanh trên tập dữ liệu mới thông qua một lượng nhỏ dữ liệu và số bước huấn luyện \citep{hospedales2021meta, vettoruzzo2024advances}. Với khả năng này, ML được sử dụng rất nhiều trong việc cải thiện tốc độ đáp ứng của mô hình trên dữ liệu như cá nhân hóa mô hình học \cite{fallah2020personalized}, giải quyết vấn đề domain shift [insert citation].

\vspace{2mm}

Một thuật toán ML cơ bản sẽ được học trên nhiều tác vụ $t$ rút ra từ cùng một phân phối tác vụ $\mathcal{T}$ \cite{hospedales2021meta}. Dữ liệu của mỗi tác vụ được chia thành tập support $\mathcal{D}_t^{support}$ (thường có kích thước nhỏ, khoảng 20\%) và tập query $\mathcal{D}_t^{query}$. Trong qua trình học, hai bước tối ưu inner và outer optimization được perform đan xen. Inner optimization cố gắng tìm ra một bộ tham số tối ưu $\theta_t^*$ cho từng mô hình học máy trên tập support của mỗi tác vụ bằng phương trình \ref{eq:inner_opt}.

\begin{equation}
  \theta_t^* = \theta_t(\phi) = \arg\min_{\theta}{\mathcal{L}_t\left( \phi, \mathcal{D}_t^{support} \right)}
  \label{eq:inner_opt}
\end{equation} Trong đó, $\phi$ là kết quả của quá trình outer optimization, đóng vai trò là giá trị khởi tạo của $\theta_t$. $\mathcal{L}_t$ là hàm lỗi của mô hình trên tác vụ $t$.

\vspace{2mm}

Sau đó, thuật toán sử dụng các bộ tham số tối ưu $\theta_t^*$ để perform trên tập query tương ứng. Lỗi của toàn bộ mô hình sau đó được tổng hợp để thực hiện quá trình outer optimization như phương trình \ref{eq:outer_opt}.

\begin{align*}
  \phi^* &= \arg\min_{\phi}\sum_{t}{\mathcal{L}_t\left[ \theta_t^*, \mathcal{D}_t^{query} \right]}\\
  &= \arg\min_{\phi}\sum_{t}{\mathcal{L}_t\left[ \theta_t(\phi), \mathcal{D}_t^{query} \right]} \numberthis
  \label{eq:outer_opt}
\end{align*}

Trong inference phase, giá trị khởi tạo cho tham số của mô hình được gán bằng $\phi^*$. Mô hình sau đó được huấn luyện nhanh trên tập support sau đó perform trên tập query. Kết quả trên tập query chính là kết quả của mô hình.

\vspace{2mm}

Bằng hình thức huấn luyện trên, mô hình $\phi^*$ sẽ có mức tổng quát hóa cao trên các tác vụ khác nhau, có thể nhanh chóng đáp ứng một tác vụ mới chỉ sau một vài bước huấn luyện. Tận dụng ưu điểm này, chúng tôi

\vspace{2mm}

Các mô hình hybrid ensemble vốn được sử dụng rất nhiều trong các bài toán xử lý time-series và được chứng minh thực nghiệm là có độ chính xác cao hơn so với các mô hình handle time-series data tiêu chuẩn vì có thể tổng hợp được sức mạnh của nhiều mô hình \cite{ayitey2023forex}. Tuy vậy, các hình thức tổng hợp của ensemble model hiện nay vẫn còn rất cứng nhắc vì chỉ có thể tổng hợp dựa trên kết quả cuối (đối với bagging models) và kết quả gần cuối (đối với stacking models). Dưới góc nhìn của ensemble model, có thể coi phương trình \ref{eq:outer_opt} là một phương pháp tổng hợp hiệu quả các sub-model, giúp tận dụng khả năng rút trích đặc trưng của từng mô hình. Nói cách khác, mô hình sau khi tổng hợp có thể rút trích đặc trưng ở mức sâu hơn, cải thiện đáng kể khả năng dự đoán so với các mô hình ensemble truyền thống.

\subsection{Neural Hierarchical Interpolation\\for Time Series (NHITS)}

NHITS được thiết kế để hướng đến việc dự đoán các long-horizon time-series data. Theo nghiên cứu \cite{challu2023nhits}, cấu trúc của \verb|NHITS| bao gồm nhiều stack liên tiếp nhau. Mỗi stack bao gồm nhiều block nối tiếp nhau. Tại mỗi block, dữ liệu lịch sử được sử dụng để dự đoán dữ liệu tương lai và dữ liệu quá khứ. Cụ thể, tại block $l$, với $L$ mẫu dữ liệu quá khứ ($\mathbf{y}_{t-L:t, l-1}$), các đặc trưng sẽ được rút trích như sau (theo \cite{challu2023nhits}):

\begin{align}
  \mathbf{y}_{t-L:t, l}^{(p)} &= Pooling\left( \mathbf{y}_{t-L:t, l-1} \right)\\
  % \mathbf{h}_l &= FullyConnected\left( \mathbf{y}_{t-L:t, l}^{(p)} \right)\\
  \mathbf{\theta}_l^b &= FullyConnected^b \left( \mathbf{y}_{t-L:t, l}^{(p)} \right)\\
  \mathbf{\theta}_l^f &= FullyConnected^f \left( \mathbf{y}_{t-L:t, l}^{(p)} \right)\\
  \mathbf{\hat{y}}_{t-L:t, l} &= g(\mathbf{\theta}_l^b)\\
  \mathbf{\hat{y}}_{t+1:t+H, l} &= g(\mathbf{\theta}_l^f)
\end{align}

Trong đó, \verb|FullyConnected| là các lớp multi-layer perception (\verb|MLP|) xếp chồng với hàm kích hoạt phi tuyến. $\mathbf{\theta}_l^f, \mathbf{\theta}_l^b$ là các hệ số nội suy forecast và backcast, được dùng để tổng hợp các giá trị đầu ra của block $l$ bằng hàm nội suy $g(\cdot)$. Đầu ra của block $l$ là giá trị forecast $\mathbf{\hat{y}}_{t+1:t+H, l}$ và giá trị backcast $\mathbf{\hat{y}}_{t-L:t, l}$. Input của block $l+1$ được tính theo phương trình \ref{eq:input_l1}.

\begin{align}
  \mathbf{y}_{t-L:t, l+1} = \mathbf{y}_{t-L:t, l} - \mathbf{\hat{y}}_{t+1:t+H, l}
  \label{eq:input_l1}
\end{align}

Tổng hợp các giá trị forecast của các block, ta được giá trị forecast của một stack. Backcast của block cuối cùng của một stack chính là đầu vào cho stack tiếp theo. Cuối cùng, tổng hợp giá trị forecast của các stack, ta được giá trị forecast dự đoán của toàn mạng.

Bằng cách xếp chồng các stack, stack sau nhận vào phần dư của stack trước, kiến trúc trên được kỳ vọng là sẽ phân rã dữ liệu thành các frequency bands khác nhau. Trong thực tế, \verb|NHITS| perform rất tốt đối với các bộ dữ liệu có tính chu kỳ cao. Tuy nhiên, tính chu kỳ của dữ liệu foreign exchange là rất thấp, thậm chí không có, gây ra khó khăn rất lớn cho \verb|NHITS|.
Bằng cách xếp chồng các stack, stack sau nhận vào phần dư của stack trước, kiến trúc trên được kỳ vọng là sẽ phân rã dữ liệu thành các frequency bands khác nhau. Trong thực tế, \verb|NHITS| perform rất tốt đối với các bộ dữ liệu có tính chu kỳ cao. Tuy nhiên, tính chu kỳ của dữ liệu foreign exchange là rất thấp, thậm chí không có, gây ra khó khăn rất lớn cho \verb|NHITS|.

% \begin{figure}
%   \large
%   \centering
%   \includegraphics[witdth=\linewidth]{./img/meo.pdf}
%   % \caption{This is a caption \cite{challu2023nhits}}
% \end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{img/meo.pdf}
  \caption{Your figure caption here.}
  \label{fig:your-label}
\end{figure}

% \mathbf{y}_{t-L:t, l+1} &= \mathbf{y}_{t-L:t, l} - \mathbf{\hat{y}}_{t+1:t+H, l}

% nói về ưu điểm, nhược điểm của fundamental analysis and technical analysis --> đây là nói chung, sẽ có phần nói riêng sau (nói về từng phương pháp, dẫn đến phương pháp của tôi)

% tình hình kinh tế, chính trị, những thứ khó có thể cấu trúc hóa thành dữ liệu cho máy học.

%----------------------------------------
% \newcommand{\scale}{0.48}
% \begin{figure}[htbp]
%   \large
%   \centering
%   (a) I-type stacking of insulating TR lattice \\
%   \begin{minipage}{\scale\hsize}
%     \includegraphics[width=\hsize]{tri_off_top.eps}
%   \end{minipage}
%   \begin{minipage}{\scale\hsize}
%     \includegraphics[width=\hsize]{tri_off_side.eps}
%   \end{minipage}

% 	\vspace{5mm}
%   (b) II-type stacking of insulating TR lattice \\
%   \begin{minipage}{\scale\hsize}
%     \includegraphics[width=\hsize]{tri_on_top.eps}
%   \end{minipage}
%   \begin{minipage}{\scale\hsize}
%     \includegraphics[width=\hsize]{tri_on_side.eps}
%   \end{minipage}

% \vspace{5mm}
%   (c) I-type stacking of metallic NTR lattice \\
%   \begin{minipage}{\scale\hsize}
%     \includegraphics[width=\hsize]{ntr_off_top.eps}
%   \end{minipage}
%   \begin{minipage}{\scale\hsize}
%     \includegraphics[width=\hsize]{ntr_off_side.eps}
%   \end{minipage}

% \vspace{5mm}
%   (d) II-type stacking of metallic NTR lattice \\
%   \begin{minipage}{\scale\hsize}
%     \includegraphics[width=\hsize]{ntr_on_top.eps}
%   \end{minipage}
%   \begin{minipage}{\scale\hsize}
%     \includegraphics[width=\hsize]{ntr_on_side.eps}
%   \end{minipage}
%   \caption{\ghost{fig.str}\label{fig.str}
%     Structure models of LiV$X_2$ (X = O, S, and Se).
%     The left figure is top-view and the right figure is side-view.
%     The green balls are Li, the blue balls are V, and red balls are O.
%   }
% \end{figure}
%----------------------------------------

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Methodology}
\label{sec.method}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Fixed-node DMC

\vspace{2mm}
Fixed-node DMC
%----------------------------------------
\begin{equation}
\left[\sum_j\frac{1}{2}\nabla^2_j
+ V_{\rm MB}\left(\vec R\right)
\right]\Psi\left(\vec R\right)
= E~\Psi\left(\vec R\right) \ .
\end{equation}
%----------------------------------------
Fixed-node DMC

%----------------------------------------
\begin{equation}
\left[\cfrac{1}{2}~\nabla^2_j
+ v_{\rm ext}\left(\vec r_j\right)
+ v_{\rm XC}\left(\vec r_j\right)
\right]\psi\left(\vec r_j\right)
= \varepsilon~\psi\left(\vec r_j\right) \ .
\end{equation}
%----------------------------------------
Fixed-node DMC

\vspace{2mm}
Fixed-node DMC
%----------------------------------------
\begin{align}
  E\left[\Psi_T\right]
  &=\frac{\int d\vec{R}\cdot \Psi_T^*\left(\vec{R}\right)
  \hat{H}\Psi_T\left(\vec{R}\right)}{\int d\vec{R}\cdot \Psi_T^*\left(\vec{R}\right)\Psi_T\left(\vec{R}\right)} ,
  \label{eq.varfunc}
\end{align}
%----------------------------------------
\ghost{eq.varfunc}
Fixed-node DMC

\vspace{2mm}
Fixed-node DMC

\vspace{2mm}
Fixed-node DMC

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Experiment}
\label{sec.experiment}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%e%%%%%%
\subsection{Density functional theory}
Density functional theory

\subsection{Fixed--node DMCTEXT}
Fixed--node DMCTEXT

\vspace{2mm}
Fixed--node DMCTEXT

\vspace{2mm}
Fixed--node DMCTEXT

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results and discussion}
\label{sec.results}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Results and discussion

%----------------------------------------
% \begin{figure}[htbp]
%   \centering
%   \includegraphics[width=\hsize]{all.eps}
%   \caption{\ghost{fig.all}\label{fig.all}
% 	FNDMC and DFT results of relative energies of
% 	the four candidate structures of LiV$X_2$
% 	(X = O, S, and Se):
% 	the I- and II-type stacking of the insulating
%     TR and metallic NTR lattices.
% 	Their structures are shown in Figure \ref{fig.str}.
% 	The lowest energy among the four structures is set
% 	to be zero and the others are given as relative values
% 	from the lowest energy.
%   }
% \end{figure}
%----------------------------------------

\vspace{2mm}
Results and discussion

\vspace{2mm}
Results and discussion

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}
\label{sec.conc}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Conclusion

% We also compared DFT results using PBE\cite{PBE}, optPBE-vdW-DF\cite{2009JK_AM,vdW-DFa,vdW-DFb}, SCAN\cite{SCAN}, and HSE06\cite{HSE06} functionals with the FNDMC results. All of them also favour non-trimerized structures in all three materials, but none of the DFT functionals do not quantitatively agree with FNDMC. Thus our results demonstrate that this series of materials: LiVO$_2$,  LiVS$_2$, and LiVSe$_2$ represent a real challenge for computational schemes and also motivates further experimental study of the ground state properties of LiVSe$_2$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Acknowledgements}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The work was carried out within the state assignment of Ministry of Science and Higher Education of the Russian Federation (No. AAAA-A18-118020190095-4, topic ``Quantum''). G.I.P is grateful for financial supports from JST SPRING, Grant Number JPMJSP2102. K. H. is grateful for financial support from MEXT-KAKENHI (JP19K05029, JP21K03400, JP21H01998, and JP22H02170). R.M. is grateful for financial supports from MEXT-KAKENHI (22H05146, 21K03400 and 19H04692).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Author contributions}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
All authors contributed to conceiving the idea. T.I., A.V.U, and G.I.P performed the calculations. T.I. made the all figures. R.M and S.V.S supervised the work. All authors contributed to the discussion and writing of the paper.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Data availability statement}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The datasets used and/or analyzed during the current study available from the corresponding authors on reasonable request.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliographystyle{elsarticle-harv} 
\bibliography{references}
\end{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section{Appendix}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%We used the VASP code\cite{VASP} for the calculations explained in this appendix.
%We relaxed the four structures of Figure \ref{fig.str}
%for every LiV$X_2$ ($X$ = O, S, and Se) with each of PBE\cite{PBE},
%optPBE-vdW-DF\cite{2009JK_AM,vdW-DFa,vdW-DFb}, and HSE06\cite{HSE06} functionals.
%We evaluated the relative energies of structures relaxed by every functional by PBE-DFT.
%%as shown in Figure \ref{fig.geomErr}.
%We found that the differences of the relative energies
%among the structures relaxed by different functionals are at most 0.018 eV/f.u.
%%the differences of the time-series in Figure \ref{fig.geomErr} are within 0.018 eV/f.u.

%%----------------------------------------
%\begin{figure}[htbp]
%  \centering
%  \includegraphics[width=\hsize]{geomErr.eps}
%  \caption{\ghost{fig.geomErr}\label{fig.geomErr}
%    FNDMC and DFT results of relative energies of the four candidate structures
%    of LiV$X_2$ (X = O, S, and Se).
%    The candidate structures are shown in Figure \ref{fig.str}.
%    TR-I and TR-II (left from the vertical line) are trimmerized
%    and NTR-I and NTR-II (right from the vertical line) are
%    non-trimmerized structures, respectively.
%    FNDMC results are same as Figure \ref{fig.dmconly}.
%    The lowest energy among the four structures is set to be zero
%    and the others are given as relative values from the lowest energy.
%  }
%\end{figure}
%%----------------------------------------

### 2024-08-25 doc/paper.tex
draft 1, result, conclusion are not done yet

@@
-\documentclass[aps,prb,groupedaddress,twocolumn,showpacs,dvipdfmx, superscriptaddress,pdftex]{revtex4-2}
+\documentclass[aps,prb,groupedaddress,twocolumn,showpacs,dvipdfmx,superscriptaddress,pdftex]{revtex4-2}
 \usepackage[whole]{bxcjkjatype}
@@
 \usepackage[usenames,dvipsnames]{color}
+\usepackage{tabularray}
+\usepackage{fvextra}
 % \usepackage{pdfpages}
@@
 \title{
-  Meta-learning for multi foreign exchange movement prediction
+    Meta-learning in movement prediction problem of aperiodic time-series data
 }
@@
 \email{mwklng2309@icloud.com}
-\affiliation{School of Information Science, JAIST, 1-1 Asahidai,
-  Nomi, Ishikawa, 923-1292, Japan}
+\affiliation{School of Information Science, JAIST, Ishikawa, Japan}
 %
@@
 \email{ichibha@icloud.com}
-\affiliation{School of Information Science, JAIST, 1-1 Asahidai,
-  Nomi, Ishikawa, 923-1292, Japan}
+\affiliation{School of Information Science, JAIST, Ishikawa, Japan}
+%
+\author{Kenta Hongo}
+\email{kenta_hongo@mac.com}
+\affiliation{Research Center for Advanced Computing Infrastructure, JAIST, Ishikawa, Japan}
 %
@@
 \email{rmaezono@mac.com}
-\affiliation{School of Information Science, JAIST, 1-1 Asahidai,
-  Nomi, Ishikawa, 923-1292, Japan}
+\affiliation{School of Information Science, JAIST, Ishikawa, Japan}
 %
@@
     \textbf{Abstract}.
-Xử lý bài toán dự đoán các chỉ số liên quan đến tài chính bằng machine learning gặp rất nhiều khó khăn vì loại dữ liệu này không có chu kỳ rõ ràng và không chỉ phụ thuộc vào lịch sử giao dịch, mà còn phụ thuộc rất nhiều vào tình hình kinh tế, chính trị, điều mà rất khó capture trong tập dữ liệu.
-Để vượt qua các thách thức này, chúng tôi sử dụng Meta-learning, kết hợp với \verb|LSTM| và \verb|CNN| để tổng hợp hiệu quả các đặc trưng ẩn của dữ liệu theo thời gian.
-Thực nghiệm trên bài toán dự đoán xu hướng tỷ giá ngoại hối của 15 cặp tiền tệ cho thấy phương pháp của chúng tôi đạt hội tụ ở mức cao khi so sánh với mô hình \verb|NHITS|.
-% đoạn này cần sửa lại xíu
+Dự đoán aperiodic time-series data (e.g.
+stock price, foreign exchange, Bitcoin price,...) là một tác vụ khó khăn đối với các mô hình học máy bởi vì loại dữ liệu này có phương sai cao và không cố định; không thể hiện chu kỳ rõ ràng nên khó rút trích đặc trưng; không chỉ phụ thuộc vào giá trị quá khứ mà còn phụ thuộc các yếu tố khác, điều khiến chúng bất ổn và không có chu kỳ như tình hình kinh tế, chính trị.
+Để khắc phục các thách thức nêu trên, chúng tôi sử dụng Meta-learning để huấn luyện kết hợp mạng \verb|LSTM| và \verb|CNN|, từ đó rút trích và tổng hợp hiệu quả các đặc trưng ẩn của dữ liệu theo thời gian.
+Thực nghiệm trên dữ liệu foreign exchange của 15 cặp tiền tệ trong vòng 24 năm (2000-2024) cho thấy phương pháp đề xuất hoạt động tốt và có độ chính xác cao hơn mô hình \verb|NHITS| - state-of-the-art (SOTA) model của năm 2023 trên dữ liệu time-series, trong bài toán dự đoán xu hướng (upward or downward) của ngày giao dịch tiếp theo.
 
@@
 %----------------------------------------
-\keywords{Meta-learning, Machine leaning, \verb|LSTM|, \verb|CNN|, Time-series, Foreign Exchange}
+\keywords{Meta-learning, \verb|LSTM|, \verb|CNN|, Aperiodic time-series data, Foreign exchange}
 \maketitle
@@
 \section{Introduction}
+\label{sec.intro}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-Foreign exchange prediction nói riêng và dự đoán các chỉ số tài chính như giá chứng khoán nói chung từ lâu đã là vấn đề đáng quan tâm của nhiều nghiên cứu \citep{li2019multi, islam2021foreign, heryadi2021foreign}.
-Hai kỹ thuật chính được sử dụng trong financial time-series prediction là fundamental analysis and technical analysis \cite{ayitey2023forex}.
-Trong khi fundamental analysis thiên về phân tích các yếu tố như chính sách, chiến lược kinh tế của công ty, quốc gia để dự đoán tương lai; technical analysis dựa hoàn toàn vào lịch sử biến động giá để phân tích xu hướng thị trường.
+Aperiodic time-series prediction nói chung hay foreign exchange (FX) prediction nói riêng từ lâu đã là vấn đề đáng quan tâm của nhiều nghiên cứu \citep{li2019multi, islam2021foreign, heryadi2021foreign}.
+Hai kỹ thuật chính được sử dụng trong aperiodic time-series prediction là fundamental analysis and technical analysis \cite{ayitey2023forex}.
+Trong khi fundamental analysis thiên về phân tích các yếu tố tác động từ bên ngoài và khó có thể capture từ các biến thiên giá trị trong quá khứ như chính sách, chiến lược kinh tế của công ty, quốc gia để dự đoán tương lai; technical analysis dựa hoàn toàn vào lịch sử biến động giá trị để phân tích xu hướng tương lai.
 
@@
 
-Việc dự đoán các chỉ số tài chính này gặp một vài thách thức cố hữu.
-Trong đó có thể kể tới: (1) - Variance của loại dữ liệu này không hề cố định, từ đó giả thuyết rằng chúng tuân theo một phân phối để có thể xấp xỉ lỗi là không thể sử dụng được, dẫn đến việc các mô hình học máy khó có thể dự đoán chính xác giá cũng như xu hướng tăng giảm trong tương lai; (2) - Các chỉ số tài chính phần lớn không tuân theo bất kỳ quy luật rõ ràng nào nên việc học các đặc trưng ẩn của dữ liệu để tiến hành dự đoán gặp rất nhiều khó khăn; (3) - Một chỉ số tài chính (e.g.
-giá cổ phiếu của một công ty) không hoàn toàn phụ thuộc vào dữ liệu quá khứ mà còn phụ thuộc nhiều vào các tin tức, tình hình kinh tế, chính trị \cite{li2019multi}.
+Việc dự đoán trên aperiodic time-series data gặp một vài thách thức cố hữu.
+Trong đó có thể kể tới: (1) - Variance của loại dữ liệu này biến thiên mạnh qua các thời kỳ.
+Từ đó giả thuyết rằng chúng tuân theo một phân phối để có thể xấp xỉ lỗi là không thể sử dụng được, dẫn đến việc các mô hình học máy khó có thể dự đoán chính xác giá trị cũng như xu hướng trong tương lai; (2) - Aperiodic data không tuân theo bất kỳ quy luật rõ ràng nào nên việc học các đặc trưng ẩn của dữ liệu để tiến hành dự đoán gặp rất nhiều khó khăn; (3) - Aperiodic time-series data (e.g.
+giá cổ phiếu của một công ty) không hoàn toàn phụ thuộc vào dữ liệu quá khứ mà còn phụ thuộc nhiều yếu tố ngoại lai (e.g.
+tin tức, tình hình kinh tế, chính trị) \cite{li2019multi}.
 
@@
 Đối với thách thức đầu tiên, các mô hình ensemble [cite something here] thường được sử dụng để hạn chế ảnh hưởng của sự biến đổi variance.
-Chúng tôi cũng tiếp cận bài toán theo hướng này nhưng ở mức cao hơn bằng cách sử dụng Meta-learning (ML) \cite{finn2017model}.
+Ensemble model giúp cung cấp một cái nhìn tổng quát, tổng hợp từ nhiều khía cạnh dựa trên các sub-models, từ đó giúp mô hình tổng quát thích ứng được với sự thay đổi mạnh của variance.
+Chúng tôi tiếp cận bài toán theo hướng này nhưng ở mức cao hơn bằng cách sử dụng Meta-learning (ML) \cite{finn2017model}.
 Phương pháp này tổng hợp hiệu quả tham số của các mô hình cục bộ, giúp giảm thiểu đáng kể variance loss.
@@
 
-Để khắc phục thách thức thứ hai, các nghiên cứu phần lớn sử dụng các đặc trưng rút trích được từ mô hình Long short-term memory neural network (\verb|LSTM|) \cite{hochreiter1997long}, Artificial neural network (\verb|ANN|), và Convolution neural network (\verb|CNN|) \cite{lecun1989handwritten}.
+Để khắc phục thách thức thứ hai, các nghiên cứu phần lớn sử dụng các đặc trưng rút trích được từ Long short-term memory neural network (\verb|LSTM|) \cite{hochreiter1997long}, Artificial neural network (\verb|ANN|), và Convolution neural network (\verb|CNN|) \cite{lecun1989handwritten}.
 Cụ thể, trong năm 2022, 20\% tổng số các bài báo liên quan đến dự đoán chỉ số tài chính sử dụng \verb|LSTM|, 20\% sử dụng \verb|ANN| và 6\% sử dụng \verb|CNN| \cite{ayitey2023forex}.
@@
 
-Đối với thách thức thứ ba, nghiên cứu \cite{fama1970efficient} đưa ra giả thuyết rằng chỉ số tài chính của các công ty ngầm phản ánh tình hình kinh tế tài chính trên thị trường.
-Các nghiên cứu \citep{overreactioncontrarian, mech1993portfolio} cũng chỉ ra sự phụ thuộc giữa chỉ số tài chính của một công ty nhất định và các chỉ số của các công ty khác.
+Đối với thách thức thứ ba, nghiên cứu \cite{fama1970efficient} đưa ra giả thuyết rằng các time-series datasets khác nhau của cùng một lĩnh vực tại cùng thời điểm phản ánh sự tác động của các yếu tố ngoại lai.
+Ví dụ, các nghiên cứu \citep{overreactioncontrarian, mech1993portfolio} cũng chỉ ra sự phụ thuộc giữa chỉ số tài chính của một công ty nhất định và các chỉ số của các công ty khác.
 Điều này càng làm tăng tính đúng đắn của giả thuyết trong nghiên cứu \cite{fama1970efficient}.
-Ngoài ra, chúng tôi cho rằng, các chỉ số tài chính còn có những phụ thuộc ngầm vào các thời điểm nhất định trong quá khứ (hidden-long-term dependency).
+Ngoài ra, chúng tôi cho rằng, loại dữ liệu này còn có những phụ thuộc ngầm vào các thời điểm nhất định trong quá khứ (hidden long-term dependency).
 Đối với cách tiếp cận truyền thống, người ta sử dụng một lượng dữ liệu quá khứ cố định (lookback window) để huấn luyện mô hình.
@@
 
-Cuối cùng, chúng tôi chứng minh tính ưu việt của thuật toán đề xuất bằng cách giải bài toán dự đoán xu hướng (tăng hoặc giảm) của tỉ giá ngoại hối và so sánh với mô hình SOTA hiện tại (\verb|NHITS| \cite{challu2023nhits}) trên hai loại dữ liệu: (1) - Dữ liệu tỉ giả của cặp tiền tệ USD/JPY; (2) - Dữ liệu tỷ giá của 60 cặp tiền tệ, cấu thành từ các nước Australia, Canada, Switzerland, Denmark, EU, United Kingdom, Hong Kong, Iceland, Japan, Norway, New Zealand, Singapore, Sweden, Turkey, United States, Mexico, China, South Africa.
+Cuối cùng, chúng tôi chứng minh tính ưu việt của thuật toán đề xuất bằng cách giải bài toán dự đoán xu hướng (tăng hoặc giảm) của tỉ giá ngoại hối và so sánh kết quả với mô hình SOTA hiện tại (\verb|NHITS| \cite{challu2023nhits}) trên hai loại dữ liệu: (1) - Dữ liệu tỉ giả của cặp tiền tệ USD/JPY; (2) - Dữ liệu tỷ giá của 60 cặp tiền tệ, cấu thành từ 18 quốc gia.
 Các tập dữ liệu này được công khai trên Internet và có thể tải về dễ dàng.
@@
 \begin{itemize}
+    \item Kết hợp đặc trưng: Rút trích đặc trưng bằng cách kết hợp các đặc trưng của \verb|LSTM| và \verb|CNN|.
+    \item Hidden long-term dependency: Chứng minh thực nghiệm rằng một aperiodic time-series data nhất định không chỉ phụ thuộc vào các yếu tố ngoại lai mà còn có các phụ thuộc ẩn với chính nó tại nhiều thời điểm khác nhau trong quá khứ.
     \item Tổng hợp hiệu quả tham số mô hình: Sử dụng ML thay cho các mô hình ensemble truyền thống trong việc tổng hợp kết quả từ các mô hình học máy.
-  \item Kết hợp đặc trưng: Rút trích đặc trưng bằng cách kết hợp các đặc trưng của \verb|LSTM| và \verb|CNN|.
-  \item Hidden-long-term dependency: Chứng minh thực nghiệm rằng một chỉ số tài chính không chỉ phụ thuộc vào các chỉ số tài chính khác mà còn có các phụ thuộc ẩn với chính nó.
-  \item Experiment: Thực nghiệm trên các bộ dữ liệu về tỉ giá hối đoái và so sánh với mô hình SOTA NHITS để chứng minh tính hiệu quả của phương pháp đề xuất.
+    \item Experiment: Thực nghiệm trên các bộ dữ liệu về tỉ giá hối đoái và so sánh với mô hình SOTA \verb|NHITS| để chứng minh tính hiệu quả của phương pháp đề xuất.
 \end{itemize}
@@
 
-\subsection{LSTM \& CNN in feature extraction}
+\subsection{LSTM \& CNN model}
 
-Như đã đề cập, \verb|LSTM| là mạng neural rất phổ biến trong việc handle các bài toán liên quan đến dự đoán chỉ số tài chính nói riêng cũng như các bài toán trên dữ liệu time-series nói chung.
-\verb|LSTM| được sử dụng phổ biến như vậy bởi vì nó xử lý tốt vấn đề vanishing gradient và có thể khai thác các mối quan hệ phi tuyến trong dữ liệu.
-Thật vậy, bằng cách duy trì cell-state trong mỗi iteration, \verb|LSTM| có thể chống lại một cách hiệu quả vấn đề vanishing gradient, từ đó bảo toàn khả năng capture các phụ thuộc dài hạn \cite{cheng2018applied}.
-\verb|LSTM| cũng thực hiện rút trích đặc trưng với các hàm kích hoạt phi tuyến, giúp học được các tham số mô hình có thể capture tính phi tuyến của dữ liệu \cite{he2016exploiting}.
+Như đã đề cập, \verb|LSTM| là mạng neural rất phổ biến trong việc handle các bài toán liên quan đến dự đoán trên dữ liệu time-series.
+\verb|LSTM| được sử dụng phổ biến như vậy bởi vì nó xử lý tốt vấn đề vanishing gradient (dễ dàng bắt gặp khi sử dụng Recurrent neural network) và có thể khai thác hiệu quả các mối quan hệ phi tuyến trong dữ liệu.
+Thật vậy, bằng cách duy trì cell-state trong mỗi iteration, \verb|LSTM| có thể khắc phục vấn đề vanishing gradient, từ đó bảo toàn khả năng capture các phụ thuộc dài hạn \cite{cheng2018applied}.
+Ngoài ra, \verb|LSTM| thực hiện rút trích đặc trưng với các hàm kích hoạt phi tuyến, giúp các tham số mô hình có thể capture tính phi tuyến của dữ liệu \cite{he2016exploiting}.
 Hai yếu tố nêu trên khiến cho \verb|LSTM| trở thành lựa chọn đầu tiên được nghĩ đến khi giải quyết các bài toán trên dữ liệu time-series.
@@
 
-\verb|CNN| được sử dụng rất nhiều trong các tác vụ xử lý hình ảnh \citep{naranjo2020review, sharma2018analysis} bởi khả năng tổng hợp các quan hệ cục bộ.
+Mạng \verb|CNN| được sử dụng rất nhiều trong các tác vụ xử lý hình ảnh \citep{naranjo2020review, sharma2018analysis} bởi khả năng tổng hợp các quan hệ cục bộ.
 Không chỉ vậy, \verb|CNN| còn được dùng rất nhiều trong các tác vụ xử lý dữ liệu time-series như speech recognition \cite{dua2022developing}, natural language processing \cite{varshitha2023natural}.
-Điều đó chứng tỏ được khả năng của \verb|CNN| trong việc khám phá mối quan hệ thời gian giữa các mẫu dữ liệu.
-Mặc dù vậy, \verb|CNN| lại rất ít được dùng trong các tác vụ dự đoán chỉ số tài chính.
+Điều đó chứng tỏ được khả năng của \verb|CNN| trong việc khám phá mối quan hệ thời gian cục bộ giữa các mẫu dữ liệu.
+Mặc dù vậy, \verb|CNN| lại rất ít được dùng trong các tác vụ dự đoán stock price hay foreign exchange.
 Trong nghiên cứu này, chúng tôi tận dụng khả năng rút trích đặc trưng cục bộ tuyệt vời của \verb|CNN| để tích hợp thêm thông tin ẩn vào quá trình huấn luyện của mô hình.
 
-\subsection{Meta-learning}
+\subsection{Model-agnostic Meta-learning (MAML)}
 
-Các thuật toán Meta-learning (ML), điển hình là Model-agnostic meta-learning (MAML) \cite{finn2017model} được biết đến với khả năng huấn luyện một mô hình có tính tổng quát cao, thích ứng nhanh trên tập dữ liệu mới thông qua một lượng nhỏ dữ liệu và số bước huấn luyện \citep{hospedales2021meta, vettoruzzo2024advances}.
-Với khả năng này, ML được sử dụng rất nhiều trong việc cải thiện tốc độ đáp ứng của mô hình trên dữ liệu như cá nhân hóa mô hình học \cite{fallah2020personalized}, giải quyết vấn đề domain shift [insert citation].
+Các thuật toán Meta-learning (ML), điển hình là MAML \cite{finn2017model} được biết đến với khả năng huấn luyện một mô hình có tính tổng quát cao, thích ứng nhanh trên tập dữ liệu mới thông qua một lượng nhỏ dữ liệu và số bước huấn luyện \citep{hospedales2021meta, vettoruzzo2024advances}.
+Với khả năng này, ML được sử dụng rất nhiều trong các tác vụ đòi hỏi khả năng đáp ứng của mô hình trên dữ liệu (e.g.
+cá nhân hóa mô hình học \citep{chen2018federated, fallah2020personalized,nguyen2022meta}, domain adaptation trong online learning \citep{hu2023meta, khoee2024domain}).
 
@@
 \begin{equation}
-  \theta_t^* = \theta_t(\phi) = \arg\min_{\theta}{\mathcal{L}_t\left( \phi, \mathcal{D}_t^{support} \right)}
+    \theta_t^* = \theta_t(\phi) = \arg\min_{\theta}{\mathcal{L}^{task}_t\left( \phi, \mathcal{D}_t^{support} \right)}
     \label{eq:inner_opt}
@@
 \begin{align*}
-  \phi^* &= \arg\min_{\phi}\sum_{t}{\mathcal{L}_t\left[ \theta_t^*, \mathcal{D}_t^{query} \right]}\\
-  &= \arg\min_{\phi}\sum_{t}{\mathcal{L}_t\left[ \theta_t(\phi), \mathcal{D}_t^{query} \right]} \numberthis
+    \phi^* &= \arg\min_{\phi}\sum_{t}{\mathcal{L}^{meta}_t\left[ \theta_t^*, \mathcal{D}_t^{query} \right]}\\
+    &= \arg\min_{\phi}\sum_{t}{\mathcal{L}^{meta}_t\left[ \theta_t(\phi), \mathcal{D}_t^{query} \right]} \numberthis
     \label{eq:outer_opt}
@@
 Bằng hình thức huấn luyện trên, mô hình $\phi^*$ sẽ có mức tổng quát hóa cao trên các tác vụ khác nhau, có thể nhanh chóng đáp ứng một tác vụ mới chỉ sau một vài bước huấn luyện.
-Tận dụng ưu điểm này, chúng tôi
 
@@
 Các mô hình hybrid ensemble vốn được sử dụng rất nhiều trong các bài toán xử lý time-series và được chứng minh thực nghiệm là có độ chính xác cao hơn so với các mô hình handle time-series data tiêu chuẩn vì có thể tổng hợp được sức mạnh của nhiều mô hình \cite{ayitey2023forex}.
-Tuy vậy, các hình thức tổng hợp của ensemble model hiện nay vẫn còn rất cứng nhắc vì chỉ có thể tổng hợp dựa trên kết quả cuối (đối với bagging models) và kết quả gần cuối (đối với stacking models).
+Tuy vậy, các hình thức tổng hợp của ensemble model hiện nay vẫn còn rất cứng nhắc vì chỉ có thể tổng hợp dựa trên kết quả cuối (cơ chế voting của bagging models) và kết quả gần cuối (đối với stacking models).
 Dưới góc nhìn của ensemble model, có thể coi phương trình \ref{eq:outer_opt} là một phương pháp tổng hợp hiệu quả các sub-model, giúp tận dụng khả năng rút trích đặc trưng của từng mô hình.
@@
 \begin{align}
-  \mathbf{y}_{t-L:t, l}^{(p)} &= Pooling\left( \mathbf{y}_{t-L:t, l-1} \right)\\
+    \mathbf{y}_{t-L:t, l}^{(p)} &= \mathbf{Pooling}\left( \mathbf{y}_{t-L:t, l-1} \right)\\
     % \mathbf{h}_l &= FullyConnected\left( \mathbf{y}_{t-L:t, l}^{(p)} \right)\\
-  \mathbf{\theta}_l^b &= FullyConnected^b \left( \mathbf{y}_{t-L:t, l}^{(p)} \right)\\
-  \mathbf{\theta}_l^f &= FullyConnected^f \left( \mathbf{y}_{t-L:t, l}^{(p)} \right)\\
-  \mathbf{\hat{y}}_{t-L:t, l} &= g(\mathbf{\theta}_l^b)\\
-  \mathbf{\hat{y}}_{t+1:t+H, l} &= g(\mathbf{\theta}_l^f)
+    \mathbf{\theta}_l^b &= \mathbf{FullyConnected}^b \left( \mathbf{y}_{t-L:t, l}^{(p)} \right)\\
+    \mathbf{\theta}_l^f &= \mathbf{FullyConnected}^f \left( \mathbf{y}_{t-L:t, l}^{(p)} \right)\\
+    \mathbf{\hat{y}}_{t-L:t, l} &= g\left(\mathbf{\theta}_l^b\right)\\
+    \mathbf{\hat{y}}_{t+1:t+H, l} &= g\left(\mathbf{\theta}_l^f\right)
 \end{align}
@@
 \begin{align}
-  \mathbf{y}_{t-L:t, l+1} = \mathbf{y}_{t-L:t, l} - \mathbf{\hat{y}}_{t+1:t+H, l}
+    \mathbf{y}_{t-L:t, l+1} = \mathbf{y}_{t-L:t, l-1} - \mathbf{\hat{y}}_{t-L:t, l}
     \label{eq:input_l1}
@@
 
-Tổng hợp các giá trị forecast của các block, ta được giá trị forecast của một stack.
+Giả sử mô hình gồm có $S$ stacks, mỗi stack có $B$ blocks.
+Tổng hợp các giá trị forecast của các block như phương trình \ref{eq:sum_block}, ta được giá trị forecast của một stack.
 Backcast của block cuối cùng của một stack chính là đầu vào cho stack tiếp theo.
-Cuối cùng, tổng hợp giá trị forecast của các stack, ta được giá trị forecast dự đoán của toàn mạng.
+Cuối cùng, tổng hợp giá trị forecast của các stack như phương trình \ref{eq:sum_stack}, ta được giá trị forecast dự đoán của toàn mạng.
 
-Bằng cách xếp chồng các stack, stack sau nhận vào phần dư của stack trước, kiến trúc trên được kỳ vọng là sẽ phân rã dữ liệu thành các frequency bands khác nhau.
-Trong thực tế, \verb|NHITS| perform rất tốt đối với các bộ dữ liệu có tính chu kỳ cao.
-Tuy nhiên, tính chu kỳ của dữ liệu foreign exchange là rất thấp, thậm chí không có, gây ra khó khăn rất lớn cho \verb|NHITS|.
-Bằng cách xếp chồng các stack, stack sau nhận vào phần dư của stack trước, kiến trúc trên được kỳ vọng là sẽ phân rã dữ liệu thành các frequency bands khác nhau.
-Trong thực tế, \verb|NHITS| perform rất tốt đối với các bộ dữ liệu có tính chu kỳ cao.
-Tuy nhiên, tính chu kỳ của dữ liệu foreign exchange là rất thấp, thậm chí không có, gây ra khó khăn rất lớn cho \verb|NHITS|.
+\begin{align}
+    \mathbf{\hat{y}}_{t+1:t+H}^s &= \sum_{l=1}^{B}{\mathbf{\hat{y}}_{t+1:t+H, l}} \label{eq:sum_block}\\
+    \mathbf{\hat{y}}_{t+1:t+H} &= \sum_{s=1}^{S}{\mathbf{\hat{y}}_{t+1:t+H}^s} \label{eq:sum_stack}
+\end{align}
 
-% \begin{figure}
-%   \large
-%   \centering
-%   \includegraphics[witdth=\linewidth]{./img/meo.pdf}
-%   % \caption{This is a caption \cite{challu2023nhits}}
-% \end{figure}
+Bằng cách xếp chồng các stack, stack sau nhận vào phần dư của stack trước, kiến trúc trên được kỳ vọng là sẽ phân rã dữ liệu thành các frequency bands khác nhau (weekly, daily, even hourly).
+Trên thực tế, \verb|NHITS| perform rất tốt đối với các bộ dữ liệu có tính chu kỳ cao như mức tiêu thụ điện, thời tiết, giao thông.
+Tuy nhiên, chúng tôi đang hướng đến aperiodic time-series dataset, vốn có tính chu kỳ rất thấp, thậm chí không có (see figure \ref{fig:fx}).
+Điều này gây ra khó khăn rất lớn cho \verb|NHITS|.
 
-\begin{figure}
+\begin{figure}[h]
     \centering
-  \includegraphics[width=\linewidth]{img/meo.pdf}
-  \caption{Your figure caption here.}
-  \label{fig:your-label}
+    \includegraphics[width=\linewidth]{img/CHF_NZD.pdf}
+    \caption{Exchange rate (close price) between Swiss franc and New Zealand dollar by day (2014-2024).}
+    \label{fig:fx}
 \end{figure}
-
-% \mathbf{y}_{t-L:t, l+1} &= \mathbf{y}_{t-L:t, l} - \mathbf{\hat{y}}_{t+1:t+H, l}
-
-% nói về ưu điểm, nhược điểm của fundamental analysis and technical analysis --> đây là nói chung, sẽ có phần nói riêng sau (nói về từng phương pháp, dẫn đến phương pháp của tôi)
-
-% tình hình kinh tế, chính trị, những thứ khó có thể cấu trúc hóa thành dữ liệu cho máy học.
-
-%----------------------------------------
-% \newcommand{\scale}{0.48}
-% \begin{figure}[htbp]
-%   \large
-%   \centering
-%   (a) I-type stacking of insulating TR lattice \\
-%   \begin{minipage}{\scale\hsize}
-%     \includegraphics[width=\hsize]{tri_off_top.eps}
-%   \end{minipage}
-%   \begin{minipage}{\scale\hsize}
-%     \includegraphics[width=\hsize]{tri_off_side.eps}
-%   \end{minipage}
-
-% 	\vspace{5mm}
-%   (b) II-type stacking of insulating TR lattice \\
-%   \begin{minipage}{\scale\hsize}
-%     \includegraphics[width=\hsize]{tri_on_top.eps}
-%   \end{minipage}
-%   \begin{minipage}{\scale\hsize}
-%     \includegraphics[width=\hsize]{tri_on_side.eps}
-%   \end{minipage}
-
-% \vspace{5mm}
-%   (c) I-type stacking of metallic NTR lattice \\
-%   \begin{minipage}{\scale\hsize}
-%     \includegraphics[width=\hsize]{ntr_off_top.eps}
-%   \end{minipage}
-%   \begin{minipage}{\scale\hsize}
-%     \includegraphics[width=\hsize]{ntr_off_side.eps}
-%   \end{minipage}
-
-% \vspace{5mm}
-%   (d) II-type stacking of metallic NTR lattice \\
-%   \begin{minipage}{\scale\hsize}
-%     \includegraphics[width=\hsize]{ntr_on_top.eps}
-%   \end{minipage}
-%   \begin{minipage}{\scale\hsize}
-%     \includegraphics[width=\hsize]{ntr_on_side.eps}
-%   \end{minipage}
-%   \caption{\ghost{fig.str}\label{fig.str}
-%     Structure models of LiV$X_2$ (X = O, S, and Se).
-%     The left figure is top-view and the right figure is side-view.
-%     The green balls are Li, the blue balls are V, and red balls are O.
-%   }
-% \end{figure}
-%----------------------------------------
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-Fixed-node DMC
+
+% nói tổng quan ở đây, vẽ cái hình và bỏ vào
+Phương pháp của chúng tôi bao gồm hai phần chính hoạt động song song nhau: (1) - Feature extraction; (2) - Parameter synthesis.
+Tổng quan phương pháp được minh họa trong hình \ref{fig:flow}.
+Trong phần feature extraction, chúng tôi kết hợp hai loại đặc trưng từ mạng \verb|CNN| và \verb|LSTM|.
+Trong phần parameter synthesis, chúng tôi sử dụng \verb|MAML| để tổng hợp tham số của các mô hình.
+Với sự góp mặt của các đặc trưng \verb|LSTM| và \verb|CNN|, chúng tôi kỳ vọng sẽ rút trích được các đặc trưng ẩn trong dữ liệu aperiodic.
+Bằng việc sử dụng \verb|MAML| trong quá trình tổng hợp trọng số, phương pháp đề xuất được kỳ vọng là một giải pháp thay thế hợp lý và hiệu quả cho các mô hình ensemble truyền thống trong việc giảm thiểu tác động của sự biến thiên variance, tổng hợp hiệu quả các yếu tố ngoại lai, cũng như giữ được các hidden long-term dependency ẩn trong quá khứ.
+
+\begin{figure*}[ht]
+    \centering
+    \includegraphics[width=\textwidth]{img/meta.png}
+    \caption{The full-flow of meta-training and meta-testing on multi-fx data. Each currency pair is regarded as a task.}
+    \label{fig:flow}
+\end{figure*}
+
+\subsection{Data preparation}
+
+Phương pháp đề xuất sử dụng các thuật toán ML để huấn luyện mô hình.
+Do đó, dữ liệu cần được tổ chức lại để các thuật toán ML có thể hoạt động được.
+Trong trường hợp dữ liệu bao gồm nhiều datasets khác nhau thuộc cùng một lĩnh vực, mỗi dataset sẽ được coi là một task của \verb|MAML|.
+Trong trường hợp dữ liệu bao gồm một dataset duy nhất, cần chia nhỏ dataset này thành các tập con ứng với các task riêng biệt.
+Tóm lại, tập dữ liệu sau khi chuẩn bị bao gồm $n$ tasks: $\mathcal{D} = \left\{ \mathcal{D}_t \right\}_{t=1}^{n}$.
+Dữ liệu tại mỗi task được chia thành tập support và query: $\mathcal{D}_t = \left\{ \mathcal{D}_t^{support}, \mathcal{D}_t^{query} \right\}$.
 
 \vspace{2mm}
-Fixed-node DMC
-%----------------------------------------
-\begin{equation}
-\left[\sum_j\frac{1}{2}\nabla^2_j
-+ V_{\rm MB}\left(\vec R\right)
-\right]\Psi\left(\vec R\right)
-= E~\Psi\left(\vec R\right) \ .
-\end{equation}
-%----------------------------------------
-Fixed-node DMC
 
-%----------------------------------------
-\begin{equation}
-\left[\cfrac{1}{2}~\nabla^2_j
-+ v_{\rm ext}\left(\vec r_j\right)
-+ v_{\rm XC}\left(\vec r_j\right)
-\right]\psi\left(\vec r_j\right)
-= \varepsilon~\psi\left(\vec r_j\right) \ .
-\end{equation}
-%----------------------------------------
-Fixed-node DMC
+Một sample dữ liệu bao gồm các cặp giá trị $(\mathbf{x}_{t-L:t}, y)$.
+Trong đó, $\mathbf{x}_{t-L:t}$ bao gồm $L$ giá trị lịch sử tính từ thời điểm $t$ trở về trước; $y\in \{0,1\}$ là nhãn dữ liệu, thể hiện xu hướng giảm, hoặc tăng của mẫu dữ liệu $x_{t+1}$ so với $x_{t}$.
+Tùy vào từng bài toán và cách cài đặt mà các phần tử trong $\mathbf{x}_{t-L:t}$ có thể là các vector hoặc các scalar number.
+Ví dụ, đối với dữ liệu chứng khoán, $\mathbf{x}_{t-L:t}$ có thể chứa các vector dữ liệu $\vec x_i = (\text{open, low, high, close})$ hoặc chỉ một giá trị close price duy nhất.
+
+% nói từng cái, có thể sẽ vẽ hình cụ thể cho từng cái
+% method phải chỉ ra được đoạn nào giải quyết thách thức nào (đã nêu ở intro)
+
+% đến đoạn thực nghiệm thì sẽ mô tả sâu quá trình tìm kiếm tham số, quá trình chia dữ liệu, tải dữ liệu ở phụ lục
+% nhận xét: nhấn mạnh vào việc giải quyết 3 thách thức nêu ở intro
+
+\subsection{Feature extraction}
+
+Lấy cảm hứng từ nghiên cứu \cite{vo2017multi}, chúng tôi đề xuất kết hợp các đặc trưng rút trích được từ mạng \verb|LSTM| và \verb|CNN|.
+Cụ thể, chúng tôi đưa từng phần tử trong vector $\mathbf{x}_{t-L:t}$ qua một lớp \verb|MLP| có đầu ra lớn hơn số chiều của $\vec x_i, i\in[t-L, t]$ để phân giải thành các đặc trưng nhỏ $\vec x'_i$.
+Các đặc trưng này sau đó được truyền qua mạng \verb|LSTM| và \verb|CNN| để lần lượt rút trích các phụ thuộc thời gian dài hạn ($\mathbf{h}_{LSTM}$) và các đặc trưng thời gian cục bộ ($\mathbf{h}_{CNN}$).
+Để có thể khai thác tối đa các ràng buộc thời gian dài hạn, chúng tôi sử dụng \verb|BidirectionalLSTM| để rút trích từ hai phía của của $\mathbf{x}_{t-L:t}$.
+Toàn bộ quy trình rút trích đặc trưng được tóm tắt như sau:
+
+\begin{align}
+    \mathbf{x'}_{t-L:t} &= \mathbf{FullyConnected}\left( \mathbf{x'}_{t-L:t} \right)\\
+    \mathbf{h}_{LSTM} &= \mathbf{BidirectionalLSTM}\left( \mathbf{x'}_{t-L:t} \right)\\
+    \mathbf{h}_{CNN} &= \mathbf{Convolution1D}\left( \mathbf{x'}_{t-L:t} \right)
+    \label{eq:feature}
+\end{align}
 
 \vspace{2mm}
-Fixed-node DMC
-%----------------------------------------
-\begin{align}
-  E\left[\Psi_T\right]
-  &=\frac{\int d\vec{R}\cdot \Psi_T^*\left(\vec{R}\right)
-  \hat{H}\Psi_T\left(\vec{R}\right)}{\int d\vec{R}\cdot \Psi_T^*\left(\vec{R}\right)\Psi_T\left(\vec{R}\right)} ,
-  \label{eq.varfunc}
-\end{align}
-%----------------------------------------
-\ghost{eq.varfunc}
-Fixed-node DMC
+
+Mạng \verb|LSTM| duy trì giá trị cell-state nhằm lưu trữ có chọn lọc các phụ thuộc dài hạn.
+Điều này rất thích hợp trong việc giải quyết các bài toán time-series data.
+Mặt khác, giá trị tương lai thường phụ thuộc rất lớn vào các giá trị lịch sử gần nhất.
+Chúng tôi đề xuất sử dụng mạng \verb|CNN| để nhấn mạnh các đặc trưng cục bộ, từ đó hướng một phần sự chú ý của mô hình vào các thời điểm nhất định.
+Do đó, phương pháp đề xuất không chỉ nhớ được các đặc trưng long-term mà còn highlight được các đặc trưng short-term.
 
 \vspace{2mm}
-Fixed-node DMC
+
+Tiếp đến, $\mathbf{h}_{LSTM}$ và $\mathbf{h}_{CNN}$ được nối với nhau (phương trình \ref{eq:concat}) sau đó chuyển đến phần phân lớp của mạng NN (phương trình \ref{eq:clf}).
+
+\begin{align}
+    \mathbf{h}_{t-L:t} &= \mathbf{Concatenate}\left( \mathbf{h}_{LSTM}, \mathbf{h}_{CNN} \right) \label{eq:concat} \\
+    \hat y &= \mathbf{FullyConnected}\left( \mathbf{h}_{t-L:t} \right) \label{eq:clf}
+\end{align}
+
+\subsection{Effective synthesis\\of models' parameters}
+
+Chúng tôi sử dụng \verb|MAML| để huấn luyện và tổng hợp trọng số của các mô hình tại các task.
+Như đã đề cập trong phần \ref{sec.relatedWork}, tối ưu tham số theo cách tiếp cận của ML chính là đi giải hai phương trình \ref{eq:inner_opt} và \ref{eq:outer_opt} bằng các phương pháp tối ưu trên dữ liệu support và query.
+Cụ thể, quá trình tối ưu bao gồm nhiều bước toàn cục (outer optimization), thực hiện trên tất cả các tasks tham gia huấn luyện.
+Mỗi bước toàn cục bao gồm nhiều bước cục bộ (inner optimization) thực hiện trên từng task riêng lẻ.
+Tại bước toàn cục $r$, quá trình tối ưu cục bộ lần thứ $e$ tại tập support của task $t$ diễn ra như sau:
+
+\begin{align}
+    \begin{cases}
+        \theta_t^{(0)} &= \phi_{r-1} \\
+        \theta_t^{(e)} &= \theta_t^{(e-1)} - \alpha \nabla_{\theta} \mathcal{L}^{task}_t\left( \theta_t^{(e-1)}, \mathcal{D}_t^{support} \right)
+    \end{cases}
+\end{align} Trong đó, $\phi_{r-1}$ là kết quả của quá trình tối ưu toàn cục lần thứ $r-1$, $\alpha$ là inner learning rate.
+
+Tiếp đó, quá trình outer optimization tại bước toàn cục được thực hiện bằng cách tổng hợp độ lỗi trên tập query của các task và tối ưu trên đó (phương trình \ref{eq:outer_sol}).
+
+\begin{align}
+    \begin{cases}
+        \phi_0 = \text{Random Initialization}\\
+        \phi_r = \phi_{r-1} - \beta \nabla_{\phi} \sum_{t=1}^n{\mathcal{L}^{meta}_t \left( \theta_t^*(\phi), \mathcal{D}_t^{query} \right)}
+    \end{cases}
+    \label{eq:outer_sol}
+\end{align} Trong đó, $\beta$ là outer learning rate.
+Giả sử thuật toán chạy $E$ steps trong inner optimization, lượng đạo hàm tại phương trình \ref{eq:outer_sol} được viết lại như sau (the notations of dataset are removed):
+
+\begin{align*}
+    &\beta \nabla_{\phi} \sum_{t=1}^n{\mathcal{L}^{meta}_t \left( \theta_t - \alpha \nabla_{\theta} \mathcal{L}^{task}_t\left( \theta_t \right) \right)}\\
+    = &\beta \sum_{t=1}^n{ \frac{\partial \mathcal{L}^{meta}_t\left(\theta_t^{(E)}\right)}{\partial \theta_t^{(E)}} \frac{\partial \theta_t^{(E)}}{\partial \phi}}\\
+    = &\beta \sum_{t=1}^n{ \nabla_{\theta} \mathcal{L}^{meta}_t\left(\theta_t^{(E)}\right) \prod_{j=0}^{E-1} {\left[\mathbb{I} - \alpha\nabla^2_{\theta}\mathcal{L}^{task}_{t}\left(\theta_t^{(j)}\right)\right]}} \numberthis
+    \label{eq:outer_sol_}
+\end{align*}
+
+Sự xuất hiện của tích các đạo hàm bậc hai trong phương trình \ref{eq:outer_sol_} khiến quá trình đạo hàm trở nên phức tạp vì phải tốn rất nhiều chi phí để duy trì các ma trận Hessian.
+Do đó, số bước tính toán để tìm ra $\theta^*$ cần phải hạn chế.
+Trên thực tế, các phương pháp sử dụng ML \citep{fallah2020personalized, chen2018federated, nguyen2022meta,finn2017model, li2017meta} thường sẽ chọn $E\in [1,5]$.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Numerical experiment}
+\label{sec.experiment}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%e%%%%%%
+
+\subsection{Dataset \& Metric}
+
+FX nói riêng và các chỉ số tài chính nói chung là dạng dữ liệu điển hình cho aperiodic time-series data.
+Do đó, chúng tôi chọn loại dữ liệu này để kiểm thử mô hình.
+Cụ thể, chúng tôi cấu hình hai bộ dữ liệu sử dụng dữ liệu FX.
+Bộ dữ liệu \verb|USD/JPY| chỉ gồm dữ liệu của cặp tiền tệ USD/JPY, được chia thành 60 tập dữ liệu con theo trình tự thời gian với kích thước bằng nhau.
+Dữ liệu được sample theo giờ từ năm 2000 đến năm 2024, bao gồm các thuộc tính open, low, high, và close price.
+Bộ dữ liệu \verb|multi-fx| bao gồm 60 cặp tiền tệ giữa 18 quốc gia Australia, Canada, Switzerland, Denmark, EU, United Kingdom, Hong Kong, Iceland, Japan, Norway, New Zealand, Singapore, Sweden, Turkey, United States, Mexico, China, South Africa.
+Dữ liệu có thuộc tính tương tự như \verb|USD/JPY| và được sample theo ngày từ năm 2014 đến 2024.
+Số lượng mẫu dữ liệu của hai tập dữ liệu là tương đương nhau và xấp xỉ $156000$ mẫu.
 
 \vspace{2mm}
-Fixed-node DMC
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Experiment}
-\label{sec.experiment}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%e%%%%%%
-\subsection{Density functional theory}
-Density functional theory
-
-\subsection{Fixed--node DMCTEXT}
-Fixed--node DMCTEXT
+Bộ dữ liệu \verb|multi-fx| được sử dụng để rút trích và tổng hợp thông tin về các yếu tố ngoại lai (i.e.
+thông tin thị trường, kinh tế, chính trị,...), vốn được cho là có ảnh hưởng đến kết quả của một chỉ số tài chính nhất định \citep{overreactioncontrarian,fama1970efficient,mech1993portfolio}.
+Bộ dữ liệu \verb|USD/JPY| được sử dụng để kiếm chứng giả thuyết của chúng tôi về việc dữ liệu tương lai ngầm phụ thuộc vào các thời điểm nhất định trong quá khứ và cần phải tổng hợp đặc trưng quá khứ một cách hiệu quả để làm rõ được các phụ thuộc này.
 
 \vspace{2mm}
-Fixed--node DMCTEXT
 
-\vspace{2mm}
-Fixed--node DMCTEXT
+Nghiên cứu sử dụng macro accuracy, macro precision, macro recall, và macro F1 để đánh giá các mô hình.
+Theo đó, trong quá trình inference, mô hình sẽ chạy trên từng task để tính metrics của mỗi task.
+Sau đó tính trung bình cộng các metrics của các task để thu được kết quả cuối.
+
+\subsection{Experiment}
+
+Chúng tôi tiến hành so sánh phương pháp đề xuất với mô hình \verb|NHITS| thông qua các metrics nêu trên.
+Dữ liệu tổng quan được cấu trúc thành training sets, validation sets, và testing sets với tỷ lệ 6:2:2.
+Tập train dùng để huấn luyện mô hình, tập validation dùng để tìm kiếm siêu tham số và tập test để đánh giá mô hình.
+Đối với \verb|NHITS|, chúng tôi chia dữ liệu như bình thường theo tỷ lệ định sẵn ở trên.
+Đối với thuật toán đề xuất, vì quá trình meta-training và meta-testing đòi hỏi phải chia dữ liệu thành các task nhỏ, và cho phép mô hình adapt trên tập support của mỗi task, chúng tôi chia dữ liệu thành 60 tasks.
+Trong mỗi task, tập support chiếm 20\% với mục đích để mô hình học cách thích ứng với dữ liệu, tập query chiếm 80\% để kiểm tra khả năng tương thích của mô hình.
+Chúng tôi sử dụng 30 tasks để meta-train, 15 tasks để meta-validate, 15 tasks để meta-test.
+Với cách chia này, chúng tôi đảm bảo được tính công bằng khi mô hình ML được huấn luyện với lượng dữ liệu tương đương mô hình \verb|NHITS|.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-Results and discussion
 
-%----------------------------------------
-% \begin{figure}[htbp]
-%   \centering
-%   \includegraphics[width=\hsize]{all.eps}
-%   \caption{\ghost{fig.all}\label{fig.all}
-% 	FNDMC and DFT results of relative energies of
-% 	the four candidate structures of LiV$X_2$
-% 	(X = O, S, and Se):
-% 	the I- and II-type stacking of the insulating
-%     TR and metallic NTR lattices.
-% 	Their structures are shown in Figure \ref{fig.str}.
-% 	The lowest energy among the four structures is set
-% 	to be zero and the others are given as relative values
-% 	from the lowest energy.
-%   }
-% \end{figure}
-%----------------------------------------
+\begin{table*}
+    \centering
+    \caption{meo meo}
+    \begin{tblr}{
+    cells = {c},
+    cell{1}{2} = {c=4}{},
+    cell{1}{6} = {c=4}{},
+    vline{2-3} = {1}{},
+    vline{2,6} = {2-4}{},
+    hline{1,5} = {-}{0.08em},
+    hline{2} = {2-9}{},
+    hline{3} = {-}{},
+    }
+        & USD/JPY    &             &          &             & multi-fx     &              &             &               \\
+        & $accuracy$ & $precision$ & $recall$ & $F_1-score$ & $accuracy$   & $precision$  & $recall$    & $F_1-score$   \\
+    NHITS &            &             &          &             &              &              &             &               \\
+    Ours  &            &             &          &             & $62.39±0.06$ & $71.11±9.34$ & $70.14±9.6$ & $69.21±11.07$ 
+    \end{tblr}
+    \label{tab:rs}
+\end{table*}
 
-\vspace{2mm}
-Results and discussion
-
-\vspace{2mm}
-Results and discussion
+% nói 1 chút về việc xài lstm+cnn/meta-learning sẽ giảm sự phụ thuộc của kết quả vào mạng NN/siêu tham số
+% cụ thể, khi xài ML, tham số finetune rất dễ, cái nào cũng cao, chỉ cần chọn ra cái cao nhất
+% giải thích tại sao cnn perform poorly: quá chú ý đến đặc trưng cục bộ mà quên mất các phụ thuộc dài hạn
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusion}
+\section{Conclusion \& Future direction}
 \label{sec.conc}
+
+% cite Meta-SGD, và 1 đống ML method vào đây để kiểu *ML có nhiều phương pháp, cần chạy thực nghiệm nhiều hơn
+% cite iMAML, Reptile vào đây để làm future direction kiểu tối ưu lại ML vì chi phí hơi đắt
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-Conclusion
 
-% We also compared DFT results using PBE\cite{PBE}, optPBE-vdW-DF\cite{2009JK_AM,vdW-DFa,vdW-DFb}, SCAN\cite{SCAN}, and HSE06\cite{HSE06} functionals with the FNDMC results.
-All of them also favour non-trimerized structures in all three materials, but none of the DFT functionals do not quantitatively agree with FNDMC.
-Thus our results demonstrate that this series of materials: LiVO$_2$,  LiVS$_2$, and LiVSe$_2$ represent a real challenge for computational schemes and also motivates further experimental study of the ground state properties of LiVSe$_2$.
+% Conclusion
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-The work was carried out within the state assignment of Ministry of Science and Higher Education of the Russian Federation (No.
+% The work was carried out within the state assignment of Ministry of Science and Higher Education of the Russian Federation (No.
 AAAA-A18-118020190095-4, topic ``Quantum'').
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-All authors contributed to conceiving the idea.
+% All authors contributed to conceiving the idea.
 T.I., A.V.U, and G.I.P performed the calculations.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-The datasets used and/or analyzed during the current study available from the corresponding authors on reasonable request.
+% The datasets used and/or analyzed during the current study available from the corresponding authors on reasonable request.
 

### 2024-08-27 doc/paper.tex
draft 2, have to run some more experiments, result, conclusion not done yet

@@
 \usepackage{amssymb}
-\usepackage{bm}
-\usepackage{dcolumn}
-\usepackage{graphicx}
-\usepackage{bm}
+% \usepackage{bm}
+% \usepackage{dcolumn}
+% \usepackage{bm}
 \usepackage{url}
-\usepackage{here}
-\usepackage{multirow}
-\usepackage{ulem}
-\usepackage{booktabs}
-\usepackage[T1]{fontenc}
+% \usepackage{here}
+% \usepackage{ulem}
 \usepackage[caption=false]{subfig}
 \usepackage[usenames,dvipsnames]{color}
-\usepackage{tabularray}
+
+\usepackage{graphicx}
+\usepackage{booktabs}
+\usepackage{rotating}
+% \usepackage{makecell}
+\usepackage{multirow}
+
 \usepackage{fvextra}
-% \usepackage{pdfpages}
-% \usepackage{epstopdf}
-% \epstopdfDeclareGraphicsRule{.pdf}{png}{.png}{convert #1 \OutputFile}
-% \AppendGraphicsExtensions{.pdf}
-% \usepackage{hyperref}
 
@@
     \textbf{Abstract}.
-Dự đoán aperiodic time-series data (e.g.
-stock price, foreign exchange, Bitcoin price,...) là một tác vụ khó khăn đối với các mô hình học máy bởi vì loại dữ liệu này có phương sai cao và không cố định; không thể hiện chu kỳ rõ ràng nên khó rút trích đặc trưng; không chỉ phụ thuộc vào giá trị quá khứ mà còn phụ thuộc các yếu tố khác, điều khiến chúng bất ổn và không có chu kỳ như tình hình kinh tế, chính trị.
+Predicting aperiodic time-series data (e.g.
+stock price, foreign exchange, Bitcoin price,...) is a difficult task for machine learning models because this type of data has high variance and is not stationary; does not show clear cycles, making it difficult to extract features; depends not only on past values but also on external factors, which make them unstable and non-cyclical such as economic and political situations.
+To overcome the above challenges, we use Meta-learning to train a combined \verb|LSTM| and \verb|CNN| network, thereby effectively extracting and synthesizing hidden features of the data over time.
+Experiments on foreign exchange data of 60 currency pairs over 24 years (2000-2024) show that the proposed method performs well and has higher accuracy than the \verb|NHITS| - the state-of-the-art model in 2023 on time-series data, in the task of predicting the trend (upward or downward) of the next trading day.
+
+    % Dự đoán aperiodic time-series data (e.g.
+stock price, foreign exchange, Bitcoin price,...) là một tác vụ khó khăn đối với các mô hình học máy bởi vì loại dữ liệu này có phương sai cao và không cố định; không thể hiện chu kỳ rõ ràng nên khó rút trích đặc trưng; không chỉ phụ thuộc vào giá trị quá khứ mà còn phụ thuộc các yếu tố ngoại lai, điều khiến chúng bất ổn và không có chu kỳ như tình hình kinh tế, chính trị.
 Để khắc phục các thách thức nêu trên, chúng tôi sử dụng Meta-learning để huấn luyện kết hợp mạng \verb|LSTM| và \verb|CNN|, từ đó rút trích và tổng hợp hiệu quả các đặc trưng ẩn của dữ liệu theo thời gian.
-Thực nghiệm trên dữ liệu foreign exchange của 15 cặp tiền tệ trong vòng 24 năm (2000-2024) cho thấy phương pháp đề xuất hoạt động tốt và có độ chính xác cao hơn mô hình \verb|NHITS| - state-of-the-art (SOTA) model của năm 2023 trên dữ liệu time-series, trong bài toán dự đoán xu hướng (upward or downward) của ngày giao dịch tiếp theo.
-
+Thực nghiệm trên dữ liệu foreign exchange của 15 cặp tiền tệ trong vòng 24 năm (2000-2024) cho thấy phương pháp đề xuất hoạt động tốt và có độ chính xác cao hơn mô hình \verb|NHITS| - state-of-the-art model của năm 2023 trên dữ liệu time-series, trong bài toán dự đoán xu hướng (upward or downward) của ngày giao dịch tiếp theo.
 \end{abstract}
@@
 
-Aperiodic time-series prediction nói chung hay foreign exchange (FX) prediction nói riêng từ lâu đã là vấn đề đáng quan tâm của nhiều nghiên cứu \citep{li2019multi, islam2021foreign, heryadi2021foreign}.
+% Aperiodic time-series prediction nói chung hay foreign exchange (FX) prediction nói riêng từ lâu đã là vấn đề đáng quan tâm của nhiều nghiên cứu \citep{li2019multi, islam2021foreign, heryadi2021foreign}.
 Hai kỹ thuật chính được sử dụng trong aperiodic time-series prediction là fundamental analysis and technical analysis \cite{ayitey2023forex}.
@@
 
+Aperiodic time-series prediction in general or foreign exchange (FX) prediction in particular has long been a matter of concern for many researchers \citep{li2019multi, islam2021foreign, heryadi2021foreign}.
+The two main techniques used in aperiodic time-series prediction are fundamental analysis and technical analysis \cite{ayitey2023forex}.
+While fundamental analysis focuses on analyzing external factors, which are difficult to be captured from past value fluctuations such as policies and economic strategies of companies and countries to predict the future; technical analysis relies entirely on historical value fluctuations to analyze future trends.
+
 \vspace{2mm}
 
-Việc dự đoán trên aperiodic time-series data gặp một vài thách thức cố hữu.
+% Việc dự đoán trên aperiodic time-series data gặp một vài thách thức cố hữu.
 Trong đó có thể kể tới: (1) - Variance của loại dữ liệu này biến thiên mạnh qua các thời kỳ.
@@
 
+Predicting on aperiodic time-series data faces several inherent challenges: (1) - The variance of this type of data varies greatly over time.
+Therefore, the assumption that they follow a distribution to approximate the error cannot be used, leading to machine learning models having difficulty accurately predicting future values and trends; (2) - Aperiodic data does not follow any explicit rules, so learning the hidden features of the data to make predictions is very difficult; (3) - Aperiodic time-series data (e.g.
+a company's stock price) depends not only on past data but also on many external factors (e.g.
+news, economic situation, politics) \cite{li2019multi}.
+
 \vspace{2mm}
 
-Đối với thách thức đầu tiên, các mô hình ensemble [cite something here] thường được sử dụng để hạn chế ảnh hưởng của sự biến đổi variance.
+% Đối với thách thức đầu tiên, các mô hình ensemble [cite something here] thường được sử dụng để hạn chế ảnh hưởng của sự biến đổi variance.
 Ensemble model giúp cung cấp một cái nhìn tổng quát, tổng hợp từ nhiều khía cạnh dựa trên các sub-models, từ đó giúp mô hình tổng quát thích ứng được với sự thay đổi mạnh của variance.
@@
 
+For the first challenge, ensemble models [cite something here] are often used to mitigate the effects of variance variation.
+Ensemble models provide a holistic, multi-perspective view based on sub-models, thereby helping the overall model adapt to strong variance changes.
+We approach the problem in a similar but higher-level way using Meta-learning (ML) \cite{finn2017model}.
+This method effectively aggregates the parameters of local models, which helps to significantly reduce variance loss.
+
 \vspace{2mm}
 
-Để khắc phục thách thức thứ hai, các nghiên cứu phần lớn sử dụng các đặc trưng rút trích được từ Long short-term memory neural network (\verb|LSTM|) \cite{hochreiter1997long}, Artificial neural network (\verb|ANN|), và Convolution neural network (\verb|CNN|) \cite{lecun1989handwritten}.
+% Để khắc phục thách thức thứ hai, các nghiên cứu phần lớn sử dụng các đặc trưng rút trích được từ Long short-term memory neural network (\verb|LSTM|) \cite{hochreiter1997long}, Artificial neural network (\verb|ANN|), và Convolution neural network (\verb|CNN|) \cite{lecun1989handwritten}.
 Cụ thể, trong năm 2022, 20\% tổng số các bài báo liên quan đến dự đoán chỉ số tài chính sử dụng \verb|LSTM|, 20\% sử dụng \verb|ANN| và 6\% sử dụng \verb|CNN| \cite{ayitey2023forex}.
@@
 
+To overcome the second challenge, most studies use features extracted from Long short-term memory neural network (\verb|LSTM|) \cite{hochreiter1997long}, Artificial neural network (\verb|ANN|), and Convolution neural network (\verb|CNN|) \cite{lecun1989handwritten}.
+Specifically, in 2022, 20\% of all publications related to financial index prediction used \verb|LSTM|, 20\% used \verb|ANN|, and 6\% used \verb|CNN| \cite{ayitey2023forex}.
+To make full use of the features extracted by the above models, we propose a method that combines these features.
+
 \vspace{2mm}
 
-Đối với thách thức thứ ba, nghiên cứu \cite{fama1970efficient} đưa ra giả thuyết rằng các time-series datasets khác nhau của cùng một lĩnh vực tại cùng thời điểm phản ánh sự tác động của các yếu tố ngoại lai.
+% Đối với thách thức thứ ba, nghiên cứu \cite{fama1970efficient} đưa ra giả thuyết rằng các time-series datasets khác nhau của cùng một lĩnh vực tại cùng thời điểm phản ánh sự tác động của các yếu tố ngoại lai.
 Ví dụ, các nghiên cứu \citep{overreactioncontrarian, mech1993portfolio} cũng chỉ ra sự phụ thuộc giữa chỉ số tài chính của một công ty nhất định và các chỉ số của các công ty khác.
@@
 
+Regarding the third challenge, \cite{fama1970efficient} research hypothesizes that different time-series datasets of the same domain at the same time point reflect the impact of extraneous factors.
+For example, studies \citep{overreactioncontrarian, mech1993portfolio} show the dependence between the financial ratios of a given company and the ratios of other companies.
+This further strengthens the hypothesis in study \cite{fama1970efficient}.
+In addition, we argue that this type of data also has hidden long-term dependencies.
+For the traditional approach, people use a fixed amount of past data (lookback window) to train the model.
+This causes a big obstacle to the learning process because long-term features over time will be forgotten.
+On the other hand, ML divides the dataset into many parts to learn and synthesize the learned parameters effectively, so it can handle this challenge well.
+
 \vspace{2mm}
 
-Cuối cùng, chúng tôi chứng minh tính ưu việt của thuật toán đề xuất bằng cách giải bài toán dự đoán xu hướng (tăng hoặc giảm) của tỉ giá ngoại hối và so sánh kết quả với mô hình SOTA hiện tại (\verb|NHITS| \cite{challu2023nhits}) trên hai loại dữ liệu: (1) - Dữ liệu tỉ giả của cặp tiền tệ USD/JPY; (2) - Dữ liệu tỷ giá của 60 cặp tiền tệ, cấu thành từ 18 quốc gia.
+% Cuối cùng, chúng tôi chứng minh tính ưu việt của thuật toán đề xuất bằng cách giải bài toán dự đoán xu hướng (tăng hoặc giảm) của tỉ giá ngoại hối và so sánh kết quả với mô hình SOTA hiện tại (\verb|NHITS| \cite{challu2023nhits}) trên hai loại dữ liệu: (1) - Dữ liệu tỉ giả của cặp tiền tệ USD/JPY; (2) - Dữ liệu tỷ giá của 60 cặp tiền tệ, cấu thành từ 18 quốc gia.
 Các tập dữ liệu này được công khai trên Internet và có thể tải về dễ dàng.
@@
 
+Finally, we demonstrate the superiority of the proposed algorithm by solving the problem of predicting the trend (up or down) of foreign exchange rates and comparing the results with the existing state-of-the-art (SOTA) model (\verb|NHITS| \cite{challu2023nhits}) on two types of data: (1) - USD/JPY exchange rate data; (2) - Exchange rate data of 60 currency pairs, comprising 18 countries.
+These data sets are publicly available on the Internet and can be easily downloaded.
+The official implementation can be found at [insert github link here].
+
 \vspace{2mm}
 
-Tóm lại, đóng góp chính của chúng tôi như sau:
+In summary, our main contributions are as follows:
 
 \begin{itemize}
-    \item Kết hợp đặc trưng: Rút trích đặc trưng bằng cách kết hợp các đặc trưng của \verb|LSTM| và \verb|CNN|.
-    \item Hidden long-term dependency: Chứng minh thực nghiệm rằng một aperiodic time-series data nhất định không chỉ phụ thuộc vào các yếu tố ngoại lai mà còn có các phụ thuộc ẩn với chính nó tại nhiều thời điểm khác nhau trong quá khứ.
-    \item Tổng hợp hiệu quả tham số mô hình: Sử dụng ML thay cho các mô hình ensemble truyền thống trong việc tổng hợp kết quả từ các mô hình học máy.
-    \item Experiment: Thực nghiệm trên các bộ dữ liệu về tỉ giá hối đoái và so sánh với mô hình SOTA \verb|NHITS| để chứng minh tính hiệu quả của phương pháp đề xuất.
+    % \item Kết hợp đặc trưng: Rút trích đặc trưng bằng cách kết hợp các đặc trưng của \verb|LSTM| và \verb|CNN|.
+    \item Feature combination: Extract feature using \verb|LSTM| and \verb|CNN| and combine them.
+
+    % \item Hidden long-term dependency: Chứng minh thực nghiệm rằng một aperiodic time-series data nhất định không chỉ phụ thuộc vào các yếu tố ngoại lai mà còn có các phụ thuộc ẩn với chính nó tại nhiều thời điểm khác nhau trong quá khứ.
+    \item Hidden long-term dependency: Experimentally demonstrate that a given aperiodic time-series data not only depends on external factors but also has hidden dependencies with itself at various points in the past.
+
+    % \item Tổng hợp hiệu quả tham số mô hình: Sử dụng ML thay cho các mô hình ensemble truyền thống trong việc tổng hợp kết quả từ các mô hình học máy.
+    \item Efficient model parameter aggregation: Use ML instead of traditional ensemble models to aggregate results from machine learning models.
+
+    % \item Experiment: Thực nghiệm trên các bộ dữ liệu về tỉ giá hối đoái và so sánh với mô hình SOTA \verb|NHITS| để chứng minh tính hiệu quả của phương pháp đề xuất.
+    \item Experiment: Experiment on exchange rate datasets and compare with \verb|NHITS| - the SOTA model to demonstrate the effectiveness of the proposed method.
 \end{itemize}
@@
 
-Như đã đề cập, \verb|LSTM| là mạng neural rất phổ biến trong việc handle các bài toán liên quan đến dự đoán trên dữ liệu time-series.
+% Như đã đề cập, \verb|LSTM| là mạng neural rất phổ biến trong việc handle các bài toán liên quan đến dự đoán trên dữ liệu time-series.
 \verb|LSTM| được sử dụng phổ biến như vậy bởi vì nó xử lý tốt vấn đề vanishing gradient (dễ dàng bắt gặp khi sử dụng Recurrent neural network) và có thể khai thác hiệu quả các mối quan hệ phi tuyến trong dữ liệu.
@@
 
+As aforementioned, \verb|LSTM| is a well-known neural network for handling prediction problems on time-series data.
+\verb|LSTM| is commonly used because it handles the problem of vanishing gradients well (easily encountered when using Recurrent neural networks) and can effectively exploit nonlinear relationships in data.
+Indeed, by maintaining the cell-state in each iteration, \verb|LSTM| can overcome the vanishing gradient problem, thereby preserving the ability to capture long-term dependencies \cite{cheng2018applied}.
+In addition, \verb|LSTM| performs feature extraction with nonlinear activation functions, which helps the model parameters capture the nonlinearity of the data \cite{he2016exploiting}.
+These factors make \verb|LSTM| to be the first choice to think of when solving problems on time-series data.
+
 \vspace{2mm}
 
-Mạng \verb|CNN| được sử dụng rất nhiều trong các tác vụ xử lý hình ảnh \citep{naranjo2020review, sharma2018analysis} bởi khả năng tổng hợp các quan hệ cục bộ.
+% Mạng \verb|CNN| được sử dụng rất nhiều trong các tác vụ xử lý hình ảnh \citep{naranjo2020review, sharma2018analysis} bởi khả năng tổng hợp các quan hệ cục bộ.
 Không chỉ vậy, \verb|CNN| còn được dùng rất nhiều trong các tác vụ xử lý dữ liệu time-series như speech recognition \cite{dua2022developing}, natural language processing \cite{varshitha2023natural}.
@@
 
+\verb|CNN| is widely used in image processing tasks \citep{naranjo2020review, sharma2018analysis} because of its ability to synthesize local relationships.
+Not only that, \verb|CNN| is also widely used in time-series data processing tasks such as speech recognition \cite{dua2022developing}, natural language processing \cite{varshitha2023natural}.
+This proves the ability of \verb|CNN| in discovering local temporal relationships between data samples.
+However, \verb|CNN| is rarely used in aperiodic time-series data prediction tasks.
+In this study, we take advantage of \verb|CNN|'s excellent local feature extraction ability to incorporate more hidden information into the model training process.
+
 \subsection{Model-agnostic Meta-learning (MAML)}
 
-Các thuật toán Meta-learning (ML), điển hình là MAML \cite{finn2017model} được biết đến với khả năng huấn luyện một mô hình có tính tổng quát cao, thích ứng nhanh trên tập dữ liệu mới thông qua một lượng nhỏ dữ liệu và số bước huấn luyện \citep{hospedales2021meta, vettoruzzo2024advances}.
+% Các thuật toán Meta-learning (ML), điển hình là MAML \cite{finn2017model} được biết đến với khả năng huấn luyện một mô hình có tính tổng quát cao, thích ứng nhanh trên tập dữ liệu mới thông qua một lượng nhỏ dữ liệu và số bước huấn luyện \citep{hospedales2021meta, vettoruzzo2024advances}.
 Với khả năng này, ML được sử dụng rất nhiều trong các tác vụ đòi hỏi khả năng đáp ứng của mô hình trên dữ liệu (e.g.
@@
 
+Meta-learning (ML) algorithms, typically MAML \cite{finn2017model} are known for their ability to train a highly general, adaptive model on new datasets with a limited amount of data and a small number of training steps \citep{hospedales2021meta, vettoruzzo2024advances}.
+With this ability, ML is widely used in tasks that require the model's ability to adapt to the data (e.g.
+personalization of learning models \citep{chen2018federated, fallah2020personalized,nguyen2022meta}, domain adaptation in online learning \citep{hu2023meta, khoee2024domain}).
+
 \vspace{2mm}
 
-Một thuật toán ML cơ bản sẽ được học trên nhiều tác vụ $t$ rút ra từ cùng một phân phối tác vụ $\mathcal{T}$ \cite{hospedales2021meta}.
+% Một thuật toán ML cơ bản sẽ được học trên nhiều tác vụ $t$ rút ra từ cùng một phân phối tác vụ $\mathcal{T}$ \cite{hospedales2021meta}.
 Dữ liệu của mỗi tác vụ được chia thành tập support $\mathcal{D}_t^{support}$ (thường có kích thước nhỏ, khoảng 20\%) và tập query $\mathcal{D}_t^{query}$.
@@
 
+A basic ML algorithm is trained on multiple tasks $t$ drawn from the same task distribution $\mathcal{T}$ \cite{hospedales2021meta}.
+The data for task $t$ is divided into a support set $\mathcal{D}_t^{support}$ (usually small, around 20\%) and a query set $\mathcal{D}_t^{query}$.
+During the learning process, two optimization steps, inner and outer optimization, are performed alternately.
+Inner optimization attempts to find an optimal set of parameters $\theta_t^*$ for each machine learning model on the support set of each task using the equation \ref{eq:inner_opt}.
+
 \begin{equation}
@@
     \label{eq:inner_opt}
-\end{equation} Trong đó, $\phi$ là kết quả của quá trình outer optimization, đóng vai trò là giá trị khởi tạo của $\theta_t$. $\mathcal{L}_t$ là hàm lỗi của mô hình trên tác vụ $t$.
+\end{equation} Where, $\phi$ is the result of the outer optimization process, which acts as the initial value of $\theta_t$.
+$\mathcal{L}^{task}_t$ is the error function of the model on the support set of task $t$.
 
@@
 
-Sau đó, thuật toán sử dụng các bộ tham số tối ưu $\theta_t^*$ để perform trên tập query tương ứng.
+% Sau đó, thuật toán sử dụng các bộ tham số tối ưu $\theta_t^*$ để perform trên tập query tương ứng.
 Lỗi của toàn bộ mô hình sau đó được tổng hợp để thực hiện quá trình outer optimization như phương trình \ref{eq:outer_opt}.
+
+The algorithm then uses the optimal parameter sets $\theta_t^*$ to perform on the corresponding query set.
+The losses of the entire models are then aggregated to perform the outer optimization process as equation \ref{eq:outer_opt}.
 
@@
 
-Trong inference phase, giá trị khởi tạo cho tham số của mô hình được gán bằng $\phi^*$.
+% Bằng hình thức huấn luyện trên, mô hình $\phi^*$ sẽ có mức tổng quát hóa cao trên các tác vụ khác nhau, có thể nhanh chóng đáp ứng một tác vụ mới chỉ sau một vài bước huấn luyện.
+
+By performing the above training method, the $\phi^*$ model will have a high level of generalization across different tasks, and can quickly respond to a new task after only a few training steps.
+
+\vspace{2mm}
+
+% Trong inference phase, giá trị khởi tạo cho tham số của mô hình được gán bằng $\phi^*$.
 Mô hình sau đó được huấn luyện nhanh trên tập support sau đó perform trên tập query.
@@
 
-\vspace{2mm}
-
-Bằng hình thức huấn luyện trên, mô hình $\phi^*$ sẽ có mức tổng quát hóa cao trên các tác vụ khác nhau, có thể nhanh chóng đáp ứng một tác vụ mới chỉ sau một vài bước huấn luyện.
+In the inference phase, the initial values for the model parameters are assigned $\phi^*$.
+The model is then adapted quickly to the support set and performed on the query set.
+The results on the query set are the model output.
 
@@
 
-Các mô hình hybrid ensemble vốn được sử dụng rất nhiều trong các bài toán xử lý time-series và được chứng minh thực nghiệm là có độ chính xác cao hơn so với các mô hình handle time-series data tiêu chuẩn vì có thể tổng hợp được sức mạnh của nhiều mô hình \cite{ayitey2023forex}.
+% Các mô hình hybrid ensemble vốn được sử dụng rất nhiều trong các bài toán xử lý time-series và được chứng minh thực nghiệm là có độ chính xác cao hơn so với các mô hình handle time-series data tiêu chuẩn vì có thể tổng hợp được sức mạnh của nhiều mô hình \cite{ayitey2023forex}.
 Tuy vậy, các hình thức tổng hợp của ensemble model hiện nay vẫn còn rất cứng nhắc vì chỉ có thể tổng hợp dựa trên kết quả cuối (cơ chế voting của bagging models) và kết quả gần cuối (đối với stacking models).
@@
 
+Hybrid ensemble models have been widely used in time-series processing problems and have been experimentally proven to be more accurate than standard time-series models because they can synthesize the strengths of many sub-models \cite{ayitey2023forex}.
+However, current ensemble model synthesis forms are still very rigid because they can only synthesize based on the final results (voting mechanism of bagging models) and near-final results (for stacking models).
+From the perspective of ensemble models, the equation \ref{eq:outer_opt} can be considered an effective method of synthesizing sub-models, which helps to take advantage of the feature extraction capabilities of each model.
+In other words, the synthesized model can extract features at a deeper level, significantly improving the prediction ability compared to traditional ensemble models.
+
 \subsection{Neural Hierarchical Interpolation\\for Time Series (NHITS)}
 
-NHITS được thiết kế để hướng đến việc dự đoán các long-horizon time-series data.
+% \verb|NHITS| được thiết kế để hướng đến việc dự đoán các long-horizon time-series data.
 Theo nghiên cứu \cite{challu2023nhits}, cấu trúc của \verb|NHITS| bao gồm nhiều stack liên tiếp nhau.
@@
 Cụ thể, tại block $l$, với $L$ mẫu dữ liệu quá khứ ($\mathbf{y}_{t-L:t, l-1}$), các đặc trưng sẽ được rút trích như sau (theo \cite{challu2023nhits}):
+
+\verb|NHITS| is designed to target the prediction of long-horizon time-series data.
+According to \cite{challu2023nhits}, the structure of \verb|NHITS| consists of multiple consecutive stacks.
+Each stack consists of multiple consecutive blocks.
+At each block, historical data is used to predict future data and past data.
+The residual of the previous block is used as input data for the following block.
+Specifically, at block $l$, with $L$ past data samples ($\mathbf{y}_{t-L:t, l-1}$), the features will be extracted as follows (from \cite{challu2023nhits}):
 
@@
 
-Trong đó, \verb|FullyConnected| là các lớp multi-layer perception (\verb|MLP|) xếp chồng với hàm kích hoạt phi tuyến.
+% Trong đó, \verb|FullyConnected| là các lớp multi-layer perception (\verb|MLP|) xếp chồng với hàm kích hoạt phi tuyến.
 $\mathbf{\theta}_l^f, \mathbf{\theta}_l^b$ là các hệ số nội suy forecast và backcast, được dùng để tổng hợp các giá trị đầu ra của block $l$ bằng hàm nội suy $g(\cdot)$.
@@
 Input của block $l+1$ được tính theo phương trình \ref{eq:input_l1}.
+
+Accordingly, \verb|FullyConnected| are stacked multi-layer perception (\verb|MLP|) layers with nonlinear activation functions.
+$\mathbf{\theta}_l^f, \mathbf{\theta}_l^b$ are the forecast and backcast interpolation coefficients, which are used to aggregate the output values of block $l$ using the interpolation function $g(\cdot)$.
+The output of block $l$ is the forecast value $\mathbf{\hat{y}}_{t+1:t+H, l}$ and the backcast value $\mathbf{\hat{y}}_{t-L:t, l}$.
+The input of block $l+1$ is calculated according to the equation \ref{eq:input_l1}.
 
@@
 
-Giả sử mô hình gồm có $S$ stacks, mỗi stack có $B$ blocks.
+% Giả sử mô hình gồm có $S$ stacks, mỗi stack có $B$ blocks.
 Tổng hợp các giá trị forecast của các block như phương trình \ref{eq:sum_block}, ta được giá trị forecast của một stack.
@@
 Cuối cùng, tổng hợp giá trị forecast của các stack như phương trình \ref{eq:sum_stack}, ta được giá trị forecast dự đoán của toàn mạng.
+
+Suppose the model consists of $S$ stacks, each stack has $B$ blocks.
+Summing the forecast values of the blocks as in equation \ref{eq:sum_block}, we get the forecast value of a stack.
+The last residual of the last block of a stack is the input for the next stack.
+Finally, summing the forecast values of the stacks as in equation \ref{eq:sum_stack}, we get the predicted forecast value of the entire network.
 
@@
 
-Bằng cách xếp chồng các stack, stack sau nhận vào phần dư của stack trước, kiến trúc trên được kỳ vọng là sẽ phân rã dữ liệu thành các frequency bands khác nhau (weekly, daily, even hourly).
+% Bằng cách xếp chồng các stack, stack sau nhận vào phần dư của stack trước, kiến trúc trên được kỳ vọng là sẽ phân rã dữ liệu thành các frequency bands khác nhau (weekly, daily, even hourly).
 Trên thực tế, \verb|NHITS| perform rất tốt đối với các bộ dữ liệu có tính chu kỳ cao như mức tiêu thụ điện, thời tiết, giao thông.
@@
 Điều này gây ra khó khăn rất lớn cho \verb|NHITS|.
+
+By concatenating stacks, each receiving the remainder of the previous stack, the above architecture is expected to decompose the data into different frequency bands (weekly, daily, even hourly).
+In practice, \verb|NHITS| performs very well for highly periodic datasets such as electricity consumption, weather, traffic.
+However, we are aiming for an aperiodic time-series dataset, which has very low, or even non-existent, periodicity (see figure \ref{fig:fx}).
+This poses a huge challenge for \verb|NHITS|.
 
@@
 % nói tổng quan ở đây, vẽ cái hình và bỏ vào
-Phương pháp của chúng tôi bao gồm hai phần chính hoạt động song song nhau: (1) - Feature extraction; (2) - Parameter synthesis.
+% Phương pháp của chúng tôi bao gồm hai phần chính hoạt động song song nhau: (1) - Feature extraction; (2) - Parameter synthesis.
 Tổng quan phương pháp được minh họa trong hình \ref{fig:flow}.
@@
 Bằng việc sử dụng \verb|MAML| trong quá trình tổng hợp trọng số, phương pháp đề xuất được kỳ vọng là một giải pháp thay thế hợp lý và hiệu quả cho các mô hình ensemble truyền thống trong việc giảm thiểu tác động của sự biến thiên variance, tổng hợp hiệu quả các yếu tố ngoại lai, cũng như giữ được các hidden long-term dependency ẩn trong quá khứ.
+
+Our method consists of two main parts that work in parallel: (1) - Feature extraction; (2) - Parameter synthesis.
+The overview of the method is illustrated in \ref{fig:flow}.
+In the feature extraction part, we combine two types of features from \verb|CNN| and \verb|LSTM| networks.
+In the parameter synthesis part, we use \verb|MAML| to synthesize the parameters of the models.
+Due to the contribution of \verb|LSTM| and \verb|CNN| features, we expect to effectively extract hidden features from aperiodic data.
+By using \verb|MAML| in the weight synthesis process, the proposed method is expected to be a reasonable and effective alternative to traditional ensemble models in minimizing the impact of variance variation, effectively synthesizing external factors, and preserving hidden long-term dependencies in the past.
 
@@
 
-Phương pháp đề xuất sử dụng các thuật toán ML để huấn luyện mô hình.
+% Phương pháp đề xuất sử dụng các thuật toán ML để huấn luyện mô hình.
 Do đó, dữ liệu cần được tổ chức lại để các thuật toán ML có thể hoạt động được.
@@
 
+The proposed method uses ML algorithms to train the model.
+Therefore, the data needs to be reorganized so that the ML algorithms can work.
+In case the data includes many different datasets belonging to the same field, each dataset will be considered a task of \verb|MAML|.
+In case the data includes a single dataset, it is necessary to divide this dataset into subsets corresponding to separate tasks.
+In summary, the prepared dataset includes $n$ tasks: $\mathcal{D} = \left\{ \mathcal{D}_t \right\}_{t=1}^{n}$.
+The data at each task is divided into support and query sets: $\mathcal{D}_t = \left\{ \mathcal{D}_t^{support}, \mathcal{D}_t^{query} \right\}$.
+
 \vspace{2mm}
 
-Một sample dữ liệu bao gồm các cặp giá trị $(\mathbf{x}_{t-L:t}, y)$.
+% Một sample dữ liệu bao gồm các cặp giá trị $(\mathbf{x}_{t-L:t}, y)$.
 Trong đó, $\mathbf{x}_{t-L:t}$ bao gồm $L$ giá trị lịch sử tính từ thời điểm $t$ trở về trước; $y\in \{0,1\}$ là nhãn dữ liệu, thể hiện xu hướng giảm, hoặc tăng của mẫu dữ liệu $x_{t+1}$ so với $x_{t}$.
@@
 Ví dụ, đối với dữ liệu chứng khoán, $\mathbf{x}_{t-L:t}$ có thể chứa các vector dữ liệu $\vec x_i = (\text{open, low, high, close})$ hoặc chỉ một giá trị close price duy nhất.
+
+A data sample consists of pairs of values $(\mathbf{x}_{t-L:t}, y)$.
+In which, $\mathbf{x}_{t-L:t}$ includes $L$ historical values from time $t$ back; $y\in \{0,1\}$ is the data label, showing the decreasing or increasing trend of the data sample $x_{t+1}$ compared to $x_{t}$.
+Depending on each problem and the implementation, the elements in $\mathbf{x}_{t-L:t}$ can be vectors or scalar numbers.
+For example, for stock data, $\mathbf{x}_{t-L:t}$ can contain $L$ data vectors $\vec x_i = (\text{open, low, high, close})$ or just a single close price value.
 
@@
 
-Lấy cảm hứng từ nghiên cứu \cite{vo2017multi}, chúng tôi đề xuất kết hợp các đặc trưng rút trích được từ mạng \verb|LSTM| và \verb|CNN|.
+% Lấy cảm hứng từ nghiên cứu \cite{vo2017multi}, chúng tôi đề xuất kết hợp các đặc trưng rút trích được từ mạng \verb|LSTM| và \verb|CNN|.
 Cụ thể, chúng tôi đưa từng phần tử trong vector $\mathbf{x}_{t-L:t}$ qua một lớp \verb|MLP| có đầu ra lớn hơn số chiều của $\vec x_i, i\in[t-L, t]$ để phân giải thành các đặc trưng nhỏ $\vec x'_i$.
@@
 Toàn bộ quy trình rút trích đặc trưng được tóm tắt như sau:
+
+Inspired by the \cite{vo2017multi} study, we propose to combine the features extracted from \verb|LSTM| and \verb|CNN| networks.
+Specifically, we pass each element in the vector $\mathbf{x}_{t-L:t}$ through a \verb|MLP| layer whose output'dimension is larger than the one of $\vec x_i, i\in[t-L, t]$ to decompose it into smaller features $\vec x'_i$.
+These features are then passed through \verb|LSTM| and \verb|CNN| networks to extract long-term temporal dependencies ($\mathbf{h}_{LSTM}$) and local temporal features ($\mathbf{h}_{CNN}$), respectively.
+To exploit the long-term temporal constraints, we use \verb|BidirectionalLSTM| to extract from both sides of $\mathbf{x}_{t-L:t}$.
+The entire feature extraction process is summarized as follows:
 
@@
 
-Mạng \verb|LSTM| duy trì giá trị cell-state nhằm lưu trữ có chọn lọc các phụ thuộc dài hạn.
+% Mạng \verb|LSTM| duy trì giá trị cell-state nhằm lưu trữ có chọn lọc các phụ thuộc dài hạn.
 Điều này rất thích hợp trong việc giải quyết các bài toán time-series data.
@@
 
+The \verb|LSTM| network maintains cell-state values ​​to selectively store long-term dependencies.
+This is very suitable for solving time-series data problems.
+On the other hand, future values ​​often depend heavily on recent historical values.
+We propose to use the \verb|CNN| network to emphasize local features, thereby directing part of the model's attention to certain time points.
+Therefore, the proposed method can not only remember long-term features but also highlight short-term features.
+
 \vspace{2mm}
 
-Tiếp đến, $\mathbf{h}_{LSTM}$ và $\mathbf{h}_{CNN}$ được nối với nhau (phương trình \ref{eq:concat}) sau đó chuyển đến phần phân lớp của mạng NN (phương trình \ref{eq:clf}).
+% Tiếp đến, $\mathbf{h}_{LSTM}$ và $\mathbf{h}_{CNN}$ được nối với nhau (phương trình \ref{eq:concat}) sau đó chuyển đến phần phân lớp của mạng NN (phương trình \ref{eq:clf}).
+
+Next, $\mathbf{h}_{LSTM}$ and $\mathbf{h}_{CNN}$ are concatenated (equation \ref{eq:concat}) and then passed to the classification part of the neural network (equation \ref{eq:clf}).
 
@@
 
-Chúng tôi sử dụng \verb|MAML| để huấn luyện và tổng hợp trọng số của các mô hình tại các task.
+% Chúng tôi sử dụng \verb|MAML| để huấn luyện và tổng hợp trọng số của các mô hình tại các task.
 Như đã đề cập trong phần \ref{sec.relatedWork}, tối ưu tham số theo cách tiếp cận của ML chính là đi giải hai phương trình \ref{eq:inner_opt} và \ref{eq:outer_opt} bằng các phương pháp tối ưu trên dữ liệu support và query.
@@
 Tại bước toàn cục $r$, quá trình tối ưu cục bộ lần thứ $e$ tại tập support của task $t$ diễn ra như sau:
+
+We use \verb|MAML| to train and aggregate the weights of the models at the tasks.
+As mentioned in \ref{sec.relatedWork}, parameter optimization in the ML approach is to solve the two equations \ref{eq:inner_opt} and \ref{eq:outer_opt} using optimization methods on the support and query data.
+Specifically, the optimization process includes many global steps (outer optimization), performed on all tasks participating in training.
+Each global step includes many local steps (inner optimization) performed on each individual task.
+At global step $r$, the $e$th local optimization process at the support set of task $t$ occurs as follows:
 
@@
     \end{cases}
-\end{align} Trong đó, $\phi_{r-1}$ là kết quả của quá trình tối ưu toàn cục lần thứ $r-1$, $\alpha$ là inner learning rate.
+\end{align} In which, $\phi_{r-1}$ is the result of the $r-1$ global optimization process, $\alpha$ is the inner learning rate.
 
-Tiếp đó, quá trình outer optimization tại bước toàn cục được thực hiện bằng cách tổng hợp độ lỗi trên tập query của các task và tối ưu trên đó (phương trình \ref{eq:outer_sol}).
+\vspace{2mm}
+
+% Tiếp đó, quá trình outer optimization tại bước toàn cục được thực hiện bằng cách tổng hợp độ lỗi trên tập query của các task và tối ưu trên đó (phương trình \ref{eq:outer_sol}).
+
+Next, the outer optimization process at the global step is performed by aggregating the losses on the query set of the tasks and optimizing on it (equation \ref{eq:outer_sol}).
 
@@
     \label{eq:outer_sol}
-\end{align} Trong đó, $\beta$ là outer learning rate.
-Giả sử thuật toán chạy $E$ steps trong inner optimization, lượng đạo hàm tại phương trình \ref{eq:outer_sol} được viết lại như sau (the notations of dataset are removed):
+\end{align} Where, $\beta$ is the outer learning rate.
+
+% Giả sử thuật toán chạy $E$ steps trong inner optimization, lượng đạo hàm tại phương trình \ref{eq:outer_sol} được viết lại như sau (the notations of dataset are removed):
+
+Assuming the algorithm runs $E$ steps in inner optimization, the derivative quantity at equation \ref{eq:outer_sol} is rewritten as follows (the notations of dataset are removed):
 
@@
 
-Sự xuất hiện của tích các đạo hàm bậc hai trong phương trình \ref{eq:outer_sol_} khiến quá trình đạo hàm trở nên phức tạp vì phải tốn rất nhiều chi phí để duy trì các ma trận Hessian.
+% Sự xuất hiện của tích các đạo hàm bậc hai trong phương trình \ref{eq:outer_sol_} khiến quá trình đạo hàm trở nên phức tạp vì phải tốn rất nhiều chi phí để duy trì các ma trận Hessian.
 Do đó, số bước tính toán để tìm ra $\theta^*$ cần phải hạn chế.
 Trên thực tế, các phương pháp sử dụng ML \citep{fallah2020personalized, chen2018federated, nguyen2022meta,finn2017model, li2017meta} thường sẽ chọn $E\in [1,5]$.
+
+The presence of the product of second order derivatives in the equation \ref{eq:outer_sol_} makes the derivation process complicated because it requires a lot of overhead to maintain the Hessian matrices.
+Therefore, the number of computational steps to find $\theta^*$ needs to be limited.
+In practice, methods using ML \citep{fallah2020personalized, chen2018federated, nguyen2022meta,finn2017model, li2017meta} often choose $E\in [1,5]$.
 
@@
 
-FX nói riêng và các chỉ số tài chính nói chung là dạng dữ liệu điển hình cho aperiodic time-series data.
+% FX nói riêng và các chỉ số tài chính nói chung là dạng dữ liệu điển hình cho aperiodic time-series data.
 Do đó, chúng tôi chọn loại dữ liệu này để kiểm thử mô hình.
@@
 
+FX in particular and financial indices in general are typical data types for aperiodic time-series data.
+Therefore, we choose this type of data to test the model.
+Specifically, we configure two datasets using FX data.
+The \verb|USD/JPY| dataset consists of only data of the USD/JPY currency pair, divided into 60 time-sequenced subsets of equal size.
+The data is sampled hourly from 2000 to 2024, including the attributes of open, low, high, and close price.
+The \verb|multi-fx| dataset consists of 60 currency pairs made up of 18 countries: Australia, Canada, Switzerland, Denmark, EU, United Kingdom, Hong Kong, Iceland, Japan, Norway, New Zealand, Singapore, Sweden, Turkey, United States, Mexico, China, South Africa.
+The data has similar attributes to \verb|USD/JPY| and sampled daily from 2014 to 2024.
+The number of data samples of these two datasets are similar and approximately $156000$ samples.
+
 \vspace{2mm}
 
-Bộ dữ liệu \verb|multi-fx| được sử dụng để rút trích và tổng hợp thông tin về các yếu tố ngoại lai (i.e.
+% Bộ dữ liệu \verb|multi-fx| được sử dụng để rút trích và tổng hợp thông tin về các yếu tố ngoại lai (i.e.
 thông tin thị trường, kinh tế, chính trị,...), vốn được cho là có ảnh hưởng đến kết quả của một chỉ số tài chính nhất định \citep{overreactioncontrarian,fama1970efficient,mech1993portfolio}.
@@
 
+The \verb|multi-fx| dataset is used to extract and aggregate information about outliers (i.e.
+market, economic, political, etc.) that are believed to influence the outcome of a given financial index \citep{overreactioncontrarian,fama1970efficient,mech1993portfolio}.
+The \verb|USD/JPY| dataset is used to test our hypothesis that future data implicitly depend on certain points in the past and that efficient past feature aggregation is needed to reveal these dependencies.
+
 \vspace{2mm}
 
-Nghiên cứu sử dụng macro accuracy, macro precision, macro recall, và macro F1 để đánh giá các mô hình.
+% Nghiên cứu sử dụng macro accuracy, macro precision, macro recall, và macro F1 để đánh giá các mô hình.
 Theo đó, trong quá trình inference, mô hình sẽ chạy trên từng task để tính metrics của mỗi task.
@@
 
+The study uses macro accuracy, macro precision, macro recall, and macro F1 score to evaluate the models.
+Accordingly, during the inference phase, the model will run on each task to calculate the metrics of each task.
+Then, the average of the metrics of the tasks is calculated to obtain the final result.
+
+\vspace{2mm}
+
 \subsection{Experiment}
 
-Chúng tôi tiến hành so sánh phương pháp đề xuất với mô hình \verb|NHITS| thông qua các metrics nêu trên.
+% Chúng tôi tiến hành so sánh phương pháp đề xuất với mô hình \verb|NHITS| thông qua các metrics nêu trên.
 Dữ liệu tổng quan được cấu trúc thành training sets, validation sets, và testing sets với tỷ lệ 6:2:2.
@@
 
+We compare the proposed method with \verb|NHITS| using the above metrics.
+The overall data is structured into training sets, validation sets, and testing sets with a ratio of 6:2:2.
+The training set is used to train the model, the validation set is used to find hyper-parameters, and the test set is used to evaluate the model.
+For \verb|NHITS|, we split the data as usual according to the above predetermined ratio.
+For the proposed algorithm, because the meta-training and meta-testing processes require dividing the data into small tasks, and allowing the model to adapt on the support set of each task, we split the data into 60 tasks.
+In each task, the support set accounts for 20\% with the purpose of letting the model adapt to the data, the query set accounts for 80\% to check the model's compatibility.
+We use 30 tasks for meta-training, 15 tasks for meta-validation, and 15 tasks for meta-testing.
+With this division, we ensure the fairness that the ML model is trained with the same amount of data as \verb|NHITS|.
+
+\vspace{2mm}
+
+% Trong cài đặt của chúng tôi, chúng tôi sử dụng một lớp \verb|FullyConnected| gồm 16 units với hàm kích hoạt \verb|ReLU| để phân giải đặc trưng ban đầu.
+Sau đó, đặc trưng này được truyền song song đến các hai khối \verb|BidirectionalLSTM| và \verb|CNN|.
+Khối \verb|BidirectionalLSTM| bao gồm 32 hidden units, các đầu ra được nối dài để tạo thành một vector cuối.
+Khối \verb|CNN| bao gồm hai layer \verb|CNN| có số filter lần lượt là 32 và 64.
+Kernel được sử dụng trong các layer có kích thước $3\times 3$.
+Theo sau mỗi layer \verb|CNN| là một layer \verb|MaxPooling| sử dụng kernel kích thước $2\times 2$.
+Kết thúc khối \verb|CNN| là một layer \verb|Flatten|.
+Đặc trưng của khối \verb|BidirectionalLSTM| và \verb|CNN| sau đó được nối lại rồi truyền qua một layer phân lớp nhị phân với hàm kích hoạt \verb|Sigmoid|.
+
+In our implementation, we use a \verb|FullyConnected| layer of 16 units with a \verb|ReLU| activation function to decompose the initial feature.
+This feature is then passed in parallel to the \verb|BidirectionalLSTM| and \verb|CNN| blocks.
+The \verb|BidirectionalLSTM| block consists of 32 hidden units, the outputs of which are concatenated to form a final vector.
+The \verb|CNN| block consists of two \verb|CNN| layers with 32 and 64 filters, respectively.
+The kernel used in the layers is of size $3\times 3$.
+Each \verb|CNN| layer is followed by a \verb|MaxPooling| layer using a kernel of size $2\times 2$.
+The \verb|CNN| block ends with a \verb|Flatten| layer.
+Features of the \verb|BidirectionalLSTM| and \verb|CNN| blocks are then concatenated and passed through a binary classification layer with a \verb|Sigmoid| activation function.
+
+% Đối với mô hình \verb|NHITS|, chúng tôi sử dụng
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\begin{table*}
+\begin{table*}[h]
     \centering
-    \caption{meo meo}
-    \begin{tblr}{
-    cells = {c},
-    cell{1}{2} = {c=4}{},
-    cell{1}{6} = {c=4}{},
-    vline{2-3} = {1}{},
-    vline{2,6} = {2-4}{},
-    hline{1,5} = {-}{0.08em},
-    hline{2} = {2-9}{},
-    hline{3} = {-}{},
-    }
-        & USD/JPY    &             &          &             & multi-fx     &              &             &               \\
-        & $accuracy$ & $precision$ & $recall$ & $F_1-score$ & $accuracy$   & $precision$  & $recall$    & $F_1-score$   \\
-    NHITS &            &             &          &             &              &              &             &               \\
-    Ours  &            &             &          &             & $62.39±0.06$ & $71.11±9.34$ & $70.14±9.6$ & $69.21±11.07$ 
-    \end{tblr}
+    \caption{Classification results (\%) of NHITS and our method using USD/JPY and multi-fx datasets. Best results per metrics are boldfaced.}
     \label{tab:rs}
+    \begin{tabular}{c|c|cccc}
+        \toprule
+        \multicolumn{1}{c}{}        &       & $accuracy$ & $precision$ & $recall$ & $F1$  \\ 
+        \hline
+        \multirow{2}{*}{\Verb|USD/JPY|}    & \Verb|NHITS| & 53.29      & 52.38    & 52.23       & 51.87      \\
+                                    & Ours  & $\mathbf{58.05±0.01}$ & $\mathbf{59.57±2.53}$ & $\mathbf{58.31±2.57}$ & $\mathbf{56.6±3.74}$    \\ 
+        \hline
+        \multirow{2}{*}{\Verb|multi-fx|} & \Verb|NHITS| & 51.12      & 51.22      & 51.16      & 50.54        \\
+                                    & Ours  & $\mathbf{62.39±0.06}$ & $\mathbf{71.11±9.34}$ & $\mathbf{70.14±9.6}$  & $\mathbf{69.21±11.07}$  \\
+        \bottomrule
+    \end{tabular}
 \end{table*}

### 2024-08-27 common/constants.py
detached compute_metrics to common.py, add val and test mode, test after train

@@ -13,3 +13,9 @@ LSTM = 'lstm'
 CNN = 'cnn'
 LSTM_CNN = 'lstm_cnn'
 ATT = 'attention'
+VAL = 'validation'
+TEST = 'test'
+
+# layer
+MAX_POOLING = 'max_pooling'
+AVERAGE_POOLING = 'average_pooling'

### 2024-08-27 execute/main.py

@@ -27,21 +27,22 @@ def main():
     input_dict = parse_args()
 
     # data preparation
-    data_dict = get_data(look_back=input_dict['look_back_window'])
+    # set data_dir='/home/s2210434/fx/fxdata/multi_USDJPY'
+    data_dict = get_data(look_back=input_dict['look_back_window'], data_dir='/home/s2210434/fx/fxdata/multi_USDJPY')
 
-    # choose randomly 50% clients for training, 25% for validating and 25% for testing (only apply for multi_fx, not multi_USDJPY)
-    # if this is from `pretrained_macro_random`
-    list_id = set(data_dict.keys())
-    list_id_train = random.sample(sorted(list_id), 30)
-    list_id_val = random.sample(sorted(list_id.difference(set(list_id_train))), 15)
-    list_id_test = list(list_id.difference(set(list_id_train).union(list_id_val)))
+    # # if pretrained models are from `pretrained_macro_random`
+    # # choose randomly 50% clients for training, 25% for validating and 25% for testing (only apply for multi_fx, not multi_USDJPY)
+    # list_id = set(data_dict.keys())
+    # list_id_train = random.sample(sorted(list_id), 30)
+    # list_id_val = random.sample(sorted(list_id.difference(set(list_id_train))), 15)
+    # list_id_test = list(list_id.difference(set(list_id_train).union(list_id_val)))
 
-    # # if pretrained models are from `pretrained_macro`
-    # # if the data is multi_USDJPY
-    # list_id = list(data_dict.keys())
-    # list_id_train = list_id[:int(len(list_id)*0.5)]
-    # list_id_val = list_id[int(len(list_id)*0.5):int(len(list_id)*0.75)]
-    # list_id_test = list_id[int(len(list_id)*0.75):]
+    # if pretrained models are from `pretrained_macro`
+    # if the data is multi_USDJPY
+    list_id = list(data_dict.keys())
+    list_id_train = list_id[:int(len(list_id)*0.5)]
+    list_id_val = list_id[int(len(list_id)*0.5):int(len(list_id)*0.75)]
+    list_id_test = list_id[int(len(list_id)*0.75):]
 
     tf.config.threading.set_inter_op_parallelism_threads(input_dict['num_cpus'])
     tf.config.threading.set_intra_op_parallelism_threads(input_dict['num_cpus'])
@@ -56,7 +57,8 @@ def main():
     )
 
[…]

### 2024-08-27 model/meta_model.py

@@ -1,6 +1,6 @@
 import tensorflow as tf
 from keras import optimizers, models, metrics
-from sklearn.metrics import precision_recall_fscore_support
+# from sklearn.metrics import precision_recall_fscore_support
 import numpy as np
 
 from copy import deepcopy
@@ -9,6 +9,7 @@ import os
 
 from .based_model import get_model
 from common.constants import *
+from common.common import compute_metrics
 
 class MAML:
     def __init__(
@@ -46,7 +47,7 @@ class MAML:
             self.loss_fn = metrics.mean_squared_error
         elif mode == CLF:
             self.loss_fn = metrics.binary_crossentropy
-            self.acc_fn = metrics.Accuracy()
+            # self.acc_fn = metrics.Accuracy()
 
     def inner_training_step(self, model:models.Model, X:list[np.ndarray], y:list[np.ndarray], num_epochs:int=2):
         """fast adapt on support set (inner step)
@@ -110,9 +111,7 @@ class MAML:
 
                 # store prediction and true label for computing acc
                 if self.mode == CLF:
-                    pred = pred.numpy()
-                    pred[pred >= 0.5] = 1
-                    pred[pred < 0.5] = 0
+                    pred = (pred.numpy()>=0.5)*1.
                     pred_values = np.vstack((pred_values, pred))
                     true_values = np.vstack((true_values, batch_y))
 
@@ -121,7 +120,7 @@ class MAML:
 
             # compute acc + loss of a task
             if self.mode == CLF:
-                acc_, precision_, recall_, f1_ = self.compute_metrics(pred_values, true_values)
+                acc_, precision_, recall_, f1_ = compute_metrics(pred_values, true_values)
                 tasks_metrics['acc'].append(acc_)
                 tasks_metrics['precision'].append(precision_)
                 tasks_metrics['recall'].append(recall_)
@@ -134,13 +133,13 @@ class MAML:
 
             tasks_metrics['loss'].append(task_loss)
 
-        tasks_metrics['loss'], tasks_metrics['std_loss'] = self.compute_metrics(tasks_metrics['loss'])
+        tasks_metrics['loss'], tasks_metrics['std_loss'] = compute_metrics(tasks_metrics['loss'])
         print(f"\n\tMean loss:\t{tasks_metrics['loss']:.5f} ± {tasks_metrics['std_loss']:.5f}")
         if self.mode == CLF:
[…]

### 2024-08-27 data/pre_process.py
baseline stuff

@@ -6,10 +6,10 @@ import numpy as np
 from common.constants import *
 
 def create_label(df:np.ndarray, mode:str=CLF) -> np.ndarray:
+    label = df[:,-1]
     if mode == CLF:
-        label = df[1:,-1] - df[:-1,-1]
-        label[label > 0] = 1
-        label[label <= 0] = 0
+        label = label[1:] - label[:-1]
+        label = (label>0)*1.
         label = np.append([-1], label) # the first value is assigned to -1, it's removed anyway
     elif mode == REG:
         pass # if it is regression problem, we don't have to do anything
@@ -50,3 +50,113 @@ def get_data(look_back:int, test_size:float=0.8, batch_size:int=32, data_dir:str
             cid += 1
 
     return data_dict
+
+def get_data_for_nhits(data_dir:str=RAW_DATA_DIR):
+    for file_name in os.listdir(data_dir):
+        # data_dir should contain only 1 *.csv file
+        if file_name.endswith('csv'):
+            # read dataset, drop the datetime/index column
+            df = pd.read_csv(os.path.join(data_dir, file_name)).to_numpy()[:,1:]
+
+            # normalize data
+            transformed_df = normalize_data(df)
+
+            # create label
+            y = create_label(transformed_df)
+
+            # concat X, y
+            transformed_df = np.concatenate([transformed_df, y.reshape(-1,1)], axis=1)
+            transformed_df = pd.DataFrame(transformed_df, columns=['open', 'high', 'low', 'close', 'y'])
+            transformed_df['unique_id'] = 0
+            transformed_df['ds'] = pd.date_range(start='2000-01-01', periods=len(transformed_df), freq='h')
+            return transformed_df[['unique_id', 'ds', 'open', 'high', 'low', 'close', 'y']]
+
+# def create_label(df:np.ndarray, mode:str=CLF) -> np.ndarray:
+#     label = df[:,-1]
+#     if mode == CLF:
+#         label = label[1:] - label[:-1]
+#         label = (label > 0)*1.
+#         label = np.append([-1], label) # the first value is assigned to -1, it's removed anyway
+#     elif mode == REG:
+#         pass # if it is regression problem, we don't have to do anything
+#     return label
+
+# def normalize_data(df:np.ndarray) -> np.ndarray:
+#     scaler = StandardScaler()
+#     return scaler.fit_transform(df), scaler
+
[…]

### 2024-08-27 model/based_model.py

@@ -41,7 +41,7 @@ def get_model(model_name:str, input_shape:tuple, mode:str):
         x = _get_cnn(input_)
     elif model_name == LSTM_CNN:
         x = _get_lstm_cnn(input_)
-    elif model_name == ATT:
+    elif model_name == ATT: # I was planning to implement attention model
         pass
     else:
         raise ValueError('wrong model name')
@@ -52,5 +52,5 @@ def get_model(model_name:str, input_shape:tuple, mode:str):
         output_ = layers.Dense(units=1, activation='sigmoid')(x)
     return models.Model(inputs=input_, outputs=output_)
 
-# bla = get_model(CNN, (20,4), CLF)
+# bla = get_model(LSTM, (20,4), CLF)
 # bla.summary()

### 2024-08-28 doc/paper.tex
draft 2

@@
 % \usepackage{ulem}
-\usepackage[caption=false]{subfig}
+% \usepackage[caption=false]{subfig}
+\usepackage{caption}
+\usepackage{subcaption}
 \usepackage[usenames,dvipsnames]{color}
@@
     % \item Kết hợp đặc trưng: Rút trích đặc trưng bằng cách kết hợp các đặc trưng của \verb|LSTM| và \verb|CNN|.
-    \item Feature combination: Extract feature using \verb|LSTM| and \verb|CNN| and combine them.
+    \item \textbf{Feature combination}: Extract feature using \verb|LSTM| and \verb|CNN| and combine them.
 
     % \item Hidden long-term dependency: Chứng minh thực nghiệm rằng một aperiodic time-series data nhất định không chỉ phụ thuộc vào các yếu tố ngoại lai mà còn có các phụ thuộc ẩn với chính nó tại nhiều thời điểm khác nhau trong quá khứ.
-    \item Hidden long-term dependency: Experimentally demonstrate that a given aperiodic time-series data not only depends on external factors but also has hidden dependencies with itself at various points in the past.
+    \item \textbf{Hidden long-term dependency}: Experimentally demonstrate that a given aperiodic time-series data not only depends on external factors but also has hidden dependencies with itself at various points in the past.
 
     % \item Tổng hợp hiệu quả tham số mô hình: Sử dụng ML thay cho các mô hình ensemble truyền thống trong việc tổng hợp kết quả từ các mô hình học máy.
-    \item Efficient model parameter aggregation: Use ML instead of traditional ensemble models to aggregate results from machine learning models.
+    \item \textbf{Efficient model parameter aggregation}: Use ML instead of traditional ensemble models to aggregate results from machine learning models.
 
     % \item Experiment: Thực nghiệm trên các bộ dữ liệu về tỉ giá hối đoái và so sánh với mô hình SOTA \verb|NHITS| để chứng minh tính hiệu quả của phương pháp đề xuất.
-    \item Experiment: Experiment on exchange rate datasets and compare with \verb|NHITS| - the SOTA model to demonstrate the effectiveness of the proposed method.
+    \item \textbf{Experiment}: Experiment on exchange rate datasets and compare with \verb|NHITS| - the SOTA model to demonstrate the effectiveness of the proposed method.
 \end{itemize}
@@
 \vspace{2mm}
-
 \subsection{Experiment}
@@
 
-We compare the proposed method with \verb|NHITS| using the above metrics.
-The overall data is structured into training sets, validation sets, and testing sets with a ratio of 6:2:2.
-The training set is used to train the model, the validation set is used to find hyper-parameters, and the test set is used to evaluate the model.
-For \verb|NHITS|, we split the data as usual according to the above predetermined ratio.
-For the proposed algorithm, because the meta-training and meta-testing processes require dividing the data into small tasks, and allowing the model to adapt on the support set of each task, we split the data into 60 tasks.
+We compare the proposed method with \verb|NHITS| using the above metrics on \verb|USD/JPY| and \verb|multi-fx| datasets.
+We also test different feature extractors (\verb|LSTM|, \verb|CNN|) for the proposed method to demonstrate the deep data understanding capabilities of \verb|LSTM+CNN|.
+The overall data is structured into training sets, validation sets, and testing sets with a ratio of 6:2:2 for training, fine-tuning and testing model, respectively.
+
+\vspace{2mm}
+
+For \verb|NHITS| model, we split the data as usual according to the above predetermined ratio.
+For the proposed algorithm, because the meta-training and meta-testing processes require dividing the data into small tasks, and allowing the model to adapt on the support set of each task, we split the data into 60 tasks (see table \ref{tab:stat_} for statistical detail).
 In each task, the support set accounts for 20\% with the purpose of letting the model adapt to the data, the query set accounts for 80\% to check the model's compatibility.
@@
 With this division, we ensure the fairness that the ML model is trained with the same amount of data as \verb|NHITS|.
+After the data configuration, we perform experiment as in appendix \ref{app:our_experiment} and \ref{app:nhits_experiment}.
+
+\begin{table}
+    \centering
+    \caption{Statistics on USD/JPY and multi-fx. \textbf{The data is not correct, have to be re-written}}
+    \label{tab:stat_}
+    \begin{tabular}{ccccccc} 
+    \toprule
+    \multirow{2}{*}{Dataset} & \multirow{2}{*}{\#task} & \multirow{2}{*}{\#samples} & \multicolumn{4}{c}{\#samples/task}  \\ 
+    \cline{4-7}
+                             &                         &                            & min & mean  & std   & max             \\ 
+    \hline
+    \Verb|USD/JPY|           & 60                      & 69,909                     & 135 & 1,398 & 1,424 & 5,201           \\
+    \Verb|multi-fx|          & 60                      & 52,497                     & 506 & 1,049 & 250   & 1,986           \\
+    \bottomrule
+    \end{tabular}
+\end{table}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Result \& Discussion}
+\label{sec.results}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% nói 1 chút về việc xài lstm+cnn/meta-learning sẽ giảm sự phụ thuộc của kết quả vào mạng NN/siêu tham số
+% cụ thể, khi xài ML, tham số finetune rất dễ, cái nào cũng cao, chỉ cần chọn ra cái cao nhất
+% giải thích tại sao cnn perform poorly: quá chú ý đến đặc trưng cục bộ mà quên mất các phụ thuộc dài hạn
+
+% Bảng \ref{tab:rs} so sánh kết quả giữa mô hình đề xuất và mô hình \verb|NHITS| trên các tập dữ liệu kể trên.
+Phương pháp của chúng tôi đạt mức hội tụ cao hơn hẳn so với mô hình \verb|NHITS| thể hiện trên tất cả các metrics.
+Cụ thể, độ chính xác cải thiện từ 5\% đến 11\% so với \verb|NHITS| trên dữ liệu \verb|USD/JPY| và \verb|multi-fx|.
+Các metrics còn lại cũng đều tăng từ 5\% đến 20\%.
+Sau đây, chúng tôi phân tích kết quả này dựa trên hai yếu tố: (1) - Khả năng rút trích đặc trưng; (2) - Khả năng tổng hợp mô hình.
+
+Table \ref{tab:rs} compares the results between the proposed model and the \verb|NHITS| model on the above datasets.
+Our method achieves a much higher convergence rate than the \verb|NHITS| model on all metrics.
+Specifically, the accuracy improves from 5\% to 11\% compared to \verb|NHITS| on \verb|USD/JPY| and \verb|multi-fx| data.
+The remaining metrics also increase from 5\% to 20\%.
+Next, we analyze this result based on two factors: (1) - Feature extraction ability; (2) - Model synthesis ability.
 
@@
 
-% Trong cài đặt của chúng tôi, chúng tôi sử dụng một lớp \verb|FullyConnected| gồm 16 units với hàm kích hoạt \verb|ReLU| để phân giải đặc trưng ban đầu.
-Sau đó, đặc trưng này được truyền song song đến các hai khối \verb|BidirectionalLSTM| và \verb|CNN|.
-Khối \verb|BidirectionalLSTM| bao gồm 32 hidden units, các đầu ra được nối dài để tạo thành một vector cuối.
-Khối \verb|CNN| bao gồm hai layer \verb|CNN| có số filter lần lượt là 32 và 64.
-Kernel được sử dụng trong các layer có kích thước $3\times 3$.
-Theo sau mỗi layer \verb|CNN| là một layer \verb|MaxPooling| sử dụng kernel kích thước $2\times 2$.
-Kết thúc khối \verb|CNN| là một layer \verb|Flatten|.
-Đặc trưng của khối \verb|BidirectionalLSTM| và \verb|CNN| sau đó được nối lại rồi truyền qua một layer phân lớp nhị phân với hàm kích hoạt \verb|Sigmoid|.
-
-In our implementation, we use a \verb|FullyConnected| layer of 16 units with a \verb|ReLU| activation function to decompose the initial feature.
-This feature is then passed in parallel to the \verb|BidirectionalLSTM| and \verb|CNN| blocks.
-The \verb|BidirectionalLSTM| block consists of 32 hidden units, the outputs of which are concatenated to form a final vector.
-The \verb|CNN| block consists of two \verb|CNN| layers with 32 and 64 filters, respectively.
-The kernel used in the layers is of size $3\times 3$.
-Each \verb|CNN| layer is followed by a \verb|MaxPooling| layer using a kernel of size $2\times 2$.
-The \verb|CNN| block ends with a \verb|Flatten| layer.
-Features of the \verb|BidirectionalLSTM| and \verb|CNN| blocks are then concatenated and passed through a binary classification layer with a \verb|Sigmoid| activation function.
-
-% Đối với mô hình \verb|NHITS|, chúng tôi sử dụng
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Results and discussion}
-\label{sec.results}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{table*}[h]
+\begin{table*}
     \centering
@@
     \hline
-        \multirow{2}{*}{\Verb|USD/JPY|}    & \Verb|NHITS| & 53.29      & 52.38    & 52.23       & 51.87      \\
-                                    & Ours  & $\mathbf{58.05±0.01}$ & $\mathbf{59.57±2.53}$ & $\mathbf{58.31±2.57}$ & $\mathbf{56.6±3.74}$    \\ 
+    \multirow{3}{*}{\Verb|USD/JPY|}  & \Verb|NHITS|          & 53.29                 & 52.38                 & 52.23                 & 51.87                   \\
+                                                & Ours(\Verb|LSTM|)     & $not\_done$           & $59.57±2.53$          & $58.31±2.57$          & $56.6±3.74$             \\
+                                                & Ours(\Verb|LSTM+CNN|) & $\mathbf{58.05±0.01}$ & $\mathbf{59.57±2.53}$ & $\mathbf{58.31±2.57}$ & $\mathbf{56.6±3.74}$    \\ 
     \hline
-        \multirow{2}{*}{\Verb|multi-fx|} & \Verb|NHITS| & 51.12      & 51.22      & 51.16      & 50.54        \\
-                                    & Ours  & $\mathbf{62.39±0.06}$ & $\mathbf{71.11±9.34}$ & $\mathbf{70.14±9.6}$  & $\mathbf{69.21±11.07}$  \\
+    \multirow{3}{*}{\Verb|multi-fx|} & \Verb|NHITS|          & 51.12                 & 51.22                 & 51.16                 & 50.54                   \\
+                                                & Ours(\Verb|LSTM|)     & $61.23±0.06$          & $70.52±9.17$          & $69.51±9.31$          & $68.9±10.0$             \\
+                                                & Ours(\Verb|LSTM+CNN|) & $\mathbf{62.39±0.06}$ & $\mathbf{71.11±9.34}$ & $\mathbf{70.14±9.6}$  & $\mathbf{69.21±11.07}$  \\
     \bottomrule
@@
 
-% nói 1 chút về việc xài lstm+cnn/meta-learning sẽ giảm sự phụ thuộc của kết quả vào mạng NN/siêu tham số
-% cụ thể, khi xài ML, tham số finetune rất dễ, cái nào cũng cao, chỉ cần chọn ra cái cao nhất
-% giải thích tại sao cnn perform poorly: quá chú ý đến đặc trưng cục bộ mà quên mất các phụ thuộc dài hạn
+\subsection{Feature extraction ability}
+
+% Chúng tôi huấn luyện ba mô hình \verb|CNN|, \verb|LSTM|, \verb|LSTM+CNN| trên 2600 mẫu dữ liệu \verb|CHF/NZD| theo cách thông thường.
+Sau khi quan sát quá trình huấn luyện của chúng (hình \ref{fig:lstm_cnn}), chúng tôi nhận thấy khả năng rút đặc trưng cục bộ tuyệt vời trên dữ liệu ngắn hạn của \verb|CNN|.
+Mặt khác, \verb|LSTM| không cải thiện sau 40 epochs, chứng tỏ rằng nếu chỉ tập trung vào khai thác các phụ thuộc ngắn/dài hạn mà không rút trích đặc trưng đủ sâu thì sẽ không mang lại kết quả cao.
+Khi kết hợp cả hai đặc trưng, mô hình cho thấy sự cải thiện trong khả năng học trên tập train.
+Mức hội tụ nhìn chung đều tăng so với việc sử dụng từng mô hình riêng lẻ.
+Điều này có thể được lý giải thông qua việc \verb|LSTM| đã tích hợp thành công các long-term feature vào mô hình, cũng như việc tận dụng các đặc trưng cục bộ của \verb|CNN|.
+
+We trained three models \verb|CNN|, \verb|LSTM|, \verb|LSTM+CNN| on 2600 samples of \verb|CHF/NZD| data in the conventional way.
+After observing their training process (figure \ref{fig:lstm_cnn}), we found that \verb|CNN| has excellent local feature extraction ability on short-term data.
+On the other hand, \verb|LSTM| does not improve after 40 epochs, indicating that focusing on exploiting short/long-term dependencies without extracting deep enough features will not yield good results.
+When combining both features, the model shows an improvement in learning ability on the training set.
+The convergence rate is generally higher than when using each model separately.
+This can be explained by the fact that \verb|LSTM| successfully integrated long-term features into the model, as well as taking advantage of the local features of \verb|CNN|.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{img/models.pdf}
+    \caption{Training process of LSTM, CNN, and LSTM+CNN using 2600 samples from CHF/NZD dataset.}
+    \label{fig:lstm_cnn}
+\end{figure}
+
+\vspace{2mm}
+
+% Tuy nhiên, khi quan sát kết quả trên tập validation, cả ba mô hình trên không đều cho kết quả rất tệ.
+Lý giải cho điều này, chúng tôi nhận định rằng dù có thể học tốt trên tập train nhưng các mô hình đều mất đi tính tổng quát hóa vì đây là loại dữ liệu phức tạp, khác với các dữ liệu như hình ảnh, video, hoặc dữ liệu có chu kỳ.
+Nếu khả năng thích ứng của mô hình không cao, kết quả tệ trên tập validation là điều có thể lường trước được.
+Mặc dù vậy, \textbf{đồ thị trong hình \ref{fig:lstm_cnn} cho thấy chúng ta không thể phủ nhận khả năng rút trích đặc trưng của} \verb|LSTM+CNN|.
+
+However, when observing the results on the validation set, all three models above give very poor results.
+To explain this, we think that although they can learn well on the training set, the models lose their generalization ability because this is a complex type of data, different from data such as images, videos, or periodic data.
+If the model's adaptability is not high, poor results on the validation set are predictable.
+However, \textbf{the graph in figure \ref{fig:lstm_cnn} shows that we cannot deny the feature extraction ability of} \verb|LSTM+CNN|.
+
+\vspace{2mm}
+
+% Sau này, khi sử dụng thuật toán ML để huấn luyện các mạng rút trích đặc trưng, chúng tôi còn nhận thấy mô hình sử dụng đặc trưng kết hợp ít bị phụ thuộc vào learning rate hơn các mô hình sử dụng đặc trưng đơn lẻ.
+Cụ thể, đối với không gian learning rate mà chúng tôi định nghĩa trong Appendix \ref{app:lr}, tất cả các mô hình đều hội tụ với độ chính xác thấp nhất từ 54\%.
+Việc này minh chứng cho khả năng ưu việt của đặc trưng do \verb|LSTM+CNN| mang lại.
+
+Later, when we used ML algorithms to train feature extraction networks, we also found that the model using combined features was less dependent on the learning rate than the model using single features.
+Specifically, for the learning rate space we defined in Appendix \ref{app:lr}, all models converged with an accuracy of at least 54\%.
+This demonstrates the superiority of the feature provided by \verb|LSTM+CNN|.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{img/acc.pdf}
+    \caption{Convergence process on validation set of neural network using LSTM+CNN feature and LSTM feature only (multi-fx dataset). The blur domains cover 99.73\% ($±3\sigma$) values of accuracies.}
+    \label{fig:lstm_cnn_feature}
+\end{figure}
+
+% Quan sát quá trình rút trích đặc trưng của \verb|NHITS|, có thể thấy mô hình này đang cố gắng mô phỏng lại quá trình phân giải tần số của Fourier transform.
+Tuy nhiên, trong toàn bộ quá trình thực thi, \verb|NHITS| sử dụng các kiến trúc rất đơn giản (chỉ bao gồm các lớp \verb|MLP| và \verb|Pooling|).
+Đối với các dữ liệu phức tạp và phi chu kỳ, sử dụng các layer \verb|MLP| là không đủ để rút trích các đặc trưng ẩn.
+
+Looking at the feature extraction process of \verb|NHITS|, it can be seen that this model tries to simulate the frequency resolution process of Fourier transform.
+However, in the whole implementation, \verb|NHITS| uses a very simple architecture (only including \verb|MLP| and \verb|Pooling| layers).
+For complex and non-periodic data, using \verb|MLP| layers is not enough to extract hidden features.
+
+\subsection{Sub-model synthesis ability}
+
+% Lý do phương pháp huấn luyện thông thường không đạt hiệu quả cao mặc cho khả năng rút trích đặc trưng đã được cải thiện là vì thiếu khả năng sử dụng đặc trưng một cách hiệu quả.
+Khả năng này trong các nghiên cứu trước đây thường được cải thiện bằng việc sử dụng các mô hình ensemble.
+Chúng tôi đã thử sử dụng các mô hình truyền thống (boosting, stacking) nhằm tăng độ chính xác của mô hình nhưng sớm nhận ra rằng phương pháp này không khả thi.
+Ngay cả khi chúng tôi vét cạn cạn trên không gian hyper-parameter để tìm ra kiến trúc mô hình con tốt nhất bằng \verb|AutoKeras| \cite{jin2023autokeras}, độ chính xác thu được trên tập \verb|USD/JPY| chỉ ở mức 52.53\% và 53.71\%, respectively.
+
+The reason why conventional training methods do not perform well despite the improved feature extraction capabilities is the lack of the ability to use features effectively.
+This ability in previous studies has often been improved by using ensemble models.
+We tried using traditional models (boosting, stacking) to increase the model accuracy but soon realized that this method was not feasible.
+Even when we performed brute-force search the hyper-parameter space to find the best sub-model architecture using \verb|AutoKeras| \cite{jin2023autokeras}, the accuracy on \verb|USD/JPY| was only 52.53\% and 53.71\%, respectively.
+
+\vspace{2mm}
+
+% Khi sử dụng \verb|MAML| để tổng hợp đặc trưng của mô hình \verb|LSTM| và \verb|LSTM+CNN|, khả năng tương thích (thể hiện qua việc hội tụ sớm) cũng như kết quả (thể hiện qua độ chính xác) của mô hình sử dụng \verb|LSTM+CNN| cao hơn hoàn toàn so với mô hình chỉ sử dụng \verb|LSTM| (xem hình \ref{fig:lstm_cnn_feature}).
+Thật vậy, chỉ sau 40 epochs, các đặc trưng \verb|LSTM+CNN| cho mức hội tụ trên 60\% và sau 100 epochs, mô hình hội tụ ở mức 62\%, vượt trên những gì các đặc trưng \verb|LSTM| có thể làm được.
+Độ biến thiên của các giá trị accuracy là rất nhỏ, cho thấy sự ổn định của \verb|MAML|.
+Đối với mô hình ML chỉ sử dụng \verb|CNN|, độ chính xác cũng như giá trị lỗi thậm chí còn không cải thiện trong toàn bộ quá trình huấn luyện.
+Giải thích cho việc này, chúng tôi cho rằng các đặc trưng của \verb|CNN| đã quá tập trung vào tính cục bộ của dữ liệu mà bỏ qua các phụ thuộc dài hạn, vốn là một đặc trưng quan trọng của dữ liệu time-series.
+Trái lại, đặc trưng của \verb|LSTM+CNN| không chỉ nắm bắt được các phụ thuộc dài hạn mà còn có thể highlight các tính chất cục bộ trong dữ liệu, khiến cho nó hoàn toàn outperform \verb|CNN| cũng như \verb|NHITS|.
+Do đó, \textbf{chúng tôi kết luận rằng ML tỏ ra linh hoạt và đạt hiệu quả cao trong việc tổng hợp các sub-model so với các mô hình ensemble cứng nhắc truyền thống}.
+
+When using \verb|MAML| to synthesize the features of the \verb|LSTM| and \verb|LSTM+CNN| models, the compatibility (indicated by early convergence) as well as the results (indicated by accuracy) of the model using \verb|LSTM+CNN| are significantly higher than those of the model using only \verb|LSTM| (see figure \ref{fig:lstm_cnn_feature}) and the \verb|NHITS| model.
+Indeed, after only 40 epochs, the \verb|LSTM+CNN| features converge to over 60\% and after 100 epochs, the model converges to 62\%, exceeding what the \verb|LSTM| features can do.
+The variation of the accuracy values ​​is very small, indicating the stability of \verb|MAML|.
+For the ML model using only \verb|CNN|, the accuracy and error values ​​do not even improve during the entire training process.
+We explain this by saying that the features of \verb|CNN| focus too much on the locality of the data and ignore the long-term dependencies, which is an important feature of time-series data.
+In contrast, the features of \verb|LSTM+CNN| not only capture the long-term dependencies but also highlight the local properties in the data, making it completely outperform \verb|CNN| as well as \verb|NHITS|.
+Therefore, \textbf{we conclude that ML is more flexible and efficient in synthesizing sub-models than traditional rigid ensemble models}.
+
+\vspace{2mm}
+
+% Khi tiếp cận dữ liệu \verb|USD/JPY| bằng ML, mô hình cũng cho độ chính xác cao.
+Bằng việc nhìn vào các khoảng thời gian trong quá khứ, \textbf{phương pháp đề xuất đã tỏ ra hiệu quả hơn trong việc nắm bắt các hidden long-term dependency thông qua việc tổng hợp hiệu quả các đặc trưng mà} \verb|LSTM| \textbf{dễ dàng quên đi trong quá trình huấn luyện}.
+Điều này đã chứng minh giả thuyết về sự tồn tại của các hidden long-term dependency mà chúng tôi nêu ra ở phần \ref{sec.intro}.
+
+When approaching the \verb|USD/JPY| data using ML, the model also shows high accuracy.
+By looking at the past time periods, \textbf{the proposed method has proven to be more effective in capturing hidden long-term dependencies through the efficient aggregation of features that the} \verb|LSTM| \textbf{easily forgets during training}.
+This proves the hypothesis of the existence of hidden long-term dependencies that we stated in section \ref{sec.intro}.
+
+% dis NHITS vì tổng hợp đặc trưng chỉ bằng phép cộng :v 
+% Trong quá trình tổng hợp các đặc trưng để đưa ra dự đoán cuối, \verb|NHITS| chỉ sử dụng các phép cộng trừ đơn giản.
+Điều này đặt trong ngữ cảnh của việc tổng hợp thông tin dựa trên các hàm nội suy và phần dư của input là hợp lý.
+Tuy nhiên, đối diện với sự phức tạp trong dữ liệu đòi hỏi một cấu trúc phức tạp hơn để có thể tổng hợp các đặc trưng một cách hợp lý hơn.
+Việc sử dụng nội suy trên dữ liệu có thể coi là rất đơn giản so với quá trình meta-optimization của các thuật toán ML.
+Hơn nữa, việc phân tách và kết hợp các dải tần số dữ liệu của \verb|NHITS| chỉ có thể hoạt động tốt nếu dữ liệu thực sự tồn tại chu kỳ.
+Dữ liệu phi chi kỳ như FX hoặc stock price có thể là điểm yếu chí mạng cho hướng tiếp cận này.
+
+In the process of synthesizing features to make the final prediction, \verb|NHITS| only uses simple addition and subtraction.
+This is reasonable in the context of synthesizing information based on interpolation functions and input residuals and trying to simulate the synthesis of frequency bands.
+However, facing complexity in data requires a more complex structure to be able to synthesize features in a more reasonable way.
+Using interpolation on the data can be considered very simple compared to the meta-optimization process of ML algorithms.
+Furthermore, \verb|NHITS|'s separation and combination of data frequency bands can only work well if the data is truly periodic.
+Non-periodic data such as FX or stock prices can be a fatal weakness for this approach.
 
@@
 \label{sec.conc}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% Các bài toán trên aperiodic time-series data đặt ra một thử thách lớn cho các mô hình học máy hiện nay bởi tính bất định trong variance cũng như tính phi chu kỳ loại dữ liệu này.
+Bằng việc kết hợp các đặc trưng cục bộ và các phụ thuộc dài hạn cũng như khai thác các phụ thuộc ẩn trong từng khoảng thời gian quá khứ, chúng tôi đã cải khả năng của mô hình học máy trên các classification metrics trong bài toán dự đoán xu hướng giá của ngày tiếp theo trên dữ liệu FX.
+Thuật toán đề xuất đã chứng minh được khả năng vượt trội của mình khi so sánh với mô hình \verb|NHITS|.
+
+Problems on aperiodic time-series data pose a great challenge to current machine learning models due to the uncertainty in variance as well as the aperiodicity of this type of data.
+By combining local features and long-term dependencies as well as exploiting hidden dependencies in each past time period, we have improved the ability of machine learning models on classification metrics in the problem of predicting the next day's price trend on FX data.
+The proposed algorithm has demonstrated its superior ability when compared with the \verb|NHITS| model.
+
+\vspace{2mm}
+
+% Trong tương lai, chúng tôi đề xuất hai hướng phát triển chính liên quan đến kiến trúc và tính cá nhân hóa của mô hình học.
+
+In the future, we propose two main directions of development related to the architecture and personalization of the learning model.
+
+\vspace{2mm}
+
+% \textbf{Model architecture}.
+Phương pháp của chúng tôi được phát triển dưới dạng module với hai module chính hoạt động song song: Module rút trích đặc trưng và Module tổng hợp mô hình.
+Điều này cung chấp cho phương pháp của chúng tôi một khả năng nâng cấp linh hoạt.
+Thí nghiệm của chúng tôi chỉ minh họa một trường hợp điển hình trong việc rút trích và tổng hợp hiệu quả các đặc trưng.
+Bằng việc thay thế các mô hình rút trích đặc trưng khác nhau và sử dụng các thuật toán ML khác nhau, hoàn toàn có thể tạo ra mô hình mới với độ chính xác cao hơn.
+
+\textbf{Model architecture}.
+Our method is modularized with two main modules operating in parallel: Feature Extraction Module and Model Synthesis Module.
+This provides our method with flexible scalability.
+Our experiment only illustrates a typical case of efficient feature extraction and synthesis.
+By replacing different feature extraction models and using different ML algorithms (\verb|Meta-SGD| \cite{li2017meta}, \verb|Reptile| \cite{nichol2018first}, \verb|iMAML| \cite{rajeswaran2019meta}), it is possible to create new models with higher accuracy.
+
+\vspace{2mm}
+
+% \textbf{Long-horizon problem}.
+Hoàn toàn có thể mở rộng phương pháp này để giải các bài toán về long-horizon prediction.
+Thật vậy, bằng việc thay đổi đầu ra và độ lỗi của mô hình, có thể tiến hành giải các bài toán này.
+Tuy nhiên, kiến trúc của các sub-model cần được nghiên cứu lại để phù hợp hơn với bài toán mới.
+
+\textbf{Long-horizon problem}.
+It is possible to extend this method to solve long-horizon prediction problems.
+Indeed, by changing the output and error of the model, it is possible to solve these problems.
+However, the architecture of the sub-models needs to be re-examined to better suit the new problem.
 
@@
 % cite iMAML, Reptile vào đây để làm future direction kiểu tối ưu lại ML vì chi phí hơi đắt
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% Conclusion
+% apply các kỹ thuật cá nhân hóa để giảm thiểu variance trong kq vì variance hơi cao
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% All authors contributed to conceiving the idea.
-T.I., A.V.U, and G.I.P performed the calculations.
-T.I.
-made the all figures.
-R.M and S.V.S supervised the work.
-All authors contributed to the discussion and writing of the paper.
+All authors contributed to conceiving the idea. Bao-Long Nguyen, Tom Ichibha performed calculations. All authors contributed to the discussion and writing of the paper.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% The datasets used and/or analyzed during the current study available from the corresponding authors on reasonable request.
+The datasets used and/or analyzed during the current study available from the corresponding authors on reasonable request.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\bibliographystyle{elsarticle-harv} 
+\clearpage
+% \bibliographystyle{elsarticle-harv} 
+\bibliographystyle{splncs04} 
 \bibliography{references}
-\end{document}
+\clearpage
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\section{Appendix}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%We used the VASP code\cite{VASP} for the calculations explained in this appendix.
-%We relaxed the four structures of Figure \ref{fig.str}
-%for every LiV$X_2$ ($X$ = O, S, and Se) with each of PBE\cite{PBE},
-%optPBE-vdW-DF\cite{2009JK_AM,vdW-DFa,vdW-DFb}, and HSE06\cite{HSE06} functionals.
-%We evaluated the relative energies of structures relaxed by every functional by PBE-DFT.
-%%as shown in Figure \ref{fig.geomErr}.
-%We found that the differences of the relative energies
-%among the structures relaxed by different functionals are at most 0.018 eV/f.u.
-%%the differences of the time-series in Figure \ref{fig.geomErr} are within 0.018 eV/f.u.
+\appendix
+\section{Experimental detail\\for proposed method}
+\label{app:our_experiment}
 
-%%----------------------------------------
-%\begin{figure}[htbp]
-%  \centering
-%  \includegraphics[width=\hsize]{geomErr.eps}
-%  \caption{\ghost{fig.geomErr}\label{fig.geomErr}
-%    FNDMC and DFT results of relative energies of the four candidate structures
-%    of LiV$X_2$ (X = O, S, and Se).
-%    The candidate structures are shown in Figure \ref{fig.str}.
-%    TR-I and TR-II (left from the vertical line) are trimmerized
-%    and NTR-I and NTR-II (right from the vertical line) are
-%    non-trimmerized structures, respectively.
-%    FNDMC results are same as Figure \ref{fig.dmconly}.
-%    The lowest energy among the four structures is set to be zero
-%    and the others are given as relative values from the lowest energy.
-%  }
-%\end{figure}
-%%----------------------------------------
+% Trong cài đặt của chúng tôi, chúng tôi sử dụng một lớp \verb|FullyConnected| gồm 16 units với hàm kích hoạt \verb|ReLU| để phân giải đặc trưng ban đầu.
+Sau đó, đặc trưng này được truyền song song đến các hai khối \verb|BidirectionalLSTM| và \verb|CNN|.
+Khối \verb|BidirectionalLSTM| bao gồm 32 hidden units, các đầu ra được nối dài để tạo thành một vector cuối.
+Khối \verb|CNN| bao gồm hai layer \verb|CNN| có số filter lần lượt là 32 và 64.
+Kernel được sử dụng trong các layer có kích thước $3\times 3$.
+Theo sau mỗi layer \verb|CNN| là một layer \verb|MaxPooling| sử dụng kernel kích thước $2\times 2$.
+Kết thúc khối \verb|CNN| là một layer \verb|Flatten|.
+Đặc trưng của khối \verb|BidirectionalLSTM| và \verb|CNN| sau đó được nối lại rồi truyền qua một layer phân lớp nhị phân với hàm kích hoạt \verb|Sigmoid|.
+
+We use a \verb|FullyConnected| layer of 16 units with a \verb|ReLU| activation function to decompose the initial feature.
+This feature is then passed in parallel to the \verb|BidirectionalLSTM| and \verb|CNN| blocks.
+The \verb|BidirectionalLSTM| block consists of 32 hidden units, the outputs of which are concatenated to form a final vector.
+The \verb|CNN| block consists of two \verb|CNN| layers with 32 and 64 filters, respectively.
+The kernel used in the layers is of size $3\times 3$.
+Each \verb|CNN| layer is followed by a \verb|MaxPooling| layer using a kernel of size $2\times 2$.
+The \verb|CNN| block ends with a \verb|Flatten| layer.
+Features of the \verb|BidirectionalLSTM| and \verb|CNN| blocks are then concatenated and passed through a binary classification layer with a \verb|Sigmoid| activation function.
+
+\vspace{2mm}
+
+The fine-tuning process of ML algorithms involves many hyper-parameters such as inner batch size, outer batch size, inner training steps, outer training steps.
+To facilitate fine-tuning, we fix most of the parameters and only fine-tune the size of lookback window, the inner and outer learning rates.
+Details are presented in table \ref{tab:our_finetune}.
+
+\begin{table}[h]
+    \centering
+    \caption{Search space for fine-tuning our method.}
+    \label{tab:our_finetune}
+    \begin{tabular}{ll} 
+    \toprule
+    \multicolumn{1}{c}{Hyper-parameter} & \multicolumn{1}{c}{Seach space}                                            \\ 
+    \hline
+    Inner batch size (samples/batch)    & \{32\}                                                                     \\
+    Inner training step                 & \{3\}                                                                      \\
+    Outer batch size (tasks/batch)      & \{5\}                                                                      \\
+    Outer training step                 & \{100\}                                                                    \\ 
+    \hline
+    Lookback window                     & \{10, 20, 30\}                                                             \\
+    Inner learning rate                 & \{0.001, 0.005, 0.01, 0.05\}                                               \\
+    Outer learning rate                 & \begin{tabular}[c]{@{}l@{}}\{0.001, 0.005, 0.0015,\\0.0055\}\end{tabular}  \\
+    \bottomrule
+    \end{tabular}
+\end{table}
+
+\section{Experimental detail for NHITS}
+\label{app:nhits_experiment}
+
+% Đối với mô hình \verb|NHITS|, chúng tôi dựa trên \cite{challu2023nhits} để định nghĩa không gian tìm kiếm cho việc fine-tune tham số, cũng như kiến trúc mô hình.
+Đối với các tham số không được đề cập trong bảng, chúng tôi sử dụng giá trị mặc định của cài đặt \verb|NHITS| trong thư viện \verb|NeuralForecast| \cite{neuralforecast}.
+Chi tiết quá trình fine-tune được trình bày trong phần phụ lục \ref{app:experiment}.
+Kết quả tốt nhất của các lần fine-tune được chọn ra và báo cáo trong nghiên cứu này.
+
+\begin{table}
+    \centering
+    \caption{Search space for fine-tuning NHITS.}
+    \label{tab:nhits_finetune}
+    \begin{tabular}{ll} 
+    \toprule
+    \multicolumn{1}{c}{Hyper-parameter} & \multicolumn{1}{c}{Seach space}                                                                         \\ 
+    \hline
+    Random seed                         & \{1\}                                                                                                   \\
+    Number of stacks                    & \{3\}                                                                                                   \\
+    Number of blocks in each stack      & \{1\}                                                                                                   \\
+    Activation function                 & \{ReLU\}                                                                                                \\
+    Batch size                          & \{256\}                                                                                                 \\
+    Epoch                               & \{500\}                                                                                                 \\
+    \hline
+    Lookback window                     & \{5, 20, 30\}                                                                                           \\
+    Pooling kernel                      & \begin{tabular}[c]{@{}l@{}}\{{[}2,2,2], [4,4,4], [8,8,8],\\{[}8,4,1], [16,8,1]\}\end{tabular}           \\
+    Stacks' coefficients                & \begin{tabular}[c]{@{}l@{}}\{{[}168,24,1], [24,12,1],\\{[}180,60,1],[40,20,1], [64,8,1]\}\end{tabular}  \\
+    Number of MLF layers                & \{1,2\}                                                                                                 \\
+    Learning rate                       & \begin{tabular}[c]{@{}l@{}}\{0.001, 0.002, 0.005, 0.01,\\0.02\}\end{tabular}                            \\
+    \bottomrule
+    \end{tabular}
+\end{table}
+
+We rely on \cite{challu2023nhits} to define the search space (table \ref{tab:nhits_finetune}) for parameter fine-tuning, as well as to fine-tune the model architecture.
+For parameters not mentioned in the table, we use the default values of the \verb|NHITS| implementation in the \verb|NeuralForecast| library \cite{neuralforecast}.
+The best results of fine-tuning process are selected and reported in this study.
+
+\section{Hyper-parameter dependence}
+\label{app:lr}
+
+% Trong quá trình thực nghiệm, chúng tôi nhận thấy việc hội tụ của mô hình ML sử dụng đặc trưng của \verb|LSTM+CNN| có thể giảm bớt sự phụ thuộc vào learning rate.
+Cụ thể, chúng tôi thử nghiệm trên 32 mô hình trên dữ liệu \verb|multi-fx|.
+Inner learning rate và outer learning rate của mỗi mô hình được lấy ngẫu nhiên trong khoảng $[0.001, 0.05]$ và $[0.001, 0.009]$.
+Cần phải hiểu rằng, đối với outer learning rate, đây là một khoảng rất rộng vì thông thường, giá trị này luôn rất nhỏ so với inner learning rate.
+
+During the experiment, we found that the convergence of ML models using the \verb|LSTM+CNN| feature can reduce the dependence on learning rate.
+Specifically, we tested 32 models on \verb|multi-fx| data.
+The inner learning rate and outer learning rate of each model were randomly selected in the range $[0.001, 0.05]$ and $[0.001, 0.009]$.
+It is important to understand that, for outer learning rate, this is a very wide range because usually, this value is always very small compared to inner learning rate.
+
+\vspace{2mm}
+
+% Quá trình hội tụ trên tập validation của các mô hình được ghi lại trong hình \ref{fig:lr}.
+Theo đó, toàn bộ mô hình đều đạt hội tụ với độ chính xác thấp nhất là trên 54\%.
+Hơn một nửa số mô hình đạt mức hội tụ cao trên 60\%.
+Các mô hình còn lại phần lớn giao động trong khoảng 58\%-60\%.
+
+The convergence process on the validation set of the models is recorded in figure \ref{fig:lr}.
+Accordingly, all models achieved convergence with the lowest accuracy of over 54\%.
+More than half of the models achieved a high convergence level of over 60\%.
+The remaining models mostly fluctuated in the range of 58\%-60\%.
+
+\vspace{2mm}
+
+% Khi thực hiện thí nghiệm tương tự với mô hình \verb|LSTM|, độ chính xác phụ thuộc rất nhiều vào việc lựa chọn các siêu tham số học.
+Khả năng tương thích trên dữ liệu cũng giảm đáng kể.
+Đối với mô hình \verb|CNN|, quá trình hội tụ thậm chí còn không xảy ra.
+
+When performing similar experiments with the \verb|LSTM| model, the accuracy depends heavily on the choice of learning hyper-parameters.
+The adaptability to the data also drops significantly.
+For the \verb|CNN| model, convergence does not even occur.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\linewidth]{img/lr.pdf}
+    \caption{Convergence of 32 models LSTM+CNN.}
+    \label{fig:lr}
+\end{figure}
+
+\vspace{2mm}
+
+% tôi có thể viết thêm 1 phần phụ lục nhằm phân tích tác động của lookback window
+% Tuy vậy, quá trình hội tụ của chúng tôi lại phụ thuộc rất lớn vào lookback window.
+Thử nghiệm trên mô hình \verb|LSTM| với dữ liệu \verb|NHITS| cho thấy, khi xét trên cùng một tập siêu tham số và cùng một kiến trúc mô hình, nếu tăng lookback window từ 10 lên 20, số lượng mô hình hội tụ tăng đột biến.
+Ngoài ra, mức hội tụ mô hình cũng được cải thiện đáng kể.
+Khi tăng lookback window từ 20 lên 30, xuất hiện một số mô hình tiềm năng, mà độ chính xác của những mô hình này có thể tiếp tục tăng nếu tiếp tục huấn luyện.
+Đây là một insight thú vị cho thấy nếu nhìn vào quá khứ đủ nhiều và đủ sâu, chúng ta có thể tăng độ chính xác trong các dự đoán tương lai.
+Nếu dữ liệu quá khứ nhiều hơn, có thể sẽ phải tăng kích thước mô hình để đáp ứng việc hiểu được dữ liệu này.
+
+However, the convergence process of the proposed method depends heavily on the lookback window.
+Experiments on the \verb|LSTM| model with \verb|NHITS| data show that, when considering the same set of hyper-parameters and the same model architecture, if the lookback window is increased from 10 to 20, the number of converged models increases dramatically.
+In addition, the model convergence level is also significantly improved.
+When increasing the lookback window from 20 to 30, some potential models appear, the accuracy of which can continue to increase if training continues.
+This is an interesting insight showing that if we look into the past enough and deeply enough, we can increase the accuracy in future predictions.
+If the past data is larger, we may have to increase the model size to meet the understanding of this data.
+
+\begin{figure}
+    \centering
+    \begin{subfigure}[b]{\linewidth}
+        \centering
+        \includegraphics[width=\textwidth]{img/lstm_10.pdf}
+        \caption{lookback window = 10}
+        \label{fig:image1}
+    \end{subfigure}
+    % \hfill
+    % \vspace{2mm}
+    \begin{subfigure}[b]{\linewidth}
+        \centering
+        \includegraphics[width=\textwidth]{img/lstm_20.pdf}
+        \caption{lookback window = 20}
+        \label{fig:image2}
+    \end{subfigure}
+    % \hfill
+    % \vspace{2mm}
+    \begin{subfigure}[b]{\linewidth}
+        \centering
+        \includegraphics[width=\textwidth]{img/lstm_30.pdf}
+        \caption{lookback window = 30}
+        \label{fig:image3}
+    \end{subfigure}
+    \caption{Change in the model's convergence process when changing the lookback window.}
+    \label{fig:three_images}
+\end{figure}
+
+\end{document}

### 2024-09-10 doc/paper/paper.tex
change directory of paper stuff

--- 2024-08-28 doc/paper.tex
@@
 % \usepackage[caption=false]{subfig}
+\usepackage{float}
 \usepackage{caption}
@@
 %
-\author{Tom Ichibha}
-\email{ichibha@icloud.com}
-\affiliation{School of Information Science, JAIST, Ishikawa, Japan}
-%
 \author{Kenta Hongo}
@@
 %
+\author{Tom Ichibha}
+\email{ichibha@icloud.com}
+\affiliation{School of Information Science, JAIST, Ishikawa, Japan}
+%
 \date{\today}
@@
 Predicting aperiodic time-series data (e.g.
-stock price, foreign exchange, Bitcoin price,...) is a difficult task for machine learning models because this type of data has high variance and is not stationary; does not show clear cycles, making it difficult to extract features; depends not only on past values but also on external factors, which make them unstable and non-cyclical such as economic and political situations.
+stock price, foreign exchange, Bitcoin price,...) is a difficult task for machine learning models because this type of data has high and not stationary variance; does not show clearly periodicity, making it difficult to extract features; depends not only on past values but also on external factors, which make them unstable and non-cyclical, such as economic and political situations.
 To overcome the above challenges, we use Meta-learning to train a combined \verb|LSTM| and \verb|CNN| network, thereby effectively extracting and synthesizing hidden features of the data over time.
-Experiments on foreign exchange data of 60 currency pairs over 24 years (2000-2024) show that the proposed method performs well and has higher accuracy than the \verb|NHITS| - the state-of-the-art model in 2023 on time-series data, in the task of predicting the trend (upward or downward) of the next trading day.
+Experiments on foreign exchange data of 60 currency pairs over 24 years (2000-2024) show that the proposed method performs well and has 5\%-11\% higher accuracy compared to the \verb|NHITS| - the state-of-the-art model in 2023 on time-series data, in the task of predicting the trend (upward or downward) of the next trading day.
 
@@
 
-Aperiodic time-series prediction in general or foreign exchange (FX) prediction in particular has long been a matter of concern for many researchers \citep{li2019multi, islam2021foreign, heryadi2021foreign}.
+Aperiodic time-series prediction in general or foreign exchange (FX) prediction in particular has been a matter of concern for many researchers \citep{li2019multi, islam2021foreign, heryadi2021foreign} for a long time.
 The two main techniques used in aperiodic time-series prediction are fundamental analysis and technical analysis \cite{ayitey2023forex}.
-While fundamental analysis focuses on analyzing external factors, which are difficult to be captured from past value fluctuations such as policies and economic strategies of companies and countries to predict the future; technical analysis relies entirely on historical value fluctuations to analyze future trends.
+Fundamental analysis focuses on analyzing external factors such as policies and economic strategies of companies and countries to predict the future.
+Meanwhile, technical analysis relies entirely on historical value fluctuations, which are difficult to capture external factors, to analyze future trends.
 
@@
 
-Predicting on aperiodic time-series data faces several inherent challenges: (1) - The variance of this type of data varies greatly over time.
-Therefore, the assumption that they follow a distribution to approximate the error cannot be used, leading to machine learning models having difficulty accurately predicting future values and trends; (2) - Aperiodic data does not follow any explicit rules, so learning the hidden features of the data to make predictions is very difficult; (3) - Aperiodic time-series data (e.g.
-a company's stock price) depends not only on past data but also on many external factors (e.g.
-news, economic situation, politics) \cite{li2019multi}.
+Aperiodic time-series data prediction faces several inherent challenges: (1) - The variance varies significantly over time, making it difficult to assume a specific distribution to approximate errors, which is a challenge for machine learning models in accurately predicting future values and trends.
+(2) - Aperiodic data does not follow explicit patterns, making it challenging to identify the latent features required for reliable predictions.
+(3) - Aperiodic time-series data is influenced not only by historical data but also by numerous external factors \cite{li2019multi}.
+For example, a company's stock price is affected by news, economic conditions, political events, and other external influences.
 
@@
 Ensemble model giúp cung cấp một cái nhìn tổng quát, tổng hợp từ nhiều khía cạnh dựa trên các sub-models, từ đó giúp mô hình tổng quát thích ứng được với sự thay đổi mạnh của variance.
-Chúng tôi tiếp cận bài toán theo hướng này nhưng ở mức cao hơn bằng cách sử dụng Meta-learning (ML) \cite{finn2017model}.
-Phương pháp này tổng hợp hiệu quả tham số của các mô hình cục bộ, giúp giảm thiểu đáng kể variance loss.
-
-For the first challenge, ensemble models [cite something here] are often used to mitigate the effects of variance variation.
-Ensemble models provide a holistic, multi-perspective view based on sub-models, thereby helping the overall model adapt to strong variance changes.
-We approach the problem in a similar but higher-level way using Meta-learning (ML) \cite{finn2017model}.
-This method effectively aggregates the parameters of local models, which helps to significantly reduce variance loss.
+% Tuy nhiên, dưới sự tổng hợp cứng nhắc của các mô hình ensemble truyền thống, kết quả thu được trên các bộ dữ liệu phi chi kỳ là rất thấp.
+
+For the first challenge, ensemble models \citep{sadeghi2021combined, zafeiriou2020intraday, ali2020complete} are often used to mitigate the effects of variance variation.
+Ensemble models provide a holistic and multi-perspective view based on sub-models, thereby helping the overall model adapt to strong variance changes.
+However, under the rigid synthesis of traditional ensemble models, the results obtained on non-periodic datasets are very poor.
 
@@
 Cụ thể, trong năm 2022, 20\% tổng số các bài báo liên quan đến dự đoán chỉ số tài chính sử dụng \verb|LSTM|, 20\% sử dụng \verb|ANN| và 6\% sử dụng \verb|CNN| \cite{ayitey2023forex}.
-Để tận dụng hết được các đặc trưng được rút trích bởi các mô hình nêu trên, chúng tôi đề xuất phương pháp kết hợp các đặc trưng này.
-
-To overcome the second challenge, most studies use features extracted from Long short-term memory neural network (\verb|LSTM|) \cite{hochreiter1997long}, Artificial neural network (\verb|ANN|), and Convolution neural network (\verb|CNN|) \cite{lecun1989handwritten}.
-Specifically, in 2022, 20\% of all publications related to financial index prediction used \verb|LSTM|, 20\% used \verb|ANN|, and 6\% used \verb|CNN| \cite{ayitey2023forex}.
-To make full use of the features extracted by the above models, we propose a method that combines these features.
+% Điều này là hoàn toàn dễ hiểu vì \verb|LSTM| và \verb|CNN| đã chứng minh được khả năng của mình trong việc rút trích các đặc trưng không gian, thời gian một cách hiệu quả.
+% Mặc dù vậy, khi đối mặt với tính phi chu kỳ, sau khoảng 100 training steps, \verb|LSTM| vẫn bị underfitting, còn \verb|CNN| bị overfitting chỉ sau 5 training steps
+
+To overcome the second challenge, most studies use features extracted from Long Short-Term Memory neural network (\verb|LSTM|) \cite{hochreiter1997long}, Artificial Neural Network (\verb|ANN|), and Convolution Neural Network (\verb|CNN|)\cite{lecun1989handwritten}.
+Specifically, in 2022, 20\% of all publications related to financial index prediction used \verb|LSTM|, 20\% used \verb|ANN|, and 6\% used \verb|CNN|\cite{ayitey2023forex}.
+This is completely understandable because \verb|LSTM| and \verb|CNN| have proven their ability to extract spatial and temporal features effectively.
+However, when faced with aperiodic data with only 2600 samples, after 40 training steps, \verb|LSTM| is still underfitting, while \verb|CNN| is overfitting after only 5 training steps.
 
@@
 Điều này càng làm tăng tính đúng đắn của giả thuyết trong nghiên cứu \cite{fama1970efficient}.
-Ngoài ra, chúng tôi cho rằng, loại dữ liệu này còn có những phụ thuộc ngầm vào các thời điểm nhất định trong quá khứ (hidden long-term dependency).
+% Do đó, bằng việc khai thác các nguồn dữ liệu có cùng domain, chúng ta có thể capture được các thông tin phản ánh các external factors.
+
+Regarding the third challenge, \cite{fama1970efficient} hypothesizes that different time-series datasets of the same domain at the same time reflect the impact of external factors.
+For example, studies \citep{overreactioncontrarian, mech1993portfolio} show the dependence between the financial ratios of a given company and the ratios of other companies.
+This further strengthens the hypothesis in study \cite{fama1970efficient}.
+Therefore, by exploiting data sources with the same domain, we can capture information reflecting external factors.
+
+\vspace{2mm}
+
+% Chúng tôi tiếp cận các thách thức nêu trên thông qua việc kết hợp đặc trưng và tổng hợp mô hình.
+Cụ thể, bằng cách sử dụng Meta-learning (ML) \cite{finn2017model}, chúng tôi tổng hợp hiệu quả tham số của các mô hình cục bộ, giúp giảm thiểu đáng kể variance loss cũng như tăng khả năng thu thập các thông tin phản ánh external factors từ các tập dữ liệu có cùng domain.
+Để học được các laten patterns, chúng tôi đề xuất phương pháp kết hợp các đặc trưng từ \verb|CNN| và \verb|LSTM|.
+Ngoài ra, chúng tôi cho rằng, aperiodic time-series data còn có những phụ thuộc ngầm vào các thời điểm nhất định trong quá khứ (hidden long-term dependency).
 Đối với cách tiếp cận truyền thống, người ta sử dụng một lượng dữ liệu quá khứ cố định (lookback window) để huấn luyện mô hình.
@@
 
-Regarding the third challenge, \cite{fama1970efficient} research hypothesizes that different time-series datasets of the same domain at the same time point reflect the impact of extraneous factors.
-For example, studies \citep{overreactioncontrarian, mech1993portfolio} show the dependence between the financial ratios of a given company and the ratios of other companies.
-This further strengthens the hypothesis in study \cite{fama1970efficient}.
-In addition, we argue that this type of data also has hidden long-term dependencies.
-For the traditional approach, people use a fixed amount of past data (lookback window) to train the model.
-This causes a big obstacle to the learning process because long-term features over time will be forgotten.
+We approach the above challenges through combination of feature combination and model synthesis.
+Specifically, by using Meta-learning (ML) \cite{finn2017model}, we efficiently synthesize the parameters of local models, which significantly reduces variance loss and increases the ability to collect information reflecting external factors from datasets in the same domain.
+To effectively learn laten patterns, we propose a method that combines features from \verb|CNN| and \verb|LSTM|.
+In addition, we assume that aperiodic time-series data also has hidden long-term dependencies.
+Traditional approaches use a fixed amount of past data (lookback window) to train the model.
+This poses a major obstacle to the learning process because long-term features will be forgotten over time.
 On the other hand, ML divides the dataset into many parts to learn and synthesize the learned parameters effectively, so it can handle this challenge well.
@@
 
-Finally, we demonstrate the superiority of the proposed algorithm by solving the problem of predicting the trend (up or down) of foreign exchange rates and comparing the results with the existing state-of-the-art (SOTA) model (\verb|NHITS| \cite{challu2023nhits}) on two types of data: (1) - USD/JPY exchange rate data; (2) - Exchange rate data of 60 currency pairs, comprising 18 countries.
+Finally, we demonstrate the superiority of the proposed algorithm by solving the problem of predicting the trend (upward or downward) of foreign exchange rates and comparing the results with the existing state-of-the-art (SOTA) model (\verb|NHITS| \cite{challu2023nhits}) on two types of data: (1) - USD/JPY exchange rate data; (2) - Exchange rate data of 60 currency pairs, comprising 18 countries.
 These data sets are publicly available on the Internet and can be easily downloaded.
-The official implementation can be found at [insert github link here].
-
-\vspace{2mm}
-
-In summary, our main contributions are as follows:
-
-\begin{itemize}
-    % \item Kết hợp đặc trưng: Rút trích đặc trưng bằng cách kết hợp các đặc trưng của \verb|LSTM| và \verb|CNN|.
-    \item \textbf{Feature combination}: Extract feature using \verb|LSTM| and \verb|CNN| and combine them.
-
-    % \item Hidden long-term dependency: Chứng minh thực nghiệm rằng một aperiodic time-series data nhất định không chỉ phụ thuộc vào các yếu tố ngoại lai mà còn có các phụ thuộc ẩn với chính nó tại nhiều thời điểm khác nhau trong quá khứ.
-    \item \textbf{Hidden long-term dependency}: Experimentally demonstrate that a given aperiodic time-series data not only depends on external factors but also has hidden dependencies with itself at various points in the past.
-
-    % \item Tổng hợp hiệu quả tham số mô hình: Sử dụng ML thay cho các mô hình ensemble truyền thống trong việc tổng hợp kết quả từ các mô hình học máy.
-    \item \textbf{Efficient model parameter aggregation}: Use ML instead of traditional ensemble models to aggregate results from machine learning models.
-
-    % \item Experiment: Thực nghiệm trên các bộ dữ liệu về tỉ giá hối đoái và so sánh với mô hình SOTA \verb|NHITS| để chứng minh tính hiệu quả của phương pháp đề xuất.
-    \item \textbf{Experiment}: Experiment on exchange rate datasets and compare with \verb|NHITS| - the SOTA model to demonstrate the effectiveness of the proposed method.
-\end{itemize}
+The official implementation can be found at \url{https://github.com/baolongnguyenmac/multi_fx}.
+
+\vspace{2mm}
+
+In summary, our main contributions include the following: First, we present a novel \textbf{feature combination technique} that extracts and combines features using both \verb|LSTM| and \verb|CNN| models.
+Second, we uncover hidden long-term dependencies, demonstrating experimentally that aperiodic time-series data not only rely on external factors but also exhibit dependencies with past period, which we call \textbf{hidden long-term dependencies}.
+Third, we introduce an efficient model parameter aggregation approach by employing ML algorithm in place of traditional ensemble models for combining parameters from multiple machine learning models.
+Finally, we validate our method through experiments on exchange rate datasets and show its effectiveness in comparison to the SOTA model, \verb|NHITS|.
 
@@
 \begin{align*}
-    \phi^* &= \arg\min_{\phi}\sum_{t}{\mathcal{L}^{meta}_t\left[ \theta_t^*, \mathcal{D}_t^{query} \right]}\\
-    &= \arg\min_{\phi}\sum_{t}{\mathcal{L}^{meta}_t\left[ \theta_t(\phi), \mathcal{D}_t^{query} \right]} \numberthis
+    \phi^* &= \arg\min_{\phi}\sum_{t}{\mathcal{L}^{meta}_t\left( \theta_t^*, \mathcal{D}_t^{query} \right)}\\
+    &= \arg\min_{\phi}\sum_{t}{\mathcal{L}^{meta}_t\left( \theta_t(\phi), \mathcal{D}_t^{query} \right)} \numberthis
     \label{eq:outer_opt}
@@
     \centering
-    \caption{Statistics on USD/JPY and multi-fx. \textbf{The data is not correct, have to be re-written}}
+    \caption{Statistics on USD/JPY and multi-fx.}
     \label{tab:stat_}
@@
     \hline
-    \Verb|USD/JPY|           & 60                      & 69,909                     & 135 & 1,398 & 1,424 & 5,201           \\
-    \Verb|multi-fx|          & 60                      & 52,497                     & 506 & 1,049 & 250   & 1,986           \\
+    \Verb|USD/JPY|           & 60                      & 150,175                     & 2,502 & 2,502.92 & 0.28 & 2,503           \\
+    \Verb|multi-fx|          & 60                      & 154,440                     & 2,467 & 2,575 & 58.61   & 2748           \\
     \bottomrule
@@
     \multirow{3}{*}{\Verb|USD/JPY|}  & \Verb|NHITS|          & 53.29                 & 52.38                 & 52.23                 & 51.87                   \\
-                                                & Ours(\Verb|LSTM|)     & $not\_done$           & $59.57±2.53$          & $58.31±2.57$          & $56.6±3.74$             \\
+                                                & Ours(\Verb|LSTM|)     & $57.12±0.08$ & $58.32±2.72$ & $57.44±2.62$ & $56.87±3.56$    \\
                                                 & Ours(\Verb|LSTM+CNN|) & $\mathbf{58.05±0.01}$ & $\mathbf{59.57±2.53}$ & $\mathbf{58.31±2.57}$ & $\mathbf{56.6±3.74}$    \\
@@
     \includegraphics[width=\linewidth]{img/lr.pdf}
-    \caption{Convergence of 32 models LSTM+CNN.}
+    \caption{Convergence processes of 32 models using LSTM+CNN as feature extractor.}
     \label{fig:lr}
@@
 
-\begin{figure}
+\begin{figure}[H]
     \centering

### 2024-09-17 execute/main.py
remove -test option in main

@@ -18,31 +18,31 @@ def parse_args():
     parser.add_argument('-rounds', '--num_rounds', type=int, default=100, required=False)
     parser.add_argument('-epochs', '--num_epochs', type=int, default=3, required=True) # we need this param in test phase
     parser.add_argument('-ncpus', '--num_cpus', type=int, default=8)
-    parser.add_argument('-test', '--test', type=str, required=False, help='Pass model_name here (e.g: lstm_0.001_0.005)')
-    # parser.add_argument('--test', action='store_true', required=False)
 
     return vars(parser.parse_args())
 
 def main():
     input_dict = parse_args()
 
-    # data preparation
-    # set data_dir='/home/s2210434/fx/fxdata/multi_USDJPY'
-    data_dict = get_data(look_back=input_dict['look_back_window'], data_dir='/home/s2210434/fx/fxdata/multi_USDJPY')
+    # config data_dir right here
+    dataset = MULTI_FX
+    data_dir = os.path.join(RAW_DATA_DIR, dataset)
+    data_dict = get_data(look_back=input_dict['look_back_window'], data_dir=data_dir)
 
-    # # if pretrained models are from `pretrained_macro_random`
-    # # choose randomly 50% clients for training, 25% for validating and 25% for testing (only apply for multi_fx, not multi_USDJPY)
-    # list_id = set(data_dict.keys())
-    # list_id_train = random.sample(sorted(list_id), 30)
-    # list_id_val = random.sample(sorted(list_id.difference(set(list_id_train))), 15)
-    # list_id_test = list(list_id.difference(set(list_id_train).union(list_id_val)))
-
-    # if pretrained models are from `pretrained_macro`
-    # if the data is multi_USDJPY
-    list_id = list(data_dict.keys())
-    list_id_train = list_id[:int(len(list_id)*0.5)]
-    list_id_val = list_id[int(len(list_id)*0.5):int(len(list_id)*0.75)]
-    list_id_test = list_id[int(len(list_id)*0.75):]
+    if 'multi' in data_dir:
+        # only apply for multi_fx
+        # choose randomly 50% clients for training, 25% for validating and 25% for testing
+        list_id = set(data_dict.keys())
+        list_id_train = random.sample(sorted(list_id), 30)
+        list_id_val = random.sample(sorted(list_id.difference(set(list_id_train))), 15)
[…]

### 2024-09-17 common/constants.py
add some constants

@@ -1,10 +1,17 @@
 import os
 
 # dirs
-RAW_DATA_DIR = os.path.join(os.path.dirname(os.path.abspath(__file__)), '../data/fx_data')
+RAW_DATA_DIR = os.path.join(os.path.dirname(os.path.abspath(__file__)), '../data')
 EXE_DIR = os.path.join(os.path.dirname(os.path.abspath(__file__)), '../execute')
 PRETRAINED_DIR = os.path.join(os.path.dirname(os.path.abspath(__file__)), '../pretrained')
 LOG_DIR = os.path.join(os.path.dirname(os.path.abspath(__file__)), '../execute/log')
+
+# dataset
+MULTI_FX = 'multi_fx'
+USD_JPY = 'USD_JPY'
+ETT = 'ett'
+ECL = 'ecl'
+WTH = 'wth'
 
 # mode
 REG = 'regression'

### 2024-09-17 data/pre_process.py

@@ -43,9 +43,10 @@ def get_data(look_back:int, test_size:float=0.8, batch_size:int=32, data_dir:str
 
             tmp_dict = {}
             tmp_dict['support_X'], tmp_dict['query_X'], tmp_dict['support_y'], tmp_dict['query_y'] = train_test_split(X, y, test_size=test_size, shuffle=False)
-            for key in tmp_dict.keys():
-                num_batch = int(np.ceil(tmp_dict[key].shape[0]/batch_size))
-                tmp_dict[key] = np.array_split(tmp_dict[key], num_batch)
+            if batch_size != -1: # do not divide data into batches
+                for key in tmp_dict.keys():
+                    num_batch = int(np.ceil(tmp_dict[key].shape[0]/batch_size))
+                    tmp_dict[key] = np.array_split(tmp_dict[key], num_batch)
             data_dict[cid] = tmp_dict
             cid += 1
 
@@ -71,92 +72,3 @@ def get_data_for_nhits(data_dir:str=RAW_DATA_DIR):
             transformed_df['ds'] = pd.date_range(start='2000-01-01', periods=len(transformed_df), freq='h')
             return transformed_df[['unique_id', 'ds', 'open', 'high', 'low', 'close', 'y']]
 
-# def create_label(df:np.ndarray, mode:str=CLF) -> np.ndarray:
-#     label = df[:,-1]
-#     if mode == CLF:
-#         label = label[1:] - label[:-1]
-#         label = (label > 0)*1.
-#         label = np.append([-1], label) # the first value is assigned to -1, it's removed anyway
-#     elif mode == REG:
-#         pass # if it is regression problem, we don't have to do anything
-#     return label
-
-# def normalize_data(df:np.ndarray) -> np.ndarray:
-#     scaler = StandardScaler()
-#     return scaler.fit_transform(df), scaler
-
-# # create dataset with look_back window
-# def create_dataset(df:np.ndarray, look_back:int, horizon:int=1, mode:str=CLF)->tuple[np.ndarray, np.ndarray]:
-#     size = df.shape[0] - look_back - horizon + 1
-#     X = np.array([df[i : i+look_back] for i in range(size)])
-#     y = create_label(df, mode)
-#     y = np.array([y[i+look_back : i+look_back+horizon] for i in range(size)])
-#     return X, y
-
[…]

### 2024-09-17 model/baseline_models.py

@@ -28,12 +28,14 @@ def add_log(log:dict, count:int, input_size:int, n_stacks:int, kernel_size:int, 
     log[count]["recall"] = recall
     log[count]["f1"] = f1
 
-def write_log(log:dict, file_name:str='NHITS_USD-JPY.json'):
+def write_log(log:dict, file_name:str):
     with open(os.path.join(PRETRAINED_DIR, 'baseline', file_name), 'w') as fo:
         json.dump(log, fo)
 
 def run():
-    data_ = get_data_for_nhits('/home/s2210434/fx/fxdata')
+    dataset = USD_JPY
+    data_dir = os.path.join(RAW_DATA_DIR, dataset, 'all')
+    data_ = get_data_for_nhits(data_dir=data_dir)
 
     pooling_sizes = [[2,2,2], [4,4,4], [8,8,8], [8,4,1], [16,8,1]]
     freqs = [[168,24,1], [24,12,1], [180,60,1], [40,20,1], [64,8,1]]
@@ -81,7 +83,7 @@ def run():
                     count += 1
 
                     add_log(log, count, input_size, 3, pooling_size, 512, lr, freq, 'linear', acc, precision, recall, f1)
-    write_log(log)
+    write_log(log=log, file_name=f'NHITS_{dataset}.json')
 
 if __name__ == '__main__':
     # data_ = prepare_data()

### 2024-09-19 doc/paper/paper.tex
Fixed untracked files

@@
 Second, we uncover hidden long-term dependencies, demonstrating experimentally that aperiodic time-series data not only rely on external factors but also exhibit dependencies with past period, which we call \textbf{hidden long-term dependencies}.
-Third, we introduce an efficient model parameter aggregation approach by employing ML algorithm in place of traditional ensemble models for combining parameters from multiple machine learning models.
+Third, we introduce an \textbf{efficient model parameter aggregation approach} by employing ML algorithm in place of traditional ensemble models for combining parameters from multiple machine learning models.
 Finally, we validate our method through experiments on exchange rate datasets and show its effectiveness in comparison to the SOTA model, \verb|NHITS|.

### 2024-09-19 model/baseline_models.py
minor change

@@ -89,146 +89,3 @@ if __name__ == '__main__':
     # data_ = prepare_data()
     # fine_tune(data_)
     run()
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-# def prepare_data(data_dir:str='./data'):
-#     data_dict = _get_data(1, batch_size=-1, data_dir=data_dir, test_size=0.4)
-#     # bỏ đi mẫu đầu của X và mẫu cuối của y để mô phòng chính xác time-series dataset
-#     # bởi vì `get_data` trả về 1 df đã construct để train bình thường còn NeuralForecast yêu cầu cái df time-series như ban đầu
-#     X = data_dict[0]['support_X'].reshape(-1, 4)[1:]
-#     y = data_dict[0]['support_y'].reshape(-1,)[:-1]
-
-#     X = X.astype(np.float32)
-#     y = y.astype(np.float32)
-
-#     data_ = pd.DataFrame(X, columns=['open', 'high', 'low', 'close'])
-#     data_['y'] = y
-#     data_['unique_id'] = 0
-#     data_['ds'] = pd.date_range(start='2000-01-01', periods=len(data_), freq='D')
-#     return data_[['unique_id', 'ds', 'open', 'high', 'low', 'close', 'y']]
-
-# def get_accuracy(y, y_hat):
-#     return np.mean(y==y_hat).item()
-
-# def add_log(log:dict, count:int, input_size:int, n_stacks:int, kernel_size:int, mlp_units:int, lr:float, freq:list[int], interpolation_mode:str, acc:float):
-#     log[count] = {}
-#     log[count]["input_size"] = input_size
-#     log[count]["n_stacks"] = n_stacks
-#     log[count]["kernel_size"] = kernel_size
-#     log[count]["mlp_units"] = mlp_units
-#     log[count]["lr"] = lr
-#     log[count]["freq"] = freq
-#     log[count]["interpolation_mode"] = interpolation_mode
-#     log[count]["acc"] = acc
-
-# def write_log(log:dict, file_name:str='log_baseline.json'):
-#     with open(os.path.join(LOG_DIR, file_name), 'w') as fo:
[…]

### 2024-09-19 model/meta_model.py

@@ -241,8 +241,10 @@ class MAML:
         self.info[f'{mode}_recall'] = f"{tasks_metrics['recall']:.5f} ± {tasks_metrics['std_recall']:.5f}"
         self.info[f'{mode}_f1'] = f"{tasks_metrics['f1']:.5f} ± {tasks_metrics['std_f1']:.5f}"
 
+        return self.info[f'{mode}_accuracy'], self.info[f'{mode}_precision'], self.info[f'{mode}_recall'], self.info[f'{mode}_f1']
+
     def save_model(self, dir_:str, model_name:str):
-        print('\nSave model\n')
+        print(f'\nSave model at {dir_}\n')
         json_file_path = os.path.join(dir_, f'{model_name}.json')
         model_file_path = os.path.join(dir_, f'{model_name}.keras')
         with open(json_file_path, 'w') as fo:

### 2024-09-20 data/pre_process.py
read the comment on github

@@ -5,8 +5,8 @@ import numpy as np
 
 from common.constants import *
 
-def create_label(df:np.ndarray, mode:str=CLF) -> np.ndarray:
-    label = df[:,-1]
+def create_label(df:np.ndarray, label:int, mode:str=CLF) -> np.ndarray:
+    label = df[:,label]
     if mode == CLF:
         label = label[1:] - label[:-1]
         label = (label>0)*1.
@@ -20,14 +20,20 @@ def normalize_data(df:np.ndarray) -> np.ndarray:
     return scaler.fit_transform(df)
 
 # create dataset with lookback window
-def create_dataset(df:np.ndarray, look_back:int)->tuple[np.ndarray, np.ndarray]:
+def create_dataset(df:np.ndarray, look_back:int, label:int)->tuple[np.ndarray, np.ndarray]:
     size = df.shape[0] - look_back
     X = np.array([df[i : i + look_back] for i in range(size)])
-    y = create_label(df)
+    y = create_label(df, label)
     y = y[look_back:]
     return X, y.reshape(-1,1)
 
-def get_data(look_back:int, test_size:float=0.8, batch_size:int=32, data_dir:str=RAW_DATA_DIR) -> dict[str, dict[str, list[np.ndarray]]]:
+def get_data(look_back:int, label:int=-1, test_size:float=0.8, batch_size:int=32, data_dir:str=RAW_DATA_DIR) -> dict[str, dict[str, list[np.ndarray]]]:
+    """
+        read data from files and return a dict whose each value contains data from a file
+
+        label (int, optional): The label to be predict the movement. Defaults to -1 (the final feature).
+        batch_size (int, optional): If `batch_size=-1`, then we don't divide data into batches
+    """
     cid = 0
     data_dict:dict[str, dict[str, pd.DataFrame]] = {}
     for file_name in os.listdir(data_dir):
@@ -39,7 +45,7 @@ def get_data(look_back:int, test_size:float=0.8, batch_size:int=32, data_dir:str
             transformed_df = normalize_data(df)
 
             # create dataset with lookback window
-            X, y = create_dataset(transformed_df, look_back)
+            X, y = create_dataset(transformed_df, look_back, label)
 
             tmp_dict = {}
             tmp_dict['support_X'], tmp_dict['query_X'], tmp_dict['support_y'], tmp_dict['query_y'] = train_test_split(X, y, test_size=test_size, shuffle=False)

### 2024-09-20 execute/main.py

@@ -2,6 +2,8 @@ import tensorflow as tf
 import random
 random.seed(84)
 import argparse
+import pandas as pd
+import json
 
 from model.meta_model import MAML
 from common.constants import *
@@ -21,15 +23,12 @@ def parse_args():
 
     return vars(parser.parse_args())
 
-def main():
-    input_dict = parse_args()
-
-    # config data_dir right here
-    dataset = MULTI_FX
+def run(dataset:str, input_dict:dict, label:int):
+    # prep data
     data_dir = os.path.join(RAW_DATA_DIR, dataset)
-    data_dict = get_data(look_back=input_dict['look_back_window'], data_dir=data_dir)
+    data_dict = get_data(look_back=input_dict['look_back_window'], data_dir=data_dir, label=label)
 
-    if 'multi' in data_dir:
+    if 'multi' in dataset:
         # only apply for multi_fx
         # choose randomly 50% clients for training, 25% for validating and 25% for testing
         list_id = set(data_dict.keys())
@@ -64,14 +63,49 @@ def main():
         maml.train(round, input_dict['num_epochs'], batch_task, list_id_val)
 
     # after training, testing is produced automatically
-    maml.valid(list_id_test, num_epochs=input_dict['num_epochs'], mode=TEST)
+    acc, precision, recall, f1 = maml.valid(list_id_test, num_epochs=input_dict['num_epochs'], mode=TEST)
 
-    # save log
-    model_name = f"{input_dict['based_model']}_{input_dict['look_back_window']}_{input_dict['inner_learning_rate']}_{input_dict['outer_learning_rate']}"
-    maml.save_model(dir_=pretrained_dir, model_name=model_name)
+    # save log and model
+    model_folder = f"{input_dict['based_model']}_{input_dict['look_back_window']}_{input_dict['inner_learning_rate']}_{input_dict['outer_learning_rate']}"
+    model_dir = os.path.join(pretrained_dir, model_folder)
+    # create a folder in `pretrained_dir` named model_name
+    if not os.path.exists(model_dir):
+        os.mkdir(model_dir)
+    maml.save_model(dir_=model_dir, model_name=label)
+
+    return acc, precision, recall, f1, model_dir
 
 if __name__ == '__main__':
-    main()
+    # config data_dir right here
+    dataset = MULTI_FX
+    data_dir = os.path.join(RAW_DATA_DIR, dataset)
+
+    # extract columns
[…]

### 2024-09-20 execute/submit_jobs.py

@@ -3,7 +3,7 @@ this file should be run from the root directory of the project
 >> python -m execute.submit_jobs
 
 it will generate `tmp.sh` to submit the job to server
-all `tmp.sh` file assume that they are called from the root directory
+all `tmp.sh` files are supposed that they are called from the root directory
 '''
 
 import subprocess
@@ -19,7 +19,7 @@ def generate_script(device:str, name_job:str, look_back:int, model_name:str, mod
 
     if device == 'cpu':
         class_name = 'DEFAULT'
-        resources = 'select=1:ncpus=64'
+        resources = f'select=1:ncpus={ncores_limit}'
         load_modules = ''
         test_device = f'echo "Using $(nproc) CPU cores" &>> {log_file_path}'
     elif device == 'gpu':
@@ -52,7 +52,8 @@ def generate_script(device:str, name_job:str, look_back:int, model_name:str, mod
     with open(os.path.join(EXE_DIR, 'tmp.sh'), "w") as fout:
         fout.write("\n".join(tmp_file) + "\n")
 
-def submit_based_model():
+# finetune meta model (based model: CNN, LSTM)
+def finetune_1():
     current_count = None
     prev_count = 96
     count = 0
@@ -81,17 +82,42 @@ def submit_based_model():
     # generate_script('gpu', WEAK_LEARNERS[0], 'boosting')
     # subprocess.run(["qsub", os.path.join(EXE_DIR, 'tmp.sh')])
 
-def submit_tunned_model():
+# finetune meta model (based model: LSTM, LSTM+CNN)
+def finetune_2():
     count = 0
     for window in [20, 30]:
-        for model in [LSTM]:
-            for inner_lr in [0.005]:
-                for outer_lr in [0.0055, 0.005]:
+        for model in [LSTM, LSTM_CNN]:
+            for inner_lr in [0.005, 0.001, 0.01, 0.05]:
+                for outer_lr in [0.0055, 0.005, 0.001, 0.0015]:
+            # for inner_lr in [0.005, 0.001, 0.01]:
+            #     for outer_lr in [0.0055, 0.005]:
                     generate_script('cpu', f'{model}_{count}', window, model, CLF, inner_lr, outer_lr, 5, 100, 3)
                     subprocess.run(["qsub", os.path.join(EXE_DIR, 'tmp.sh')])
                     time.sleep(1)
                     count += 1
 
+'''
[…]

### 2024-10-04 doc/paper/paper.tex
draft 3

@@
 \usepackage{amssymb}
-% \usepackage{bm}
-% \usepackage{dcolumn}
-% \usepackage{bm}
 \usepackage{url}
-% \usepackage{here}
-% \usepackage{ulem}
-% \usepackage[caption=false]{subfig}
 \usepackage{float}
@@
 \usepackage{rotating}
-% \usepackage{makecell}
 \usepackage{multirow}
@@
 \def\green{\color{green}}
+
+% \usepackage{makecell}
+% \usepackage{bm}
+% \usepackage{dcolumn}
+% \usepackage{here}
+% \usepackage{ulem}
+% \usepackage[caption=false]{subfig}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \author{Bao-Long Nguyen}
@@
 
-For the first challenge, ensemble models \citep{sadeghi2021combined, zafeiriou2020intraday, ali2020complete} are often used to mitigate the effects of variance variation.
+For the variance variation challenge, ensemble models \citep{sadeghi2021combined, zafeiriou2020intraday, ali2020complete} are often used to mitigate the effects of variance variation.
 Ensemble models provide a holistic and multi-perspective view based on sub-models, thereby helping the overall model adapt to strong variance changes.
@@
 
-To overcome the second challenge, most studies use features extracted from Long Short-Term Memory neural network (\verb|LSTM|) \cite{hochreiter1997long}, Artificial Neural Network (\verb|ANN|), and Convolution Neural Network (\verb|CNN|)\cite{lecun1989handwritten}.
-Specifically, in 2022, 20\% of all publications related to financial index prediction used \verb|LSTM|, 20\% used \verb|ANN|, and 6\% used \verb|CNN|\cite{ayitey2023forex}.
+To overcome the pattern variation challenge, most studies use features extracted from Long Short-Term Memory neural network (\verb|LSTM|) \cite{hochreiter1997long}, Artificial Neural Network (\verb|ANN|), and Convolution Neural Network (\verb|CNN|) \cite{lecun1989handwritten}.
+Specifically, in 2022, 20\% of all publications related to financial index prediction used \verb|LSTM|, 20\% used \verb|ANN|, and 6\% used \verb|CNN| \cite{ayitey2023forex}.
 This is completely understandable because \verb|LSTM| and \verb|CNN| have proven their ability to extract spatial and temporal features effectively.
@@
 
-Regarding the third challenge, \cite{fama1970efficient} hypothesizes that different time-series datasets of the same domain at the same time reflect the impact of external factors.
+Regarding the external influences challenge, \cite{fama1970efficient} hypothesizes that different time-series datasets of the same domain at the same time reflect the impact of external factors.
 For example, studies \citep{overreactioncontrarian, mech1993portfolio} show the dependence between the financial ratios of a given company and the ratios of other companies.
@@
 
-In summary, our main contributions include the following: First, we present a novel \textbf{feature combination technique} that extracts and combines features using both \verb|LSTM| and \verb|CNN| models.
+In summary, our main contributions include the following: First, we present a novel \textbf{feature combination technique} for aperiodic time-series data that extracts and combines features using both \verb|LSTM| and \verb|CNN| models.
 Second, we uncover hidden long-term dependencies, demonstrating experimentally that aperiodic time-series data not only rely on external factors but also exhibit dependencies with past period, which we call \textbf{hidden long-term dependencies}.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Related work}
+\section{Related works}
 \label{sec.relatedWork}
@@
 
-\subsection{Model-agnostic Meta-learning (MAML)}
+\subsection{Optimization-based Meta-learning}
 
@@
 
-Meta-learning (ML) algorithms, typically MAML \cite{finn2017model} are known for their ability to train a highly general, adaptive model on new datasets with a limited amount of data and a small number of training steps \citep{hospedales2021meta, vettoruzzo2024advances}.
+Optimization-based Meta-learning (ML) algorithms, typically MAML \cite{finn2017model} are known for their ability to train a highly general, adaptive model on new datasets with a limited amount of data and a small number of training steps \citep{hospedales2021meta, vettoruzzo2024advances}.
 With this ability, ML is widely used in tasks that require the model's ability to adapt to the data (e.g.
@@
 Hybrid ensemble models have been widely used in time-series processing problems and have been experimentally proven to be more accurate than standard time-series models because they can synthesize the strengths of many sub-models \cite{ayitey2023forex}.
-However, current ensemble model synthesis forms are still very rigid because they can only synthesize based on the final results (voting mechanism of bagging models) and near-final results (for stacking models).
+However, current ensemble model synthesis forms are still very rigid because they can only synthesize based on the final results (e.g.
+voting mechanism of bagging models) and near-final results (e.g.
+stacking models).
 From the perspective of ensemble models, the equation \ref{eq:outer_opt} can be considered an effective method of synthesizing sub-models, which helps to take advantage of the feature extraction capabilities of each model.
@@
 $\mathbf{\theta}_l^f, \mathbf{\theta}_l^b$ là các hệ số nội suy forecast và backcast, được dùng để tổng hợp các giá trị đầu ra của block $l$ bằng hàm nội suy $g(\cdot)$.
-Đầu ra của block $l$ là giá trị forecast $\mathbf{\hat{y}}_{t+1:t+H, l}$ và giá trị backcast $\mathbf{\hat{y}}_{t-L:t, l}$.
-Input của block $l+1$ được tính theo phương trình \ref{eq:input_l1}.
-
-Accordingly, \verb|FullyConnected| are stacked multi-layer perception (\verb|MLP|) layers with nonlinear activation functions.
+Số chiều của các hệ số nội suy trong mỗi stack được quy định bởi expressiveness ratio $r_l$: $\left| \theta_l \right| =  \left \lceil r_l H \right \rceil$.
+Thông thường, expressiveness ratio sẽ rất nhỏ ở các stack đầu tiên và tăng dần về cuối.
+Theo đó, các stack có thể mô phỏng lại các tần số từ thấp đến cao.
+Ngoài ra, bằng việc sử dụng hàm nội suy, \verb|NHITS| không cần quá nhiều phần cứng để huấn luyện mạng neural trong trường hợp H lớn.
+
+Accordingly, \verb|Pooling| compresses the original data into $\mathbf{y}_{t-L:t, l}^{(p)}$.
+\verb|FullyConnected| contains many perceptron layers with nonlinear activation functions.
 $\mathbf{\theta}_l^f, \mathbf{\theta}_l^b$ are the forecast and backcast interpolation coefficients, which are used to aggregate the output values of block $l$ using the interpolation function $g(\cdot)$.
+The dimensionality of the interpolation coefficients in each stack is determined by the expressiveness ratio: $\left| \theta_l \right| = \left \lceil r_l H \right \rceil$.
+Typically, the expressiveness ratio will be very small in the first stacks and increase towards the end.
+Accordingly, the stacks can simulate frequencies from low to high.
+In addition, by using $g(\cdot)$, \verb|NHITS| does not require too much hardware to train the neural network in the case of large $H$.
+
+\vspace{2mm}
+
+% Đầu ra của block $l$ là giá trị forecast $\mathbf{\hat{y}}_{t+1:t+H, l}$ và giá trị backcast $\mathbf{\hat{y}}_{t-L:t, l}$. Input của block $l+1$ được tính theo phương trình \ref{eq:input_l1}.
+
 The output of block $l$ is the forecast value $\mathbf{\hat{y}}_{t+1:t+H, l}$ and the backcast value $\mathbf{\hat{y}}_{t-L:t, l}$.
@@
     \centering
-    \includegraphics[width=\linewidth]{img/CHF_NZD.pdf}
-    \caption{Exchange rate (close price) between Swiss franc and New Zealand dollar by day (2014-2024).}
+    \includegraphics[width=\linewidth]{img/USD_JPY.pdf}
+    \caption{Exchange rate (close price) between US dollar and Japanese yen by day (2014-2024).}
     \label{fig:fx}
@@
 Inspired by the \cite{vo2017multi} study, we propose to combine the features extracted from \verb|LSTM| and \verb|CNN| networks.
-Specifically, we pass each element in the vector $\mathbf{x}_{t-L:t}$ through a \verb|MLP| layer whose output'dimension is larger than the one of $\vec x_i, i\in[t-L, t]$ to decompose it into smaller features $\vec x'_i$.
+Specifically, we pass each element in the vector $\mathbf{x}_{t-L:t}$ through \verb|FullyConnected| layers whose output'dimension is larger than the one of $\vec x_i, i\in[t-L, t]$ to decompose it into smaller features $\vec x'_i$.
 These features are then passed through \verb|LSTM| and \verb|CNN| networks to extract long-term temporal dependencies ($\mathbf{h}_{LSTM}$) and local temporal features ($\mathbf{h}_{CNN}$), respectively.
@@
 
-The \verb|LSTM| network maintains cell-state values ​​to selectively store long-term dependencies.
+The \verb|LSTM| network maintains cell-state values to selectively store long-term dependencies.
 This is very suitable for solving time-series data problems.
-On the other hand, future values ​​often depend heavily on recent historical values.
+On the other hand, future values often depend heavily on recent historical values.
 We propose to use the \verb|CNN| network to emphasize local features, thereby directing part of the model's attention to certain time points.
@@
 Each global step includes many local steps (inner optimization) performed on each individual task.
-At global step $r$, the $e$th local optimization process at the support set of task $t$ occurs as follows:
+At global step $r$-th, the $e$-th local optimization process at the support set of task $t$ occurs as follows:
 
@@
 \end{align} Where, $\beta$ is the outer learning rate.
+
+\vspace{2mm}
 
@@
 Specifically, we configure two datasets using FX data.
-The \verb|USD/JPY| dataset consists of only data of the USD/JPY currency pair, divided into 60 time-sequenced subsets of equal size.
+\verb|USD/JPY| dataset consists of only data of the exchange rate between US dollar and Japanese yen, is divided into 60 time-sequenced subsets of equal size.
 The data is sampled hourly from 2000 to 2024, including the attributes of open, low, high, and close price.
-The \verb|multi-fx| dataset consists of 60 currency pairs made up of 18 countries: Australia, Canada, Switzerland, Denmark, EU, United Kingdom, Hong Kong, Iceland, Japan, Norway, New Zealand, Singapore, Sweden, Turkey, United States, Mexico, China, South Africa.
+\verb|multi-fx| dataset consists of 60 currency pairs made of 18 countries: Australia, Canada, Switzerland, Denmark, EU, United Kingdom, Hong Kong, Iceland, Japan, Norway, New Zealand, Singapore, Sweden, Turkey, United States, Mexico, China, South Africa.
 The data has similar attributes to \verb|USD/JPY| and sampled daily from 2014 to 2024.
-The number of data samples of these two datasets are similar and approximately $156000$ samples.
 
@@
 
-The \verb|multi-fx| dataset is used to extract and aggregate information about outliers (i.e.
+Our goal is to use \verb|multi-fx| dataset to extract and aggregate information about outliers (i.e.
 market, economic, political, etc.) that are believed to influence the outcome of a given financial index \citep{overreactioncontrarian,fama1970efficient,mech1993portfolio}.
-The \verb|USD/JPY| dataset is used to test our hypothesis that future data implicitly depend on certain points in the past and that efficient past feature aggregation is needed to reveal these dependencies.
+Meanwhile \verb|USD/JPY| dataset is used to test our hypothesis that future data implicitly depend on certain points in the past and that efficient past feature aggregation is needed to reveal these dependencies.
+
+\vspace{2mm}
+
+% Ngoài ra, chúng tôi sử dụng thêm 2 bộ dữ liệu: Electricity Transformer Temperature (\verb|ETT-m2|) và Weather (\verb|WTH|).
+Tập dữ liệu \verb|ETT-m2| bao gồm 7 trường dữ liệu, đo đạc các thông số của máy biến áp tại một tỉnh của Trung Quốc sau mỗi 15 phút trong khoảng thời gian từ July 2016 to July 2018.
+Tập dữ liệu \verb|WTH| bao gồm 12 trường dữ liệu, ghi nhận các thông số của thời tiết tại Weather Station of the Max Planck Biogeochemistry Institute in Jena, Germany sau mỗi 10 phút trong năm 2020.
+Đây là hai tập dữ liệu thể hiện tính chu kỳ rất mạnh mà \verb|NHITS| đã rất thành công trong việc dự đoán.
+Thực nghiệm trên chúng giúp so sánh toàn diện hơn về khả năng của phương pháp đề xuất đối với \verb|NHITS|.
+
+In addition, we used two periodic datasets: Electricity Transformer Temperature (\verb|ETT-m2|) \cite{zhou2021informer} and Weather (\verb|WTH|) \cite{Kolle}.
+\verb|ETT-m2| dataset consists of 7 data fields, measuring the parameters of a transformer in a province of China every 15 minutes from July 2016 to July 2018.
+\verb|WTH| dataset consists of 12 data fields, recording the weather parameters at the Weather Station of the Max Planck Biogeochemistry Institute in Jena, Germany every 10 minutes in 2020.
+These two datasets exhibit very strong periodicity, which \verb|NHITS| has been very successful in predicting.
+Experiments on them provide a more comprehensive comparison of the proposed method's capabilities against \verb|NHITS|.
 
@@
 The study uses macro accuracy, macro precision, macro recall, and macro F1 score to evaluate the models.
-Accordingly, during the inference phase, the model will run on each task to calculate the metrics of each task.
+Accordingly, during the inference phase, the model will run on all tasks to calculate the metrics of each one.
 Then, the average of the metrics of the tasks is calculated to obtain the final result.
@@
 
-We compare the proposed method with \verb|NHITS| using the above metrics on \verb|USD/JPY| and \verb|multi-fx| datasets.
+We compare the proposed method with \verb|NHITS| using the above metrics and dataset.
+For each dataset, each attribute will be predicted for future trends based on past data.
+The final result is calculated by averaging the metrics across the attributes.
 We also test different feature extractors (\verb|LSTM|, \verb|CNN|) for the proposed method to demonstrate the deep data understanding capabilities of \verb|LSTM+CNN|.
@@
 For \verb|NHITS| model, we split the data as usual according to the above predetermined ratio.
-For the proposed algorithm, because the meta-training and meta-testing processes require dividing the data into small tasks, and allowing the model to adapt on the support set of each task, we split the data into 60 tasks (see table \ref{tab:stat_} for statistical detail).
+For the proposed algorithm, because the meta-training and meta-testing processes require dividing the data into small tasks, and allowing the model to adapt on the support set of each task, we split the data into 40-60 tasks (see table \ref{tab:stat_} for statistical detail).
 In each task, the support set accounts for 20\% with the purpose of letting the model adapt to the data, the query set accounts for 80\% to check the model's compatibility.
-We use 30 tasks for meta-training, 15 tasks for meta-validation, and 15 tasks for meta-testing.
+We use 50\% of tasks for meta-training, 25\% for meta-validation, and 25\% for meta-testing.
 With this division, we ensure the fairness that the ML model is trained with the same amount of data as \verb|NHITS|.
@@
     \Verb|multi-fx|          & 60                      & 154,440                     & 2,467 & 2,575 & 58.61   & 2748           \\
+    \Verb|ETT-m2|            & 48                      & 69,680                      & 1,451 & 1,451.67 & 0.47   & 1,452           \\
+    \Verb|WTH|               & 40                      & 35,064                      & 876 & 876.6 & 0.49   & 877           \\
     \bottomrule
@@
 \label{sec.results}
+% TODO: phải viết lại đống nhận xét dựa trên kết quả mới
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 % Bảng \ref{tab:rs} so sánh kết quả giữa mô hình đề xuất và mô hình \verb|NHITS| trên các tập dữ liệu kể trên.
-Phương pháp của chúng tôi đạt mức hội tụ cao hơn hẳn so với mô hình \verb|NHITS| thể hiện trên tất cả các metrics.
+Trên dữ liệu phi chu kỳ (\verb|multi-fx| and \verb|USD/JPY|), phương pháp của chúng tôi đạt mức hội tụ cao hơn hẳn so với mô hình \verb|NHITS| thể hiện trên tất cả các metrics.
 Cụ thể, độ chính xác cải thiện từ 5\% đến 11\% so với \verb|NHITS| trên dữ liệu \verb|USD/JPY| và \verb|multi-fx|.
 Các metrics còn lại cũng đều tăng từ 5\% đến 20\%.
+Trên dữ liệu có chu kỳ, phương pháp của chúng tôi ghi nhận các kết quả cao hơn hoặc xấp xỉ kết quả của \verb|NHITS|.
+Cụ thể, trên tập \verb|WTH|, phương pháp của chúng tôi cho giá trị cao hơn khoảng 1\% trên tất cả các độ đo, trên tập \verb|ETT-m2|, chúng tôi thu được độ chính xác cao hơn khoảng 1\% trong khi các metrics khác thấp hơn khoảng 3\%.
 Sau đây, chúng tôi phân tích kết quả này dựa trên hai yếu tố: (1) - Khả năng rút trích đặc trưng; (2) - Khả năng tổng hợp mô hình.
@@
 Table \ref{tab:rs} compares the results between the proposed model and the \verb|NHITS| model on the above datasets.
-Our method achieves a much higher convergence rate than the \verb|NHITS| model on all metrics.
-Specifically, the accuracy improves from 5\% to 11\% compared to \verb|NHITS| on \verb|USD/JPY| and \verb|multi-fx| data.
+On aperiodic data, our method achieves a much higher convergence rate than the \verb|NHITS| model on all metrics.
+Indeed, the accuracy improves from 5\% to 11\% compared to \verb|NHITS| on \verb|USD/JPY| and \verb|multi-fx| data.
 The remaining metrics also increase from 5\% to 20\%.
-Next, we analyze this result based on two factors: (1) - Feature extraction ability; (2) - Model synthesis ability.
-
-\vspace{2mm}
+On periodic data (\verb|ETT-m2| and \verb|WTH| dataset), our method obtains results that are higher than or close to those of \verb|NHITS|.
+Specifically, our results differ by 1-3\% on all metrics compared to the results of \verb|NHITS|.
+In the following subsections, we analyze these results on aperiodic datasets based on two factors: (1) - Feature extraction ability; (2) - Model synthesis ability; and relate to periodic datasets.
 
@@
     \toprule
-    \multicolumn{1}{c}{}                          &                                    & $accuracy$            & $precision$           & $recall$              & $F1$                    \\
+    \multicolumn{1}{c}{}                &                        & $accuracy$            & $precision$           & $recall$              & $F1-score$                    \\
     \hline
-    \multirow{3}{*}{\Verb|USD/JPY|}  & \Verb|NHITS|          & 53.29                 & 52.38                 & 52.23                 & 51.87                   \\
-                                                & Ours(\Verb|LSTM|)     & $57.12±0.08$ & $58.32±2.72$ & $57.44±2.62$ & $56.87±3.56$    \\
-                                                & Ours(\Verb|LSTM+CNN|) & $\mathbf{58.05±0.01}$ & $\mathbf{59.57±2.53}$ & $\mathbf{58.31±2.57}$ & $\mathbf{56.6±3.74}$    \\
+    \multirow{3}{*}{\Verb|USD/JPY|}     & \Verb|NHITS|           & $58.09$               & $58.32$               & $57.53$               & $56.53$              \\
+                                        & Ours(\Verb|LSTM|)      & $69.45±1.2$           & $69.87±1.04$          & $69.49±0.59$          & $68.89±1.43$            \\
+                                        & Ours(\Verb|LSTM+CNN|)  & $\mathbf{69.67±1.07}$ & $\mathbf{70.12±1.02}$ & $\mathbf{69.74±0.33}$ & $\mathbf{69.08±0.34}$    \\
     \hline
-    \multirow{3}{*}{\Verb|multi-fx|} & \Verb|NHITS|          & 51.12                 & 51.22                 & 51.16                 & 50.54                   \\
-                                                & Ours(\Verb|LSTM|)     & $61.23±0.06$          & $70.52±9.17$          & $69.51±9.31$          & $68.9±10.0$             \\
-                                                & Ours(\Verb|LSTM+CNN|) & $\mathbf{62.39±0.06}$ & $\mathbf{71.11±9.34}$ & $\mathbf{70.14±9.6}$  & $\mathbf{69.21±11.07}$  \\
+    \multirow{3}{*}{\Verb|multi-fx|}    & \Verb|NHITS|           & $52.63$               & $52.62$               & $52.59$               & $52.49$              \\
+                                        & Ours(\Verb|LSTM|)      & $63.63±2.79$          & $64.58±2.65$          & $63.87±1.93$          & $62.33±5.07$             \\
+                                        & Ours(\Verb|LSTM+CNN|)  & $\mathbf{63.78±2.4}$  & $\mathbf{64.84±0.76}$ & $\mathbf{64.06±1.67}$ & $\mathbf{62.37±3.89}$  \\
+    \hline
+    \multirow{3}{*}{\Verb|ETT-m2|}      & \Verb|NHITS|           & $71.72$               & $\mathbf{67.22}$      & $\mathbf{63.69}$      & $\mathbf{63.28}$             \\
+                                        & Ours(\Verb|LSTM|)      & $71.41±6.36$          & $63.48±2.41$          & $60.57±3.16$          & $59.77±3.42$             \\
+                                        & Ours(\Verb|LSTM+CNN|)  & $\mathbf{72.09±5.37}$ & $64.12±2.66$          & $61.13±2.78$          & $60.45±3.19$              \\
+    \hline
+    \multirow{3}{*}{\Verb|WTH|}         & \Verb|NHITS|           & $74.17$               & $68.18$               & $66.77$               & $67.08$              \\
+                                        & Ours(\Verb|LSTM|)      & $74.65±2.31$          & $68.76±2.38$          & $65.68±1.8$           & $65.84±2.03$             \\
+                                        & Ours(\Verb|LSTM+CNN|)  & $\mathbf{75.45±2.07}$ & $\mathbf{69.64±1.96}$ & $\mathbf{67.18±1.49}$ & $\mathbf{67.09±1.82}$  \\
     \bottomrule
@@
 
-We trained three models \verb|CNN|, \verb|LSTM|, \verb|LSTM+CNN| on 2600 samples of \verb|CHF/NZD| data in the conventional way.
+We trained three models \verb|CNN|, \verb|LSTM|, \verb|LSTM+CNN| on 2600 samples of \verb|USD/JPY| data in the conventional way.
 After observing their training process (figure \ref{fig:lstm_cnn}), we found that \verb|CNN| has excellent local feature extraction ability on short-term data.
@@
     \includegraphics[width=\linewidth]{img/models.pdf}
-    \caption{Training process of LSTM, CNN, and LSTM+CNN using 2600 samples from CHF/NZD dataset.}
+    \caption{Training process of LSTM, CNN, and LSTM+CNN using 2600 samples from USD/JPY dataset.}
     \label{fig:lstm_cnn}
@@
 
-Later, when we used ML algorithms to train feature extraction networks, we also found that the model using combined features was less dependent on the learning rate than the model using single features.
+% % phải lột đoạn này đi, mà chắc tôi đi ăn trước đã
+% Later, when we used ML algorithms to train feature extraction networks, we also found that the model using combined features was less dependent on the learning rate than the model using single features.
 Specifically, for the learning rate space we defined in Appendix \ref{app:lr}, all models converged with an accuracy of at least 54\%.
 This demonstrates the superiority of the feature provided by \verb|LSTM+CNN|.
+
+% \vspace{2mm}
 
@@
     \includegraphics[width=\linewidth]{img/acc.pdf}
-    \caption{Convergence process on validation set of neural network using LSTM+CNN feature and LSTM feature only (multi-fx dataset). The blur domains cover 99.73\% ($±3\sigma$) values of accuracies.}
+    \caption{Convergence process on validation set of ML model using LSTM+CNN feature and LSTM feature only (predict close price of multi-fx dataset).
+The blur domains cover 99.73\% ($±3\sigma$) values of accuracies.}
     \label{fig:lstm_cnn_feature}
@@
 
+% Quá trình rút trích và tổng hợp đặc trưng của \verb|NHITS| thể hiện rằng phương pháp này đang cố gắng mô phỏng lại quá trình phân giải tần số của Fourier transform.
+Đặc trưng về các giải tần là thông tin tối quan trọng đối với dữ liệu có chu kỳ.
+Do đó, các dự đoán của \verb|NHITS| có thể dễ dàng đạt được độ chính xác cao trên loại dữ liệu này.
+Tuy nhiên, rất khó để có thể phân giải tần số trên dữ liệu phi chu kỳ.
+Hơn nữa, không khó để nhận ra các kỹ thuật rút trích đặc trưng của \verb|NHITS| là vô cùng đơn giản (chỉ gồm các lớp \verb|FullyConnected| và \verb|Pooling|).
+Chính những điều này đã khiến cho \verb|NHITS| gặp trở ngại lớn trên \verb|multi-fx|, \verb|USD/JPY| datasets và các dữ liệu phi chu kỳ nói chung.
+Mặt khác, phương pháp đề xuất của chúng tôi sử dụng các mạng deep learning để rút trích đặc trưng nên có thể dễ dàng thu được các deep hidden feature.
+Do đó, trên cả periodic data lẫn aperiodic data, phương pháp của chúng tôi đều đạt được kết quả xấp xỉ hoặc cao hơn \verb|NHITS|.
+
+The feature extraction process of \verb|NHITS| shows that this method is trying to simulate the frequency resolution process of Fourier transform.
+The frequency features are the most important information for periodic data.
+Therefore, the predictions of \verb|NHITS| can easily achieve high accuracy on this type of data.
+However, it is very difficult to resolve the frequency on non-periodic data.
+Moreover, it is not difficult to see that the feature extraction techniques of \verb|NHITS| are extremely simple (only including \verb|FullyConnected| and \verb|Pooling| classes).
+These are the reasons why \verb|NHITS| encountered great difficulties on \verb|multi-fx|, \verb|USD/JPY| dataset and non-periodic data in general.
+On the other hand, our proposed method uses deep learning networks for feature extraction, so it can easily obtain deep hidden features.
+Therefore, on both periodic and aperiodic data, our method achieves results approximately equal to or better than \verb|NHITS|.
+
 % Quan sát quá trình rút trích đặc trưng của \verb|NHITS|, có thể thấy mô hình này đang cố gắng mô phỏng lại quá trình phân giải tần số của Fourier transform.
+Đối với dữ liệu thể hiện tính chu kỳ rõ rệt như \verb|ETT-m2| và \verb|WTH|, hiệu quả mà \verb|NHITS| mang lại là rất lớn.
 Tuy nhiên, trong toàn bộ quá trình thực thi, \verb|NHITS| sử dụng các kiến trúc rất đơn giản (chỉ bao gồm các lớp \verb|MLP| và \verb|Pooling|).
 Đối với các dữ liệu phức tạp và phi chu kỳ, sử dụng các layer \verb|MLP| là không đủ để rút trích các đặc trưng ẩn.
-
-Looking at the feature extraction process of \verb|NHITS|, it can be seen that this model tries to simulate the frequency resolution process of Fourier transform.
-However, in the whole implementation, \verb|NHITS| uses a very simple architecture (only including \verb|MLP| and \verb|Pooling| layers).
-For complex and non-periodic data, using \verb|MLP| layers is not enough to extract hidden features.
+Trái lại, phương pháp của chúng tôi có thể dễ dàng rút trích các deep hidden features, thể hiện ở việc đạt độ chính xác cao trên dữ liệu phi chu kỳ.
+Nhưng khi đối
+
+% Observing the feature extraction process of \verb|NHITS|, it can be seen that this model is trying to simulate the frequency resolution process of Fourier transform.
+For data with obvious periodicity such as \verb|ETT-m2| and \verb|WTH|, the efficiency that \verb|NHITS| brings is very good.
+However, in the whole implementation, \verb|NHITS| uses very simple architectures (only including \verb|MLP| and \verb|Pooling| layers).
+For complex and aperiodic data, using \verb|FullyConnected| layers is not enough to extract hidden features.
 
@@
 
-The reason why conventional training methods do not perform well despite the improved feature extraction capabilities is the lack of the ability to use features effectively.
+The reason why conventional training methods do not perform well despite the improved feature extraction capabilities is due to the lack of the ability to use features effectively.
 This ability in previous studies has often been improved by using ensemble models.
@@
 
-When using \verb|MAML| to synthesize the features of the \verb|LSTM| and \verb|LSTM+CNN| models, the compatibility (indicated by early convergence) as well as the results (indicated by accuracy) of the model using \verb|LSTM+CNN| are significantly higher than those of the model using only \verb|LSTM| (see figure \ref{fig:lstm_cnn_feature}) and the \verb|NHITS| model.
-Indeed, after only 40 epochs, the \verb|LSTM+CNN| features converge to over 60\% and after 100 epochs, the model converges to 62\%, exceeding what the \verb|LSTM| features can do.
-The variation of the accuracy values ​​is very small, indicating the stability of \verb|MAML|.
-For the ML model using only \verb|CNN|, the accuracy and error values ​​do not even improve during the entire training process.
+When using \verb|MAML| to synthesize the features of the \verb|LSTM| and \verb|LSTM+CNN| models, the compatibility (indicated by early convergence) as well as the results (indicated by accuracy) of the model using \verb|LSTM+CNN| are significantly higher than those of the model using only \verb|LSTM| (see figure \ref{fig:lstm_cnn_feature}) and \verb|NHITS| model.
+Indeed, on \verb|multi-fx| dataset, after only 40 epochs, the \verb|LSTM+CNN| features converge to over 60\% and after 100 epochs, the model converges to 62\%, exceeding what the \verb|LSTM| features can do.
+The variation of the accuracy values is very small, indicating the stability of \verb|MAML|.
+For the ML model using only \verb|CNN|, the accuracy and error values do not even improve during the entire training process.
 We explain this by saying that the features of \verb|CNN| focus too much on the locality of the data and ignore the long-term dependencies, which is an important feature of time-series data.
@@
 
-When approaching the \verb|USD/JPY| data using ML, the model also shows high accuracy.
+When approaching \verb|USD/JPY|, \verb|ETT-m2|, and \verb|WTH| dataset using ML, the model also shows accuracies higher than or close to \verb|NHITS|.
 By looking at the past time periods, \textbf{the proposed method has proven to be more effective in capturing hidden long-term dependencies through the efficient aggregation of features that the} \verb|LSTM| \textbf{easily forgets during training}.
@@
 By combining local features and long-term dependencies as well as exploiting hidden dependencies in each past time period, we have improved the ability of machine learning models on classification metrics in the problem of predicting the next day's price trend on FX data.
-The proposed algorithm has demonstrated its superior ability when compared with the \verb|NHITS| model.
+The proposed algorithm has demonstrated its superior performance when compared to \verb|NHITS| on aperiodic data as well as achieving an accuracy close to \verb|NHITS| on periodic data.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\clearpage
+% \clearpage
 % \bibliographystyle{elsarticle-harv}
@@
 \bibliography{references}
-\clearpage
+% \clearpage
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\begin{table}[h]
+\begin{table}[H]
     \centering
@@
 \label{app:nhits_experiment}
-
-% Đối với mô hình \verb|NHITS|, chúng tôi dựa trên \cite{challu2023nhits} để định nghĩa không gian tìm kiếm cho việc fine-tune tham số, cũng như kiến trúc mô hình.
-Đối với các tham số không được đề cập trong bảng, chúng tôi sử dụng giá trị mặc định của cài đặt \verb|NHITS| trong thư viện \verb|NeuralForecast| \cite{neuralforecast}.
-Chi tiết quá trình fine-tune được trình bày trong phần phụ lục \ref{app:experiment}.
-Kết quả tốt nhất của các lần fine-tune được chọn ra và báo cáo trong nghiên cứu này.
 
@@
 
+% Đối với mô hình \verb|NHITS|, chúng tôi dựa trên \cite{challu2023nhits} để định nghĩa không gian tìm kiếm cho việc fine-tune tham số, cũng như kiến trúc mô hình.
+Đối với các tham số không được đề cập trong bảng, chúng tôi sử dụng giá trị mặc định của cài đặt \verb|NHITS| trong thư viện \verb|NeuralForecast| \cite{neuralforecast}.
+Chi tiết quá trình fine-tune được trình bày trong phần phụ lục \ref{app:experiment}.
+Kết quả tốt nhất của các lần fine-tune được chọn ra và báo cáo trong nghiên cứu này.
+
 We rely on \cite{challu2023nhits} to define the search space (table \ref{tab:nhits_finetune}) for parameter fine-tuning, as well as to fine-tune the model architecture.
@@
 
-\section{Hyper-parameter dependence}
-\label{app:lr}
-
-% Trong quá trình thực nghiệm, chúng tôi nhận thấy việc hội tụ của mô hình ML sử dụng đặc trưng của \verb|LSTM+CNN| có thể giảm bớt sự phụ thuộc vào learning rate.
+% \section{Hyper-parameter dependence}
+% \label{app:lr}
+
+% % Trong quá trình thực nghiệm, chúng tôi nhận thấy việc hội tụ của mô hình ML sử dụng đặc trưng của \verb|LSTM+CNN| có thể giảm bớt sự phụ thuộc vào learning rate.
 Cụ thể, chúng tôi thử nghiệm trên 32 mô hình trên dữ liệu \verb|multi-fx|.
@@
 
-During the experiment, we found that the convergence of ML models using the \verb|LSTM+CNN| feature can reduce the dependence on learning rate.
+% During the experiment, we found that the convergence of ML models using the \verb|LSTM+CNN| feature can reduce the dependence on learning rate.
 Specifically, we tested 32 models on \verb|multi-fx| data.
@@
 
-\vspace{2mm}
-
-% Quá trình hội tụ trên tập validation của các mô hình được ghi lại trong hình \ref{fig:lr}.
+% \vspace{2mm}
+
+% % Quá trình hội tụ trên tập validation của các mô hình được ghi lại trong hình \ref{fig:lr}.
 Theo đó, toàn bộ mô hình đều đạt hội tụ với độ chính xác thấp nhất là trên 54\%.
@@
 
-The convergence process on the validation set of the models is recorded in figure \ref{fig:lr}.
+% The convergence process on the validation set of the models is recorded in figure \ref{fig:lr}.
 Accordingly, all models achieved convergence with the lowest accuracy of over 54\%.
@@
 
-\vspace{2mm}
-
-% Khi thực hiện thí nghiệm tương tự với mô hình \verb|LSTM|, độ chính xác phụ thuộc rất nhiều vào việc lựa chọn các siêu tham số học.
+% \vspace{2mm}
+
+% % Khi thực hiện thí nghiệm tương tự với mô hình \verb|LSTM|, độ chính xác phụ thuộc rất nhiều vào việc lựa chọn các siêu tham số học.
 Khả năng tương thích trên dữ liệu cũng giảm đáng kể.
@@
 
-When performing similar experiments with the \verb|LSTM| model, the accuracy depends heavily on the choice of learning hyper-parameters.
+% When performing similar experiments with the \verb|LSTM| model, the accuracy depends heavily on the choice of learning hyper-parameters.
 The adaptability to the data also drops significantly.
@@
 
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=\linewidth]{img/lr.pdf}
-    \caption{Convergence processes of 32 models using LSTM+CNN as feature extractor.}
-    \label{fig:lr}
-\end{figure}
-
-\vspace{2mm}
-
-% tôi có thể viết thêm 1 phần phụ lục nhằm phân tích tác động của lookback window
-% Tuy vậy, quá trình hội tụ của chúng tôi lại phụ thuộc rất lớn vào lookback window.
+% \begin{figure}[h]
+%     \centering
+%     \includegraphics[width=\linewidth]{img/lr.pdf}
+%     \caption{Convergence processes of 32 models using LSTM+CNN as feature extractor.}
+%     \label{fig:lr}
+% \end{figure}
+
+% \vspace{2mm}
+
+% % tôi có thể viết thêm 1 phần phụ lục nhằm phân tích tác động của lookback window
+% % Tuy vậy, quá trình hội tụ của chúng tôi lại phụ thuộc rất lớn vào lookback window.
 Thử nghiệm trên mô hình \verb|LSTM| với dữ liệu \verb|NHITS| cho thấy, khi xét trên cùng một tập siêu tham số và cùng một kiến trúc mô hình, nếu tăng lookback window từ 10 lên 20, số lượng mô hình hội tụ tăng đột biến.
@@
 
-However, the convergence process of the proposed method depends heavily on the lookback window.
+% However, the convergence process of the proposed method depends heavily on the lookback window.
 Experiments on the \verb|LSTM| model with \verb|NHITS| data show that, when considering the same set of hyper-parameters and the same model architecture, if the lookback window is increased from 10 to 20, the number of converged models increases dramatically.
@@
 
-\begin{figure}[H]
-    \centering
-    \begin{subfigure}[b]{\linewidth}
-        \centering
-        \includegraphics[width=\textwidth]{img/lstm_10.pdf}
-        \caption{lookback window = 10}
-        \label{fig:image1}
-    \end{subfigure}
-    % \hfill
-    % \vspace{2mm}
-    \begin{subfigure}[b]{\linewidth}
-        \centering
-        \includegraphics[width=\textwidth]{img/lstm_20.pdf}
-        \caption{lookback window = 20}
-        \label{fig:image2}
-    \end{subfigure}
-    % \hfill
-    % \vspace{2mm}
-    \begin{subfigure}[b]{\linewidth}
-        \centering
-        \includegraphics[width=\textwidth]{img/lstm_30.pdf}
-        \caption{lookback window = 30}
-        \label{fig:image3}
-    \end{subfigure}
-    \caption{Change in the model's convergence process when changing the lookback window.}
-    \label{fig:three_images}
-\end{figure}
+% \begin{figure}[H]
+%     \centering
+%     \begin{subfigure}[b]{\linewidth}
+%         \centering
+%         \includegraphics[width=\textwidth]{img/lstm_10.pdf}
+%         \caption{lookback window = 10}
+%         \label{fig:image1}
+%     \end{subfigure}
+%     % \hfill
+%     % \vspace{2mm}
+%     \begin{subfigure}[b]{\linewidth}
+%         \centering
+%         \includegraphics[width=\textwidth]{img/lstm_20.pdf}
+%         \caption{lookback window = 20}
+%         \label{fig:image2}
+%     \end{subfigure}
+%     % \hfill
+%     % \vspace{2mm}
+%     \begin{subfigure}[b]{\linewidth}
+%         \centering
+%         \includegraphics[width=\textwidth]{img/lstm_30.pdf}
+%         \caption{lookback window = 30}
+%         \label{fig:image3}
+%     \end{subfigure}
+%     \caption{Change in the model's convergence process when changing the lookback window.}
+%     \label{fig:three_images}
+% \end{figure}
 

### 2024-10-08 common/common.py
see README.md

@@ -2,13 +2,20 @@ import numpy as np
 from sklearn.metrics import precision_recall_fscore_support
 from keras import metrics
 
-def compute_metrics(pred_values:list[float], true_values=None):
-    acc_fn = metrics.Accuracy()
+from .constants import *
+
+def compute_metrics(pred_values:list[float], true_values=None, mode:str=CLF):
     if true_values is None:
         # compute metrics across tasks or simply take the average of a list of metric
         return float(np.mean(pred_values)), float(np.std(pred_values))
     else:
         # compute metrics in a task
-        acc_fn.update_state(true_values, pred_values)
-        precision, recall, f1, _ = precision_recall_fscore_support(true_values, pred_values, average='macro')
-        return float(acc_fn.result()), float(precision), float(recall), float(f1)
+        if mode == CLF:
+            metric_fn = metrics.Accuracy()
+            metric_fn.update_state(y_true=true_values, y_pred=pred_values)
+            precision, recall, f1, _ = precision_recall_fscore_support(y_true=true_values, y_pred=pred_values, average='macro')
+            return float(metric_fn.result()), float(precision), float(recall), float(f1)
+        elif mode == REG:
+            metric_fn = metrics.MeanSquaredError()
+            metric_fn.update_state(true_values, pred_values)
+            return float(metric_fn.result())

### 2024-10-08 common/constants.py

@@ -12,14 +12,14 @@ USD_JPY = 'USD_JPY'
 ETT = 'ett'
 ECL = 'ecl'
 WTH = 'wth'
-FACTORY_1 = 'important_1'
-FACTORY_2 = 'important_2'
+FACTORY_1 = 'factory_1'
+FACTORY_2 = 'factory_2'
 
 # mode
 REG = 'regression'
 CLF = 'classification'
-VAL = 'validation'
-TEST = 'test'
+VAL = 'Validation'
+TEST = 'Test'
 
 # based model
 LSTM = 'lstm'

### 2024-10-08 data/dataloader.py

@@ -1,0 +1,149 @@
+import pandas as pd
+from sklearn.preprocessing import StandardScaler
+from sklearn.model_selection import train_test_split
+import numpy as np
+
+from common.constants import *
+
+class DataLoader:
+    def __init__( self, dataset:str, look_back:int=None, batch_size:int=32, test_size:float=0.8, mode:str=CLF) -> None:
+        """initialize for dataloader. in case of NHITS, we need nothing but the `dataset`
+
+        Args:
+            dataset (str): name of dataset
+            look_back (int, optional): number of historical data-points used to predict, None means loading data for NHITS. Defaults to None.
+            batch_size (int, optional): batch size. Defaults to 32. If `batch_size=-1`, then we don't divide data into batches
+            test_size (float, optional): the amount of data in query set. Defaults to 0.8.
+            mode (str, optional): either classification problem or regression problem. Defaults to CLF.
+        """
+        self.look_back = look_back
+        self.batch_size = batch_size
+        self.test_size = test_size
+        self.dataset = dataset
+        self.data_dir = os.path.join(RAW_DATA_DIR, dataset)
+        self.mode = mode
+
+        # extract columns by reading 1 *.csv file
+        # this is not used for NHITS, the one for NHITS is written at `get_data`
+        for filename in os.listdir(self.data_dir):
+            if filename.endswith('.csv'):
+                tmp_df = pd.read_csv(os.path.join(self.data_dir, filename)).to_numpy()[:,1:]
+                self.columns = list(range(tmp_df.shape[1]))
+                break
+
+    def _create_label(self, df:np.ndarray, label:int) -> np.ndarray:
+        """create label for a given dataframe
+        if this is REG problem, we remove that column in the given dataframe
+
+        Args:
+            df (np.ndarray): given dataframe
+            label (int): index of column chosen to be the label
+
+        Returns:
+            np.ndarray: the label column
+        """
+        label_col = df[:,label]
+        if self.mode == CLF:
+            label_col = label_col[1:] - label_col[:-1]
+            label_col = (label_col>0)*1.
+            label_col = np.append([-1], label_col) # the first value is assigned to -1, it's removed anyway
+        elif self.mode == REG:
+            # if it is regression problem, we have to drop the label column in df
+            df = np.delete(df, label, axis=1)
+        return label_col
+
+    def _normalize_data(self, df:np.ndarray, label:int=-1) -> np.ndarray:
+        scaler = StandardScaler()
[…]

### 2024-10-08 data/pre_process.py

@@ -1,81 +1,0 @@
-import pandas as pd
-from sklearn.preprocessing import StandardScaler
-from sklearn.model_selection import train_test_split
-import numpy as np
-
-from common.constants import *
-
-def create_label(df:np.ndarray, label:int, mode:str=CLF) -> np.ndarray:
-    label = df[:,label]
-    if mode == CLF:
-        label = label[1:] - label[:-1]
-        label = (label>0)*1.
-        label = np.append([-1], label) # the first value is assigned to -1, it's removed anyway
-    elif mode == REG:
-        pass # if it is regression problem, we don't have to do anything
-    return label
-
-def normalize_data(df:np.ndarray) -> np.ndarray:
-    scaler = StandardScaler()
-    return scaler.fit_transform(df)
-
-# create dataset with lookback window
-def create_dataset(df:np.ndarray, look_back:int, label:int)->tuple[np.ndarray, np.ndarray]:
-    size = df.shape[0] - look_back
-    X = np.array([df[i : i + look_back] for i in range(size)])
-    y = create_label(df, label)
-    y = y[look_back:]
-    return X, y.reshape(-1,1)
-
-def get_data(look_back:int, label:int=-1, test_size:float=0.8, batch_size:int=32, data_dir:str=RAW_DATA_DIR) -> dict[str, dict[str, list[np.ndarray]]]:
-    """
-        read data from files and return a dict whose each value contains data from a file
-
-        label (int, optional): The label to be predict the movement. Defaults to -1 (the final feature).
-        batch_size (int, optional): If `batch_size=-1`, then we don't divide data into batches
-    """
-    cid = 0
-    data_dict:dict[str, dict[str, pd.DataFrame]] = {}
-    for file_name in os.listdir(data_dir):
-        if file_name.endswith('.csv'):
-            # read dataset, drop the datetime/index column
-            df = pd.read_csv(os.path.join(data_dir, file_name)).to_numpy()[:,1:]
-
-            # normalize data
-            transformed_df = normalize_data(df)
-
-            # create dataset with lookback window
-            X, y = create_dataset(transformed_df, look_back, label)
-
-            tmp_dict = {}
-            tmp_dict['support_X'], tmp_dict['query_X'], tmp_dict['support_y'], tmp_dict['query_y'] = train_test_split(X, y, test_size=test_size, shuffle=False)
-            if batch_size != -1: # do not divide data into batches
-                for key in tmp_dict.keys():
[…]

### 2024-10-08 execute/main.py

@@ -2,12 +2,12 @@ import tensorflow as tf
 import random
 random.seed(84)
 import argparse
-import pandas as pd
 import json
+import numpy as np
 
 from model.meta_model import MAML
 from common.constants import *
-from data.pre_process import get_data
+from data.dataloader import DataLoader
 
 def parse_args():
     parser = argparse.ArgumentParser()
@@ -23,22 +23,28 @@ def parse_args():
 
     return vars(parser.parse_args())
 
-def run(dataset:str, input_dict:dict, label:int):
-    # prep data
-    data_dir = os.path.join(RAW_DATA_DIR, dataset)
-    data_dict = get_data(look_back=input_dict['look_back_window'], data_dir=data_dir, label=label)
+def run(data_dict:dict[str, dict[str, list[np.ndarray]]], input_dict:dict, model_dir:str, is_multi:bool=False) -> dict[str:float]:
+    """train model with given hyper-param in `input_dict`
 
-    if 'multi' in dataset:
+    Args:
+        data_dict (dict[str, dict[str, list[np.ndarray]]]): dictionary containing all tasks
+        input_dict (dict): hyper-param
+        model_dir (str): directory to save model and its log
+        is_multi (bool, optional): if True, shuffle tasks. Defaults to False.
+
+    Returns:
+        metric_dict_: a dictionary containing all metrics for CLF or REG problem
+    """
+    list_id = set(data_dict.keys())
+    if is_multi:
         # only apply for multi_fx
         # choose randomly 50% clients for training, 25% for validating and 25% for testing
-        list_id = set(data_dict.keys())
         list_id_train = random.sample(sorted(list_id), len(list_id)//2)
         list_id_val = random.sample(sorted(list_id.difference(set(list_id_train))), len(list_id)//4)
         list_id_test = list(list_id.difference(set(list_id_train).union(list_id_val)))
     else:
         # bởi vì các dataset khác không multi
         # nên phải đảm bảo tính thứ tự trong huấn luyện
-        list_id = list(data_dict.keys())
         list_id_train = list_id[:int(len(list_id)*0.5)]
         list_id_val = list_id[int(len(list_id)*0.5):int(len(list_id)*0.75)]
         list_id_test = list_id[int(len(list_id)*0.75):]
@@ -55,57 +61,69 @@ def run(dataset:str, input_dict:dict, label:int):
         mode=input_dict['mode']
     )
 
[…]

### 2024-10-08 execute/submit_jobs.py

@@ -8,6 +8,7 @@ all `tmp.sh` files are supposed that they are called from the root directory
 
 import subprocess
 import time
+import itertools
 
 from common.constants import *
 
@@ -54,29 +55,21 @@ def generate_script(device:str, name_job:str, look_back:int, model_name:str, mod
 
 # finetune meta model (based model: CNN, LSTM)
 def finetune_1():
-    current_count = None
-    prev_count = 96
+    windows = [10, 20, 30]
+    models = [LSTM, CNN]
+    inner_lrs = [0.001, 0.005, 0.01, 0.05]
+    outer_lrs = [0.001, 0.005, 0.0015, 0.0055]
+
     count = 0
-    for window in [10, 20, 30]:
-        for model in [LSTM, CNN]:
-            for inner_lr in [0.001, 0.005, 0.01, 0.05]:
-                for outer_lr in [0.001, 0.005, 0.0015, 0.0055]:
-                    if count < 20:
-                        count += 1
-                        continue
+    for window, model, inner_lr, outer_lr in itertools.product(windows, models, inner_lrs, outer_lrs):
+        if count < 20:
+            count += 1
+            continue
 
-                    generate_script('cpu', f'{model}_{count}', window, model, CLF, inner_lr, outer_lr, 5, 100, 3)
-                    subprocess.run(["qsub", os.path.join(EXE_DIR, 'tmp.sh')])
-                    time.sleep(1)
-                    count += 1
-
-                    if count == current_count:
-                        return
-
-    # for project_name in [LSTM, CNN]:
-    #     generate_script('cpu', project_name)
-    #     subprocess.run(["qsub", os.path.join(EXE_DIR, 'tmp.sh')])
-    #     time.sleep(3)
+        generate_script('cpu', f'{model}_{count}', window, model, CLF, inner_lr, outer_lr, 5, 100, 3)
+        subprocess.run(["qsub", os.path.join(EXE_DIR, 'tmp.sh')])
+        time.sleep(1)
+        count += 1
 
     # # demo
     # generate_script('gpu', WEAK_LEARNERS[0], 'boosting')
@@ -84,17 +77,19 @@ def finetune_1():
 
 # finetune meta model (based model: LSTM, LSTM+CNN)
 def finetune_2():
+    windows = [20, 30]
[…]

### 2024-10-08 model/based_model.py

@@ -1,6 +1,5 @@
 # libraries
-from keras import Sequential, layers, models
-import tensorflow as tf
+from keras import layers, models
 
 from common.constants import *
 

### 2024-10-08 model/baseline_models.py

@@ -1,17 +1,18 @@
 from neuralforecast import NeuralForecast
 from neuralforecast.models import NHITS
-from neuralforecast.losses.pytorch import DistributionLoss
+from neuralforecast.losses.pytorch import DistributionLoss, MSE
 
 import os
 import json
 import pandas as pd
 import json
+import itertools
 
-from data.pre_process import get_data_for_nhits
+from data.dataloader import DataLoader
 from common.constants import *
 from common.common import compute_metrics
 
-def add_log(log:dict, count:int, input_size:int, n_stacks:int, kernel_size:int, mlp_units:int, lr:float, freq:list[int], interpolation_mode:str, acc:float, precision:float, recall:float, f1:float):
+def add_log(log:dict, count:int, input_size:int, n_stacks:int, kernel_size:int, mlp_units:int, lr:float, freq:list[int], interpolation_mode:str, metric_log:dict):
     # init
     log[count] = {}
 
@@ -25,82 +26,97 @@ def add_log(log:dict, count:int, input_size:int, n_stacks:int, kernel_size:int, 
     log[count]["interpolation_mode"] = interpolation_mode
 
     # metrics
-    log[count]["acc"] = acc
-    log[count]["precision"] = precision
-    log[count]["recall"] = recall
-    log[count]["f1"] = f1
+    for key in metric_log:
+        log[count][key] = metric_log[key]
 
 def write_log(log:dict, file_name:str):
     with open(os.path.join(PRETRAINED_DIR, 'baseline', file_name), 'w') as fo:
         json.dump(log, fo)
 
-def run(dataset:str, label:int, columns:list[str]):
-    data_dir = os.path.join(RAW_DATA_DIR, dataset, 'all')
-    data_dict = get_data_for_nhits(data_dir=data_dir, label=label)
+def run(df:pd.DataFrame, param_dict:dict, columns:list[str]):
+    loss_fn = param_dict['loss_fn']
+    model = NHITS(
+        # const
+        h=1, loss=loss_fn, valid_loss=loss_fn, max_steps=500, hist_exog_list=columns, val_check_steps=5,
+        # early_stop_patience_steps=5,
 
+        # finetune
+        n_pool_kernel_size = param_dict['n_pool_kernel_size'],
+        n_freq_downsample = param_dict['n_freq_downsample'],
+        learning_rate = param_dict['learning_rate'],
+        input_size = param_dict['input_size'],
+
[…]

### 2024-10-08 model/meta_model.py
meta_model.py and analyze.py should get back to the closest version

@@ -1,6 +1,5 @@
 import tensorflow as tf
 from keras import optimizers, models, metrics
-# from sklearn.metrics import precision_recall_fscore_support
 import numpy as np
 
 from copy import deepcopy
@@ -35,19 +34,18 @@ class MAML:
         self.info['inner_lr'] = inner_lr
         self.info['outer_lr'] = outer_lr
         self.info['train_loss'] = []
-        self.info['train_acc'] = []
         self.info['val_loss'] = []
-        self.info['val_acc'] = []
         self.info['val_std_loss'] = []
-        self.info['val_std_acc'] = []
 
         # init metrics
         self.mode = mode
         if mode == REG:
             self.loss_fn = metrics.mean_squared_error
         elif mode == CLF:
+            self.info['train_acc'] = []
+            self.info['val_acc'] = []
+            self.info['val_std_acc'] = []
             self.loss_fn = metrics.binary_crossentropy
-            # self.acc_fn = metrics.Accuracy()
 
     def inner_training_step(self, model:models.Model, X:list[np.ndarray], y:list[np.ndarray], num_epochs:int=2):
         """fast adapt on support set (inner step)
@@ -118,20 +116,18 @@ class MAML:
             # compute sum loss for back-propagation
             sum_task_losses += task_loss
 
-            # compute acc + loss of a task
+            # compute acc + loss of a task then print out
+            task_loss /= len(X)
+            tasks_metrics['loss'].append(task_loss)
             if self.mode == CLF:
                 acc_, precision_, recall_, f1_ = compute_metrics(pred_values, true_values)
                 tasks_metrics['acc'].append(acc_)
                 tasks_metrics['precision'].append(precision_)
                 tasks_metrics['recall'].append(recall_)
                 tasks_metrics['f1'].append(f1_)
-                task_loss /= len(true_values)
                 print(f'Outer compute on task {task_id}: loss={task_loss:.5f}\t acc={acc_:.5f}')
             elif self.mode == REG:
-                task_loss /= len(X)
                 print(f'Outer compute on task {task_id}: loss={task_loss:.5f}')
-
-            tasks_metrics['loss'].append(task_loss)
 
[…]

### 2024-10-08 doc/paper/paper.tex

@@
 Predicting aperiodic time-series data (e.g.
-stock price, foreign exchange, Bitcoin price,...) is a difficult task for machine learning models because this type of data has high and not stationary variance; does not show clearly periodicity, making it difficult to extract features; depends not only on past values but also on external factors, which make them unstable and non-cyclical, such as economic and political situations.
-To overcome the above challenges, we use Meta-learning to train a combined \verb|LSTM| and \verb|CNN| network, thereby effectively extracting and synthesizing hidden features of the data over time.
-Experiments on foreign exchange data of 60 currency pairs over 24 years (2000-2024) show that the proposed method performs well and has 5\%-11\% higher accuracy compared to the \verb|NHITS| - the state-of-the-art model in 2023 on time-series data, in the task of predicting the trend (upward or downward) of the next trading day.
+stock price, foreign exchange, Bitcoin price, etc.) is a difficult task for machine learning models because this type of data has high and not stationary variance; does not show clearly periodicity, making it difficult to extract features; depends not only on past values but also on external factors such as economic and political situations, which make them unstable and non-cyclical.
+To overcome the above challenges, we use Meta-learning to train a combined \verb|LSTM| and \verb|CNN| network, thereby effectively extracting and synthesizing deep hidden features of the data over time.
+Experiments on aperiodic data (foreign exchange data of 60 currency pairs over 24 years from 2000 to 2024) show that the proposed method performs well and has 5\%-11\% higher accuracy compared to the \verb|NHITS| - the state-of-the-art model in 2023 on time-series data, in the task of predicting the trend (upward or downward) of the next trading day.
+Meanwhile, the results obtained on periodic data (Electricity Transformer Temperature, Weather) are approximately \verb|NHITS|.
 
@@
 
-For the variance variation challenge, ensemble models \citep{sadeghi2021combined, zafeiriou2020intraday, ali2020complete} are often used to mitigate the effects of variance variation.
+Ensemble models \citep{sadeghi2021combined, zafeiriou2020intraday, ali2020complete} are often used to mitigate the effects of variance variation challenge.
 Ensemble models provide a holistic and multi-perspective view based on sub-models, thereby helping the overall model adapt to strong variance changes.
@@
 This is completely understandable because \verb|LSTM| and \verb|CNN| have proven their ability to extract spatial and temporal features effectively.
-However, when faced with aperiodic data with only 2600 samples, after 40 training steps, \verb|LSTM| is still underfitting, while \verb|CNN| is overfitting after only 5 training steps.
+However, when faced with aperiodic data with only 2600 samples, after 40 training steps, our experiment shows that \verb|LSTM| is still underfitting, while \verb|CNN| is overfitting after only 5 training steps.
 
@@
 
-Regarding the external influences challenge, \cite{fama1970efficient} hypothesizes that different time-series datasets of the same domain at the same time reflect the impact of external factors.
-For example, studies \citep{overreactioncontrarian, mech1993portfolio} show the dependence between the financial ratios of a given company and the ratios of other companies.
+Regarding the challenge of external influences, \cite{fama1970efficient} hypothesizes that different time-series datasets of the same domain at the same time reflect the impact of external factors.
+Studies \citep{overreactioncontrarian, mech1993portfolio} also show the dependence between the financial ratios of a given company and the ones of other companies.
 This further strengthens the hypothesis in study \cite{fama1970efficient}.
@@
 
-We approach the above challenges through combination of feature combination and model synthesis.
+We approach the above challenges through the combination of feature and model synthesis.
 Specifically, by using Meta-learning (ML) \cite{finn2017model}, we efficiently synthesize the parameters of local models, which significantly reduces variance loss and increases the ability to collect information reflecting external factors from datasets in the same domain.
@@
 Traditional approaches use a fixed amount of past data (lookback window) to train the model.
-This poses a major obstacle to the learning process because long-term features will be forgotten over time.
+This poses a great obstacle to the learning process because long-term features will be forgotten over time.
 On the other hand, ML divides the dataset into many parts to learn and synthesize the learned parameters effectively, so it can handle this challenge well.
@@
 
-Finally, we demonstrate the superiority of the proposed algorithm by solving the problem of predicting the trend (upward or downward) of foreign exchange rates and comparing the results with the existing state-of-the-art (SOTA) model (\verb|NHITS| \cite{challu2023nhits}) on two types of data: (1) - USD/JPY exchange rate data; (2) - Exchange rate data of 60 currency pairs, comprising 18 countries.
-These data sets are publicly available on the Internet and can be easily downloaded.
+Finally, we demonstrate the superiority of the proposed algorithm by solving the problem of predicting the trend (upward or downward) of time-series data and comparing the results with the existing state-of-the-art (SOTA) model (\verb|NHITS| \cite{challu2023nhits}) on two types of data: (1) - Aperiodic data, which contains USD/JPY exchange rate data and the exchange rate data of 60 currency pairs, comprising 18 countries from 2000 to 2024; (2) - Periodic data which contains Electricity Transformer Temperature \cite{zhou2021informer} and Weather \cite{Kolle} dataset.
+These datasets are publicly available on the Internet and can be downloaded easily.
 The official implementation can be found at \url{https://github.com/baolongnguyenmac/multi_fx}.
@@
 In summary, our main contributions include the following: First, we present a novel \textbf{feature combination technique} for aperiodic time-series data that extracts and combines features using both \verb|LSTM| and \verb|CNN| models.
-Second, we uncover hidden long-term dependencies, demonstrating experimentally that aperiodic time-series data not only rely on external factors but also exhibit dependencies with past period, which we call \textbf{hidden long-term dependencies}.
+Second, we discover the \textbf{hidden long-term dependencies} and demonstrate experimentally that aperiodic time-series data not only rely on external factors but also exhibit dependencies with past period.
 Third, we introduce an \textbf{efficient model parameter aggregation approach} by employing ML algorithm in place of traditional ensemble models for combining parameters from multiple machine learning models.
-Finally, we validate our method through experiments on exchange rate datasets and show its effectiveness in comparison to the SOTA model, \verb|NHITS|.
+Finally, we validate our method through experiments on exchange rate datasets and show its effectiveness in comparison to a SOTA model, \verb|NHITS|.
 
@@
 As aforementioned, \verb|LSTM| is a well-known neural network for handling prediction problems on time-series data.
-\verb|LSTM| is commonly used because it handles the problem of vanishing gradients well (easily encountered when using Recurrent neural networks) and can effectively exploit nonlinear relationships in data.
+\verb|LSTM| is commonly used because it efficiently handles the problem of vanishing gradients (easily encountered when using Recurrent neural networks) and can effectively exploit nonlinear relationships in data.
 Indeed, by maintaining the cell-state in each iteration, \verb|LSTM| can overcome the vanishing gradient problem, thereby preserving the ability to capture long-term dependencies \cite{cheng2018applied}.
@@
 stacking models).
-From the perspective of ensemble models, the equation \ref{eq:outer_opt} can be considered an effective method of synthesizing sub-models, which helps to take advantage of the feature extraction capabilities of each model.
+From the perspective of ensemble models, the equation \ref{eq:outer_opt} can be considered an optimization-based method of synthesizing sub-models, which helps to take advantage of the feature extraction capabilities of each model.
 In other words, the synthesized model can extract features at a deeper level, significantly improving the prediction ability compared to traditional ensemble models.
@@
 Each stack consists of multiple consecutive blocks.
-At each block, historical data is used to predict future data and past data.
-The residual of the previous block is used as input data for the following block.
+At each block, historical data is used to predict future data and approximate past data.
+The difference between historical data and the approximate one of the previous block is used as input data for the following block.
 Specifically, at block $l$, with $L$ past data samples ($\mathbf{y}_{t-L:t, l-1}$), the features will be extracted as follows (from \cite{challu2023nhits}):
@@
 $\mathbf{\theta}_l^f, \mathbf{\theta}_l^b$ are the forecast and backcast interpolation coefficients, which are used to aggregate the output values of block $l$ using the interpolation function $g(\cdot)$.
-The dimensionality of the interpolation coefficients in each stack is determined by the expressiveness ratio: $\left| \theta_l \right| = \left \lceil r_l H \right \rceil$.
+The dimensionality of the interpolation coefficients in each stack is determined by the expressiveness ratio: $\left| \theta_l^f \right| = \left \lceil r_l H \right \rceil$.
 Typically, the expressiveness ratio will be very small in the first stacks and increase towards the end.
@@
 Our method consists of two main parts that work in parallel: (1) - Feature extraction; (2) - Parameter synthesis.
-The overview of the method is illustrated in \ref{fig:flow}.
+The overview of the method is a ML process as mentioned in section \ref{sec.relatedWork} and is illustrated in figure \ref{fig:flow}.
 In the feature extraction part, we combine two types of features from \verb|CNN| and \verb|LSTM| networks.
@@
 In which, $\mathbf{x}_{t-L:t}$ includes $L$ historical values from time $t$ back; $y\in \{0,1\}$ is the data label, showing the decreasing or increasing trend of the data sample $x_{t+1}$ compared to $x_{t}$.
-Depending on each problem and the implementation, the elements in $\mathbf{x}_{t-L:t}$ can be vectors or scalar numbers.
-For example, for stock data, $\mathbf{x}_{t-L:t}$ can contain $L$ data vectors $\vec x_i = (\text{open, low, high, close})$ or just a single close price value.
+Depending on each problem and the implementation, the elements in $\mathbf{x}_{t-L:t}$ can be a matrix or a vector.
+For example, for stock data, $\mathbf{x}_{t-L:t}$ can contain $L$ vectors $\vec x_i = (\text{open, low, high, close})$ or can be a vector of close price values only.
 
@@
 \begin{align}
-    \mathbf{x'}_{t-L:t} &= \mathbf{FullyConnected}\left( \mathbf{x'}_{t-L:t} \right)\\
+    \mathbf{x'}_{t-L:t} &= \mathbf{FullyConnected}\left( \mathbf{x}_{t-L:t} \right)\\
     \mathbf{h}_{LSTM} &= \mathbf{BidirectionalLSTM}\left( \mathbf{x'}_{t-L:t} \right)\\
@@
 We propose to use the \verb|CNN| network to emphasize local features, thereby directing part of the model's attention to certain time points.
-Therefore, the proposed method can not only remember long-term features but also highlight short-term features.
+Therefore, the proposed method not only remembers long-term features but also highlights short-term features.
 
@@
 Each global step includes many local steps (inner optimization) performed on each individual task.
-At global step $r$-th, the $e$-th local optimization process at the support set of task $t$ occurs as follows:
+At outer optimization step $r$-th, the $e$-th inner optimization step at the support set of task $t$ occurs as follows:
 
@@
     \end{cases}
-\end{align} In which, $\phi_{r-1}$ is the result of the $r-1$ global optimization process, $\alpha$ is the inner learning rate.
+\end{align} In which, $\phi_{r-1}$ is the result of the $r-1$ outer optimization process, $\alpha$ is the inner learning rate.
 
@@
 
-Next, the outer optimization process at the global step is performed by aggregating the losses on the query set of the tasks and optimizing on it (equation \ref{eq:outer_sol}).
+Next, the outer optimization process is performed by aggregating the losses on the query set of the tasks and optimizing on it (equation \ref{eq:outer_sol}).
 
@@
 
-The presence of the product of second order derivatives in the equation \ref{eq:outer_sol_} makes the derivation process complicated because it requires a lot of overhead to maintain the Hessian matrices.
+The presence of the product of second order derivatives in the equation \ref{eq:outer_sol_} makes the derivation process complicated because it requires a lot of overhead to maintain Hessian matrices.
 Therefore, the number of computational steps to find $\theta^*$ needs to be limited.
@@
 
-Our goal is to use \verb|multi-fx| dataset to extract and aggregate information about outliers (i.e.
+Our goal is to use \verb|multi-fx| dataset to check the information extraction and aggregation of external factors (i.e.
 market, economic, political, etc.) that are believed to influence the outcome of a given financial index \citep{overreactioncontrarian,fama1970efficient,mech1993portfolio}.
@@
 We compare the proposed method with \verb|NHITS| using the above metrics and dataset.
-For each dataset, each attribute will be predicted for future trends based on past data.
+For each dataset, each attribute will be predicted the future trends based on past data.
 The final result is calculated by averaging the metrics across the attributes.
 We also test different feature extractors (\verb|LSTM|, \verb|CNN|) for the proposed method to demonstrate the deep data understanding capabilities of \verb|LSTM+CNN|.
-The overall data is structured into training sets, validation sets, and testing sets with a ratio of 6:2:2 for training, fine-tuning and testing model, respectively.
-
-\vspace{2mm}
-
-For \verb|NHITS| model, we split the data as usual according to the above predetermined ratio.
+
+\vspace{2mm}
+
+For \verb|NHITS| model, data is structured into training sets, validation sets, and testing sets with a ratio of 6:2:2 for training, fine-tuning and testing model, respectively.
 For the proposed algorithm, because the meta-training and meta-testing processes require dividing the data into small tasks, and allowing the model to adapt on the support set of each task, we split the data into 40-60 tasks (see table \ref{tab:stat_} for statistical detail).
@@
 We use 50\% of tasks for meta-training, 25\% for meta-validation, and 25\% for meta-testing.
-With this division, we ensure the fairness that the ML model is trained with the same amount of data as \verb|NHITS|.
-After the data configuration, we perform experiment as in appendix \ref{app:our_experiment} and \ref{app:nhits_experiment}.
+With this division, we ensure the fairness that the ML model is trained with the same amount of data as \verb|NHITS| (see appendix \ref{app:data}).
+The detail of experiments can be found at appendix \ref{app:our_experiment} and \ref{app:nhits_experiment}.
 
@@
 
-Table \ref{tab:rs} compares the results between the proposed model and the \verb|NHITS| model on the above datasets.
+Table \ref{tab:rs} compares the results between the proposed model and the \verb|NHITS| model using the above datasets.
 On aperiodic data, our method achieves a much higher convergence rate than the \verb|NHITS| model on all metrics.
@@
 The convergence rate is generally higher than when using each model separately.
-This can be explained by the fact that \verb|LSTM| successfully integrated long-term features into the model, as well as taking advantage of the local features of \verb|CNN|.
+This can be explained by the fact that \verb|LSTM| \textbf{successfully integrated long-term features into the model, as well as taking advantage of the local features highlighting ability of} \verb|CNN|.
 
@@
 However, when observing the results on the validation set, all three models above give very poor results.
-To explain this, we think that although they can learn well on the training set, the models lose their generalization ability because this is a complex type of data, different from data such as images, videos, or periodic data.
-If the model's adaptability is not high, poor results on the validation set are predictable.
-However, \textbf{the graph in figure \ref{fig:lstm_cnn} shows that we cannot deny the feature extraction ability of} \verb|LSTM+CNN|.
+To explain this, we think that although they can learn well on the training set, the models lose their generalization ability because this is a complex type of data, which is different from data such as images, videos, or periodic data.
+If the model's adaptability is not good, poor results on the validation set are predictable.
+This suggests that there is a need for a method that efficiently utilizes and synthesizes the obtained features, which is discuss in detail in the next subsection.
 
@@
 However, it is very difficult to resolve the frequency on non-periodic data.
-Moreover, it is not difficult to see that the feature extraction techniques of \verb|NHITS| are extremely simple (only including \verb|FullyConnected| and \verb|Pooling| classes).
+Moreover, it is not difficult to see that the feature extraction techniques of \verb|NHITS| are extremely simple (only including \verb|FullyConnected| and \verb|Pooling| layers).
 These are the reasons why \verb|NHITS| encountered great difficulties on \verb|multi-fx|, \verb|USD/JPY| dataset and non-periodic data in general.
 On the other hand, our proposed method uses deep learning networks for feature extraction, so it can easily obtain deep hidden features.
-Therefore, on both periodic and aperiodic data, our method achieves results approximately equal to or better than \verb|NHITS|.
+Therefore, on both periodic and aperiodic data, our method achieves results close to or better than \verb|NHITS|.
 
@@
 For the ML model using only \verb|CNN|, the accuracy and error values do not even improve during the entire training process.
-We explain this by saying that the features of \verb|CNN| focus too much on the locality of the data and ignore the long-term dependencies, which is an important feature of time-series data.
+To explain this, we argue that the features of \verb|CNN| have focused too much on the locality of the data and ignored the long-term dependencies, which is an important feature of time-series data.
 In contrast, the features of \verb|LSTM+CNN| not only capture the long-term dependencies but also highlight the local properties in the data, making it completely outperform \verb|CNN| as well as \verb|NHITS|.
-Therefore, \textbf{we conclude that ML is more flexible and efficient in synthesizing sub-models than traditional rigid ensemble models}.
+Therefore, \textbf{we conclude that ML is more flexible and efficient in synthesizing sub-models as well as external factors than traditional rigid ensemble models}.
 
@@
 This proves the hypothesis of the existence of hidden long-term dependencies that we stated in section \ref{sec.intro}.
+
+\vspace{2mm}
 
@@
 \appendix
+\section{Fairness in data}
+\label{app:data}
+
+% Với \verb|NHITS|, chúng tôi sử dụng các chia dữ liệu truyền thống. Tập dữ liệu ban đầu được chia thành training set, validation set, và testing set theo tỷ lệ 6:2:2.
+
+For \verb|NHITS|, we use the traditional data split. The dataset is split into a training set, validation set, and testing set in a 6:2:2 ratio.
+
+\vspace{2mm}
+
+% Đối với phương pháp đề xuất, chúng tôi chia dữ liệu thành các tập dữ liệu nhỏ, mỗi tập dữ liệu lại được chia thành tập support và tập query.
+Gọi số lượng task sau khi chia là $n$, chúng tôi sử dụng $0.5n$ tasks cho meta-training, $0.25n$ tasks cho meta-validating, và $0.25n$ tasks cho meta-testing.
+Trong mỗi task, tập support chiếm 20\%, tập query chiếm 80\%.
+
+For the proposed method, we split the data into small datasets, each of which is divided into a support set and a query set.
+Let the number of tasks after splitting be $n$, we use $0.5n$ tasks for meta-training, $0.25n$ tasks for meta-validating, and $0.25n$ tasks for meta-testing.
+In each task, the support set accounts for 20\%, and the query set accounts for 80\%.
+
+\vspace{2mm}
+
+% Gọi $M$ là số mẫu dữ liệu trong tập dữ liệu ban đầu, số mẫu dữ liệu sử dụng để train, validate, và test cho \verb|NHITS| lần lượt là $0.6M, 0.2M, 0.2M$.
+Đối chiếu với lượng dữ liệu sử dụng cho phương pháp đề xuất, mỗi task sẽ bao gồm $\frac{M}{n}$ mẫu dữ liệu.
+Chúng tôi sử dụng $0.5M$ cho meta-training (vì số task cho meta-training là $0.5n$).
+Tiếp đến, trong quá trình validate hoặc test, chúng tôi sử dụng 20\% dữ liệu mỗi task cho quá trình model adaptation và 80\% dữ liệu còn lại cho validate hoặc test.
+Theo đó, tổng số mẫu dùng cho quá trình huấn luyện và adaptation là $0.5M + 0.25n \times 0.2\frac{M}{n} + 0.25n \times 0.2\frac{M}{n} = 0.6M$.
+Số mẫu dùng để đánh giá mô hình trong quá trình validation và testing đều là $0.25n \times 0.8\frac{M}{n} = 0.2M$.
+Như vậy, chúng tôi đã bảo đảm sự công bằng trong huấn luyện và kiểm thử cho các thí nghiệm.
+
+Let $M$ be the number of data samples in the original dataset, the number of data samples used for training, validation, and testing for \verb|NHITS| are $0.6M, 0.2M, 0.2M$ respectively.
+Compared to the amount of data used for the proposed method, each task will include $\frac{M}{n}$ data samples.
+We use $0.5M$ samples for meta-training (because the number of tasks for meta-training is $0.5n$).
+Next, during the validation or testing process, we use 20\% of the data of each task for the model adaptation process and the remaining 80\% of the data for validation or testing.
+Accordingly, the total number of samples used for training and adaptation is $0.5M + 0.25n \times 0.2\frac{M}{n} + 0.25n \times 0.2\frac{M}{n} = 0.6M$.
+The number of samples used to evaluate the model in the validation and testing process is both $0.25n \times 0.8\frac{M}{n} = 0.2M$.
+Thus, we have ensured fairness in training and testing for the experiments.
+
 \section{Experimental detail\\for proposed method}
@@
 
-\begin{table}[H]
+\begin{table}[h]
     \centering
@@
 
-\begin{table}
+% Đối với mô hình \verb|NHITS|, chúng tôi dựa trên \cite{challu2023nhits} để định nghĩa không gian tìm kiếm cho việc fine-tune tham số, cũng như kiến trúc mô hình.
+Đối với các tham số không được đề cập trong bảng, chúng tôi sử dụng giá trị mặc định của cài đặt \verb|NHITS| trong thư viện \verb|NeuralForecast| \cite{neuralforecast}.
+Chi tiết quá trình fine-tune được trình bày trong phần phụ lục \ref{app:experiment}.
+Kết quả tốt nhất của các lần fine-tune được chọn ra và báo cáo trong nghiên cứu này.
+
+We rely on \cite{challu2023nhits} to define the search space (table \ref{tab:nhits_finetune}) for parameter fine-tuning, as well as to fine-tune the model architecture.
+For parameters not mentioned in the table, we use the default values of the \verb|NHITS| implementation in the \verb|NeuralForecast| library \cite{neuralforecast}.
+The best results of fine-tuning process are selected and reported in this study.
+
+\begin{table}[h]
     \centering
@@
 \end{table}
-
-% Đối với mô hình \verb|NHITS|, chúng tôi dựa trên \cite{challu2023nhits} để định nghĩa không gian tìm kiếm cho việc fine-tune tham số, cũng như kiến trúc mô hình.
-Đối với các tham số không được đề cập trong bảng, chúng tôi sử dụng giá trị mặc định của cài đặt \verb|NHITS| trong thư viện \verb|NeuralForecast| \cite{neuralforecast}.
-Chi tiết quá trình fine-tune được trình bày trong phần phụ lục \ref{app:experiment}.
-Kết quả tốt nhất của các lần fine-tune được chọn ra và báo cáo trong nghiên cứu này.
-
-We rely on \cite{challu2023nhits} to define the search space (table \ref{tab:nhits_finetune}) for parameter fine-tuning, as well as to fine-tune the model architecture.
-For parameters not mentioned in the table, we use the default values of the \verb|NHITS| implementation in the \verb|NeuralForecast| library \cite{neuralforecast}.
-The best results of fine-tuning process are selected and reported in this study.
 

### 2024-12-08 doc/paper/paper.tex
paper vn

--- 2024-10-04 doc/paper/paper.tex
@@
 \usepackage{multirow}
+\usepackage{tabularray}
+% \usepackage{color}
 
@@
 
+    % \textbf{Abstract}.
+Predicting aperiodic time-series data (e.g.
+stock price, foreign exchange, Bitcoin price, etc.) is a difficult task for machine learning models because this type of data has high and not stationary variance; does not show clearly periodicity, making it difficult to extract features; depends not only on past values but also on external factors such as economic and political situations, which make them unstable and non-cyclical.
+To overcome the above challenges, we use Meta-learning to train a combined \verb|LSTM| and \verb|CNN| network, thereby effectively extracting and synthesizing deep hidden features of the data over time.
+Experiments on aperiodic data (foreign exchange data of 60 currency pairs over 24 years from 2000 to 2024) show that the proposed method performs well and has 5\%-11\% higher accuracy compared to the \verb|NHITS| - the state-of-the-art model in 2023 on time-series data, in the task of predicting the trend (upward or downward) of the next trading day.
+Meanwhile, the results obtained on periodic data (Electricity Transformer Temperature, Weather) are approximately \verb|NHITS|.
+
     \textbf{Abstract}.
-Predicting aperiodic time-series data (e.g.
-stock price, foreign exchange, Bitcoin price,...) is a difficult task for machine learning models because this type of data has high and not stationary variance; does not show clearly periodicity, making it difficult to extract features; depends not only on past values but also on external factors, which make them unstable and non-cyclical, such as economic and political situations.
-To overcome the above challenges, we use Meta-learning to train a combined \verb|LSTM| and \verb|CNN| network, thereby effectively extracting and synthesizing hidden features of the data over time.
-Experiments on foreign exchange data of 60 currency pairs over 24 years (2000-2024) show that the proposed method performs well and has 5\%-11\% higher accuracy compared to the \verb|NHITS| - the state-of-the-art model in 2023 on time-series data, in the task of predicting the trend (upward or downward) of the next trading day.
-
-    % Dự đoán aperiodic time-series data (e.g.
-stock price, foreign exchange, Bitcoin price,...) là một tác vụ khó khăn đối với các mô hình học máy bởi vì loại dữ liệu này có phương sai cao và không cố định; không thể hiện chu kỳ rõ ràng nên khó rút trích đặc trưng; không chỉ phụ thuộc vào giá trị quá khứ mà còn phụ thuộc các yếu tố ngoại lai, điều khiến chúng bất ổn và không có chu kỳ như tình hình kinh tế, chính trị.
+Dự đoán dữ liệu chuỗi thời gian phi chu kỳ (ví dụ: giá cổ phiếu, tỷ giá ngoại hối, giá Bitcoin,...) là một tác vụ khó khăn đối với các mô hình học máy bởi vì loại dữ liệu này có phương sai cao và không cố định; không thể hiện chu kỳ rõ ràng nên khó rút trích đặc trưng; không chỉ phụ thuộc vào giá trị quá khứ mà còn phụ thuộc các tác động bên ngoài như tình hình kinh tế, chính trị, khiến chúng bất ổn và không có chu kỳ.
 Để khắc phục các thách thức nêu trên, chúng tôi sử dụng Meta-learning để huấn luyện kết hợp mạng \verb|LSTM| và \verb|CNN|, từ đó rút trích và tổng hợp hiệu quả các đặc trưng ẩn của dữ liệu theo thời gian.
-Thực nghiệm trên dữ liệu foreign exchange của 15 cặp tiền tệ trong vòng 24 năm (2000-2024) cho thấy phương pháp đề xuất hoạt động tốt và có độ chính xác cao hơn mô hình \verb|NHITS| - state-of-the-art model của năm 2023 trên dữ liệu time-series, trong bài toán dự đoán xu hướng (upward or downward) của ngày giao dịch tiếp theo.
+Các thử nghiệm trên dữ liệu phi chu kỳ (dữ liệu ngoại hối của 60 cặp tiền tệ trong 24 năm từ 2000 đến 2024) cho thấy phương pháp đề xuất hoạt động tốt và có độ chính xác cao hơn 5\%-11\% so với \verb|NHITS| - mô hình SOTA 2023 trên dữ liệu chuỗi thời gian, trong nhiệm vụ dự đoán xu hướng (tăng hoặc giảm) của ngày giao dịch tiếp theo.
+Trong khi đó, kết quả thu được trên dữ liệu chu kỳ (Nhiệt độ máy biến áp điện, Thời tiết) đạt xấp xỉ \verb|NHITS|.
 \end{abstract}
@@
 
-% Aperiodic time-series prediction nói chung hay foreign exchange (FX) prediction nói riêng từ lâu đã là vấn đề đáng quan tâm của nhiều nghiên cứu \citep{li2019multi, islam2021foreign, heryadi2021foreign}.
+Dự đoán dữ liệu chuỗi thời gian phi chu kỳ nói chung hay dự đoán tỷ giá ngoại hối (foreign exchange, FX) nói riêng từ lâu đã là vấn đề đáng quan tâm của nhiều nghiên cứu \citep{li2019multi, islam2021foreign, heryadi2021foreign}.
 Hai kỹ thuật chính được sử dụng trong aperiodic time-series prediction là fundamental analysis and technical analysis \cite{ayitey2023forex}.
@@
 
-Aperiodic time-series prediction in general or foreign exchange (FX) prediction in particular has been a matter of concern for many researchers \citep{li2019multi, islam2021foreign, heryadi2021foreign} for a long time.
+% Aperiodic time-series prediction in general or foreign exchange (FX) prediction in particular has been a matter of concern for many researchers \citep{li2019multi, islam2021foreign, heryadi2021foreign} for a long time.
 The two main techniques used in aperiodic time-series prediction are fundamental analysis and technical analysis \cite{ayitey2023forex}.
@@
 
-% Việc dự đoán trên aperiodic time-series data gặp một vài thách thức cố hữu.
+Theo \cite{li2019multi}, việc dự đoán trên aperiodic time-series data gặp một vài thách thức cố hữu.
 Trong đó có thể kể tới: (1) - Variance của loại dữ liệu này biến thiên mạnh qua các thời kỳ.
@@
 giá cổ phiếu của một công ty) không hoàn toàn phụ thuộc vào dữ liệu quá khứ mà còn phụ thuộc nhiều yếu tố ngoại lai (e.g.
-tin tức, tình hình kinh tế, chính trị) \cite{li2019multi}.
-
-Aperiodic time-series data prediction faces several inherent challenges: (1) - The variance varies significantly over time, making it difficult to assume a specific distribution to approximate errors, which is a challenge for machine learning models in accurately predicting future values and trends.
+tin tức, tình hình kinh tế, chính trị).
+
+% Aperiodic time-series data prediction faces several inherent challenges: (1) - The variance varies significantly over time, making it difficult to assume a specific distribution to approximate errors, which is a challenge for machine learning models in accurately predicting future values and trends.
 (2) - Aperiodic data does not follow explicit patterns, making it challenging to identify the latent features required for reliable predictions.
@@
 
-% Đối với thách thức đầu tiên, các mô hình ensemble [cite something here] thường được sử dụng để hạn chế ảnh hưởng của sự biến đổi variance.
+Đối với thách thức đầu tiên, các mô hình ensemble \citep{sadeghi2021combined, zafeiriou2020intraday, ali2020complete} thường được sử dụng để hạn chế ảnh hưởng của sự biến đổi variance.
 Ensemble model giúp cung cấp một cái nhìn tổng quát, tổng hợp từ nhiều khía cạnh dựa trên các sub-models, từ đó giúp mô hình tổng quát thích ứng được với sự thay đổi mạnh của variance.
-% Tuy nhiên, dưới sự tổng hợp cứng nhắc của các mô hình ensemble truyền thống, kết quả thu được trên các bộ dữ liệu phi chi kỳ là rất thấp.
-
-For the variance variation challenge, ensemble models \citep{sadeghi2021combined, zafeiriou2020intraday, ali2020complete} are often used to mitigate the effects of variance variation.
+Tuy nhiên, dưới sự tổng hợp cứng nhắc của các mô hình ensemble truyền thống, kết quả thu được trên các bộ dữ liệu phi chu kỳ là rất thấp.
+
+% Ensemble models \citep{sadeghi2021combined, zafeiriou2020intraday, ali2020complete} are often used to mitigate the effects of variance variation challenge.
 Ensemble models provide a holistic and multi-perspective view based on sub-models, thereby helping the overall model adapt to strong variance changes.
@@
 
-% Để khắc phục thách thức thứ hai, các nghiên cứu phần lớn sử dụng các đặc trưng rút trích được từ Long short-term memory neural network (\verb|LSTM|) \cite{hochreiter1997long}, Artificial neural network (\verb|ANN|), và Convolution neural network (\verb|CNN|) \cite{lecun1989handwritten}.
+Để khắc phục thách thức thứ hai, các nghiên cứu phần lớn sử dụng các đặc trưng rút trích được từ Long short-term memory neural network (\verb|LSTM|) \cite{hochreiter1997long}, Artificial neural network (\verb|ANN|), và Convolution neural network (\verb|CNN|) \cite{lecun1989handwritten}.
 Cụ thể, trong năm 2022, 20\% tổng số các bài báo liên quan đến dự đoán chỉ số tài chính sử dụng \verb|LSTM|, 20\% sử dụng \verb|ANN| và 6\% sử dụng \verb|CNN| \cite{ayitey2023forex}.
-% Điều này là hoàn toàn dễ hiểu vì \verb|LSTM| và \verb|CNN| đã chứng minh được khả năng của mình trong việc rút trích các đặc trưng không gian, thời gian một cách hiệu quả.
-% Mặc dù vậy, khi đối mặt với tính phi chu kỳ, sau khoảng 100 training steps, \verb|LSTM| vẫn bị underfitting, còn \verb|CNN| bị overfitting chỉ sau 5 training steps
-
-To overcome the pattern variation challenge, most studies use features extracted from Long Short-Term Memory neural network (\verb|LSTM|) \cite{hochreiter1997long}, Artificial Neural Network (\verb|ANN|), and Convolution Neural Network (\verb|CNN|) \cite{lecun1989handwritten}.
+Điều này là hoàn toàn dễ hiểu vì \verb|LSTM| và \verb|CNN| đã chứng minh được khả năng của mình trong việc rút trích các đặc trưng không gian, thời gian một cách hiệu quả.
+Mặc dù vậy, khi đối mặt với tính phi chu kỳ trên 2600 mẫu dữ liệu FX, sau khoảng 100 training steps, \verb|LSTM| vẫn bị underfitting, còn \verb|CNN| bị overfitting chỉ sau 5 training steps
+
+% To overcome the pattern variation challenge, most studies use features extracted from Long Short-Term Memory neural network (\verb|LSTM|) \cite{hochreiter1997long}, Artificial Neural Network (\verb|ANN|), and Convolution Neural Network (\verb|CNN|) \cite{lecun1989handwritten}.
 Specifically, in 2022, 20\% of all publications related to financial index prediction used \verb|LSTM|, 20\% used \verb|ANN|, and 6\% used \verb|CNN| \cite{ayitey2023forex}.
 This is completely understandable because \verb|LSTM| and \verb|CNN| have proven their ability to extract spatial and temporal features effectively.
-However, when faced with aperiodic data with only 2600 samples, after 40 training steps, \verb|LSTM| is still underfitting, while \verb|CNN| is overfitting after only 5 training steps.
-
-\vspace{2mm}
-
-% Đối với thách thức thứ ba, nghiên cứu \cite{fama1970efficient} đưa ra giả thuyết rằng các time-series datasets khác nhau của cùng một lĩnh vực tại cùng thời điểm phản ánh sự tác động của các yếu tố ngoại lai.
-Ví dụ, các nghiên cứu \citep{overreactioncontrarian, mech1993portfolio} cũng chỉ ra sự phụ thuộc giữa chỉ số tài chính của một công ty nhất định và các chỉ số của các công ty khác.
+However, when faced with aperiodic data with only 2600 samples, after 40 training steps, our experiment shows that \verb|LSTM| is still underfitting, while \verb|CNN| is overfitting after only 5 training steps.
+
+\vspace{2mm}
+
+Đối với thách thức thứ ba, nghiên cứu \cite{fama1970efficient} đưa ra giả thuyết rằng các time-series datasets khác nhau của cùng một lĩnh vực tại cùng thời điểm phản ánh sự tác động của các yếu tố ngoại lai.
+Các nghiên cứu \citep{overreactioncontrarian, mech1993portfolio} cũng chỉ ra sự phụ thuộc giữa chỉ số tài chính của một công ty nhất định và các chỉ số của các công ty khác.
 Điều này càng làm tăng tính đúng đắn của giả thuyết trong nghiên cứu \cite{fama1970efficient}.
-% Do đó, bằng việc khai thác các nguồn dữ liệu có cùng domain, chúng ta có thể capture được các thông tin phản ánh các external factors.
-
-Regarding the external influences challenge, \cite{fama1970efficient} hypothesizes that different time-series datasets of the same domain at the same time reflect the impact of external factors.
-For example, studies \citep{overreactioncontrarian, mech1993portfolio} show the dependence between the financial ratios of a given company and the ratios of other companies.
+Do đó, bằng việc khai thác các nguồn dữ liệu có cùng domain, chúng ta có thể capture được các thông tin phản ánh tác động bên ngoài.
+
+% Regarding the challenge of external influences, \cite{fama1970efficient} hypothesizes that different time-series datasets of the same domain at the same time reflect the impact of external factors.
+Studies \citep{overreactioncontrarian, mech1993portfolio} also show the dependence between the financial ratios of a given company and the ones of other companies.
 This further strengthens the hypothesis in study \cite{fama1970efficient}.
@@
 
-% Chúng tôi tiếp cận các thách thức nêu trên thông qua việc kết hợp đặc trưng và tổng hợp mô hình.
-Cụ thể, bằng cách sử dụng Meta-learning (ML) \cite{finn2017model}, chúng tôi tổng hợp hiệu quả tham số của các mô hình cục bộ, giúp giảm thiểu đáng kể variance loss cũng như tăng khả năng thu thập các thông tin phản ánh external factors từ các tập dữ liệu có cùng domain.
+Chúng tôi tiếp cận các thách thức nêu trên thông qua việc kết hợp đặc trưng và tổng hợp mô hình.
+Cụ thể, bằng cách sử dụng Meta-learning (ML) \cite{finn2017model}, chúng tôi tổng hợp hiệu quả tham số của các mô hình cục bộ, giúp giảm thiểu đáng kể variance loss cũng như tăng khả năng thu thập các thông tin phản ánh tác động bên ngoài từ các tập dữ liệu có cùng domain.
 Để học được các laten patterns, chúng tôi đề xuất phương pháp kết hợp các đặc trưng từ \verb|CNN| và \verb|LSTM|.
@@
 
-We approach the above challenges through combination of feature combination and model synthesis.
+% We approach the above challenges through the combination of feature and model synthesis.
 Specifically, by using Meta-learning (ML) \cite{finn2017model}, we efficiently synthesize the parameters of local models, which significantly reduces variance loss and increases the ability to collect information reflecting external factors from datasets in the same domain.
@@
 
-% Cuối cùng, chúng tôi chứng minh tính ưu việt của thuật toán đề xuất bằng cách giải bài toán dự đoán xu hướng (tăng hoặc giảm) của tỉ giá ngoại hối và so sánh kết quả với mô hình SOTA hiện tại (\verb|NHITS| \cite{challu2023nhits}) trên hai loại dữ liệu: (1) - Dữ liệu tỉ giả của cặp tiền tệ USD/JPY; (2) - Dữ liệu tỷ giá của 60 cặp tiền tệ, cấu thành từ 18 quốc gia.
-Các tập dữ liệu này được công khai trên Internet và có thể tải về dễ dàng.
-Bản cài đặt chính thức có thể xem tại [bỏ cái link github vào đây].
-
-Finally, we demonstrate the superiority of the proposed algorithm by solving the problem of predicting the trend (upward or downward) of foreign exchange rates and comparing the results with the existing state-of-the-art (SOTA) model (\verb|NHITS| \cite{challu2023nhits}) on two types of data: (1) - USD/JPY exchange rate data; (2) - Exchange rate data of 60 currency pairs, comprising 18 countries.
-These data sets are publicly available on the Internet and can be easily downloaded.
+Cuối cùng, chúng tôi chứng minh tính ưu việt của thuật toán đề xuất bằng cách giải quyết bài toán dự đoán xu hướng (tăng hoặc giảm) của dữ liệu chuỗi thời gian và so sánh kết quả với mô hình SOTA năm 2023 (\verb|NHITS| \cite{challu2023nhits}) trên hai loại dữ liệu: (1) - Dữ liệu phi chu kỳ, bao gồm dữ liệu tỷ giá hối đoái USD/JPY và dữ liệu tỷ giá hối đoái của 60 cặp tiền tệ, bao gồm 18 quốc gia từ năm 2000 đến năm 2024; (2) - Dữ liệu có chu kỳ bao gồm tập dữ liệu Electricity Transformer Temperature \cite{zhou2021informer} và Weather \cite{Kolle}.
+Các tập dữ liệu này được cung cấp công khai trên Internet và có thể tải xuống dễ dàng.
+Bản cài đặt chính thức có thể xem tại \url{https://github.com/baolongnguyenmac/multi_fx}.
+
+% Finally, we demonstrate the superiority of the proposed algorithm by solving the problem of predicting the trend (upward or downward) of time-series data and comparing the results with the existing state-of-the-art (SOTA) model (\verb|NHITS| \cite{challu2023nhits}) on two types of data: (1) - Aperiodic data, which contains USD/JPY exchange rate data and the exchange rate data of 60 currency pairs, comprising 18 countries from 2000 to 2024; (2) - Periodic data which contains Electricity Transformer Temperature \cite{zhou2021informer} and Weather \cite{Kolle} dataset.
+These datasets are publicly available on the Internet and can be downloaded easily.
 The official implementation can be found at \url{https://github.com/baolongnguyenmac/multi_fx}.
@@
 
-In summary, our main contributions include the following: First, we present a novel \textbf{feature combination technique} for aperiodic time-series data that extracts and combines features using both \verb|LSTM| and \verb|CNN| models.
-Second, we uncover hidden long-term dependencies, demonstrating experimentally that aperiodic time-series data not only rely on external factors but also exhibit dependencies with past period, which we call \textbf{hidden long-term dependencies}.
+Tóm lại, đóng góp chính của chúng tôi như sau:
+
+\begin{itemize}
+    \item Kết hợp đặc trưng: Rút trích đặc trưng bằng cách kết hợp các đặc trưng của \verb|LSTM| và \verb|CNN|.
+    \item Hidden long-term dependency: Chứng minh thực nghiệm rằng dữ liệu time-series tồn tại các phụ thuộc ẩn với chính nó tại nhiều thời điểm khác nhau trong quá khứ.
+    \item Tổng hợp hiệu quả tham số mô hình: Sử dụng ML thay cho các mô hình ensemble truyền thống trong việc tổng hợp kết quả từ các mô hình học máy.
+    \item Experiment: Thực nghiệm trên dữ liệu phi chu kỳ (tỷ giá hối đoái của 60 cặp tiền tệ) và dữ liệu có chu kỳ (Electricity Transformer Temperature, Weather) rồi so sánh với mô hình SOTA 2023 \verb|NHITS| để chứng minh tính hiệu quả của phương pháp đề xuất.
+\end{itemize}
+
+% In summary, our main contributions include the following: First, we present a novel \textbf{feature combination technique} for aperiodic time-series data that extracts and combines features using both \verb|LSTM| and \verb|CNN| models.
+Second, we discover the \textbf{hidden long-term dependencies} and demonstrate experimentally that aperiodic time-series data not only rely on external factors but also exhibit dependencies with past period.
 Third, we introduce an \textbf{efficient model parameter aggregation approach} by employing ML algorithm in place of traditional ensemble models for combining parameters from multiple machine learning models.
-Finally, we validate our method through experiments on exchange rate datasets and show its effectiveness in comparison to the SOTA model, \verb|NHITS|.
+Finally, we validate our method through experiments on exchange rate datasets and show its effectiveness in comparison to a SOTA model, \verb|NHITS|.
 
@@
 
-% Như đã đề cập, \verb|LSTM| là mạng neural rất phổ biến trong việc handle các bài toán liên quan đến dự đoán trên dữ liệu time-series.
+Như đã đề cập, \verb|LSTM| là mạng neural rất phổ biến trong việc handle các bài toán liên quan đến dự đoán trên dữ liệu time-series.
 \verb|LSTM| được sử dụng phổ biến như vậy bởi vì nó xử lý tốt vấn đề vanishing gradient (dễ dàng bắt gặp khi sử dụng Recurrent neural network) và có thể khai thác hiệu quả các mối quan hệ phi tuyến trong dữ liệu.
@@
 
-As aforementioned, \verb|LSTM| is a well-known neural network for handling prediction problems on time-series data.
-\verb|LSTM| is commonly used because it handles the problem of vanishing gradients well (easily encountered when using Recurrent neural networks) and can effectively exploit nonlinear relationships in data.
+% As aforementioned, \verb|LSTM| is a well-known neural network for handling prediction problems on time-series data.
+\verb|LSTM| is commonly used because it efficiently handles the problem of vanishing gradients (easily encountered when using Recurrent neural networks) and can effectively exploit nonlinear relationships in data.
 Indeed, by maintaining the cell-state in each iteration, \verb|LSTM| can overcome the vanishing gradient problem, thereby preserving the ability to capture long-term dependencies \cite{cheng2018applied}.
@@
 
-% Mạng \verb|CNN| được sử dụng rất nhiều trong các tác vụ xử lý hình ảnh \citep{naranjo2020review, sharma2018analysis} bởi khả năng tổng hợp các quan hệ cục bộ.
+Mạng \verb|CNN| được sử dụng rất nhiều trong các tác vụ xử lý hình ảnh \citep{naranjo2020review, sharma2018analysis} bởi khả năng tổng hợp các quan hệ cục bộ.
 Không chỉ vậy, \verb|CNN| còn được dùng rất nhiều trong các tác vụ xử lý dữ liệu time-series như speech recognition \cite{dua2022developing}, natural language processing \cite{varshitha2023natural}.
@@
 
-\verb|CNN| is widely used in image processing tasks \citep{naranjo2020review, sharma2018analysis} because of its ability to synthesize local relationships.
+% \verb|CNN| is widely used in image processing tasks \citep{naranjo2020review, sharma2018analysis} because of its ability to synthesize local relationships.
 Not only that, \verb|CNN| is also widely used in time-series data processing tasks such as speech recognition \cite{dua2022developing}, natural language processing \cite{varshitha2023natural}.
@@
 
-% Các thuật toán Meta-learning (ML), điển hình là MAML \cite{finn2017model} được biết đến với khả năng huấn luyện một mô hình có tính tổng quát cao, thích ứng nhanh trên tập dữ liệu mới thông qua một lượng nhỏ dữ liệu và số bước huấn luyện \citep{hospedales2021meta, vettoruzzo2024advances}.
+Các thuật toán Meta-learning (ML) dựa trên tối ưu, điển hình là MAML \cite{finn2017model} được biết đến với khả năng huấn luyện một mô hình có tính tổng quát cao, thích ứng nhanh trên tập dữ liệu mới thông qua một lượng nhỏ dữ liệu và số bước huấn luyện \citep{hospedales2021meta, vettoruzzo2024advances}.
 Với khả năng này, ML được sử dụng rất nhiều trong các tác vụ đòi hỏi khả năng đáp ứng của mô hình trên dữ liệu (e.g.
@@
 
-Optimization-based Meta-learning (ML) algorithms, typically MAML \cite{finn2017model} are known for their ability to train a highly general, adaptive model on new datasets with a limited amount of data and a small number of training steps \citep{hospedales2021meta, vettoruzzo2024advances}.
+% Optimization-based Meta-learning (ML) algorithms, typically MAML \cite{finn2017model} are known for their ability to train a highly general, adaptive model on new datasets with a limited amount of data and a small number of training steps \citep{hospedales2021meta, vettoruzzo2024advances}.
 With this ability, ML is widely used in tasks that require the model's ability to adapt to the data (e.g.
@@
 
-% Một thuật toán ML cơ bản sẽ được học trên nhiều tác vụ $t$ rút ra từ cùng một phân phối tác vụ $\mathcal{T}$ \cite{hospedales2021meta}.
+Một thuật toán ML cơ bản sẽ được học trên nhiều tác vụ $t$ rút ra từ cùng một phân phối tác vụ $\mathcal{T}$ \cite{hospedales2021meta}.
 Dữ liệu của mỗi tác vụ được chia thành tập support $\mathcal{D}_t^{support}$ (thường có kích thước nhỏ, khoảng 20\%) và tập query $\mathcal{D}_t^{query}$.
@@
 
-A basic ML algorithm is trained on multiple tasks $t$ drawn from the same task distribution $\mathcal{T}$ \cite{hospedales2021meta}.
+% A basic ML algorithm is trained on multiple tasks $t$ drawn from the same task distribution $\mathcal{T}$ \cite{hospedales2021meta}.
 The data for task $t$ is divided into a support set $\mathcal{D}_t^{support}$ (usually small, around 20\%) and a query set $\mathcal{D}_t^{query}$.
@@
     \label{eq:inner_opt}
-\end{equation} Where, $\phi$ is the result of the outer optimization process, which acts as the initial value of $\theta_t$.
+\end{equation} Trong đó, $\phi$ là kết quả của quá trình outer optimization, được sử dụng làm tham số khởi tạo cho $\theta_t$.
+$\mathcal{L}^{task}_t$ là hàm lỗi của mô hình trên tập support của task $t$.
+
+% Where, $\phi$ is the result of the outer optimization process, which acts as the initial value of $\theta_t$.
 $\mathcal{L}^{task}_t$ is the error function of the model on the support set of task $t$.
@@
 
-% Sau đó, thuật toán sử dụng các bộ tham số tối ưu $\theta_t^*$ để perform trên tập query tương ứng.
+Sau đó, thuật toán sử dụng các bộ tham số tối ưu $\theta_t^*$ để perform trên tập query tương ứng.
 Lỗi của toàn bộ mô hình sau đó được tổng hợp để thực hiện quá trình outer optimization như phương trình \ref{eq:outer_opt}.
 
-The algorithm then uses the optimal parameter sets $\theta_t^*$ to perform on the corresponding query set.
+% The algorithm then uses the optimal parameter sets $\theta_t^*$ to perform on the corresponding query set.
 The losses of the entire models are then aggregated to perform the outer optimization process as equation \ref{eq:outer_opt}.
@@
 
-% Bằng hình thức huấn luyện trên, mô hình $\phi^*$ sẽ có mức tổng quát hóa cao trên các tác vụ khác nhau, có thể nhanh chóng đáp ứng một tác vụ mới chỉ sau một vài bước huấn luyện.
-
-By performing the above training method, the $\phi^*$ model will have a high level of generalization across different tasks, and can quickly respond to a new task after only a few training steps.
-
-\vspace{2mm}
-
-% Trong inference phase, giá trị khởi tạo cho tham số của mô hình được gán bằng $\phi^*$.
+Bằng hình thức huấn luyện trên, mô hình $\phi^*$ sẽ có mức tổng quát hóa cao trên các tác vụ khác nhau, có thể nhanh chóng đáp ứng một tác vụ mới chỉ sau một vài bước huấn luyện.
+
+% By performing the above training method, the $\phi^*$ model will have a high level of generalization across different tasks, and can quickly respond to a new task after only a few training steps.
+
+\vspace{2mm}
+
+Trong inference phase, giá trị khởi tạo cho tham số của mô hình được gán bằng $\phi^*$.
 Mô hình sau đó được huấn luyện nhanh trên tập support sau đó perform trên tập query.
@@
 
-In the inference phase, the initial values for the model parameters are assigned $\phi^*$.
+% In the inference phase, the initial values for the model parameters are assigned $\phi^*$.
 The model is then adapted quickly to the support set and performed on the query set.
@@
 
-% Các mô hình hybrid ensemble vốn được sử dụng rất nhiều trong các bài toán xử lý time-series và được chứng minh thực nghiệm là có độ chính xác cao hơn so với các mô hình handle time-series data tiêu chuẩn vì có thể tổng hợp được sức mạnh của nhiều mô hình \cite{ayitey2023forex}.
+Các mô hình hybrid ensemble vốn được sử dụng rất nhiều trong các bài toán xử lý time-series và được chứng minh thực nghiệm là có độ chính xác cao hơn so với các mô hình handle time-series data tiêu chuẩn vì có thể tổng hợp được sức mạnh của nhiều mô hình \cite{ayitey2023forex}.
 Tuy vậy, các hình thức tổng hợp của ensemble model hiện nay vẫn còn rất cứng nhắc vì chỉ có thể tổng hợp dựa trên kết quả cuối (cơ chế voting của bagging models) và kết quả gần cuối (đối với stacking models).
@@
 
-Hybrid ensemble models have been widely used in time-series processing problems and have been experimentally proven to be more accurate than standard time-series models because they can synthesize the strengths of many sub-models \cite{ayitey2023forex}.
+% Hybrid ensemble models have been widely used in time-series processing problems and have been experimentally proven to be more accurate than standard time-series models because they can synthesize the strengths of many sub-models \cite{ayitey2023forex}.
 However, current ensemble model synthesis forms are still very rigid because they can only synthesize based on the final results (e.g.
@@
 
-% \verb|NHITS| được thiết kế để hướng đến việc dự đoán các long-horizon time-series data.
-Theo nghiên cứu \cite{challu2023nhits}, cấu trúc của \verb|NHITS| bao gồm nhiều stack liên tiếp nhau.
+\verb|NHITS| được thiết kế để hướng đến việc dự đoán một khoảng dữ liệu time-series tương lai.
+Cấu trúc của \verb|NHITS| bao gồm nhiều stack liên tiếp nhau.
 Mỗi stack bao gồm nhiều block nối tiếp nhau.
 Tại mỗi block, dữ liệu lịch sử được sử dụng để dự đoán dữ liệu tương lai và dữ liệu quá khứ.
-Cụ thể, tại block $l$, với $L$ mẫu dữ liệu quá khứ ($\mathbf{y}_{t-L:t, l-1}$), các đặc trưng sẽ được rút trích như sau (theo \cite{challu2023nhits}):
-
-\verb|NHITS| is designed to target the prediction of long-horizon time-series data.
+Cụ thể, tại block $l$, với $L$ mẫu dữ liệu quá khứ ($\mathbf{y}_{t-L:t, l-1}$), các đặc trưng sẽ được rút trích như sau \cite{challu2023nhits}:
+
+% \verb|NHITS| is designed to target the prediction of long-horizon time-series data.
 According to \cite{challu2023nhits}, the structure of \verb|NHITS| consists of multiple consecutive stacks.
 Each stack consists of multiple consecutive blocks.
-At each block, historical data is used to predict future data and past data.
-The residual of the previous block is used as input data for the following block.
+At each block, historical data is used to predict future data and approximate past data.
+The difference between historical data and the approximate one of the previous block is used as input data for the following block.
 Specifically, at block $l$, with $L$ past data samples ($\mathbf{y}_{t-L:t, l-1}$), the features will be extracted as follows (from \cite{challu2023nhits}):
@@
 
-% Trong đó, \verb|FullyConnected| là các lớp multi-layer perception (\verb|MLP|) xếp chồng với hàm kích hoạt phi tuyến.
+Trong đó, \verb|Pooling| nén dữ liệu ban đầy vào $\mathbf{y}_{t-L:t, l}^{(p)}$.
+\verb|FullyConnected| gồm lớp multi-layer perception xếp chồng với hàm kích hoạt phi tuyến.
 $\mathbf{\theta}_l^f, \mathbf{\theta}_l^b$ là các hệ số nội suy forecast và backcast, được dùng để tổng hợp các giá trị đầu ra của block $l$ bằng hàm nội suy $g(\cdot)$.
@@
 Theo đó, các stack có thể mô phỏng lại các tần số từ thấp đến cao.
-Ngoài ra, bằng việc sử dụng hàm nội suy, \verb|NHITS| không cần quá nhiều phần cứng để huấn luyện mạng neural trong trường hợp H lớn.
-
-Accordingly, \verb|Pooling| compresses the original data into $\mathbf{y}_{t-L:t, l}^{(p)}$.
+Ngoài ra, bằng việc sử dụng hàm nội suy, \verb|NHITS| không cần quá nhiều phần cứng để huấn luyện mạng neural trong trường hợp $H$ lớn.
+
+% Accordingly, \verb|Pooling| compresses the original data into $\mathbf{y}_{t-L:t, l}^{(p)}$.
 \verb|FullyConnected| contains many perceptron layers with nonlinear activation functions.
 $\mathbf{\theta}_l^f, \mathbf{\theta}_l^b$ are the forecast and backcast interpolation coefficients, which are used to aggregate the output values of block $l$ using the interpolation function $g(\cdot)$.
-The dimensionality of the interpolation coefficients in each stack is determined by the expressiveness ratio: $\left| \theta_l \right| = \left \lceil r_l H \right \rceil$.
+The dimensionality of the interpolation coefficients in each stack is determined by the expressiveness ratio: $\left| \theta_l^f \right| = \left \lceil r_l H \right \rceil$.
 Typically, the expressiveness ratio will be very small in the first stacks and increase towards the end.
@@
 
-% Đầu ra của block $l$ là giá trị forecast $\mathbf{\hat{y}}_{t+1:t+H, l}$ và giá trị backcast $\mathbf{\hat{y}}_{t-L:t, l}$. Input của block $l+1$ được tính theo phương trình \ref{eq:input_l1}.
-
-The output of block $l$ is the forecast value $\mathbf{\hat{y}}_{t+1:t+H, l}$ and the backcast value $\mathbf{\hat{y}}_{t-L:t, l}$.
+Đầu ra của block $l$ là giá trị forecast $\mathbf{\hat{y}}_{t+1:t+H, l}$ và giá trị backcast $\mathbf{\hat{y}}_{t-L:t, l}$. Input của block $l+1$ được tính theo phương trình \ref{eq:input_l1}.
+
+% The output of block $l$ is the forecast value $\mathbf{\hat{y}}_{t+1:t+H, l}$ and the backcast value $\mathbf{\hat{y}}_{t-L:t, l}$.
 The input of block $l+1$ is calculated according to the equation \ref{eq:input_l1}.
@@
 
-% Giả sử mô hình gồm có $S$ stacks, mỗi stack có $B$ blocks.
+Giả sử mô hình gồm có $S$ stacks, mỗi stack có $B$ blocks.
 Tổng hợp các giá trị forecast của các block như phương trình \ref{eq:sum_block}, ta được giá trị forecast của một stack.
@@
 
-Suppose the model consists of $S$ stacks, each stack has $B$ blocks.
+% Suppose the model consists of $S$ stacks, each stack has $B$ blocks.
 Summing the forecast values of the blocks as in equation \ref{eq:sum_block}, we get the forecast value of a stack.
@@
 
-% Bằng cách xếp chồng các stack, stack sau nhận vào phần dư của stack trước, kiến trúc trên được kỳ vọng là sẽ phân rã dữ liệu thành các frequency bands khác nhau (weekly, daily, even hourly).
+Bằng cách xếp chồng các stack, stack sau nhận vào phần dư của stack trước, kiến trúc trên được kỳ vọng là sẽ phân rã dữ liệu thành các frequency bands khác nhau (weekly, daily, even hourly).
 Trên thực tế, \verb|NHITS| perform rất tốt đối với các bộ dữ liệu có tính chu kỳ cao như mức tiêu thụ điện, thời tiết, giao thông.
@@
 
-By concatenating stacks, each receiving the remainder of the previous stack, the above architecture is expected to decompose the data into different frequency bands (weekly, daily, even hourly).
+% By concatenating stacks, each receiving the remainder of the previous stack, the above architecture is expected to decompose the data into different frequency bands (weekly, daily, even hourly).
 In practice, \verb|NHITS| performs very well for highly periodic datasets such as electricity consumption, weather, traffic.
@@
 % nói tổng quan ở đây, vẽ cái hình và bỏ vào
-% Phương pháp của chúng tôi bao gồm hai phần chính hoạt động song song nhau: (1) - Feature extraction; (2) - Parameter synthesis.
+Phương pháp của chúng tôi bao gồm hai phần chính hoạt động song song nhau: (1) - Feature extraction; (2) - Parameter synthesis.
 Tổng quan phương pháp được minh họa trong hình \ref{fig:flow}.
@@
 
-Our method consists of two main parts that work in parallel: (1) - Feature extraction; (2) - Parameter synthesis.
-The overview of the method is illustrated in \ref{fig:flow}.
+% Our method consists of two main parts that work in parallel: (1) - Feature extraction; (2) - Parameter synthesis.
+The overview of the method is a ML process as mentioned in section \ref{sec.relatedWork} and is illustrated in figure \ref{fig:flow}.
 In the feature extraction part, we combine two types of features from \verb|CNN| and \verb|LSTM| networks.
@@
 
-% Phương pháp đề xuất sử dụng các thuật toán ML để huấn luyện mô hình.
+Phương pháp đề xuất sử dụng các thuật toán ML để huấn luyện mô hình.
 Do đó, dữ liệu cần được tổ chức lại để các thuật toán ML có thể hoạt động được.
@@
 
-The proposed method uses ML algorithms to train the model.
+% The proposed method uses ML algorithms to train the model.
 Therefore, the data needs to be reorganized so that the ML algorithms can work.
@@
 
-% Một sample dữ liệu bao gồm các cặp giá trị $(\mathbf{x}_{t-L:t}, y)$.
+Một sample dữ liệu bao gồm các cặp giá trị $(\mathbf{x}_{t-L:t}, y)$.
 Trong đó, $\mathbf{x}_{t-L:t}$ bao gồm $L$ giá trị lịch sử tính từ thời điểm $t$ trở về trước; $y\in \{0,1\}$ là nhãn dữ liệu, thể hiện xu hướng giảm, hoặc tăng của mẫu dữ liệu $x_{t+1}$ so với $x_{t}$.
@@
 
-A data sample consists of pairs of values $(\mathbf{x}_{t-L:t}, y)$.
+% A data sample consists of pairs of values $(\mathbf{x}_{t-L:t}, y)$.
 In which, $\mathbf{x}_{t-L:t}$ includes $L$ historical values from time $t$ back; $y\in \{0,1\}$ is the data label, showing the decreasing or increasing trend of the data sample $x_{t+1}$ compared to $x_{t}$.
-Depending on each problem and the implementation, the elements in $\mathbf{x}_{t-L:t}$ can be vectors or scalar numbers.
-For example, for stock data, $\mathbf{x}_{t-L:t}$ can contain $L$ data vectors $\vec x_i = (\text{open, low, high, close})$ or just a single close price value.
+Depending on each problem and the implementation, the elements in $\mathbf{x}_{t-L:t}$ can be a matrix or a vector.
+For example, for stock data, $\mathbf{x}_{t-L:t}$ can contain $L$ vectors $\vec x_i = (\text{open, low, high, close})$ or can be a vector of close price values only.
 
@@
 
-% Lấy cảm hứng từ nghiên cứu \cite{vo2017multi}, chúng tôi đề xuất kết hợp các đặc trưng rút trích được từ mạng \verb|LSTM| và \verb|CNN|.
-Cụ thể, chúng tôi đưa từng phần tử trong vector $\mathbf{x}_{t-L:t}$ qua một lớp \verb|MLP| có đầu ra lớn hơn số chiều của $\vec x_i, i\in[t-L, t]$ để phân giải thành các đặc trưng nhỏ $\vec x'_i$.
+Lấy cảm hứng từ nghiên cứu \cite{vo2017multi}, chúng tôi đề xuất kết hợp các đặc trưng rút trích được từ mạng \verb|LSTM| và \verb|CNN|.
+Cụ thể, chúng tôi đưa từng phần tử trong vector $\mathbf{x}_{t-L:t}$ qua một lớp \verb|FullyConnected| có đầu ra lớn hơn số chiều của $\vec x_i, i\in[t-L, t]$ để phân giải thành các đặc trưng nhỏ $\vec x'_i$.
 Các đặc trưng này sau đó được truyền qua mạng \verb|LSTM| và \verb|CNN| để lần lượt rút trích các phụ thuộc thời gian dài hạn ($\mathbf{h}_{LSTM}$) và các đặc trưng thời gian cục bộ ($\mathbf{h}_{CNN}$).
@@
 
-Inspired by the \cite{vo2017multi} study, we propose to combine the features extracted from \verb|LSTM| and \verb|CNN| networks.
+% Inspired by the \cite{vo2017multi} study, we propose to combine the features extracted from \verb|LSTM| and \verb|CNN| networks.
 Specifically, we pass each element in the vector $\mathbf{x}_{t-L:t}$ through \verb|FullyConnected| layers whose output'dimension is larger than the one of $\vec x_i, i\in[t-L, t]$ to decompose it into smaller features $\vec x'_i$.
@@
 
-% Mạng \verb|LSTM| duy trì giá trị cell-state nhằm lưu trữ có chọn lọc các phụ thuộc dài hạn.
+Mạng \verb|LSTM| duy trì giá trị cell-state nhằm lưu trữ có chọn lọc các phụ thuộc dài hạn.
 Điều này rất thích hợp trong việc giải quyết các bài toán time-series data.
@@
 
-The \verb|LSTM| network maintains cell-state values to selectively store long-term dependencies.
+% The \verb|LSTM| network maintains cell-state values to selectively store long-term dependencies.
 This is very suitable for solving time-series data problems.
@@
 We propose to use the \verb|CNN| network to emphasize local features, thereby directing part of the model's attention to certain time points.
-Therefore, the proposed method can not only remember long-term features but also highlight short-term features.
-
-\vspace{2mm}
-
-% Tiếp đến, $\mathbf{h}_{LSTM}$ và $\mathbf{h}_{CNN}$ được nối với nhau (phương trình \ref{eq:concat}) sau đó chuyển đến phần phân lớp của mạng NN (phương trình \ref{eq:clf}).
-
-Next, $\mathbf{h}_{LSTM}$ and $\mathbf{h}_{CNN}$ are concatenated (equation \ref{eq:concat}) and then passed to the classification part of the neural network (equation \ref{eq:clf}).
+Therefore, the proposed method not only remembers long-term features but also highlights short-term features.
+
+\vspace{2mm}
+
+Tiếp đến, $\mathbf{h}_{LSTM}$ và $\mathbf{h}_{CNN}$ được nối với nhau (phương trình \ref{eq:concat}) sau đó chuyển đến phần phân lớp của mạng NN (phương trình \ref{eq:clf}).
+
+% Next, $\mathbf{h}_{LSTM}$ and $\mathbf{h}_{CNN}$ are concatenated (equation \ref{eq:concat}) and then passed to the classification part of the neural network (equation \ref{eq:clf}).
 
@@
 
-% Chúng tôi sử dụng \verb|MAML| để huấn luyện và tổng hợp trọng số của các mô hình tại các task.
+Chúng tôi sử dụng \verb|MAML| để huấn luyện và tổng hợp trọng số của các mô hình tại các task.
 Như đã đề cập trong phần \ref{sec.relatedWork}, tối ưu tham số theo cách tiếp cận của ML chính là đi giải hai phương trình \ref{eq:inner_opt} và \ref{eq:outer_opt} bằng các phương pháp tối ưu trên dữ liệu support và query.
@@
 
-We use \verb|MAML| to train and aggregate the weights of the models at the tasks.
+% We use \verb|MAML| to train and aggregate the weights of the models at the tasks.
 As mentioned in \ref{sec.relatedWork}, parameter optimization in the ML approach is to solve the two equations \ref{eq:inner_opt} and \ref{eq:outer_opt} using optimization methods on the support and query data.
@@
     \end{cases}
-\end{align} In which, $\phi_{r-1}$ is the result of the $r-1$ global optimization process, $\alpha$ is the inner learning rate.
-
-\vspace{2mm}
-
-% Tiếp đó, quá trình outer optimization tại bước toàn cục được thực hiện bằng cách tổng hợp độ lỗi trên tập query của các task và tối ưu trên đó (phương trình \ref{eq:outer_sol}).
-
-Next, the outer optimization process at the global step is performed by aggregating the losses on the query set of the tasks and optimizing on it (equation \ref{eq:outer_sol}).
+\end{align} In which, $\phi_{r-1}$ is the result of the $r-1$ outer optimization process, $\alpha$ is the inner learning rate.
+
+\vspace{2mm}
+
+Tiếp đó, quá trình outer optimization tại bước toàn cục được thực hiện bằng cách tổng hợp độ lỗi trên tập query của các task và tối ưu trên đó (phương trình \ref{eq:outer_sol}).
+
+% Next, the outer optimization process is performed by aggregating the losses on the query set of the tasks and optimizing on it (equation \ref{eq:outer_sol}).
 
@@
 
-% Giả sử thuật toán chạy $E$ steps trong inner optimization, lượng đạo hàm tại phương trình \ref{eq:outer_sol} được viết lại như sau (the notations of dataset are removed):
-
-Assuming the algorithm runs $E$ steps in inner optimization, the derivative quantity at equation \ref{eq:outer_sol} is rewritten as follows (the notations of dataset are removed):
+Giả sử thuật toán chạy $E$ steps trong inner optimization, lượng đạo hàm tại phương trình \ref{eq:outer_sol} được viết lại như sau (the notations of dataset are removed):
+
+% Assuming the algorithm runs $E$ steps in inner optimization, the derivative quantity at equation \ref{eq:outer_sol} is rewritten as follows (the notations of dataset are removed):
 
@@
 
-% Sự xuất hiện của tích các đạo hàm bậc hai trong phương trình \ref{eq:outer_sol_} khiến quá trình đạo hàm trở nên phức tạp vì phải tốn rất nhiều chi phí để duy trì các ma trận Hessian.
+Sự xuất hiện của tích các đạo hàm bậc hai trong phương trình \ref{eq:outer_sol_} khiến quá trình đạo hàm trở nên phức tạp vì phải tốn rất nhiều chi phí để duy trì các ma trận Hessian.
 Do đó, số bước tính toán để tìm ra $\theta^*$ cần phải hạn chế.
@@
 
-The presence of the product of second order derivatives in the equation \ref{eq:outer_sol_} makes the derivation process complicated because it requires a lot of overhead to maintain the Hessian matrices.
+% The presence of the product of second order derivatives in the equation \ref{eq:outer_sol_} makes the derivation process complicated because it requires a lot of overhead to maintain Hessian matrices.
 Therefore, the number of computational steps to find $\theta^*$ needs to be limited.
@@
 
-% FX nói riêng và các chỉ số tài chính nói chung là dạng dữ liệu điển hình cho aperiodic time-series data.
+FX nói riêng và các chỉ số tài chính nói chung là dạng dữ liệu điển hình cho aperiodic time-series data.
 Do đó, chúng tôi chọn loại dữ liệu này để kiểm thử mô hình.
@@
 Dữ liệu có thuộc tính tương tự như \verb|USD/JPY| và được sample theo ngày từ năm 2014 đến 2024.
-Số lượng mẫu dữ liệu của hai tập dữ liệu là tương đương nhau và xấp xỉ $156000$ mẫu.
-
-FX in particular and financial indices in general are typical data types for aperiodic time-series data.
+
+% FX in particular and financial indices in general are typical data types for aperiodic time-series data.
 Therefore, we choose this type of data to test the model.
@@
 
-% Bộ dữ liệu \verb|multi-fx| được sử dụng để rút trích và tổng hợp thông tin về các yếu tố ngoại lai (i.e.
+Bộ dữ liệu \verb|multi-fx| được sử dụng để rút trích và tổng hợp thông tin về các yếu tố ngoại lai (i.e.
 thông tin thị trường, kinh tế, chính trị,...), vốn được cho là có ảnh hưởng đến kết quả của một chỉ số tài chính nhất định \citep{overreactioncontrarian,fama1970efficient,mech1993portfolio}.
@@
 
-Our goal is to use \verb|multi-fx| dataset to extract and aggregate information about outliers (i.e.
+% Our goal is to use \verb|multi-fx| dataset to check the information extraction and aggregation of external factors (i.e.
 market, economic, political, etc.) that are believed to influence the outcome of a given financial index \citep{overreactioncontrarian,fama1970efficient,mech1993portfolio}.
@@
 
-% Ngoài ra, chúng tôi sử dụng thêm 2 bộ dữ liệu: Electricity Transformer Temperature (\verb|ETT-m2|) và Weather (\verb|WTH|).
+Ngoài ra, chúng tôi sử dụng thêm 2 bộ dữ liệu: Electricity Transformer Temperature (\verb|ETT-m2|) và Weather (\verb|WTH|).
 Tập dữ liệu \verb|ETT-m2| bao gồm 7 trường dữ liệu, đo đạc các thông số của máy biến áp tại một tỉnh của Trung Quốc sau mỗi 15 phút trong khoảng thời gian từ July 2016 to July 2018.
@@
 
-In addition, we used two periodic datasets: Electricity Transformer Temperature (\verb|ETT-m2|) \cite{zhou2021informer} and Weather (\verb|WTH|) \cite{Kolle}.
+% In addition, we used two periodic datasets: Electricity Transformer Temperature (\verb|ETT-m2|) \cite{zhou2021informer} and Weather (\verb|WTH|) \cite{Kolle}.
 \verb|ETT-m2| dataset consists of 7 data fields, measuring the parameters of a transformer in a province of China every 15 minutes from July 2016 to July 2018.
@@
 
-% Nghiên cứu sử dụng macro accuracy, macro precision, macro recall, và macro F1 để đánh giá các mô hình.
+Nghiên cứu sử dụng macro accuracy, macro precision, macro recall, và macro F1 để đánh giá các mô hình.
 Theo đó, trong quá trình inference, mô hình sẽ chạy trên từng task để tính metrics của mỗi task.
@@
 
-The study uses macro accuracy, macro precision, macro recall, and macro F1 score to evaluate the models.
+% The study uses macro accuracy, macro precision, macro recall, and macro F1 score to evaluate the models.
 Accordingly, during the inference phase, the model will run on all tasks to calculate the metrics of each one.
@@
 
-We compare the proposed method with \verb|NHITS| using the above metrics and dataset.
-For each dataset, each attribute will be predicted for future trends based on past data.
+Chúng tôi so sánh phương pháp đề xuất với \verb|NHITS| bằng cách sử dụng các số liệu và tập dữ liệu trên.
+Đối với mỗi tập dữ liệu, mỗi trường dữ liệu sẽ được dự đoán xu hướng tương lai dựa trên dữ liệu trong quá khứ.
+Kết quả cuối cùng được tính bằng cách lấy trung bình kết quả trên các trường dữ liệu.
+Chúng tôi cũng thử nghiệm các bộ trích xuất tính năng khác nhau (\verb|LSTM|, \verb|CNN|) cho phương pháp đề xuất để chứng minh khả năng nắm bắt dữ liệu của \verb|LSTM+CNN|.
+
+% We compare the proposed method with \verb|NHITS| using the above metrics and dataset.
+For each dataset, each attribute will be predicted the future trends based on past data.
 The final result is calculated by averaging the metrics across the attributes.
 We also test different feature extractors (\verb|LSTM|, \verb|CNN|) for the proposed method to demonstrate the deep data understanding capabilities of \verb|LSTM+CNN|.
-The overall data is structured into training sets, validation sets, and testing sets with a ratio of 6:2:2 for training, fine-tuning and testing model, respectively.
-
-\vspace{2mm}
-
-For \verb|NHITS| model, we split the data as usual according to the above predetermined ratio.
+
+\vspace{2mm}
+
+Đối với mô hình \verb|NHITS|, dữ liệu được cấu trúc thành các tập huấn luyện, tập xác thực và tập kiểm tra với tỷ lệ 6:2:2.
+Đối với thuật toán đề xuất, vì quy trình meta-training và meta-testing yêu cầu chia dữ liệu thành các tác vụ nhỏ và cho phép mô hình thích ứng trên tập hỗ trợ của từng tác vụ, chúng tôi chia dữ liệu thành 40-60 tác vụ (xem bảng \ref{tab:stat_} để biết chi tiết thống kê).
+Trong mỗi tác vụ, tập support chiếm 20\% với mục đích cho phép mô hình thích ứng với dữ liệu, tập query chiếm 80\% để kiểm tra tính tương thích của mô hình.
+Chúng tôi sử dụng 50\% tác vụ cho meta-training, 25\% cho meta-validating và 25\% cho meta-testing.
+Với cách chia này, chúng tôi đảm bảo tính công bằng rằng mô hình ML được huấn luyện với cùng lượng dữ liệu như \verb|NHITS| (xem phần phụ lục \ref{app:data}).
+Chi tiết về các thí nghiệm có thể được tìm thấy ở phần phụ lục \ref{app:our_experiment} và \ref{app:nhits_experiment}.
+
+% For \verb|NHITS| model, data is structured into training sets, validation sets, and testing sets with a ratio of 6:2:2 for training, fine-tuning and testing model, respectively.
 For the proposed algorithm, because the meta-training and meta-testing processes require dividing the data into small tasks, and allowing the model to adapt on the support set of each task, we split the data into 40-60 tasks (see table \ref{tab:stat_} for statistical detail).
@@
     \centering
-    \caption{Statistics on USD/JPY and multi-fx.}
+    \caption{Statistics on datasets.}
     \label{tab:stat_}
-    \begin{tabular}{ccccccc}
-    \toprule
-    \multirow{2}{*}{Dataset} & \multirow{2}{*}{\#task} & \multirow{2}{*}{\#samples} & \multicolumn{4}{c}{\#samples/task}  \\
-    \cline{4-7}
-                             &                         &                            & min & mean  & std   & max             \\
-    \hline
-    \Verb|USD/JPY|           & 60                      & 150,175                     & 2,502 & 2,502.92 & 0.28 & 2,503           \\
-    \Verb|multi-fx|          & 60                      & 154,440                     & 2,467 & 2,575 & 58.61   & 2748           \\
-    \Verb|ETT-m2|            & 48                      & 69,680                      & 1,451 & 1,451.67 & 0.47   & 1,452           \\
-    \Verb|WTH|               & 40                      & 35,064                      & 876 & 876.6 & 0.49   & 877           \\
-    \bottomrule
-    \end{tabular}
+    \begin{tblr}{
+        cells = {c},
+        hline{1,6} = {-}{0.08em},
+        hline{2} = {-}{},
+    }
+    Dataset                      & Attribute & Task & Sample    & Sample/Task    \\
+    \Verb|USD/JPY|               & 4         & 60   & 150,175   & 2,503          \\
+    \Verb|multi-fx|              & 4         & 60   & 154,440   & 2,575          \\
+    \Verb|ETT-m2|                & 7         & 48   & 69,680    & 1,452          \\
+    \Verb|WTH|                   & 12        & 40   & 35,064    & 877            
+    \end{tblr}
 \end{table}
@@
 
-% Bảng \ref{tab:rs} so sánh kết quả giữa mô hình đề xuất và mô hình \verb|NHITS| trên các tập dữ liệu kể trên.
+Bảng \ref{tab:rs} so sánh kết quả giữa mô hình đề xuất và mô hình \verb|NHITS| trên các tập dữ liệu kể trên.
 Trên dữ liệu phi chu kỳ (\verb|multi-fx| and \verb|USD/JPY|), phương pháp của chúng tôi đạt mức hội tụ cao hơn hẳn so với mô hình \verb|NHITS| thể hiện trên tất cả các metrics.
@@
 
-Table \ref{tab:rs} compares the results between the proposed model and the \verb|NHITS| model on the above datasets.
+% Table \ref{tab:rs} compares the results between the proposed model and the \verb|NHITS| model using the above datasets.
 On aperiodic data, our method achieves a much higher convergence rate than the \verb|NHITS| model on all metrics.
@@
     \centering
-    \caption{Classification results (\%) of NHITS and our method using USD/JPY and multi-fx datasets. Best results per metrics are boldfaced.}
+    \caption{Classification results (\%) of NHITS and our method. Best results per metrics are boldfaced.}
     \label{tab:rs}
@@
 
-% Chúng tôi huấn luyện ba mô hình \verb|CNN|, \verb|LSTM|, \verb|LSTM+CNN| trên 2600 mẫu dữ liệu \verb|CHF/NZD| theo cách thông thường.
+Chúng tôi huấn luyện ba mô hình \verb|CNN|, \verb|LSTM|, \verb|LSTM+CNN| trên 2600 mẫu dữ liệu \verb|CHF/NZD| theo cách thông thường.
 Sau khi quan sát quá trình huấn luyện của chúng (hình \ref{fig:lstm_cnn}), chúng tôi nhận thấy khả năng rút đặc trưng cục bộ tuyệt vời trên dữ liệu ngắn hạn của \verb|CNN|.
@@
 
-We trained three models \verb|CNN|, \verb|LSTM|, \verb|LSTM+CNN| on 2600 samples of \verb|USD/JPY| data in the conventional way.
+% We trained three models \verb|CNN|, \verb|LSTM|, \verb|LSTM+CNN| on 2600 samples of \verb|USD/JPY| data in the conventional way.
 After observing their training process (figure \ref{fig:lstm_cnn}), we found that \verb|CNN| has excellent local feature extraction ability on short-term data.
@@
 
-% Tuy nhiên, khi quan sát kết quả trên tập validation, cả ba mô hình trên không đều cho kết quả rất tệ.
+Tuy nhiên, khi quan sát kết quả trên tập validation, cả ba mô hình trên không đều cho kết quả rất tệ.
 Lý giải cho điều này, chúng tôi nhận định rằng dù có thể học tốt trên tập train nhưng các mô hình đều mất đi tính tổng quát hóa vì đây là loại dữ liệu phức tạp, khác với các dữ liệu như hình ảnh, video, hoặc dữ liệu có chu kỳ.
@@
 
-However, when observing the results on the validation set, all three models above give very poor results.
-To explain this, we think that although they can learn well on the training set, the models lose their generalization ability because this is a complex type of data, different from data such as images, videos, or periodic data.
-If the model's adaptability is not high, poor results on the validation set are predictable.
-However, \textbf{the graph in figure \ref{fig:lstm_cnn} shows that we cannot deny the feature extraction ability of} \verb|LSTM+CNN|.
+% However, when observing the results on the validation set, all three models above give very poor results.
+To explain this, we think that although they can learn well on the training set, the models lose their generalization ability because this is a complex type of data, which is different from data such as images, videos, or periodic data.
+If the model's adaptability is not good, poor results on the validation set are predictable.
+This suggests that there is a need for a method that efficiently utilizes and synthesizes the obtained features, which is discuss in detail in the next subsection.
 
@@
 
-% Quá trình rút trích và tổng hợp đặc trưng của \verb|NHITS| thể hiện rằng phương pháp này đang cố gắng mô phỏng lại quá trình phân giải tần số của Fourier transform.
+Quá trình rút trích và tổng hợp đặc trưng của \verb|NHITS| thể hiện rằng phương pháp này đang cố gắng mô phỏng lại quá trình phân giải tần số của Fourier transform.
 Đặc trưng về các giải tần là thông tin tối quan trọng đối với dữ liệu có chu kỳ.
@@
 
-The feature extraction process of \verb|NHITS| shows that this method is trying to simulate the frequency resolution process of Fourier transform.
+% The feature extraction process of \verb|NHITS| shows that this method is trying to simulate the frequency resolution process of Fourier transform.
 The frequency features are the most important information for periodic data.
@@
 
-% Lý do phương pháp huấn luyện thông thường không đạt hiệu quả cao mặc cho khả năng rút trích đặc trưng đã được cải thiện là vì thiếu khả năng sử dụng đặc trưng một cách hiệu quả.
+Lý do phương pháp huấn luyện thông thường không đạt hiệu quả cao mặc cho khả năng rút trích đặc trưng đã được cải thiện là vì thiếu khả năng sử dụng đặc trưng một cách hiệu quả.
 Khả năng này trong các nghiên cứu trước đây thường được cải thiện bằng việc sử dụng các mô hình ensemble.
@@
 
-The reason why conventional training methods do not perform well despite the improved feature extraction capabilities is due to the lack of the ability to use features effectively.
+% The reason why conventional training methods do not perform well despite the improved feature extraction capabilities is due to the lack of the ability to use features effectively.
 This ability in previous studies has often been improved by using ensemble models.
@@
 
-% Khi sử dụng \verb|MAML| để tổng hợp đặc trưng của mô hình \verb|LSTM| và \verb|LSTM+CNN|, khả năng tương thích (thể hiện qua việc hội tụ sớm) cũng như kết quả (thể hiện qua độ chính xác) của mô hình sử dụng \verb|LSTM+CNN| cao hơn hoàn toàn so với mô hình chỉ sử dụng \verb|LSTM| (xem hình \ref{fig:lstm_cnn_feature}).
+Khi sử dụng \verb|MAML| để tổng hợp đặc trưng của mô hình \verb|LSTM| và \verb|LSTM+CNN|, khả năng tương thích (thể hiện qua việc hội tụ sớm) cũng như kết quả (thể hiện qua độ chính xác) của mô hình sử dụng \verb|LSTM+CNN| cao hơn hoàn toàn so với mô hình chỉ sử dụng \verb|LSTM| (xem hình \ref{fig:lstm_cnn_feature}).
 Thật vậy, chỉ sau 40 epochs, các đặc trưng \verb|LSTM+CNN| cho mức hội tụ trên 60\% và sau 100 epochs, mô hình hội tụ ở mức 62\%, vượt trên những gì các đặc trưng \verb|LSTM| có thể làm được.
@@
 
-When using \verb|MAML| to synthesize the features of the \verb|LSTM| and \verb|LSTM+CNN| models, the compatibility (indicated by early convergence) as well as the results (indicated by accuracy) of the model using \verb|LSTM+CNN| are significantly higher than those of the model using only \verb|LSTM| (see figure \ref{fig:lstm_cnn_feature}) and \verb|NHITS| model.
+% When using \verb|MAML| to synthesize the features of the \verb|LSTM| and \verb|LSTM+CNN| models, the compatibility (indicated by early convergence) as well as the results (indicated by accuracy) of the model using \verb|LSTM+CNN| are significantly higher than those of the model using only \verb|LSTM| (see figure \ref{fig:lstm_cnn_feature}) and \verb|NHITS| model.
 Indeed, on \verb|multi-fx| dataset, after only 40 epochs, the \verb|LSTM+CNN| features converge to over 60\% and after 100 epochs, the model converges to 62\%, exceeding what the \verb|LSTM| features can do.
@@
 For the ML model using only \verb|CNN|, the accuracy and error values do not even improve during the entire training process.
-We explain this by saying that the features of \verb|CNN| focus too much on the locality of the data and ignore the long-term dependencies, which is an important feature of time-series data.
+To explain this, we argue that the features of \verb|CNN| have focused too much on the locality of the data and ignored the long-term dependencies, which is an important feature of time-series data.
 In contrast, the features of \verb|LSTM+CNN| not only capture the long-term dependencies but also highlight the local properties in the data, making it completely outperform \verb|CNN| as well as \verb|NHITS|.
-Therefore, \textbf{we conclude that ML is more flexible and efficient in synthesizing sub-models than traditional rigid ensemble models}.
-
-\vspace{2mm}
-
-% Khi tiếp cận dữ liệu \verb|USD/JPY| bằng ML, mô hình cũng cho độ chính xác cao.
+Therefore, \textbf{we conclude that ML is more flexible and efficient in synthesizing sub-models as well as external factors than traditional rigid ensemble models}.
+
+\vspace{2mm}
+
+Khi tiếp cận dữ liệu \verb|USD/JPY| bằng ML, mô hình cũng cho độ chính xác cao.
 Bằng việc nhìn vào các khoảng thời gian trong quá khứ, \textbf{phương pháp đề xuất đã tỏ ra hiệu quả hơn trong việc nắm bắt các hidden long-term dependency thông qua việc tổng hợp hiệu quả các đặc trưng mà} \verb|LSTM| \textbf{dễ dàng quên đi trong quá trình huấn luyện}.
@@
 
-When approaching \verb|USD/JPY|, \verb|ETT-m2|, and \verb|WTH| dataset using ML, the model also shows accuracies higher than or close to \verb|NHITS|.
+% When approaching \verb|USD/JPY|, \verb|ETT-m2|, and \verb|WTH| dataset using ML, the model also shows accuracies higher than or close to \verb|NHITS|.
 By looking at the past time periods, \textbf{the proposed method has proven to be more effective in capturing hidden long-term dependencies through the efficient aggregation of features that the} \verb|LSTM| \textbf{easily forgets during training}.
@@
 
+\vspace{2mm}
+
 % dis NHITS vì tổng hợp đặc trưng chỉ bằng phép cộng :v
-% Trong quá trình tổng hợp các đặc trưng để đưa ra dự đoán cuối, \verb|NHITS| chỉ sử dụng các phép cộng trừ đơn giản.
+Trong quá trình tổng hợp các đặc trưng để đưa ra dự đoán cuối, \verb|NHITS| chỉ sử dụng các phép cộng trừ đơn giản.
 Điều này đặt trong ngữ cảnh của việc tổng hợp thông tin dựa trên các hàm nội suy và phần dư của input là hợp lý.
@@
 
-In the process of synthesizing features to make the final prediction, \verb|NHITS| only uses simple addition and subtraction.
+% In the process of synthesizing features to make the final prediction, \verb|NHITS| only uses simple addition and subtraction.
 This is reasonable in the context of synthesizing information based on interpolation functions and input residuals and trying to simulate the synthesis of frequency bands.
@@
 
-% Các bài toán trên aperiodic time-series data đặt ra một thử thách lớn cho các mô hình học máy hiện nay bởi tính bất định trong variance cũng như tính phi chu kỳ loại dữ liệu này.
+Các bài toán trên aperiodic time-series data đặt ra một thử thách lớn cho các mô hình học máy hiện nay bởi tính bất định trong variance cũng như tính phi chu kỳ loại dữ liệu này.
 Bằng việc kết hợp các đặc trưng cục bộ và các phụ thuộc dài hạn cũng như khai thác các phụ thuộc ẩn trong từng khoảng thời gian quá khứ, chúng tôi đã cải khả năng của mô hình học máy trên các classification metrics trong bài toán dự đoán xu hướng giá của ngày tiếp theo trên dữ liệu FX.
@@
 
-Problems on aperiodic time-series data pose a great challenge to current machine learning models due to the uncertainty in variance as well as the aperiodicity of this type of data.
+% Problems on aperiodic time-series data pose a great challenge to current machine learning models due to the uncertainty in variance as well as the aperiodicity of this type of data.
 By combining local features and long-term dependencies as well as exploiting hidden dependencies in each past time period, we have improved the ability of machine learning models on classification metrics in the problem of predicting the next day's price trend on FX data.
@@
 
-% \textbf{Model architecture}.
+\textbf{Model architecture}.
 Phương pháp của chúng tôi được phát triển dưới dạng module với hai module chính hoạt động song song: Module rút trích đặc trưng và Module tổng hợp mô hình.
@@
 
-\textbf{Model architecture}.
+% \textbf{Model architecture}.
 Our method is modularized with two main modules operating in parallel: Feature Extraction Module and Model Synthesis Module.
@@
 
-% \textbf{Long-horizon problem}.
+\textbf{Long-horizon problem}.
 Hoàn toàn có thể mở rộng phương pháp này để giải các bài toán về long-horizon prediction.
@@
 
-\textbf{Long-horizon problem}.
+\vspace{2mm}
+
+% \textbf{Long-horizon problem}.
 It is possible to extend this method to solve long-horizon prediction problems.
@@
 \bibliography{references}
-% \clearpage
+\clearpage
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \appendix
+\section{Fairness in data}
+\label{app:data}
+
+Với \verb|NHITS|, chúng tôi sử dụng các chia dữ liệu truyền thống. Tập dữ liệu ban đầu được chia thành training set, validation set, và testing set theo tỷ lệ 6:2:2.
+
+% For \verb|NHITS|, we use the traditional data split. The dataset is split into a training set, validation set, and testing set in a 6:2:2 ratio.
+
+\vspace{2mm}
+
+Đối với phương pháp đề xuất, chúng tôi chia dữ liệu thành các tập dữ liệu nhỏ, mỗi tập dữ liệu lại được chia thành tập support và tập query.
+Gọi số lượng task sau khi chia là $n$, chúng tôi sử dụng $0.5n$ tasks cho meta-training, $0.25n$ tasks cho meta-validating, và $0.25n$ tasks cho meta-testing.
+Trong mỗi task, tập support chiếm 20\%, tập query chiếm 80\%.
+
+% For the proposed method, we split the data into small datasets, each of which is divided into a support set and a query set.
+Let the number of tasks after splitting be $n$, we use $0.5n$ tasks for meta-training, $0.25n$ tasks for meta-validating, and $0.25n$ tasks for meta-testing.
+In each task, the support set accounts for 20\%, and the query set accounts for 80\%.
+
+\vspace{2mm}
+
+Gọi $M$ là số mẫu dữ liệu trong tập dữ liệu ban đầu, số mẫu dữ liệu sử dụng để train, validate, và test cho \verb|NHITS| lần lượt là $0.6M, 0.2M, 0.2M$.
+Đối chiếu với lượng dữ liệu sử dụng cho phương pháp đề xuất, mỗi task sẽ bao gồm $\frac{M}{n}$ mẫu dữ liệu.
+Chúng tôi sử dụng $0.5M$ cho meta-training (vì số task cho meta-training là $0.5n$).
+Tiếp đến, trong quá trình validate hoặc test, chúng tôi sử dụng 20\% dữ liệu mỗi task cho quá trình model adaptation và 80\% dữ liệu còn lại cho validate hoặc test.
+Theo đó, tổng số mẫu dùng cho quá trình huấn luyện và adaptation là $0.5M + 0.25n \times 0.2\frac{M}{n} + 0.25n \times 0.2\frac{M}{n} = 0.6M$.
+Số mẫu dùng để đánh giá mô hình trong quá trình validation và testing đều là $0.25n \times 0.8\frac{M}{n} = 0.2M$.
+Như vậy, chúng tôi đã bảo đảm sự công bằng trong huấn luyện và kiểm thử cho các thí nghiệm.
+
+% Let $M$ be the number of data samples in the original dataset, the number of data samples used for training, validation, and testing for \verb|NHITS| are $0.6M, 0.2M, 0.2M$ respectively.
+Compared to the amount of data used for the proposed method, each task will include $\frac{M}{n}$ data samples.
+We use $0.5M$ samples for meta-training (because the number of tasks for meta-training is $0.5n$).
+Next, during the validation or testing process, we use 20\% of the data of each task for the model adaptation process and the remaining 80\% of the data for validation or testing.
+Accordingly, the total number of samples used for training and adaptation is $0.5M + 0.25n \times 0.2\frac{M}{n} + 0.25n \times 0.2\frac{M}{n} = 0.6M$.
+The number of samples used to evaluate the model in the validation and testing process is both $0.25n \times 0.8\frac{M}{n} = 0.2M$.
+Thus, we have ensured fairness in training and testing for the experiments.
+
 \section{Experimental detail\\for proposed method}
@@
 
-% Trong cài đặt của chúng tôi, chúng tôi sử dụng một lớp \verb|FullyConnected| gồm 16 units với hàm kích hoạt \verb|ReLU| để phân giải đặc trưng ban đầu.
+Trong cài đặt của chúng tôi, chúng tôi sử dụng một lớp \verb|FullyConnected| gồm 16 units với hàm kích hoạt \verb|ReLU| để phân giải đặc trưng ban đầu.
 Sau đó, đặc trưng này được truyền song song đến các hai khối \verb|BidirectionalLSTM| và \verb|CNN|.
@@
 
-We use a \verb|FullyConnected| layer of 16 units with a \verb|ReLU| activation function to decompose the initial feature.
+% We use a \verb|FullyConnected| layer of 16 units with a \verb|ReLU| activation function to decompose the initial feature.
 This feature is then passed in parallel to the \verb|BidirectionalLSTM| and \verb|CNN| blocks.
@@
 
-The fine-tuning process of ML algorithms involves many hyper-parameters such as inner batch size, outer batch size, inner training steps, outer training steps.
+Quá trình fine-tune của thuật toán ML liên quan đến nhiều siêu tham số như inner batch size, outer batch size, inner training step, outer training step,...
+Để dễ dàng fine-tune, chúng tôi cố định hầu hết các tham số và chỉ fine-tune kích thước của lookback window, inner và outer learning rate.
+Chi tiết được trình bày trong bảng \ref{tab:our_finetune}.
+
+% The fine-tuning process of ML algorithms involves many hyper-parameters such as inner batch size, outer batch size, inner training steps, outer training steps.
 To facilitate fine-tuning, we fix most of the parameters and only fine-tune the size of lookback window, the inner and outer learning rates.
@@
 
-\begin{table}
+Đối với mô hình \verb|NHITS|, chúng tôi dựa trên \cite{challu2023nhits} để định nghĩa không gian tìm kiếm cho việc fine-tune tham số, cũng như kiến trúc mô hình.
+Đối với các tham số không được đề cập trong bảng, chúng tôi sử dụng giá trị mặc định của cài đặt \verb|NHITS| trong thư viện \verb|NeuralForecast| \cite{neuralforecast}.
+Chi tiết quá trình fine-tune được trình bày trong phần phụ lục \ref{app:experiment}.
+Kết quả tốt nhất của các lần fine-tune được chọn ra và báo cáo trong nghiên cứu này.
+
+% We rely on \cite{challu2023nhits} to define the search space (table \ref{tab:nhits_finetune}) for parameter fine-tuning, as well as to fine-tune the model architecture.
+For parameters not mentioned in the table, we use the default values of the \verb|NHITS| implementation in the \verb|NeuralForecast| library \cite{neuralforecast}.
+The best results of fine-tuning process are selected and reported in this study.
+
+\begin{table}[h]
     \centering

### 2024-12-10 model/baseline_models.py
baseline model: missing multi-fx

@@ -7,6 +7,7 @@ import json
 import pandas as pd
 import json
 import itertools
+import argparse
 
 from data.dataloader import DataLoader
 from common.constants import *
@@ -29,8 +30,8 @@ def add_log(log:dict, count:int, input_size:int, n_stacks:int, kernel_size:int, 
     for key in metric_log:
         log[count][key] = metric_log[key]
 
-def write_log(log:dict, file_name:str='NHITS_USD-JPY.json'):
-    with open(os.path.join(PRETRAINED_DIR, 'baseline', file_name), 'w') as fo:
+def write_log(log:dict, dataset_dir:str, label:int):
+    with open(os.path.join(dataset_dir, f'{label}.json'), 'w') as fo:
         json.dump(log, fo)
 
 def run(df:pd.DataFrame, param_dict:dict, columns:list[str]):
@@ -59,10 +60,25 @@ def run(df:pd.DataFrame, param_dict:dict, columns:list[str]):
     return rs_frame
 
 if __name__ == '__main__':
-    # config data and problem
-    dataset = MULTI_FX
+    # get dataset from command line
+    parser = argparse.ArgumentParser()
+    parser.add_argument('-dataset', '--dataset', type=str, required=True)
+    input_dict = vars(parser.parse_args())
+
+    # config dataset and model
+    dataset = input_dict['dataset']
     mode = CLF
     dataloader = DataLoader(dataset=dataset)
+
+    # create baseline_dir if needed
+    baseline_dir = os.path.join(PRETRAINED_DIR, 'baseline')
+    if not os.path.exists(baseline_dir):
+        os.mkdir(baseline_dir)
+
+    # create a folder for each dataset in baseline_dir
+    dataset_dir = os.path.join(baseline_dir, dataset)
+    if not os.path.exists(dataset_dir):
+        os.mkdir(dataset_dir)
 
     # hyper-param search space
     pooling_sizes = [[2,2,2], [4,4,4], [8,8,8], [8,4,1], [16,8,1]]
@@ -74,7 +90,7 @@ if __name__ == '__main__':
     # predict all columns (1 by 1) in the dataset
     # since NHITS doesn't take much time to run
     # i decided not to split the fine-tune process into jobs
-    # 75 models will be run simultaneously
+    # 75 models will be run sequentially
     for idx, label in enumerate(dataloader.columns):
         # get data and the given label
         print(f'\nPredict on label {idx+1}/{len(dataloader.columns)}\n')
[…]

### 2024-12-10 common/constants.py
minor changes

@@ -3,7 +3,7 @@ import os
 # dirs
 RAW_DATA_DIR = os.path.join(os.path.dirname(os.path.abspath(__file__)), '../data')
 EXE_DIR = os.path.join(os.path.dirname(os.path.abspath(__file__)), '../execute')
-PRETRAINED_DIR = os.path.join(os.path.dirname(os.path.abspath(__file__)), '../pretrained')
+PRETRAINED_DIR = os.path.join(os.path.dirname(os.path.abspath(__file__)), '../pretrained_new')
 LOG_DIR = os.path.join(os.path.dirname(os.path.abspath(__file__)), '../execute/log')
 
 # dataset

### 2024-12-10 execute/main.py

@@ -10,6 +10,7 @@ from data.dataloader import DataLoader
 
 def parse_args():
     parser = argparse.ArgumentParser()
+    parser.add_argument('-dataset', '--dataset', type=str, required=True)
     parser.add_argument('-window', '--look_back_window', type=int, required=True)
     parser.add_argument('-model', '--based_model', type=str, choices=[LSTM, CNN, LSTM_CNN, ATT], default=LSTM, required=False)
     parser.add_argument('-mode', '--mode', type=str, choices=[REG, CLF], default=CLF, required=True)
@@ -96,7 +97,7 @@ if __name__ == '__main__':
     input_dict = parse_args()
 
     # config data
-    dataset = MULTI_FX
+    dataset = input_dict['dataset']
     dataloader = DataLoader(dataset=dataset, look_back=input_dict['look_back_window'], mode=input_dict['mode'], is_multi='multi' in dataset)
 
     # specify model_dir to save model and its log
@@ -130,4 +131,4 @@ if __name__ == '__main__':
     # write summary of models
     write_summary_log(summary, model_dir)
 
-# python -m execute.main -window 10 -model lstm_cnn -mode classification -out_shape 2 -inner_lr 0.001 -outer_lr 0.0055 -bt_size 3 -rounds 3 -epochs 3 -ncpus 8
+# python -m execute.main -dataset USD_JPY -window 10 -model lstm_cnn -mode classification -out_shape 2 -inner_lr 0.001 -outer_lr 0.0055 -bt_size 3 -rounds 3 -epochs 3 -ncpus 8

### 2024-12-10 execute/submit_jobs.py

@@ -12,22 +12,21 @@ import itertools
 
 from common.constants import *
 
-def generate_script(device:str, name_job:str, look_back:int, model_name:str, mode:str, out_shape:int, inner_lr:float, outer_lr:float, bt_size:int, rounds:int, epochs:int):
-    ncores_limit = 64 if device=='cpu' else 26 if device=='gpu' else 0
+def generate_script(device:str, name_job:str, command:str):
     log_file_path = os.path.join(LOG_DIR, f"{name_job}.log")
-    command = f'python -m execute.main -window {look_back} -model {model_name} -mode {mode} -out_shape {out_shape} -inner_lr {inner_lr} -outer_lr {outer_lr} -bt_size {bt_size} -rounds {rounds} -epochs {epochs} -ncpus {ncores_limit} &>> {log_file_path}'
+    command = f'{command} &>> {log_file_path}'
     name_job = f'{name_job}_job'
 
     if device == 'cpu':
         class_name = 'DEFAULT'
         resources = 'select=1:ncpus=64'
-        timeout = 'walltime=24:00:00'
+        timeout = 'walltime=72:00:00'
         load_modules = ''
         test_device = f'echo "Using $(nproc) CPU cores" &>> {log_file_path}'
     elif device == 'gpu':
         class_name = 'GPU-1A'
         resources = 'select=1:ngpus=1'
-        timeout = 'walltime=24:00:00'
+        timeout = 'walltime=72:00:00'
         load_modules = 'module load cuda'
         test_device = f'nvidia-smi &>> {log_file_path}'
     else:
@@ -46,7 +45,7 @@ def generate_script(device:str, name_job:str, look_back:int, model_name:str, mod
 
         '\n# Commands to run your job',
 
-        f'echo "Starting {model_name} at `date`" &> {log_file_path}',
+        f'echo "Starting {name_job} at `date`" &> {log_file_path}',
         f'conda activate fx_env',
         f'conda env list &>> {log_file_path}',
         test_device,
@@ -57,36 +56,52 @@ def generate_script(device:str, name_job:str, look_back:int, model_name:str, mod
         fout.write("\n".join(tmp_file) + "\n")
 
 '''
-function finetune_3 fine-tunes all of meta-model (LSTM, LSTM+CNN) on all dataset
-    [v] multi-fx
[…]

### 2024-12-31 doc/thesis/chapters/1.intro.tex
thesis draft

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Chapter1: Introduction 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\chapter{Introduction}
\label{chap:intro}

\section{Background}

% viết từ time-series forecasting và các phương pháp tiếp cận (sơ bộ)
% Cùng với numerical, categorical, and text data, time-series data đang ngày càng trở nên phổ biến không chỉ ở nguồn sinh dữ liệu (e.g., lĩnh vực tài chính, năng lượng, vận tải,\dots) mà còn ở các phương pháp để phân tích và dự đoán chúng (e.g., phân rã tần số, các phương pháp dựa trên ghi nhớ dữ liệu quá khứ,\dots). Điều này là dễ hiểu vì phân tích và dự đoán time-series data mang lại những lợi ích lớn trong việc tổ chức, hoạch định chiến lược sản xuất và kinh doanh.

Time-series data is gaining popularity alongside numerical, categorical, and text data not only in terms of data sources (like finance, energy, and transportation), but also in terms of methods for analyzing and forecasting them (like frequency decomposition and methods based on historical data memorization). This makes sense as planning, organizing, and developing business strategies are greatly aided by the analysis and prediction of time-series data.

% Hai hướng tiếp cận chính được sử dụng trong phân tích và dự đoán time-series data là fundamental analysis và technical analysis \cite{ayitey2023forex}. Trong khi fundamental analysis thiên về phân tích các yếu tố tác động từ bên ngoài, vốn rất khó có thể capture từ các biến thiên giá trị trong quá khứ, như chính sách, chiến lược kinh tế của công ty, quốc gia để dự đoán tương lai; technical analysis dựa hoàn toàn vào lịch sử biến động giá trị để phân tích xu hướng tương lai.

The two main approaches used in analyzing and predicting time-series data are fundamental analysis and technical analysis \cite{ayitey2023forex}. While fundamental analysis concentrates on examining external factors that are hard to capture from past price changes, such as the economic strategies and policies of nations and businesses, technical analysis relies entirely on historical price changes to forecast future trends

% Chính vì tính phi cấu trúc trong các thông tin của tin tức, đạt được sự tự động hóa hiệu quả trong các phương pháp fundamental analysis là rất khó. Do đó, các nhà nghiên cứu thường tập trung phát triển các phương pháp technical analysis. Cùng với sự phát triển của machine learning và deep learning, các phương pháp trải dài từ Auto-regressive model (năm 1951) \cite{moran1951hypothesis}, Moving average model (năm 2000) \cite{rosenblatt2000gaussian}, đến Long Short-term memory model - \verb|LSTM| (năm 1997) \cite{hochreiter1997long} và Transformers \cite{vaswani2017attention} (năm 2017), được phát triển nhằm tự động hóa một cách thông minh quá trình dự báo.

It is challenging to achieve efficient automation of basic analytical methods due to the unstructured nature of news data. As a result, academics frequently concentrate on creating techniques for technical analysis. The forecasting process can now be intelligently automated with the help of machine learning and deep learning techniques, such as Auto-regressive model (1951) \cite{moran1951hypothesis}, Moving average model (2000) \cite{rosenblatt2000gaussian}, Long short-term memory model - \verb|LSTM| (1997) \cite{hochreiter1997long}, and Transformers (2017) \cite{vaswani2017attention}.

% thu hẹp lại aperiodic time-series data (phân tích 3 thách thức chính)
% Các phương pháp nêu trên và các biến thể chủ yếu hoạt động trên periodic time-series data (e.g., thời tiết, giao thông,...). Loại dữ liệu này thể hiện rõ tính chu kỳ, hoặc tính mùa vụ theo thời gian. Do đó, việc phân tích và dự đoán được diễn ra rất dễ dàng với độ chính xác cao. Điều này càng đúng hơn với các phương pháp hướng đến phân rã chu kỳ bằng các deep neural network hiện nay \cite{liu2022scinet,chen2021autoformer,zhou2022fedformer}. Tuy nhiên, rất ít nghiên cứu chú trọng đến \textbf{aperiodic time-series data} (e.g., foreign exchange rate, stock price,...).

The aforementioned techniques, together with their variations, are mostly applicable to periodic time-series data, such as traffic, weather, etc. Over time, this kind of data clearly shows a seasonal or cyclical character. As a result, analysis and prediction may be done quickly and accurately. This is especially true for existing techniques that use deep neural networks to try to decompose periods \cite{liu2022scinet,chen2021autoformer,zhou2022fedformer}. Nevertheless, \textbf{aperiodic time-series data} (e.g., stock prices, foreign exchange rates, etc.) has been the subject of very few investigations.

% Điểm khác biệt rất dễ nhận thấy ở aperiodic time-series data so với periodic time-series data nằm ở chỗ chúng không thể hiện một chu kỳ rõ ràng (see figure \ref{fig:ill_aperiodic}). Thật vậy, trong hình \ref{fig:periodic}, thuộc tính \textit{Temperature oil} (target variable) của tập dữ liệu Electricity Transformer Temperature (\verb|ETT-m2|) \cite{zhou2021informer} thể hiện tính chu kỳ rất rõ ràng theo thời gian vì nhu cầu sử dụng điện biến thiên theo mùa. Trong khi đó, \textit{Close price} between US dollar and Japanese yen (figure \ref{fig:aperiodic}) bị chi phối bởi rất nhiều yếu tố. Một tin tức mới về tình hình kinh tế cũng có thể gây ra tác động lớn lên tỷ giá này. Do đó, tỷ giá này không thể hiện bất kỳ tính chu kỳ nào.

As can be seen in figure \ref{fig:ill_aperiodic}, the primary distinction between aperiodic and periodic time-series data is their lack of a distinct periodicity. In fact, because of seasonal fluctuations in electricity demand, the \textit{Temperature oil} (target variable) attribute of the Electricity Transformer Temperature dataset (\verb|ETT-m2|) \cite{zhou2021informer} shows a very noticeable periodicity over time in figure \ref{fig:periodic}. In the meanwhile, a number of external factors influence the \textit{Close price} between the US dollar and the Japanese yen (figure \ref{fig:aperiodic}). This exchange rate can be significantly impacted by recent economic news. As a result, there is no periodicity in this exchange rate.

\begin{figure}[H]
    \centering
    \begin{subfigure}[b]{0.5\textwidth}
        \centering
        \includegraphics[width=\textwidth]{usd_close.pdf}
        \caption{Exchange ratio between USD and JPY.}
        \label{fig:aperiodic}
    \end{subfigure}%
    ~
    \begin{subfigure}[b]{0.5\textwidth}
        \centering
        \includegraphics[width=\textwidth]{ett_OT.pdf}
        \cprotect\caption{\textit{Oil temperature} in \verb|ETT-m2| dataset.}
        \label{fig:periodic}
    \end{subfigure}

    \caption{70.000 samples of aperiodic (\ref{fig:aperiodic}) and periodic (\ref{fig:periodic}) time-series dataset.}
    \label{fig:ill_aperiodic}
\end{figure}

% Đặc điểm chính của aperiodic time-series data nằm ở chỗ dữ liệu này bị chi phối rất nhiều bởi các yếu tố bên ngoài như chiến lược sản xuất, chính sách kinh tế, chính trị của các công ty hay quốc gia. Chính điều này đã gây ra sự thay đổi đáng kể trong tính chất dữ liệu, hay còn biết đến là concept drift scenarios \cite{liu2025onsitnet}. Về mặt toán học, điều này khiến cho phương sai của aperiodic time-series data biến thiên liên tục với khoảng biến thiên rất rộng. Cũng chính điều này đã khiến cho việc rút trích đặc trưng trên dữ liệu thông qua một cửa sổ trượt gặp rất nhiều khó khăn do dữ liệu đã không còn thể hiện tính chu kỳ.

Aperiodic time-series data's primary feature is that it is heavily impacted by external factors, including production strategy and the political and economic policies of nations or businesses. This has resulted in notable modifications to the data's characteristics, which are also referred to as concept drift scenarios \cite{liu2025onsitnet}. Mathematically, this results in aperiodic time-series data with a continuous and incredibly wide variance. Additionally, because the data no longer exhibits periodicity, it becomes extremely challenging to extract characteristics from it using a sliding window.

% Tóm lại, làm việc trên aperiodic time-series data đối mặt với ba thách thức chính: (1) Dữ liệu bị ảnh hưởng rất nhiều bởi các yếu tố kinh tế, chính trị, xã hội; (2) Dữ liệu có phương sai lớn và không cố định; (3) Khó rút trích đặc trưng vì dữ liệu thể hiện tính phi chu kỳ rất mạnh. \textbf{Trong nghiên cứu này, chúng tôi tập trung giải quyết ba thách thức vừa nêu trong bài toán dự đoán xu hướng (tăng hoặc giảm) trên aperiodic time-series data.}

In conclusion, there are three primary obstacles to overcome when dealing with aperiodic time-series data: (1) External influences including politics, economy, and society have a significant impact on the data; (2) Data has a massive and non-stationary variance; and (3) Data's strong non-periodicity makes feature extraction challenging. \textbf{In this study, we focus on solving the three challenges mentioned above in the problem of trend prediction (upward or downward) on aperiodic time-series data.}

\section{Motivation}
% cần paraphrase lại

% viết về việc 3 thách thức chính được giải quyết nhưng chưa hiệu quả
% Trong ba thách thức nêu trên, thách thức đầu tiên là thách thức khó giải quyết nhất dựa trên các phương pháp technical analysis vì các thông tin về chính trị, xã hội rất khó để rút trích thông qua dữ liệu lịch sử của một tập dữ liệu. Tuy nhiên, theo efficient market hypothesis \cite{fama1970efficient}, chính các thông số trong dữ liệu lịch sử lại phản ánh rất rõ các thông tin về sự biến động của thị trường. Điều này nghĩa là, \textbf{bằng việc xem xét kỹ lưỡng lịch sử giao dịch, người ta hoàn toàn có thể nắm bắt được các biến động thị trường và dự báo được tương lai} mà không cần sử dụng đến các thông tin về chính trị, xã hội. Các nghiên cứu \cite{overreactioncontrarian, mech1993portfolio} cũng chỉ ra sự phụ thuộc giữa chỉ số tài chính của một công ty nhất định và chỉ số của các công ty khác. Nói cách khác, \textbf{việc tích hợp phân tích nhiều nguồn thông tin khác nhau thuộc cùng một lĩnh vực có thể mang lại những thông tin phân tích đáng giá cho các chỉ số đang quan tâm.}

In three aforementioned challenges, the first challenge is the most difficult to solve based on technical analysis methods because political and social information is difficult to extract through historical data of a data set. However, according to the efficient market hypothesis \cite{fama1970efficient}, historical data itself clearly reflects information about market fluctuations. That means, \textbf{by carefully examining the transaction history, one can completely grasp market fluctuations and forecast the future} without using political and social information. In addition, \cite{overreactioncontrarian, mech1993portfolio} studies also point out the dependence between the financial indicators of a certain company and the indicators of other companies. In other words, \textbf{integrating analysis of multiple sources of information within the same domain can yield valuable analytical insights for the indicator of interest.}

% Khi đối mặt với thách thức thứ hai, các mô hình ensemble thường được sử dụng để hạn chế ảnh hưởng của sự biến đổi variance \cite{ali2020complete, zafeiriou2020intraday, sadeghi2021combined}. Về cơ bản, các mô hình ensemble hoạt động dựa trên việc chia nhỏ dữ liệu thành nhiều phần và giả định một mức phương sai cố định trên các thành phần này sau đó mô phỏng lại sự biến thiên đó. Hướng tiếp cận này là hoàn toàn hợp lý vì một mô hình khó có thể nắm bắt được toàn bộ sự biến thiên của dữ liệu xuyên suốt một khoảng thời gian dài. \textbf{Bằng việc phân tích và tổng hợp thông tin từ nhiều sub-model, ensemble learning cung cấp một cái nhìn đa chiều về dữ liệu, giúp mô hình tổng quát thích ứng tốt với sự thay đổi mạnh mẽ của phương sai.}

When faced with the second challenge, ensemble learning is often used to mitigate the effects of variance \cite{ali2020complete, zafeiriou2020intraday, sadeghi2021combined}. Basically, ensemble learning works by splitting the data into multiple parts and assuming a fixed level of variance across these parts, then simulating that variation. This approach is reasonable because it is difficult for a single model to capture all the variability in the data over a long period of time. \textbf{By analyzing and synthesizing information from multiple sub-models, ensemble learning provides a multidimensional view of the data, which helps the overall model adapt well to strong variance changes.}

% Cho đến nay, rất nhiều mô hình được đề xuất để có thể rút trích đặc trưng trên dữ liệu nói chung và time-series data nói riêng. Điển hình trong số đó, Convolutional neural network - \verb|CNN| \cite{lecun1989handwritten}, \verb|LSTM| \cite{hochreiter1997long}, và cơ chế attention \cite{bahdanau2014neural, luong2015effective, vaswani2017attention} lần lượt được đề xuất để rút trích các đặc trưng cục bộ, ghi nhớ các đặc trưng ngắn-dài hạn, và nhấn mạnh các vector trong ma trận đặc trưng. \textbf{Bằng việc sử dụng hiệu quả các phương pháp này, các ràng buộc ẩn trong aperiodic time-series data có thể được rút trích một cách hiệu quả}, giúp vượt qua thách thức thứ ba.

Up to now, many models have been proposed to be able to extract features on data in general and time-series data in particular. Typically, Convolutional neural network - \verb|CNN| \cite{lecun1989handwritten}, \verb|LSTM| \cite{hochreiter1997long}, and attention mechanism \cite{bahdanau2014neural, luong2015effective, vaswani2017attention} have been proposed to extract local features, remember short-term and long-term features, and emphasize vectors in the feature matrix, respectively. \textbf{By selectively using these methods, hidden constraints in aperiodic time-series data can be extracted effectively}, helping to overcome the third challenge.

% tôi ở đây để đề xuất và giải quyết 3 thách thức này bằng meta-learning, LSTM (gáy các đặc điểm của 2 thằng này ra)
% Mặt khác, các thuật toán Meta-learning (ML) được biết đến với khả năng tăng tính tổng quát cũng như khả năng tương thích của một mô hình trên một tập dữ liệu hữu hạn \cite{hospedales2021meta, vettoruzzo2024advances}. Trong khi huấn luyện, quá trình tối ưu hóa của ML có thể coi là một quá trình tổng hợp hiệu quả các mô hình trên các sub-tasks (các sub-datasets). Dựa trên khả năng này, \textbf{các mô hình máy học được huấn luyện bằng ML được kỳ vọng là sẽ tổng hợp hiệu quả thông tin từ nhiều nguồn dữ liệu/khoảng thời gian cũng như thích ứng tốt với sự thay đổi liên tục của phương sai.}

On the other hand, Meta-learning (ML) algorithms are known for their ability to increase the generalization and adaptability of a model on a limited dataset \cite{hospedales2021meta, vettoruzzo2024advances}. During training, the optimization process of ML can be viewed as an efficient synthesis of models across sub-tasks (sub-datasets). Based on this ability, \textbf{a machine learning models trained by ML algorithms are expected to efficiently synthesize information from multiple data sources/time periods as well as adapt well to the continuous change of variance.}

\section{Outline}

This work is divided into 6 chapters:

\begin{itemize}
    \item Chapter \ref{chap:intro} provides the research context, problem statement, challenges and motivation.
    \item Chapter \ref{chap:related_work} presents related works, advantages and disadvantages of each study. The mentioned studies belong to two popular approaches in analyzing time-series data: Deep feature-based approach and Frequency decomposition approach. In addition, Chapter \ref{chap:related_work} also presents ML, providing the foundation for the proposed method in Chapter \ref{chap:method}.
    \item Chapter \ref{chap:method} presents a new approach for aperiodic time-series data based on the \verb|LSTM|/\verb|LSTM+CNN| feature extraction method and ML model optimization method.
    \item Chapter \ref{chap:experiment} sets up experiments, evaluation methods for the proposed method and baseline model.
    \item Chapter \ref{chap:result} analyzes the results achieved during the experiment.
    \item Chapter \ref{chap:conclusion} summarizes the main contributions of the study as well as presents future development directions.
\end{itemize}

### 2024-12-31 doc/thesis/chapters/2.related_works.tex
thesis draft

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Chapter2: Related Work
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\chapter{Related Works}
\label{chap:related_work}

% % tôi sẽ chuyển cái này thành 1 idea để viết phương pháp
% % kiểu: Lấy ý tưởng từ ensemble model này kia
% \section{Ensemble models}

\section{Deep feature-based approach}

\subsection{Convolutional neural network}

\subsection{Recurrent Neural Network}

\subsection{Long Short-Term Memory}

\section{Frequency decomposition approach}

\subsection{A transformer-based method}

\subsection{NHITS}

% Neural Hierarchical Interpolation for Time Series Forecasting (\verb|NHITS|) \cite{challu2023nhits} được thiết kế để hướng đến việc dự đoán một long-horizon time-series data bằng cách phân rã tín hiệu đầu vào thành các dải tần số riêng biệt. Cấu trúc của \verb|NHITS| bao gồm $S$ stack, mỗi stack bao gồm $B$ block nối tiếp nhau. Tại mỗi block, một Multi-layer perceptron (\verb|MLP|) sử dụng dữ liệu lịch sử để dự đoán chính nó và dữ liệu tương lai (see figure \ref{fig:nhits}).

Neural Hierarchical Interpolation for Time Series Forecasting (\verb|NHITS|) \cite{challu2023nhits} is designed to target the prediction of long-horizon time-series data by decomposing the input signal into discrete frequency bands. The structure of \verb|NHITS| consists of $S$ stacks, each stack consisting of $B$ consecutive blocks. At each block, a Multi-layer perceptron (\verb|MLP|) uses historical data to predict itself and future data (see figure \ref{fig:nhits}).

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{nhits-arch.png}
    \cprotect\caption{\verb|NHITS| architecture \cite{challu2023nhits}.}
    \label{fig:nhits}
\end{figure}

% Cụ thể, tại block $l$, với $L$ mẫu dữ liệu quá khứ ($\mathbf{y}_{t-L:t, l-1}$), một \verb|MLP| sẽ sử dụng lần lượt ba kỹ thuật Multi-rate signal sampling, Non-linear regression, và Hierarchical interpolation để hồi quy dữ liệu quá khứ và dự đoán dữ liệu tương lai.

Specifically, at block $l$, with $L$ historical samples ($\mathbf{y}_{t-L:t, l-1}$), an \verb|MLP| use three techniques, Multi-rate signal sampling, Non-linear regression, and Hierarchical interpolation, respectively, to regress past data and predict future data.

% \textbf{Multi-rate signal sampling}. \verb|MaxPool| layer với kernel size $l$ nén dữ liệu ban đầu vào $\mathbf{y}_{t-L:t, l}^{(p)}$ (equation \ref{eq:maxPool}). Khi $k_l$ lớn, lượng dữ liệu được xét trong một cửa sổ trượt lớn, mạng sẽ chú ý đến các sóng tín hiệu với bước sóng dài (tần số thấp). Khi $k_l$ nhỏ, các đặc trưng thu được là của các bước sóng nhỏ (tần số cao). Bằng việc sử dụng lớp \verb|MaxPool|, các \verb|MLP| sẽ có khả năng làm việc trên một dải tần nhất định, giúp tăng hiệu quả của việc phân rã tần số. Không chỉ dừng lại ở việc phân tách dải tần, \verb|MaxPool| về cơ bản làm giảm kích thước của tín hiệu đầu vào, giúp tiết kiệm bộ nhớ trong quá trình training và inference.

\textbf{Multi-rate signal sampling}. \verb|MaxPool| layer with kernel size $l$ compresses the original data into $\mathbf{y}_{t-L:t, l}^{(p)}$ (equation \ref{eq:maxPool}). When $k_l$ is large, the amount of data considered in a sliding window is also large, the network will pay more attention to input signal with long wavelengths (low frequencies). When $k_l$ is small, the obtained features are of short wavelengths (high frequencies). By using \verb|MaxPool| layer, \verb|MLP| will be able to work on a certain frequency band, which helps to increase the efficiency of frequency decomposition. Not only stopping at frequency band decomposition, \verb|MaxPool| essentially reduces the size of the input signal, helping to save memory during training and inference.

\begin{align}
    \mathbf{y}_{t-L:t, l}^{(p)} = \mathbf{MaxPool}\left( \mathbf{y}_{t-L:t, l-1}, k_l \right)
    \label{eq:maxPool}
\end{align}

% \textbf{Non-linear regression}. Sau khi nén dữ liệu, \verb|NHITS| tiến hành học các hệ số nội suy bằng các perceptron layer xếp chồng với hàm kích hoạt phi tuyến (\verb|FullyConnected|). Mục tiêu của \verb|FullyConnected| là tạo ra các vector $\mathbf{\theta}_l^f, \mathbf{\theta}_l^b$ (equation \ref{eq:non_linear}). Đây là hai vector nội suy forecast và backcast, được dùng để tổng hợp các giá trị đầu ra của block $l$ bằng hàm nội suy $g(\cdot)$. Trong đó, $\mathbf{\theta}_l^f$ được dùng để dự báo các giá trị tương lai còn $\mathbf{\theta}_l^b$ được dùng để hồi quy các giá trị input.

\textbf{Non-linear regression}. After compressing the data, \verb|NHITS| learns the interpolation coefficients using stacked perceptron layers with a nonlinear activation function (\verb|FullyConnected|). The goal of \verb|FullyConnected| is to generate the vectors $\mathbf{\theta}_l^f, \mathbf{\theta}_l^b$ (equation \ref{eq:non_linear}). These are the two interpolation vectors forecast and backcast, which are used to aggregate the output values of block $l$ using the interpolation function $g(\cdot)$. In which, $\mathbf{\theta}_l^f$ is used to forecast future values and $\mathbf{\theta}_l^b$ is used to regress the input values.

\begin{align}
    \mathbf{\theta}_l^b &= \mathbf{FullyConnected}^b \left( \mathbf{y}_{t-L:t, l}^{(p)} \right)\\
    \mathbf{\theta}_l^f &= \mathbf{FullyConnected}^f \left( \mathbf{y}_{t-L:t, l}^{(p)} \right)\\
    \label{eq:non_linear}
\end{align}

% \textbf{Hierarchical interpolation}. Để dự đoán $H$ giá trị tương lai, một mạng neural thông thường phải thiết kế lại số neuron đầu ra. Đôi với mô hình transformer, muốn tăng số mẫu đầu ra, cần phải tăng số mẫu đầu vào để các layer cross-attention hoạt động hiệu quả. Điều này khiến cho quá trình huấn luyện truyền thống tiêu tốn rất nhiều tài nguyên khi có nhu cầu dự đoán $H$ lớn. \verb|NHITS| giải quyết vấn đề này bằng cách sử dụng phương trình nội suy với các hệ số nội suy đã được chuẩn bị trước ở bước \textbf{Non-linear regression} (equation \ref{eq:interpolation}). Số chiều của các hệ số nội suy trong mỗi stack được quy định bởi expressiveness ratio $r_l$: $\left| \theta_l \right| =  \left \lceil r_l H \right \rceil$. Thông thường, expressiveness ratio sẽ rất nhỏ ở các stack đầu tiên và tăng dần về cuối. Theo đó, các stack có thể mô phỏng lại các tần số từ thấp đến cao. Ngoài ra, bằng việc sử dụng hàm nội suy, \verb|NHITS| không cần quá nhiều phần cứng để huấn luyện mạng neural trong trường hợp $H$ lớn.

\textbf{Hierarchical interpolation}. To forecast $H$ future values, a conventional neural network must redesign the number of output neurons. For transformer models, to increase the number of output samples, it is necessary to increase the number of input samples so that the cross-attention layers can work effectively. This makes the traditional training process very resource-consuming when there is a need to predict large $H$. \verb|NHITS| solves this problem by using an interpolation equation with pre-prepared interpolation coefficients in the \textbf{Non-linear regression} step (equation \ref{eq:interpolation}). The dimensionality of the interpolation coefficients in each stack is determined by the expressiveness ratio $r_l$: $\left| \theta_l \right| = \left \lceil r_l H \right \rceil$. Typically, the expressiveness ratio will be very small in the first stacks and gradually increase towards the end. Accordingly, the stacks can simulate frequencies from low to high. In addition, by using the interpolation function, \verb|NHITS| does not require too much hardware to train the neural network in the case of large $H$.

\begin{align}
    \mathbf{\hat{y}}_{t-L:t, l} &= g\left(\mathbf{\theta}_l^b\right)\\
    \mathbf{\hat{y}}_{t+1:t+H, l} &= g\left(\mathbf{\theta}_l^f\right)
    \label{eq:interpolation}
\end{align}

% Đầu ra của block $l$ là giá trị forecast $\mathbf{\hat{y}}_{t+1:t+H, l}$ và giá trị backcast $\mathbf{\hat{y}}_{t-L:t, l}$. Input of block $l+1$ is the difference between the its backcast and its output (\ref{eq:input_l1}).

The output of block $l$ is the forecast value $\mathbf{\hat{y}}_{t+1:t+H, l}$ and the backcast value $\mathbf{\hat{y}}_{ t-L:t, l}$. Input of block $l+1$ is the difference between its backcast and its output (\ref{eq:input_l1}).

\begin{align}
    \mathbf{y}_{t-L:t, l+1} = \mathbf{y}_{t-L:t, l-1} - \mathbf{\hat{y}}_{t-L:t, l}
    \label{eq:input_l1}
\end{align}

% Tổng hợp các giá trị forecast của $B$ block như phương trình \ref{eq:sum_block}, ta được giá trị forecast của một stack. Cuối cùng, tổng hợp giá trị forecast của các stack như phương trình \ref{eq:sum_stack}, ta được giá trị forecast dự đoán của toàn mạng.

Summing the forecast values of $B$ blocks as in equation \ref{eq:sum_block}, we get the forecast value of a stack. Finally, summing the forecast values of the stacks as in equation \ref{eq:sum_stack}, we get the predicted forecast value of the entire network.

\begin{align}
    \mathbf{\hat{y}}_{t+1:t+H}^s &= \sum_{l=1}^{B}{\mathbf{\hat{y}}_{t+1:t+H, l}} \label{eq:sum_block}\\
    \mathbf{\hat{y}}_{t+1:t+H} &= \sum_{s=1}^{S}{\mathbf{\hat{y}}_{t+1:t+H}^s} \label{eq:sum_stack}
\end{align}

% Bằng cách xếp chồng các stack, stack sau nhận vào phần dư của stack trước, kiến trúc trên được kỳ vọng là sẽ phân rã dữ liệu thành các frequency bands khác nhau (weekly, daily, even hourly). Trong thực tế, \verb|NHITS| perform rất tốt đối với các bộ dữ liệu có tính chu kỳ cao như mức tiêu thụ điện, thời tiết, giao thông. Tuy nhiên, chúng tôi đang hướng đến aperiodic time-series dataset, vốn có tính chu kỳ rất thấp. Điều này gây ra khó khăn rất lớn cho \verb|NHITS|.

By concatenating stacks, each receiving the remainder of the previous stack, the above architecture is expected to decompose the data into different frequency bands (weekly, daily, even hourly). In practice, \verb|NHITS| performs very well for highly periodic datasets such as electricity consumption, weather, traffic. However, we are aiming for an aperiodic time-series dataset, which has very low, or even non-existent, periodicity. This poses a huge challenge for \verb|NHITS|.

\section{Optimization-based Meta-learning}
\label{sec:ml}

% Meta-learning (ML) là một phương pháp huấn luyện cho phép mô hình học có thể gia tăng kinh nghiệm qua việc thực hiện nhiều task khác nhau trong cùng một phân phối task. Việc này trạng bị cho mô hình học máy khả năng tổng quát hóa cao, thích ứng nhanh trên task mới chỉ sau một vài bước huấn luyện với dữ liệu huấn luyện giới hạn \cite{hospedales2021meta, vettoruzzo2024advances}. Với khả năng này, ML được sử dụng rất nhiều trong các tác vụ đòi hỏi khả năng đáp ứng của mô hình trên dữ liệu (e.g. cá nhân hóa mô hình học \cite{chen2018federated, fallah2020personalized,nguyen2022meta}, domain adaptation trong online learning \cite{hu2023meta, khoee2024domain}).

Meta-learning (ML) is a training method that allows a learning model to gain experience by performing many different tasks in the same task distribution. This equips the machine learning model with the ability to generalize highly and adapt quickly to new tasks after only a few training steps with limited training data \cite{hospedales2021meta, vettoruzzo2024advances}. With this ability, ML is widely used in tasks that require the ability to fast adapt to new data (e.g. personalization of learning models \cite{chen2018federated, fallah2020personalized,nguyen2022meta}, domain adaptation in online learning \cite{hu2023meta, khoe2024domain}).

% Đối với phương pháp huấn luyện mô hình học truyền thống, chúng ta huấn luyện mô hình dự đoán $\hat{y} = f_\theta(\mathbf{x})$ trên tập dữ liệu $\mathcal{D}_t = \left\{ \left(\mathbf{x}_i, y_i \right) \right\}$ của task $t$. Ký hiệu $\mathcal{L}$ là hàm lỗi, $\phi$ là prior knowledge, mục tiêu của quá trình huấn luyện là tối thiểu hóa hàm lỗi trên tập dữ liệu $\mathcal{D}$ bằng cách tìm tham số $\theta$ thỏa mãn:

For the traditional learning model training method, we train the prediction model $\hat{y} = f_\theta(\mathbf{x})$ on the dataset $\mathcal{D}_t = \left\{ \left(\mathbf{x}_i, y_i \right) \right\}$ of task $t$. Denote $\mathcal{L}$ is the error function, $\phi$ is the prior knowledge, the goal of the training process is to minimize the error function on the dataset $\mathcal{D}$ by finding the parameter $\theta$ that satisfies:

\begin{equation}
    \theta^* = \arg\min_{\theta}{\mathcal{L}(\mathcal{D}_t; \theta, \phi)}
\end{equation}

% Hướng tiếp cận của ML nằm ở chỗ cố gắng học một prior knowledge $\phi$ thật tốt. Điều này đạt được thông qua việc học một phân phối task $\mathcal{T}$ \cite{hospedales2021meta}. Sau khi học được một prior knowledge tốt, có thể sử dụng prior knowledge này cho các task mới trong cùng phân phối task $\mathcal{T}$. Về mặt toán học, ký hiệu $\mathcal{L}(\mathcal{D}_t, \phi)$ là hàm lỗi biểu diễn sự hiệu quả của việc sử dụng $\phi$ trong huấn luyện mô hình trên task $T$, mục tiêu của ML được biểu diễn như sau:

The ML approach is to try to learn a good prior knowledge $\phi$. This is achieved by learning a task distribution $\mathcal{T}$ \cite{hospedales2021meta}. Once a good prior knowledge is learned, it can be used for new tasks in the same task distribution $\mathcal{T}$. Mathematically, denote $\mathcal{L}(\mathcal{D}_t, \phi)$ is the error function that represents the effectiveness of using $\phi$ in training the model on task $T$, the ML goal is expressed as follows:

\begin{equation}
    \min_{\phi} \mathop{\mathbb{E}}_{t\sim \mathcal{T}} \mathcal{L}(\mathcal{D}_t, \phi)
\end{equation}

\subsection{Model-Agnostic Meta-Learning (MAML)}

% Đối với hướng tiếp cận dựa trên tối ưu, một thuật toán ML cơ bản, điển hình là \verb|MAML|, sẽ được học trên nhiều tác vụ $t$ rút ra từ cùng một phân phối tác vụ $\mathcal{T}$ \cite{hospedales2021meta}. Dữ liệu của mỗi tác vụ được chia thành tập support $\mathcal{D}_t^{support}$ (thường có kích thước nhỏ, khoảng 20\%) và tập query $\mathcal{D}_t^{query}$. Trong qua trình học, inner và outer optimization được perform đan xen. Trong đó, mục tiêu của inner optimization là cố gắng giải quyết task $t$ bằng cách tìm ra một bộ tham số tối ưu $\theta_t^*$ trên tập support thông qua $\phi$:

For the optimization-based approach, a basic ML algorithm, typically \verb|MAML|, is learned on multiple tasks $t$ drawn from the same task distribution $\mathcal{T}$ \cite{hospedales2021meta}. The data for each task is divided into a support set $\mathcal{D}_t^{support}$ (usually small, around 20\%) and a query set $\mathcal{D}_t^{query}$. During the learning process, inner and outer optimization are performed alternately. The goal of inner optimization is to attempt to solve task $t$ by finding an optimal set of parameters $\theta_t^*$ on the support set via $\phi$:

\begin{equation}
    \theta_t^* = \theta_t(\phi) = \arg\min_{\theta}{\mathcal{L}^{task}_t\left( \phi, \mathcal{D}_t^{support} \right)}
    \label{eq:inner_opt}
\end{equation} Where, $\phi$ is the result of the outer optimization process, which acts as the initial value of $\theta_t$. $\mathcal{L}^{task}_t$ is the error function of the model on the support set of task $t$.

% Mục tiêu của outer optimization là tìm ra prior knowledge tối ưu $\phi^*$, giúp cho việc học một task mới trong phân phối $\mathcal{T}$ trở nên nhanh chóng và hiệu quả. Cụ thể, thuật toán sử dụng các bộ tham số tối ưu $\theta_t^*$ để perform trên tập query tương ứng. Lỗi của toàn bộ mô hình sau đó được tổng hợp để thực hiện quá trình outer optimization:

The goal of outer optimization is to find the optimal prior knowledge $\phi^*$, which makes learning a new task in the distribution $\mathcal{T}$ fast and efficient. Specifically, the algorithm uses the optimal parameter sets $\theta_t^*$ to perform on the corresponding query set. The errors of the entire model are then aggregated to perform the outer optimization process:

\begin{align*}
    \phi^* &= \arg\min_{\phi}\sum_{t}{\mathcal{L}^{meta}_t\left( \theta_t^*, \mathcal{D}_t^{query} \right)}\\
    &= \arg\min_{\phi}\sum_{t}{\mathcal{L}^{meta}_t\left( \theta_t(\phi), \mathcal{D}_t^{query} \right)} \numberthis
    \label{eq:outer_opt}
\end{align*}

% Bằng hình thức huấn luyện trên, mô hình $\phi^*$ sẽ có mức tổng quát hóa cao trên các tác vụ khác nhau, có thể nhanh chóng đáp ứng một tác vụ mới chỉ sau một vài bước huấn luyện.

By performing the above training method, the $\phi^*$ model will have a high level of generalization across different tasks, and can quickly respond to a new task after only a few training steps.

% Trong inference phase, giá trị khởi tạo cho tham số của mô hình được gán bằng $\phi^*$. Mô hình sau đó được huấn luyện nhanh trên tập support sau đó perform trên tập query. Kết quả trên tập query chính là kết quả của mô hình.

In the inference phase, the initial values for the model parameters are assigned $\phi^*$. The model is then adapted quickly to the support set and performed on the query set. The results on the query set are the model output.

\subsection{Meta-SGD}

% Về phần thuật toán \verb|Meta-SGD|, nghiên cứu \cite{li2017meta} chỉ ra rằng việc sử dụng một siêu tham số học cục bộ nhỏ, được cố định theo thời gian hoặc một một siêu tham số cục bộ giảm dần theo thời gian chỉ phù hợp cho ngữ cảnh huấn luyện một mô hình với bộ dữ liệu lớn trong thời gian dài. Trong ngữ cảnh dữ liệu gắn nhãn có ít nhưng mô hình cần phải thích ứng nhanh với tập dữ liệu mới, phương pháp chọn siêu tham số như vậy không còn phù hợp nữa.

As for the \verb|Meta-SGD| algorithm, study \cite{li2017meta} shows that using a small, fixed local learning rate over time or a decreasing local learning rate over time is only suitable for the context of training a model with a large dataset over a long period of time. In the context of limited labeled data but the model needs to adapt quickly to new data sets, such hyperparameter selection method is no longer suitable.

% Nghiên cứu \cite{li2017meta} cũng đề ra một hướng tiếp cận mới cho phép tự điều chỉnh và tối ưu siêu tham số cục bộ. Theo đó, ngoài việc tối ưu prior knowledge ($\omega$), thuật toán coi inner learning rate $\alpha$ cũng là một tham số có thể tối ưu. Bằng việc khởi tạo $\alpha$ là một ma trận có kích thước giống như $\theta$, thuật toán hướng tới việc cập nhật cả hướng đi lẫn learning rate cho từng phần tử trọng số trong $\theta$ bằng cách điều chỉnh $\alpha$ tại outer optimization. Các task sau đó sử dụng $\alpha$ bằng cách nhân vô hướng đại lượng này với đạo hàm hàm lỗi cục bộ. Theo đó, \verb|Meta-SGD| vừa khiến cho các mô hình học thích ứng nhanh trên các tập dữ liệu cục bộ, vừa đóng góp lớn vào việc cá nhân hóa mô hình học cho từng task.

The \cite{li2017meta} study also proposes a new approach that allows self-tuning and local hyperparameter optimization. Accordingly, in addition to optimizing prior knowledge ($\omega$), the algorithm also considers the inner learning rate $\alpha$ as an learnable parameter. By initializing $\alpha$ as a matrix of the same size as $\theta$, the algorithm aims to update both the direction and the learning rate for each weight element in $\theta$ by adjusting $\alpha$ at the outer optimization. Subsequently, tasks use $\alpha$ by multiplying (element-wise product) this quantity by the local error function derivative. Accordingly, \verb|Meta-SGD| not only makes learning models adapt quickly on local data sets, but also contributes greatly to personalizing the learning model for each task.

One disadvantage of optimization-based ML method lies in solving equation \ref{eq:outer_opt} which requires massive overhead to compute and maintain a Hessian matrix. Even so, ML algorithms still achieve a high accuracy in handling many problem in which the quick adaptation or effective model synthesis are required.

### 2024-12-31 doc/thesis/chapters/3.method.tex
thesis draft

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Chapter 3: Method
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\chapter{Methodology}
\label{chap:method}

% Phương pháp của chúng tôi bao gồm hai phần chính hoạt động song song nhau: (1) - Feature extraction; (2) - Model synthesis. Tổng quan phương pháp được minh họa trong hình \ref{fig:flow}. Trong phần feature extraction, chúng tôi kết hợp hai loại đặc trưng từ mạng \verb|LSTM| và \verb|LSTM+CNN|. Trong phần Model synthesis, chúng tôi sử dụng \verb|MAML| để tổng hợp tham số của các mô hình. Với sự góp mặt của các đặc trưng từ các mạng học sâu, chúng tôi kỳ vọng sẽ rút trích được các đặc trưng ẩn trong dữ liệu aperiodic. Thông qua sự kết hợp giữa quá trình tổng hợp mô hình bằng thuật toán \verb|MAML| và quá trình rút trích đặc trưng bằng \verb|LSTM|/\verb|LSTM+CNN|, phương pháp đề xuất được kỳ vọng là một giải pháp thay thế hợp lý và hiệu quả cho các mô hình ensemble truyền thống trong việc tổng hợp hiệu quả external factors, giảm thiểu tác động của sự biến thiên phương sai, cũng như rút trích được các hidden long-term dependency trong quá khứ.

We propose \verb|Temporal-ML|, a ML-based method which consists of two main parts that work in parallel: (1) - Temporal feature extraction; (2) - Models' parameter synthesis. The overview of the method is illustrated in figure \ref{fig:flow} and the detail is presented in algorithm \ref{alg:maml}. In the \textit{Temporal feature extraction} section, we use two types of features from \verb|LSTM| and \verb|LSTM+CNN| networks. In the \textit{Effective synthesis of model's parameters} section, we use \verb|MAML| to synthesize the parameters of the models.

Due to the contribution of \verb|LSTM| and \verb|LSTM+CNN| features, we expect to effectively extract hidden features from aperiodic data. By using \verb|MAML| in the weight aggregation process, the proposed method is expected to be a reasonable and effective alternative to traditional ensemble models in effectively synthesizing external factors, minimizing the impact of variance variation, and efficiency capturing hidden long-term dependencies in the past.

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{meta.png}
    \caption{The full-flow of meta-training and meta-testing on multi-fx data. Each currency pair is regarded as a task.}
    \label{fig:flow}
\end{figure}

\begin{algorithm}[H]
    \caption{Temporal-ML} \label{alg:maml}
    \begin{algorithmic}[1]
        \State Initialize $\phi_0$
        \For{round $r=1,2,...$} \Comment{Outer loop}
            \State Sample a subset $T_r$ of $m$ tasks
            \For{task $t \in T_r$} \Comment{Inner loop}
                \State Initialize inner weight $\theta_t^{(0)} \leftarrow \phi_{r-1}$
                \State Forward: For all data point $\mathbf{x}_{}$ in $\mathcal{D}_t^{support}$
                \begin{align*}
                    \mathbf{x'} &= \mathbf{FullyConnected}\left( \mathbf{x} \right)\\
                    \mathbf{h}_{LSTM} &= \mathbf{BidirectionalLSTM}\left( \mathbf{x'} \right)\\
                    \mathbf{h}_{CNN} &= \mathbf{Convolution1D}\left( \mathbf{x'} \right)\\
                    \hat{y} &= \begin{cases}
                        \mathbf{FullyConnected}\left( \mathbf{h}_{LSTM} \right)\\
                        \mathbf{FullyConnected}\left( \mathbf{Concatenate}\left( \mathbf{h}_{LSTM}, \mathbf{h}_{CNN} \right)\right)
                    \end{cases}
                \end{align*}

                \State Backward and aggregate meta loss:\Comment{Inner optimization}
                \begin{align*}
                    &\mathcal{L}_t^{task}\left( \theta_t^{(e-1)}, \mathcal{D}_t^{support} \right) = \mathbf{CrossEntropyLoss}\left( \mathbf{y}, \hat{\mathbf{y}} \right)\\
                    &\theta_t^{(e)} \leftarrow \theta_t^{(e-1)} - \alpha\nabla_\theta\mathcal{L}_t^{task}\left( \theta_t^{(e-1)}, \mathcal{D}_t^{support} \right)\\
                    &\mathcal{L}^{meta} \leftarrow \mathcal{L}^{meta} + \mathcal{L}_t^{meta}\left( \theta_t^{(e)}, \mathcal{D}_t^{query} \right)
                \end{align*}
            \EndFor

            \State Outer optimization at round $r$:\Comment{Outer optimization}
            \begin{align*}
                \phi_{r+1} \leftarrow \phi_r - \beta\nabla_\phi\mathcal{L}^{meta}
            \end{align*}
        \EndFor
    \end{algorithmic}
\end{algorithm}

\section{Data preparation}
\label{sec:data_prep}

% Phương pháp đề xuất sử dụng các thuật toán ML để huấn luyện mô hình. Do đó, dữ liệu cần được tổ chức lại để các thuật toán ML có thể hoạt động được. Trong trường hợp dữ liệu bao gồm nhiều datasets khác nhau thuộc cùng một lĩnh vực, mỗi dataset sẽ được coi là một task của \verb|MAML|. Trong trường hợp dữ liệu bao gồm một dataset duy nhất, cần chia nhỏ dataset này thành các tập con ứng với các task riêng biệt. Tóm lại, tập dữ liệu sau khi chuẩn bị bao gồm $n$ tasks: $\mathcal{D} = \left\{ \mathcal{D}_t \right\}_{t=1}^{n}$. Dữ liệu tại mỗi task được chia thành tập support và query: $\mathcal{D}_t = \left\{ \mathcal{D}_t^{support}, \mathcal{D}_t^{query} \right\}$.

\verb|Temporal-ML| uses ML algorithms to train the model. Therefore, the data needs to be reorganized so that the ML algorithms can work. In case the data includes many different datasets belonging to the same field, each dataset will be considered a task of \verb|MAML|. In case the data includes a single dataset, it is necessary to divide this dataset into subsets corresponding to separate tasks. In summary, the prepared dataset includes $n$ tasks: $\mathcal{D} = \left\{ \mathcal{D}_t \right\}_{t=1}^{n}$. The data at each task is divided into support and query sets: $\mathcal{D}_t = \left\{ \mathcal{D}_t^{support}, \mathcal{D}_t^{query} \right\}$.

% Một sample dữ liệu bao gồm các cặp giá trị $(\mathbf{x}_{t-L:t}, y)$. Trong đó, $\mathbf{x}_{t-L:t}$ bao gồm $L$ giá trị lịch sử tính từ thời điểm $t$ trở về trước; $y\in \{0,1\}$ là nhãn dữ liệu, thể hiện xu hướng giảm, hoặc tăng của mẫu dữ liệu $x_{t+1}$ so với $x_{t}$. Tùy vào từng bài toán và cách cài đặt mà các phần tử trong $\mathbf{x}_{t-L:t}$ có thể là các vector hoặc các scalar number. Ví dụ, đối với dữ liệu chứng khoán, $\mathbf{x}_{t-L:t}$ có thể chứa các vector dữ liệu $\vec x_i = (\text{open, low, high, close})$ hoặc chỉ một giá trị close price duy nhất.

A data sample consists of pairs of values $(\mathbf{x}_{t-L:t}, y)$. In which, $\mathbf{x}_{t-L:t}$ includes $L$ historical values from time $t$ back; $y\in \{0,1\}$ is the data label, showing the decreasing or increasing trend of the data sample $x_{t+1}$ compared to $x_{t}$. Depending on each problem and the implementation, the elements in $\mathbf{x}_{t-L:t}$ can be a matrix or a vector. For example, for stock data, $\mathbf{x}_{t-L:t}$ can contain $L$ vectors $\vec x_i = (\text{open, low, high, close})$ or can be a vector of close price values only.

\section{Temporal feature extraction}

% Quá trình rút trích đặc trưng được thực hiện bằng hai phương pháp khác nhau: thông qua \verb|LSTM| và \verb|LSTM+CNN|. Đối với phương pháp sử dụng \verb|LSTM|, từng phần tử trong ma trận $\mathbf{x}_{t-L:t}$ (viết gọn là $\mathbf{x}$) được đưa qua một lớp \verb|FullyConnected| có đầu ra lớn hơn số chiều của $\vec x_i, i\in[t-L, t]$ to present the input by a larger vector $\vec x'_i$ (equation \ref{eq:decompose}). Theo đó, các đặc trưng được thể hiện sâu sắc và rõ ràng hơn. Các đặc trưng này sau đó được truyền qua mạng \verb|LSTM| rút trích có chọn lọc các phụ thuộc thời gian dài hạn ($\mathbf{h}_{LSTM}$). Để có thể khai thác tối đa các ràng buộc thời gian dài hạn, chúng tôi sử dụng \verb|BidirectionalLSTM| để rút trích từ hai phía của của $\mathbf{x}$ (equation \ref{eq:feature_lstm})

Feature extraction is performed using two different methods: via \verb|LSTM| and \verb|LSTM+CNN|. For the method using \verb|LSTM|, each element in the matrix $\mathbf{x}_{t-L:t}$ (abbreviated as $\mathbf{x}$) is passed through a \verb|FullyConnected| layer whose output is larger than the dimension of $\vec x_i, i\in[t-L, t]$ to obtain vector $\vec x'_i$ (equation \ref{eq:decompose}). Accordingly, the data characteristics are expressed more deeply and clearly. These features are then passed through the \verb|LSTM| network to selectively extract long-term temporal dependencies ($\mathbf{h}_{LSTM}$). To fully exploit the long-term time constraints, we use \verb|BidirectionalLSTM| to extract from both sides of $\mathbf{x}$ (equation \ref{eq:feature_lstm})

\begin{align}
    \mathbf{x'} &= \mathbf{FullyConnected}\left( \mathbf{x} \right) \label{eq:decompose}\\
    \mathbf{h}_{LSTM} &= \mathbf{BidirectionalLSTM}\left( \mathbf{x'} \right) \label{eq:feature_lstm}
\end{align}

% Lấy cảm hứng từ nghiên cứu \cite{vo2017multi}, chúng tôi đề xuất \verb|LSTM+CNN|, kết hợp các đặc trưng rút trích được từ mạng \verb|LSTM| và \verb|CNN|. Cụ thể, ngoài việc sử dụng \verb|LSTM|, $\mathbf{x'}$ còn được đưa vào \verb|CNN| để rút trích đặc trưng thời gian cục bộ ($\mathbf{h}_{CNN}$) (equation \ref{eq:feature_cnn}). Tiếp đến, $\mathbf{h}_{LSTM}$ và $\mathbf{h}_{CNN}$ được nối với nhau (equation \ref{eq:concat}) sau đó chuyển đến phần phân lớp của neural network (equation \ref{eq:clf}).

Inspired by study \cite{vo2017multi}, we use \verb|LSTM+CNN| features, which combines the features extracted from \verb|LSTM| and \verb|CNN|. Specifically, in addition to using \verb|LSTM|, $\mathbf{x'}$ is also fed into \verb|CNN| to extract local temporal features $\mathbf{h}_{CNN}$ (equation \ref{eq:feature_cnn}). Next, $\mathbf{h}_{LSTM}$ and $\mathbf{h}_{CNN}$ are concatenated (equation \ref{eq:concat}) and then passed to the classification head of the neural network (equation \ref{eq:clf}).

\begin{align}
    \mathbf{h}_{CNN} &= \mathbf{Convolution1D}\left( \mathbf{x'} \right) \label{eq:feature_cnn}\\
    \mathbf{h} &= \mathbf{Concatenate}\left( \mathbf{h}_{LSTM}, \mathbf{h}_{CNN} \right) \label{eq:concat} \\
    \hat y &= \mathbf{FullyConnected}\left( \mathbf{h} \right) \label{eq:clf}
\end{align}

% Mạng \verb|LSTM| duy trì giá trị cell-state nhằm lưu trữ có chọn lọc các phụ thuộc dài hạn. Điều này rất thích hợp trong việc giải quyết các bài toán time-series data. Mặt khác, giá trị tương lai thường phụ thuộc rất lớn vào các giá trị lịch sử gần nhất. Chúng tôi đề xuất sử dụng mạng \verb|CNN| để nhấn mạnh các đặc trưng cục bộ, từ đó hướng một phần sự chú ý của mô hình vào các thời điểm nhất định. Do đó, phương pháp đề xuất không chỉ nhớ được các đặc trưng long-term mà còn highlight được các đặc trưng short-term.

The \verb|LSTM| network maintains cell-state values to selectively store long-term dependencies. This is very suitable for solving time-series data problems. On the other hand, future values often depend heavily on recent historical values. We propose to use the \verb|CNN| network to emphasize local features, thereby directing part of the model's attention to certain time points. Therefore, the proposed method not only remembers long-term features but also highlights short-term features.

% Next, $\mathbf{h}_{LSTM}$ and $\mathbf{h}_{CNN}$ are concatenated (equation \ref{eq:concat}) and then passed to the classification part of the neural network (equation \ref{eq:clf}).

\section{Effective synthesis of models' parameters}

% Chúng tôi sử dụng \verb|MAML| như trình bày trong thuật toán \ref{alg:maml} để huấn luyện và tổng hợp trọng số của các mô hình tại các task. Như đã đề cập trong phần \ref{sec:ml}, tối ưu tham số theo cách tiếp cận của ML chính là đi giải hai phương trình \ref{eq:inner_opt} và \ref{eq:outer_opt} bằng các phương pháp tối ưu trên dữ liệu support và query. Cụ thể, quá trình tối ưu bao gồm nhiều bước toàn cục (outer optimization), thực hiện trên tất cả các tasks tham gia huấn luyện. Mỗi bước toàn cục bao gồm nhiều bước cục bộ (inner optimization) thực hiện trên từng task riêng lẻ. Tại bước toàn cục $r$, quá trình tối ưu cục bộ lần thứ $e$ tại tập support của task $t$ diễn ra như sau:

We use \verb|MAML| as presented in the algorithm \ref{alg:maml} to train and aggregate the weights of the models at the tasks. As mentioned in section \ref{sec:ml}, parameter optimization in the ML approach is to solve the two equations \ref{eq:inner_opt} and \ref{eq:outer_opt} using optimization methods on the support set and query set. Specifically, the optimization process includes many global steps (outer optimization), performed on all tasks participating in training. Each global step includes many local steps (inner optimization) performed on each individual task. At global step $r$, the $e$th local optimization process on the support set of task $t$ proceeds as follows:

\begin{align}
    \begin{cases}
        \theta_t^{(0)} &= \phi_{r-1} \\
        \theta_t^{(e)} &= \theta_t^{(e-1)} - \alpha \nabla_{\theta} \mathcal{L}^{task}_t\left( \theta_t^{(e-1)}, \mathcal{D}_t^{support} \right)
    \end{cases}
\end{align} In which, $\phi_{r-1}$ is the result of the $r-1$ outer optimization process, $\alpha$ is the inner learning rate.

% Tiếp đó, quá trình outer optimization tại bước toàn cục được thực hiện bằng cách tổng hợp độ lỗi trên tập query của các task và tối ưu trên đó (equation \ref{eq:outer_sol}).

Next, the outer optimization process is performed by aggregating the losses on the query set of the tasks and optimizing on it (equation \ref{eq:outer_sol}).

\begin{align}
    \begin{cases}
        \phi_0 = \text{Random Initialization}\\
        \phi_r = \phi_{r-1} - \beta \nabla_{\phi} \sum_{t=1}^n{\mathcal{L}^{meta}_t \left( \theta_t^*(\phi), \mathcal{D}_t^{query} \right)}
    \end{cases}
    \label{eq:outer_sol}
\end{align} Where, $\beta$ is the outer learning rate.

% Giả sử thuật toán chạy $E$ steps trong inner optimization, lượng đạo hàm tại phương trình \ref{eq:outer_sol} được viết lại như sau (the notations of dataset are removed):

Assuming the algorithm runs $E$ steps in inner optimization, the derivative quantity at equation \ref{eq:outer_sol} is rewritten as follows (the notations of dataset are removed):

\begin{align*}
    \nabla_{\phi} \sum_{t=1}^n{\mathcal{L}^{meta}_t \left( \theta_t^*(\phi)\right)} =&\nabla_{\phi} \sum_{t=1}^n{\mathcal{L}^{meta}_t \left( \theta_t - \alpha \nabla_{\theta} \mathcal{L}^{task}_t\left( \theta_t \right) \right)}\\
    = &\sum_{t=1}^n{ \frac{\partial \mathcal{L}^{meta}_t\left(\theta_t^{(E)}\right)}{\partial \theta_t^{(E)}} \frac{\partial \theta_t^{(E)}}{\partial \phi}}\\
    = &\sum_{t=1}^n{ \nabla_{\theta} \mathcal{L}^{meta}_t\left(\theta_t^{(E)}\right) \prod_{j=0}^{E-1} {\left[\mathbb{I} - \alpha\nabla^2_{\theta}\mathcal{L}^{task}_{t}\left(\theta_t^{(j)}\right)\right]}} \numberthis
    \label{eq:outer_grad}
\end{align*}

% Sự xuất hiện của tích các đạo hàm bậc hai trong phương trình \ref{eq:outer_grad} khiến quá trình đạo hàm trở nên phức tạp vì phải tốn rất nhiều chi phí để duy trì các ma trận Hessian. Do đó, số bước tính toán để tìm ra $\theta^*$ cần phải hạn chế. Trên thực tế, các phương pháp sử dụng ML \cite{fallah2020personalized, chen2018federated, nguyen2022meta,finn2017model, li2017meta} thường sẽ chọn $E\in [1,5]$.

The presence of the product of second order derivatives in the equation \ref{eq:outer_grad} makes the derivation process complicated because it requires a lot of overhead to maintain Hessian matrices. Therefore, the number of computational steps to find $\theta^*$ needs to be limited. In practice, methods using ML \cite{fallah2020personalized, chen2018federated, nguyen2022meta,finn2017model, li2017meta} often choose $E\in [1,5]$.

% phải cmt thêm về phương pháp tổng hợp mô hình
% Các mô hình hybrid ensemble vốn được sử dụng rất nhiều trong các bài toán xử lý time-series và được chứng minh thực nghiệm là có độ chính xác cao hơn so với các mô hình handle time-series data tiêu chuẩn vì có thể tổng hợp được sức mạnh của nhiều mô hình \cite{ayitey2023forex}. Tuy vậy, các hình thức tổng hợp của ensemble model hiện nay vẫn còn rất cứng nhắc vì chỉ có thể tổng hợp dựa trên kết quả cuối (cơ chế voting của bagging models) và kết quả gần cuối (đối với stacking models). Dưới góc nhìn của ensemble model, có thể coi outer optimization nói chung hay phương trình \ref{eq:outer_opt} nói riêng, là một phương pháp tổng hợp hiệu quả các sub-model thông qua quá trình tối ưu, giúp tận dụng hiệu quả khả năng rút trích đặc trưng của từng mô hình. Nói cách khác, mô hình sau khi tổng hợp có thể rút trích đặc trưng ở mức sâu hơn, cải thiện đáng kể khả năng dự đoán so với các mô hình ensemble truyền thống.

Hybrid ensemble models have been widely used in time-series processing problems and have been experimentally proven to be more accurate than standard time-series models because they can synthesize the strengths of many sub-models \cite{ayitey2023forex}. However, current ensemble model synthesis forms are still very rigid because they can only synthesize based on the final results (e.g. voting mechanism of bagging models) and semi-final results (e.g. stacking models). From the perspective of ensemble models, the equation \ref{eq:outer_sol} can be considered an optimization-based method of synthesizing sub-models, which helps to take advantage of the feature extraction capabilities of each model. In other words, the synthesized model can extract features at a deeper level, significantly improving the prediction ability compared to traditional ensemble models.

### 2024-12-31 doc/thesis/chapters/4.experiment.tex
thesis draft

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Chapter 4: Numerical Experiment
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\chapter{Numerical Experiment}
\label{chap:experiment}

\section{Dataset description}

% mô tả dữ liệu
% FX nói riêng và các chỉ số tài chính nói chung là dạng dữ liệu điển hình cho aperiodic time-series data. Do đó, chúng tôi chọn loại dữ liệu này để kiểm thử mô hình. Cụ thể, chúng tôi cấu hình hai bộ dữ liệu sử dụng dữ liệu FX. Bộ dữ liệu \verb|USD/JPY| chỉ gồm dữ liệu của cặp tiền tệ USD/JPY, được chia thành 60 tập dữ liệu con theo trình tự thời gian với kích thước bằng nhau. Dữ liệu được sample theo giờ từ năm 2000 đến năm 2024, bao gồm các thuộc tính open, low, high, và close price. Bộ dữ liệu \verb|multi-fx| bao gồm 60 cặp tiền tệ giữa 18 quốc gia Australia, Canada, Switzerland, Denmark, EU, United Kingdom, Hong Kong, Iceland, Japan, Norway, New Zealand, Singapore, Sweden, Turkey, United States, Mexico, China, South Africa. Dữ liệu có thuộc tính tương tự như \verb|USD/JPY| và được sample theo ngày từ năm 2014 đến 2024.

Foreign exchange (FX) in particular and financial indices in general are typical data types for aperiodic time-series data. Therefore, we choose this type of data to test the model. Specifically, we configure two datasets using FX data. \verb|USD/JPY| dataset consists of only data of the exchange rate between US dollar and Japanese yen. The data is sampled hourly from 2000 to 2024, including the attributes of open, low, high, and close price. \textbf{By carefully examining the transaction history as well as effectively aggregating data characteristics, we expect to forecast the movement of price without utilizing external information.}

\verb|multi-fx| dataset consists of 60 currency pairs made of 18 countries: Australia, Canada, Switzerland, Denmark, EU, United Kingdom, Hong Kong, Iceland, Japan, Norway, New Zealand, Singapore, Sweden, Turkey, United States, Mexico, China, South Africa. The data has similar attributes to \verb|USD/JPY| and sampled daily from 2014 to 2024. \textbf{By integrating analysis of multiple sources of information in the domain of FX, we expect to obtain valuable analytical insights for the indicator of interest.}

% Ngoài ra, chúng tôi sử dụng thêm 2 bộ dữ liệu: Electricity Transformer Temperature (\verb|ETT-m2|) và Weather (\verb|WTH|). Tập dữ liệu \verb|ETT-m2| bao gồm 7 trường dữ liệu, đo đạc các thông số của máy biến áp tại một tỉnh của Trung Quốc sau mỗi 15 phút trong khoảng thời gian từ July 2016 to July 2018. Tập dữ liệu \verb|WTH| bao gồm 12 trường dữ liệu, ghi nhận các thông số của thời tiết tại Weather Station of the Max Planck Biogeochemistry Institute in Jena, Germany sau mỗi 10 phút trong năm 2020. Đây là hai tập dữ liệu thể hiện tính chu kỳ rất mạnh mà \verb|NHITS| đã rất thành công trong việc dự đoán. Thực nghiệm trên chúng giúp so sánh toàn diện hơn về khả năng của phương pháp đề xuất đối với \verb|NHITS|.

In addition, we used two periodic datasets: Electricity Transformer Temperature (\verb|ETT-m2|) \cite{zhou2021informer} and Weather (\verb|WTH|) \cite{Kolle}. \verb|ETT-m2| dataset consists of 7 data fields, measuring the parameters of a transformer in a province of China every 15 minutes from July 2016 to July 2018. \verb|WTH| dataset consists of 12 data fields, recording the weather parameters at the Weather Station of the Max Planck Biogeochemistry Institute in Jena, Germany every 10 minutes in 2020. These two datasets exhibit very strong periodicity, which \verb|NHITS| has been very successful in predicting. Experiments on them provide a more comprehensive comparison of the proposed method's capabilities against \verb|NHITS|.

\section{Data pre-process}

\subsection{Temporal-ML}
\label{subsec:ml_experiment}

% viết về cách tổng hợp metric ở đây
% xét trên 1 thuộc tính cố định: chạy 60 thằng -> 60 bộ metric. lấy TBC trên 60 bộ này -> metric ± std
% chạy hết các attribute -> 4 thằng metric ± std
% -> trung bình theo kiểu mean(metric), \sqrt(mean(std^2))
Như đã đề cập trong section \ref{sec:data_prep}, để sử dụng ML, các tập dữ liệu cần phải được cấu trúc thành các task riêng biệt. Trong mỗi task, tập support chiếm 20\% dữ liệu, tập query chiếm 80\% dữ liệu. Đối với \verb|multi-fx|, mỗi cặp tiền tệ được chia thành 2 tập dữ liệu, ứng với 2 tasks. Do đó, 60 cặp tiền tệ tạo thành 120 tasks. Đối với các tập dữ liệu còn lại, mỗi tập được chia nhỏ ra thành các chuỗi thời gian, mỗi chuỗi tương ứng với một task. Trong cả quá trình, 50\% số tasks được sử dụng trong quá trình meta-training, 25\% số tasks được sử dụng trong quá trình meta-validation, số tasks còn lại được sử dụng trong meta-testing. Thống kê dữ liệu cho các task theo tập dữ liệu được trình bày trong bảng \ref{tab:stat_}.

\begin{table}
    \centering
    \caption{Statistics on datasets.}
    \label{tab:stat_}
    \begin{tblr}{
        cells = {c},
        hline{1,6} = {-}{0.08em},
        hline{2} = {-}{},
    }
    Dataset                      & Attribute & Task & Sample    & Sample/Task    \\
    \Verb|USD/JPY|               & 4         & 60   & 150,175   & 2,503          \\
    \Verb|multi-fx|              & 4         & 120   & 154,440  & 1,287          \\
    \Verb|ETT-m2|                & 7         & 48   & 69,680    & 1,452          \\
    \Verb|WTH|                   & 12        & 40   & 35,064    & 877            
    \end{tblr}
\end{table}

Vì các tập dữ liệu đều là time-series, dữ liệu trong quá trình huấn luyện không được tiết lộ bất cứ thông tin gì về tương lai. Do đó, 120 tasks của dữ liệu \verb|multi-fx| được chia như hình \ref{fig:multi_fx_split}. Đối với các tập dữ liệu còn lại, dữ liệu sử dụng trong quá trình huấn luyện, kiểm thử, kiểm tra phải đảm bảo thứ tự thời gian từ quá khứ đến tương lai.

\begin{figure}
    \centering
    \begin{subfigure}[b]{0.5\textwidth}
        \centering
        \includegraphics[width=\textwidth]{multi_fx (1).png}
        \caption{}
        \label{fig:multi_fx_split}
    \end{subfigure}%
    ~
    \begin{subfigure}[b]{0.5\textwidth}
        \centering
        \includegraphics[width=0.92\textwidth]{multi_fx_prep.png}
        \caption{}
        \label{fig:multi_fx_prep}
    \end{subfigure}

    \cprotect\caption{(a): Splitting data for \verb|multi-fx|. (b): Pre-process for \verb|multi-fx| (dash line rectangle accounts for 20\% data of validation and testing tasks).}
    % \label{fig:ill_aperiodic}
\end{figure}

Quá trình tiền xử lý bằng \verb|z-score| theo đó được thực hiện trên toàn bộ dữ liệu huấn luyện và các tập support của dữ liệu kiểm thử và kiểm tra của từng cặp tiền tệ (see figure \ref{fig:multi_fx_prep}) vì đây là các dữ liệu đã được biết trước. Cụ thể, tiền xử lý thuộc tính $X$ được thực hiện như sau:

\begin{align}
    \mu_{X} &= \frac{1}{n} \sum_{i=1}^n{x_i} \quad \text{($n$ là số mẫu của thuộc tính $X$)} \label{eq:mean}\\
    \sigma_{X} &= \sqrt{\frac{1}{n} \sum_{i=1}^n{(x_i - \mu_X)^2}} \label{eq:std}\\
    X_{new} &= \frac{X - \mu_{X}}{\sigma_{X}} \label{eq:z_score}
\end{align}

$\mu_{X}$ và $\sigma_{X}$ thu được từ phương trình \ref{eq:mean} và \ref{eq:std} được sử dụng để chuẩn hóa cho thuộc tính $X$ trong query set của quá trình meta-validation và meta-testing.

\subsection{Baseline model}
\label{subsec:baseline_experiment}

% viết về cách tổng hợp metric ở đây
Nghiên cứu này sử dụng thuật toán \verb|NHITS| \cite{challu2023nhits} làm baseline model. \verb|NHITS| sử dụng toàn bộ thuộc tính để dự đoán xu hướng tiếp theo cho một thuộc tính trong một tập dữ liệu. Thuật toán hướng đến phân rã các giải tần và kết hợp lại để dự đoán tương lai. Đối với các tập dữ liệu \verb|USD/JPY|, \verb|ETT-m2|, và \verb|WTH|, dữ liệu tại mỗi tập được chia thành training set, validation set, và test set với tỷ lệ tương ứng 6:2:2.

Chúng tôi tiền xử lý dữ liệu huấn luyện bằng \verb|z-score| như đã đề cập ở subsection \ref{subsec:ml_experiment}. Sau đó, $\mu_{X}$ và $\sigma_{X}$ thu được từ phương trình \ref{eq:mean} và \ref{eq:std} được sử dụng để chuẩn hóa cho thuộc tính $X$ trong validation set trong quá trình tuning. Khi đã chọn được các siêu tham số tốt nhất, dữ liệu của training set và validation set được chuẩn hóa lại từ đầu, sau đó được huấn luyện với bộ siêu tham số tốt nhất để dự đoán test set.

Đối với tập dữ liệu \verb|multi-fx|, vì \verb|NHITS| không có cơ chế tổng hợp mô hình trên các tập dữ liệu khác nhau, chúng tôi lặp lại quy trình huấn luyện (huấn luyện, tuning, testing trên 60\%, 20\%, 20\% dữ liệu, respectively) cho từng cặp tiền tệ, sau đó tổng hợp metrics trên các cặp tiền tệ để thu được đánh giá cuối.

\section{Metric evaluation}

Nghiên cứu sử dụng accuracy, precision, recall, và F1 để đánh giá các mô hình:

\begin{align*}
    acc &= \frac{TP+TN}{TP+FP+TN+FN}\\
    P &= \frac{TP}{TP+FP}\\
    R &= \frac{TP}{TP+FN}\\
    F1 &= \frac{2PR}{P+R}
\end{align*}Trong đó, $TP, TN, FP, FN$ lần lượt là số mẫu \textit{true\_positive, true\_negative, false\_positive, false\_negative}.

% The study uses accuracy, precision, recall, and F1 score to evaluate the models. Accordingly, during the inference phase, the model will run on all tasks to calculate the metrics of each one. Then, the average of the metrics of the tasks is calculated to obtain the final result.

\subsection{Temporal-ML}

Giả sử thuật toán \verb|Temporal-ML| được đánh giá trên tập dữ liệu $\mathcal{D}$ gồm $a$ thuộc tính và được chia thành $n$ tasks. Quá trình đánh giá được thực hiện bằng cách để mô hình dự đoán xu hướng của từng thuộc tính với đầu vào là tất cả các thuộc tính, sau đó thực hiện hai bước tổng hợp: (1) - Tổng hợp trên task; (2) - Tổng hợp trên thuộc tính.

\textbf{Tổng hợp trên task}. Xét metric $m$ khi mô hình dự đoán thuộc tính $k$ bất kỳ ($1\leq k \leq a$). Sau khi dự đoán thuộc tính $k$, chúng ta thu được $n$ giá trị: $\{m^{(k)}_1,\dots,m^{(k)}_n\}$. Quá trình tổng hợp metrics $m$ của $n$ tasks được thực hiện như sau:

\begin{align}
    \bar{m}^{(k)} &= \frac{1}{n}\sum_{i=1}^n{m^{(k)}_i} \label{eq:mean_task}\\
    s^{(k)}_m &= \sqrt{\frac{1}{n} \sum_{i=1}^n{(m^{(k)}_i - \bar{m}^{(k)})^2}} \label{eq:std_task}
\end{align}

\textbf{Tổng hợp trên thuộc tính}. Mô hình thực hiện dự đoán tất cả $a$ thuộc tính và thu được $a$ metrics: $\left\{ \left( \bar{m}^{(k)}\pm s^{(k)}_m \right) \right\}_{k=1}^a$. Quá trình tổng hợp metric $m$ của $a$ thuộc tính diễn ra như sau:

\begin{align}
    \bar{m} &= \frac{1}{a}\sum_{k=1}^a{\bar{m}^{(k)}} \label{eq:mean_att}\\
    s_m &= \sqrt{\frac{1}{a} \sum_{k=1}^a{(s^{(k)}_m)^2}} \label{eq:std:att}
\end{align}

\subsection{Baseline model}

\verb|NHITS| cũng được đánh giá trên tập dữ liệu $\mathcal{D}$ với $a$ thuộc tính và phải thực hiện dự đoán trên từng thuộc tính. Xét metric $m$, sau quá trình dự đoán thuộc tính $k$, ta thu được $a$ giá trị: $\{m^{(k)}_1,\dots,m^{(k)}_a\}$. Quá trình tổng hợp metric $m$ của $a$ thuộc tính được thực hiện như sau:

\begin{align*}
    \bar{m} &= \frac{1}{a}\sum_{k=1}^a{m^{(k)}}\\
    s_m &= \sqrt{\frac{1}{a} \sum_{i=1}^n{(m^{(k)} - \bar{m})^2}}
\end{align*}

Như đã đề cập ở trên, \verb|NHITS| không có cơ chế tổng hợp mô hình khi kiểm thử trên dữ liệu \verb|multi-fx|. Do đó, chúng tôi kiểm thử \verb|NHITS| trên từng cặp tiền tệ trong \verb|multi-fx| và tổng hợp metrics giống như \verb|Temporal-ML|.

\subsection{Fairness in evaluation}

Gọi $m$ là số mẫu dữ liệu trong mỗi tasks, tổng số dữ liệu của quá trình huấn luyện, kiểm thử, và kiểm tra là: $mn$. Đối với \verb|NHITS|, mô hình được huấn luyện, kiểm thử, kiểm tra trên 60\%, 20\%, 20\% dữ liệu, respectively:

\begin{align*}
    \text{Train: } &0.6mn\\
    \text{Validation: } &0.2mn\\
    \text{Testing: } &0.2mn
\end{align*}

Đối với \verb|Temporal-ML|, quá trình làm giàu knowledge cho mô hình sử dụng toàn bộ dữ liệu huấn luyện và tập support của dữ liệu kiểm thử/kiểm tra. Quá trình kiểm thử/kiểm tra được thực hiện trên tập query:

\begin{align*}
    \text{Train: } &0.5mn + 20\%\left( \frac{m}{2}n \right) = 0.6mn\\
    \text{Validation: } &80\% \left( \frac{m}{2}\frac{n}{2} \right)  =0.2mn\\
    \text{Testing: } &80\% \left( \frac{m}{2}\frac{n}{2} \right) =0.2mn
\end{align*}

Do đó, với cách chia dữ liệu như trên, chúng tôi đạt được sự công bằng về số lượng mẫu dữ liệu trong quá trình kiểm thử.

\section{Hyper parameters tuning}
% quá trình fine-tune (kiểu cố định outer và quan sát inner) --> chuyển xuống phần phụ lục?

\subsection{Temporal-ML}

Trong cài đặt của chúng tôi, chúng tôi sử dụng một lớp \verb|FullyConnected| gồm 16 units với hàm kích hoạt \verb|ReLU| để biểu diễn các đặc trưng sâu hơn và rõ ràng hơn. Sau đó, đặc trưng này được truyền song song đến các hai khối \verb|BidirectionalLSTM| và \verb|CNN|. Khối \verb|BidirectionalLSTM| bao gồm 32 hidden units, các đầu ra được nối dài để tạo thành một vector cuối. Khối \verb|CNN| bao gồm hai layer \verb|CNN| có số filter lần lượt là 32 và 64. Kernel được sử dụng trong các layer có kích thước $3\times 3$. Theo sau mỗi layer \verb|CNN| là một layer \verb|MaxPooling| sử dụng kernel kích thước $2\times 2$. Kết thúc khối \verb|CNN| là một layer \verb|Flatten|. Trong trường hợp chỉ sử dụng các đặc trưng dài hạn, đầu ra của khối \verb|BidirectionalLSTM| sẽ được sử dụng để phân lớp. Nếu sử dụng thêm các đặc trưng cục bộ, đầu ra của khối \verb|BidirectionalLSTM| và \verb|CNN| sẽ được nối lại rồi truyền qua một layer phân lớp nhị phân với hàm kích hoạt \verb|Sigmoid|.

% We use a \verb|FullyConnected| layer of 16 units with a \verb|ReLU| activation function to decompose the initial feature. This feature is then passed in parallel to the \verb|BidirectionalLSTM| and \verb|CNN| blocks. The \verb|BidirectionalLSTM| block consists of 32 hidden units, the outputs of which are concatenated to form a final vector. The \verb|CNN| block consists of two \verb|CNN| layers with 32 and 64 filters, respectively. The kernel used in the layers is of size $3\times 3$. Each \verb|CNN| layer is followed by a \verb|MaxPooling| layer using a kernel of size $2\times 2$. The \verb|CNN| block ends with a \verb|Flatten| layer. Features of the \verb|BidirectionalLSTM| and \verb|CNN| blocks are then concatenated and passed through a binary classification layer with a \verb|Sigmoid| activation function.

\begin{table}[H]
    \centering
    \caption{Search space for fine-tuning our method.}
    \label{tab:our_finetune}
    \begin{tabular}{ll} 
    \toprule
    \multicolumn{1}{c}{Hyper-parameter} & \multicolumn{1}{c}{Seach space}   \\ 
    \hline
    Inner batch size (samples/batch)    & \{32\}                            \\
    Inner training step                 & \{3\}                             \\
    Outer batch size (tasks/batch)      & \{5\}                             \\
    Outer training step                 & \{100\}                           \\ 
    \hline
    Lookback window                     & \{10, 20, 30\}                    \\
    Inner learning rate                 & \{0.001, 0.005, 0.01, 0.05\}      \\
    Outer learning rate                 & \{0.001, 0.005, 0.0015, 0.0055\}  \\
    \bottomrule
    \end{tabular}
\end{table}

Quá trình fine-tune của thuật toán ML liên quan đến nhiều siêu tham số như inner batch size, outer batch size, inner training step, outer training step,... Để dễ dàng fine-tune, chúng tôi cố định hầu hết các tham số và chỉ fine-tune kích thước của lookback window, inner và outer learning rate. Chi tiết được trình bày trong bảng \ref{tab:our_finetune}.

% The fine-tuning process of ML algorithms involves many hyper-parameters such as inner batch size, outer batch size, inner training steps, outer training steps. To facilitate fine-tuning, we fix most of the parameters and only fine-tune the size of lookback window, the inner and outer learning rates. Details are presented in table \ref{tab:our_finetune}.

\subsection{Baseline model}

Đối với mô hình \verb|NHITS|, chúng tôi dựa trên \cite{challu2023nhits} để định nghĩa không gian tìm kiếm cho việc fine-tune tham số, cũng như kiến trúc mô hình. Đối với các tham số không được đề cập trong bảng, chúng tôi sử dụng giá trị mặc định của cài đặt \verb|NHITS| trong thư viện \verb|NeuralForecast| \cite{neuralforecast}. Kết quả tốt nhất của các lần fine-tune được chọn ra và báo cáo trong nghiên cứu này.

% We rely on \cite{challu2023nhits} to define the search space (table \ref{tab:nhits_finetune}) for parameter fine-tuning, as well as to fine-tune the model architecture. For parameters not mentioned in the table, we use the default values of the \verb|NHITS| implementation in the \verb|NeuralForecast| library \cite{neuralforecast}. The best results of fine-tuning process are selected and reported in this study.

\begin{table}[H]
    \centering
    \caption{Search space for fine-tuning NHITS.}
    \label{tab:nhits_finetune}
    \begin{tabular}{ll} 
    \toprule
    \multicolumn{1}{c}{Hyper-parameter} & \multicolumn{1}{c}{Seach space}                              \\ 
    \hline
    Random seed                         & \{1\}                                                        \\
    Number of stacks                    & \{3\}                                                        \\
    Number of blocks in each stack      & \{1\}                                                        \\
    Activation function                 & \{ReLU\}                                                     \\
    Batch size                          & \{256\}                                                      \\
    Epoch                               & \{500\}                                                      \\ 
    \hline
    Lookback window                     & \{5, 20, 30\}                                                \\
    Pooling kernel                      & \{{[}2,2,2], [4,4,4], [8,8,8], [8,4,1], [16,8,1]\}           \\
    Stacks' coefficients                & \{{[}168,24,1], [24,12,1], [180,60,1],[40,20,1], [64,8,1]\}  \\
    Number of MLP layers                & \{1,2\}                                                      \\
    Learning rate                       & \{0.001, 0.002, 0.005, 0.01, 0.02\}                          \\
    \bottomrule
    \end{tabular}
\end{table}

### 2024-12-31 doc/thesis/chapters/5.results.tex
thesis draft

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Chapter 5: Results
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\chapter{Results}
\label{chap:result}

\section{Main results}

% vẽ, mô tả cái bảng
% bẻ nhỏ từng attribute trong fx dataset --> thành 4 bảng, 1 bảng tổng hợp là 5
% bảng tổng hợp vẽ ở đây
% chắc phải viết thêm cả CNN

\section{Feature extraction ability}

% 4 bảng phụ vẽ ở đây
% kiểu: sử dụng mô hình để dự đoán từng attribute, ta có bảng kq sau
% phân tích bảng kết quả
% giải thích lý do tại sao LSTM lại ngon hơn LSTM+CNN

\section{Sub-model aggregation ability}

% vẽ về quá trình hội tụ
% mô tả về quá trình hội tụ của các thuật toán
% múa

### 2024-12-31 doc/thesis/chapters/6.conclusion.tex
thesis draft

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Chapter 6: Conclusion
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\chapter{Conclusion \& Future works}
\label{chap:conclusion}

\section{Conclusion}

% viết về contribution

\section{Future works}

### 2024-12-31 doc/thesis/pre_info/abstract.tex
thesis draft

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Abstract
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{center}
    {\LARGE\bf Abstract}
\end{center}
\vspace{20pt}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Keywords
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\vspace{5mm}
\noindent
\textbf{Keywords:} Meta-learning, \verb|LSTM|, \verb|CNN|, Aperiodic time-series data, Foreign exchange

### 2024-12-31 doc/thesis/pre_info/acknowledge.tex
thesis draft

\chapter*{Acknowledgements}

I would like to express my gratitude to my thesis advisors, Prof. Ryo Maezono and Associate Prof. Kenta Hongo for their advice, support, and the opportunities they have given to me during my Master's year in JAIST.

The research topic on movement prediction of aperiodic time-series data was brought to me by Assistant Prof. Ichiba Tomohiro, so I would also like to thank him for introducing me to this problem and for the guidance he provided.

This journey would never have begun without the enthusiastic support and guidance from Prof. Le Hoai Bac. Thank you for leading me to JAIST and for everything.

I would like to thank everyone in Maezono and Hongo laboratory, fellow students and the laboratory staff, for helping me out whenever I run into trouble with the administrative documents or health.

Lastly, I wish to thank my parents for their never-ending support, motivation, and phone calls, and for helping me to stay connected with my family and friends at home.

### 2024-12-31 doc/thesis/clean.sh

@@ -1,0 +1,3 @@
+rm *.aux *.blg *.bbl *.fdb_latexmk *.lof *.lot *.out *.fls *.toc *.gz *.xml *.log
+rm *.cpt
+rm main-blx.bib

### 2025-01-04 doc/thesis/chapters/1.intro.tex
minor changes

@@
     \item Chapter \ref{chap:related_work} presents related works, advantages and disadvantages of each study.
-The mentioned studies belong to two popular approaches in analyzing time-series data: Deep feature-based approach and Frequency decomposition approach.
-In addition, Chapter \ref{chap:related_work} also presents ML, providing the foundation for the proposed method in Chapter \ref{chap:method}.
+In addition, ML is also presented in this chapter, providing the foundation for the proposed method in Chapter \ref{chap:method}.
     \item Chapter \ref{chap:method} presents a new approach for aperiodic time-series data based on the \verb|LSTM|/\verb|LSTM+CNN| feature extraction method and ML model optimization method.

### 2025-01-04 doc/thesis/chapters/2.related_works.tex
minor changes

@@
 
-% % tôi sẽ chuyển cái này thành 1 idea để viết phương pháp
-% % kiểu: Lấy ý tưởng từ ensemble model này kia
-% \section{Ensemble models}
+In this chapter, we analyze several approaches in handling time-series data, indicating the advantages and disadvantages of each approach.
+The related studies range from old methods such as ensemble model, deep feature-based approach, to recent methods which aim to decompose the input signal to frequency bands.
+
+\section{Ensemble models}
+
+% cơ chế tổng hợp cùi bắp
 
@@
 
+% không có cơ chế tổng hợp multiple source
+
 \subsection{Convolutional neural network}
+
+% viết về conv1d, 2d, 3d...
 
@@
 
+% viết 1 chút về bidirectional lstm
+
 \section{Frequency decomposition approach}
 
+% không có cơ chế tổng hợp multiple source
+
 \subsection{A transformer-based method}
+
+% require massive overhead
 
@@
 
-By concatenating stacks, each receiving the remainder of the previous stack, the above architecture is expected to decompose the data into different frequency bands (weekly, daily, even hourly).
+By concatenating stacks, each receiving the remainder of the previous stack, \textbf{the above architecture is expected to decompose the data into different frequency bands (weekly, daily, even hourly)}.
 In practice, \verb|NHITS| performs very well for highly periodic datasets such as electricity consumption, weather, traffic.
-However, we are aiming for an aperiodic time-series dataset, which has very low, or even non-existent, periodicity.
+\textbf{However, we are aiming for an aperiodic time-series dataset, which has very low, or even non-existent, periodicity}.
 This poses a huge challenge for \verb|NHITS|.
@@
 
-One disadvantage of optimization-based ML method lies in solving equation \ref{eq:outer_opt} which requires massive overhead to compute and maintain a Hessian matrix.
-Even so, ML algorithms still achieve a high accuracy in handling many problem in which the quick adaptation or effective model synthesis are required.
+\textbf{One disadvantage of optimization-based ML method lies in solving equation \ref{eq:outer_opt} which requires massive overhead to compute and maintain a Hessian matrix.
+Even so, ML algorithms still achieve a high accuracy in handling many problem in which the quick adaptation or effective model synthesis are required}.

### 2025-01-04 doc/thesis/chapters/3.method.tex
minor changes

@@
 
-We propose \verb|Temporal-ML|, a ML-based method which consists of two main parts that work in parallel: (1) - Temporal feature extraction; (2) - Models' parameter synthesis.
+We propose \verb|Temporal-ML|, a ML-based method which consists of two main components that work in parallel: (1) - Temporal feature extraction; (2) - Models' parameter synthesis.
 The overview of the method is illustrated in figure \ref{fig:flow} and the detail is presented in algorithm \ref{alg:maml}.
-In the \textit{Temporal feature extraction} section, we use two types of features from \verb|LSTM| and \verb|LSTM+CNN| networks.
-In the \textit{Effective synthesis of model's parameters} section, we use \verb|MAML| to synthesize the parameters of the models.
+In the \textit{Temporal feature extraction} section, we propose to use two types of features from \verb|LSTM| and \verb|LSTM+CNN| networks.
+In the \textit{Effective synthesis of model's parameters} section, \verb|MAML| is utilized to synthesize the parameters of the models.
 
@@
 Therefore, the data needs to be reorganized so that the ML algorithms can work.
-In case the data includes many different datasets belonging to the same field, each dataset will be considered a task of \verb|MAML|.
+In case the data includes many different datasets belonging to the same field, each dataset will be considered a task.
+\textbf{The goal of using multiple datasets of the same domain is to utilize the corelation between them} as indicated in study \cite{overreactioncontrarian, mech1993portfolio}.
 In case the data includes a single dataset, it is necessary to divide this dataset into subsets corresponding to separate tasks.
+\textbf{In this way, we aim to enhance the learning of information about external factor reflected in single dataset} \cite{fama1970efficient}.
 In summary, the prepared dataset includes $n$ tasks: $\mathcal{D} = \left\{ \mathcal{D}_t \right\}_{t=1}^{n}$.
@@
 We propose to use the \verb|CNN| network to emphasize local features, thereby directing part of the model's attention to certain time points.
-Therefore, the proposed method not only remembers long-term features but also highlights short-term features.
+\textbf{Therefore, the proposed method not only remembers long-term features but also highlights short-term features.
+In the ideal situation, hidden constraints in aperiodic time-series data can be extracted efficiently}.
 
@@
 stacking models).
-From the perspective of ensemble models, the equation \ref{eq:outer_sol} can be considered an optimization-based method of synthesizing sub-models, which helps to take advantage of the feature extraction capabilities of each model.
-In other words, the synthesized model can extract features at a deeper level, significantly improving the prediction ability compared to traditional ensemble models.
+\textbf{From the perspective of ensemble models, equation \ref{eq:outer_sol} can be considered an optimization-based method of synthesizing sub-models, which helps to take advantage of the feature extraction capabilities of each model.
+Moreover, equation \ref{eq:outer_sol} produces a fast-adaptive model which is extremely suitable for the context of unstable variance in data}.
+In other words, the synthesized model can quickly adapt to data with new variance as well as extract features at a deeper level, which significantly improves the prediction ability compared to traditional ensemble models.

### 2025-01-04 doc/thesis/chapters/4.experiment.tex
split chap.4 into proposed method and baseline model

@@
 \label{chap:experiment}
+
+Chương 4 mô tả các thí nghiệm để đánh giá \verb|Temporal-ML| và so sánh với \verb|NHITS| trên các tập dữ liệu lớn và phổ biến.
+Các thí nghiệm được thực hiện trên aperiodic data bao gồm một dataset duy nhất và aperiodic data bao gồm nhiều datasets để kiểm chứng khả năng rút trích các thông tin về thị trường cũng như khả năng tổng hợp thông tin từ nhiều nguồn.
+Dữ liệu có chu kỳ cũng được sử dụng để tăng tính khách quan trong các đánh giá.
+
+% ablation study
+% Ngoài ra, so sánh giữa \verb|Temporal-ML| và các mô hình đơn lẻ \verb|LSTM|, \verb|LSTM+CNN| cũng được thực hiện như một ablation study.
 
@@
 
-Foreign exchange (FX) in particular and financial indices in general are typical data types for aperiodic time-series data.
+Foreign exchange in particular and financial indices in general are typical data types for aperiodic time-series data.
 Therefore, we choose this type of data to test the model.
-Specifically, we configure two datasets using FX data.
+Specifically, we configure two datasets using foreign exchange data.
 \verb|USD/JPY| dataset consists of only data of the exchange rate between US dollar and Japanese yen.
@@
 
-\section{Data pre-process}
+\section{Temporal-ML}
 
-\subsection{Temporal-ML}
-\label{subsec:ml_experiment}
+\subsection{Data pre-processing}
+\label{subsec:ml_preprocess}
 
@@
 
-\begin{table}
+\begin{table}[H]
     \centering
@@
         \centering
-        \includegraphics[width=\textwidth]{multi_fx (1).png}
+        \includegraphics[width=\textwidth]{multi_fx.png}
         \caption{}
@@
 
-\subsection{Baseline model}
-\label{subsec:baseline_experiment}
+\subsection{Metric evaluation}
 
-% viết về cách tổng hợp metric ở đây
-Nghiên cứu này sử dụng thuật toán \verb|NHITS| \cite{challu2023nhits} làm baseline model.
-\verb|NHITS| sử dụng toàn bộ thuộc tính để dự đoán xu hướng tiếp theo cho một thuộc tính trong một tập dữ liệu.
-Thuật toán hướng đến phân rã các giải tần và kết hợp lại để dự đoán tương lai.
-Đối với các tập dữ liệu \verb|USD/JPY|, \verb|ETT-m2|, và \verb|WTH|, dữ liệu tại mỗi tập được chia thành training set, validation set, và test set với tỷ lệ tương ứng 6:2:2.
+Nghiên cứu sử dụng \textit{Accuracy, Precision, Recall}, và \textit{F1-score} để đánh giá các mô hình (equations \ref{eq:acc}, \ref{eq:P}, \ref{eq:R}, \ref{eq:F1}).
+Trong đó, \textit{Accuracy} được sử dụng để tính tỷ lệ giữa số mẫu phân lớp đúng và tổng số mẫu được dự đoán, \textit{Precision} đo mức độ tin cậy của mô hình khi nó gán nhãn cho một mẫu dữ liệu, \textit{Recall} kiểm định tỷ lệ bỏ sót các mẫu dữ liệu của một phân lớp, \textit{F1-score} kiểm tra mức độ bias của mô hình trên các nhãn có số lượng lớn bằng cách lấy trung bình điều hòa trên \textit{Precision} và \textit{Recall}.
 
-Chúng tôi tiền xử lý dữ liệu huấn luyện bằng \verb|z-score| như đã đề cập ở subsection \ref{subsec:ml_experiment}.
-Sau đó, $\mu_{X}$ và $\sigma_{X}$ thu được từ phương trình \ref{eq:mean} và \ref{eq:std} được sử dụng để chuẩn hóa cho thuộc tính $X$ trong validation set trong quá trình tuning.
-Khi đã chọn được các siêu tham số tốt nhất, dữ liệu của training set và validation set được chuẩn hóa lại từ đầu, sau đó được huấn luyện với bộ siêu tham số tốt nhất để dự đoán test set.
-
-Đối với tập dữ liệu \verb|multi-fx|, vì \verb|NHITS| không có cơ chế tổng hợp mô hình trên các tập dữ liệu khác nhau, chúng tôi lặp lại quy trình huấn luyện (huấn luyện, tuning, testing trên 60\%, 20\%, 20\% dữ liệu, respectively) cho từng cặp tiền tệ, sau đó tổng hợp metrics trên các cặp tiền tệ để thu được đánh giá cuối.
-
-\section{Metric evaluation}
-
-Nghiên cứu sử dụng accuracy, precision, recall, và F1 để đánh giá các mô hình:
-
-\begin{align*}
-    acc &= \frac{TP+TN}{TP+FP+TN+FN}\\
-    P &= \frac{TP}{TP+FP}\\
-    R &= \frac{TP}{TP+FN}\\
-    F1 &= \frac{2PR}{P+R}
-\end{align*}Trong đó, $TP, TN, FP, FN$ lần lượt là số mẫu \textit{true\_positive, true\_negative, false\_positive, false\_negative}.
+\begin{align}
+    acc &= \frac{TP+TN}{TP+FP+TN+FN}\label{eq:acc}\\
+    P &= \frac{TP}{TP+FP}\label{eq:P}\\
+    R &= \frac{TP}{TP+FN}\label{eq:R}\\
+    F1 &= \frac{2PR}{P+R}\label{eq:F1}\\
+\end{align}Trong đó, $TP, TN, FP, FN$ lần lượt là số mẫu \textit{true\_positive, true\_negative, false\_positive, false\_negative}.
 
@@
 
-\subsection{Temporal-ML}
-
 Giả sử thuật toán \verb|Temporal-ML| được đánh giá trên tập dữ liệu $\mathcal{D}$ gồm $a$ thuộc tính và được chia thành $n$ tasks.
 Quá trình đánh giá được thực hiện bằng cách để mô hình dự đoán xu hướng của từng thuộc tính với đầu vào là tất cả các thuộc tính, sau đó thực hiện hai bước tổng hợp: (1) - Tổng hợp trên task; (2) - Tổng hợp trên thuộc tính.
+Quá trình đánh giá này được thiết kế dựa trên \cite{challu2023nhits}.
 
@@
 
-\subsection{Baseline model}
-
-\verb|NHITS| cũng được đánh giá trên tập dữ liệu $\mathcal{D}$ với $a$ thuộc tính và phải thực hiện dự đoán trên từng thuộc tính.
-Xét metric $m$, sau quá trình dự đoán thuộc tính $k$, ta thu được $a$ giá trị: $\{m^{(k)}_1,\dots,m^{(k)}_a\}$.
-Quá trình tổng hợp metric $m$ của $a$ thuộc tính được thực hiện như sau:
-
-\begin{align*}
-    \bar{m} &= \frac{1}{a}\sum_{k=1}^a{m^{(k)}}\\
-    s_m &= \sqrt{\frac{1}{a} \sum_{i=1}^n{(m^{(k)} - \bar{m})^2}}
-\end{align*}
-
-Như đã đề cập ở trên, \verb|NHITS| không có cơ chế tổng hợp mô hình khi kiểm thử trên dữ liệu \verb|multi-fx|.
-Do đó, chúng tôi kiểm thử \verb|NHITS| trên từng cặp tiền tệ trong \verb|multi-fx| và tổng hợp metrics giống như \verb|Temporal-ML|.
-
-\subsection{Fairness in evaluation}
-
-Gọi $m$ là số mẫu dữ liệu trong mỗi tasks, tổng số dữ liệu của quá trình huấn luyện, kiểm thử, và kiểm tra là: $mn$.
-Đối với \verb|NHITS|, mô hình được huấn luyện, kiểm thử, kiểm tra trên 60\%, 20\%, 20\% dữ liệu, respectively:
-
-\begin{align*}
-    \text{Train: } &0.6mn\\
-    \text{Validation: } &0.2mn\\
-    \text{Testing: } &0.2mn
-\end{align*}
-
-Đối với \verb|Temporal-ML|, quá trình làm giàu knowledge cho mô hình sử dụng toàn bộ dữ liệu huấn luyện và tập support của dữ liệu kiểm thử/kiểm tra.
-Quá trình kiểm thử/kiểm tra được thực hiện trên tập query:
-
-\begin{align*}
-    \text{Train: } &0.5mn + 20\%\left( \frac{m}{2}n \right) = 0.6mn\\
-    \text{Validation: } &80\% \left( \frac{m}{2}\frac{n}{2} \right)  =0.2mn\\
-    \text{Testing: } &80\% \left( \frac{m}{2}\frac{n}{2} \right) =0.2mn
-\end{align*}
-
-Do đó, với cách chia dữ liệu như trên, chúng tôi đạt được sự công bằng về số lượng mẫu dữ liệu trong quá trình kiểm thử.
-
-\section{Hyper parameters tuning}
-% quá trình fine-tune (kiểu cố định outer và quan sát inner) --> chuyển xuống phần phụ lục?
-
-\subsection{Temporal-ML}
+\subsection{Fine-tuning}
 
@@
 
-\subsection{Baseline model}
+\section{Baseline model}
+
+\subsection{Data pre-processing}
+\label{subsec:baseline_prepocess}
+
+% viết về cách tổng hợp metric ở đây
+Nghiên cứu này sử dụng thuật toán \verb|NHITS| (AAAI 2023) làm baseline model.
+\verb|NHITS| sử dụng toàn bộ thuộc tính để dự đoán xu hướng tiếp theo cho một thuộc tính trong một tập dữ liệu.
+Thuật toán hướng đến phân rã các giải tần và kết hợp lại để dự đoán tương lai.
+Đối với các tập dữ liệu \verb|USD/JPY|, \verb|ETT-m2|, và \verb|WTH|, dữ liệu tại mỗi tập được chia thành training set, validation set, và test set với tỷ lệ tương ứng 6:2:2.
+
+Chúng tôi tiền xử lý dữ liệu huấn luyện bằng \verb|z-score| như đã đề cập ở subsection \ref{subsec:ml_preprocess}.
+Sau đó, $\mu_{X}$ và $\sigma_{X}$ thu được từ phương trình \ref{eq:mean} và \ref{eq:std} được sử dụng để chuẩn hóa cho thuộc tính $X$ trong validation set trong quá trình tuning.
+Khi đã chọn được các siêu tham số tốt nhất, dữ liệu của training set và validation set được chuẩn hóa lại từ đầu, sau đó được huấn luyện với bộ siêu tham số tốt nhất để dự đoán test set.
+
+Đối với tập dữ liệu \verb|multi-fx|, vì \verb|NHITS| không có cơ chế tổng hợp mô hình trên các tập dữ liệu khác nhau, chúng tôi lặp lại quy trình huấn luyện (huấn luyện, tuning, testing trên 60\%, 20\%, 20\% dữ liệu, respectively) cho từng cặp tiền tệ, sau đó tổng hợp metrics trên các cặp tiền tệ để thu được đánh giá cuối.
+
+\subsection{Metric evaluation}
+
+\verb|NHITS| cũng sử dụng các metrics tượng tự \verb|Temporal-ML| và được đánh giá trên tập dữ liệu $\mathcal{D}$ với $a$ thuộc tính và phải thực hiện dự đoán trên từng thuộc tính.
+Xét metric $m$, sau quá trình dự đoán thuộc tính $k$, ta thu được $a$ giá trị: $\{m^{(k)}_1,\dots,m^{(k)}_a\}$.
+Quá trình tổng hợp metric $m$ của $a$ thuộc tính được thực hiện như sau:
+
+\begin{align*}
+    \bar{m} &= \frac{1}{a}\sum_{k=1}^a{m^{(k)}}\\
+    s_m &= \sqrt{\frac{1}{a} \sum_{i=1}^n{(m^{(k)} - \bar{m})^2}}
+\end{align*}
+
+Như đã đề cập, \verb|NHITS| không có cơ chế tổng hợp mô hình khi kiểm thử trên dữ liệu \verb|multi-fx|.
+Do đó, chúng tôi kiểm thử \verb|NHITS| trên từng cặp tiền tệ trong \verb|multi-fx| và tổng hợp metrics giống như \verb|Temporal-ML|.
+
+\subsection{Fine-tuning}
 
@@
     \label{tab:nhits_finetune}
-    \begin{tabular}{ll} 
+    \resizebox{\columnwidth}{!}{\begin{tabular}{ll} 
     \toprule
@@
     \bottomrule
-    \end{tabular}
+    \end{tabular}}
 \end{table}
+
+\section{Fairness in evaluation}
+
+Gọi $m$ là số mẫu dữ liệu trong mỗi tasks, tổng số dữ liệu của quá trình huấn luyện, kiểm thử, và kiểm tra là: $mn$.
+Đối với \verb|NHITS|, mô hình được huấn luyện, kiểm thử, kiểm tra trên 60\%, 20\%, 20\% dữ liệu, respectively:
+
+\begin{align*}
+    \text{Train: } &0.6mn\\
+    \text{Validation: } &0.2mn\\
+    \text{Testing: } &0.2mn
+\end{align*}
+
+Đối với \verb|Temporal-ML|, quá trình làm giàu knowledge cho mô hình sử dụng toàn bộ dữ liệu huấn luyện và tập support của dữ liệu kiểm thử/kiểm tra.
+Quá trình kiểm thử/kiểm tra được thực hiện trên tập query:
+
+\begin{align*}
+    \text{Train: } &0.5mn + 20\%\left( \frac{m}{2}n \right) = 0.6mn\\
+    \text{Validation: } &80\% \left( \frac{m}{2}\frac{n}{2} \right)  =0.2mn\\
+    \text{Testing: } &80\% \left( \frac{m}{2}\frac{n}{2} \right) =0.2mn
+\end{align*}
+
+Do đó, với cách chia dữ liệu như trên, chúng tôi đạt được sự công bằng về số lượng mẫu dữ liệu trong quá trình kiểm thử.
+
+% i am not sure about this
+\section{Ablation study}
+
+% có kq rồi mình tính tiếp
+
+Chúng tôi thực hiện đánh giá từng component trong \verb|Temporal-ML| bằng cách loại bỏ chúng khỏi thuật toán và so sánh kết quả với thuật toán gốc.
+This includes three models without \verb|MAML| (\verb|LSTM|, \verb|CNN|, \verb|LSTM+CNN|), and

### 2025-01-04 doc/thesis/chapters/5.results.tex
add table chap. 5

@@
 % vẽ, mô tả cái bảng
-% bẻ nhỏ từng attribute trong fx dataset --> thành 4 bảng, 1 bảng tổng hợp là 5
 % bảng tổng hợp vẽ ở đây
+\begin{table}[H]
+    \centering
+    \cprotect\caption{Classification results (\%) of \verb|Temporal-ML| and \verb|NHITS|. Best results per metrics are boldfaced.}
+    \label{tab:rs}
+    \resizebox{\columnwidth}{!}{\begin{tabular}{c|c|cccc}
+    \toprule
+    \multicolumn{1}{c}{}                &                        & $accuracy$              & $precision$             & $recall$                & $F1-score$                 \\
+    \hline
+    \multirow{3}{*}{\Verb|USD/JPY|}     & \Verb|NHITS|           & $58.46\pm9.60$          & $58.24\pm9.89$          & $57.65\pm10.08$         & $55.82\pm11.22$            \\
+                                        & Ours(\Verb|LSTM|)      & $\mathbf{70.33\pm1.69}$ & $\mathbf{70.73\pm1.82}$ & $\mathbf{70.26\pm1.93}$ & $\mathbf{69.28\pm2.50}$    \\
+                                        & Ours(\Verb|LSTM+CNN|)  & $68.75\pm1.78$          & $69.15\pm1.55$          & $68.73\pm1.77$          & $69.14\pm2.17$             \\
+    \hline
+    \multirow{3}{*}{\Verb|multi-fx|}    & \Verb|NHITS|           & $-$                 & $-$                 & $-$                 & $-$              \\
+                                        & Ours(\Verb|LSTM|)      & $\mathbf{66.26\pm7.45}$ & $\mathbf{67.08\pm8.97}$ & $\mathbf{65.76\pm7.79}$ & $\mathbf{64.06\pm10.00}$   \\
+                                        & Ours(\Verb|LSTM+CNN|)  & $66.08\pm8.19$          & $66.56\pm9.23$          & $65.41\pm8.66$          & $63.78\pm10.83$            \\
+    \hline
+    \multirow{3}{*}{\Verb|ETT-m2|}      & \Verb|NHITS|           & $\mathbf{71.88\pm14.06}$& $\mathbf{66.86\pm10.21}$& $\mathbf{62.49\pm10.40}$& $\mathbf{62.69\pm10.72}$   \\
+                                        & Ours(\Verb|LSTM|)      & $71.14\pm9.13$          & $62.21\pm7.51$          & $58.75\pm5.30$          & $57.61\pm6.74$             \\
+                                        & Ours(\Verb|LSTM+CNN|)  & $71.40\pm9.10$          & $63.25\pm8.82$          & $60.22\pm8.42$          & $59.43\pm9.24$             \\
+    \hline
+    \multirow{3}{*}{\Verb|WTH|}         & \Verb|NHITS|           & $74.18\pm8.11$          & $68.05\pm6.65$          & $65.04\pm7.67$          & $65.40\pm7.97$             \\
+                                        & Ours(\Verb|LSTM|)      & $\mathbf{74.97\pm2.63}$ & $\mathbf{69.13\pm3.44}$ & $\mathbf{65.98\pm3.96}$ & $\mathbf{66.00\pm3.80}$    \\
+                                        & Ours(\Verb|LSTM+CNN|)  & $74.41\pm2.71$          & $68.17\pm4.60$          & $65.50\pm4.82$          & $65.63\pm4.47$             \\
+    \bottomrule
+    \end{tabular}}
+\end{table}
+
 % chắc phải viết thêm cả CNN
@@
 
+% bẻ nhỏ từng attribute trong fx dataset --> thành 4 bảng, 1 bảng tổng hợp là 5
 % 4 bảng phụ vẽ ở đây

### 2025-01-21 doc/thesis/chapters/2.related_works.tex
cnn, ensemble, rnn

@@
 
+% Phương sai trên time-series data là không cố định và thường thay đổi rất lớn qua các thời điểm.
+Sự thay đổi này thường được gọi là concept drift scenarios \cite{liu2025onsitnet}.
+Để tăng tính ổn định cho hệ thống dự đoán trong ngữ cảnh phương sai của dữ liệu biến thiên liên tục, thay vì sử dụng một mô hình, ensemble approach đề xuất sử dụng nhiều mô hình và kết hợp chúng lại để tạo ra các dự đoán.
+
+The variance on time-series data is not constant and often varies greatly over time.
+This variation is often referred to as the concept of drift scenarios \cite{liu2025onsitnet}.
+To increase the stability of the prediction system in the context of continuously varying data variance, instead of using a single model, the ensemble approach proposes using multiple models and combining them to generate predictions.
+
+% Thay vì huấn luyện một learner từ training data, ensemble methods huấn luyện một set of learners để giải quyết cùng một vấn đề.
+Dạng tổng quát của ensemble model $g(\mathbf{x})$ được thể hiện thông qua việc tổng hợp các mô hình $\{h_i, i=1\dots T\}$ như sau:
+
+Instead of training a learner from training data, ensemble methods train a set of learners to solve the same problem.
+The general form of the ensemble model $g(\mathbf{x})$ is expressed by the aggregation of models $\{h_i, i=1\dots T\}$ as follows:
+
+\begin{equation}
+    g(\mathbf{x}) = \sum_{i=1}^T{\alpha_i h_i(\mathbf{x})} \text{, with } \sum_{i=1}^T{\alpha_i}=1
+    \label{eq:ensemble}
+\end{equation}
+
+% Trong phương trình \ref{eq:ensemble} phép tính tổng được đề cập một cách trừu tượng, ám chỉ quá trình tổng hợp.
+$h_i$ được gọi là base learner.
+Base learner có thể được tạo ra từ bất kỳ base learning algorithm nào (e.g., decision tree, neural network, etc.,).
+Hầu hết các ensemble methods sử dụng cùng một base learning algorithm với training data khác nhau để huấn luyện base learners.
+Điều này tạo ra các base learners cùng loại (i.e., các base learners đều là decision tree hoặc neural network, etc.,).
+
+In equation \ref{eq:ensemble} the summation operation is mentioned abstractly, referring to the process of ensemble.
+$h_i$ is called the base learner.
+The base learner can be created from any base learning algorithm (e.g., decision tree, neural network, etc.,).
+Most ensemble methods use the same base learning algorithm with different training data to train the base learners.
+This results in base learners of the same type (i.e., the base learners are all decision trees or neural networks, etc.,).
+
+% Khả năng tổng quát hóa của ensemble model thường tốt hơn các base learners vì chúng có kết hợp các weak learners, vốn chỉ có thể dự đoán tốt hơn mức ngẫu nhiên một chút, thành một strong learners, có khả năng dự đoán với độ chính xác tốt hơn nhiều lần.
+Do đó, base learners thường được gọi là weak learners.
+Thật vậy, đối với mức noise tăng cần trong dữ liệu, sự kết hợp các mô hình luôn cho dự đoán tốt hơn so với việc sử dụng một mô hình đơn lẻ để dự đoán.
+
+Ensemble models are often better at generalizing than base learners because they combine weak learners, which can predict only slightly better than random, to built a strong learner, which can predict with much better accuracy.
+Hence, base learners are often called weak learners.
+Indeed, when increasing levels of noise in the data, ensemble models always give better predictions than a single model.
+
+% Ngoài ra, chi phí tính toán cho quá trình tạo ra nhiều base learners và kết hợp chúng không quá lớn so với việc xây dựng một single model bởi vì khi tạo ra một single model, chúng ta thường phải tạo ra rất nhiều phiên bản của mô hình đó trong quá trình tuning.
+Việc này tương đương với việc tạo ra nhiều base learners trong ensemble approach.
+Quá trình kết hợp các đặc trưng của ensemble model thông thường không tốn quá nhiều chi phí vì chiến lược kết hợp được sử dụng thường rất đơn giản (e.g., voting, averaging).
+
+In addition, the computational cost of creating multiple base learners and combining them is not too large compared to building a single model because when creating a single model, we often have to create many versions of that model during tuning.
+This is equivalent to generating multiple base learners in the ensemble approach.
+The process of combining features of an ensemble model is usually not too expensive because the combination strategy is usually very simple (e.g., voting, averaging).
+
+% Dựa trên quá trình tạo ra các base learners, chúng tôi đề cập đến hai hướng tiếp cận chính các phương pháp ensemble: \verb|Bagging| and \verb|Boosting| model.
+Trong đó, \verb|Bagging| tạo ra các base learners một cách song song còn \verb|Boosting| tạo ra các base learners một cách tuần tự.
+
+Based on the process of generating base learners, we refer to two main approaches to ensemble methods: \verb|Bagging| and \verb|Boosting| models.
+Where, \verb|Bagging| generates base learners in parallel while \verb|Boosting| generates base learners sequentially.
+
+\subsection{Bagging model}
+
+% Động lực chính của \verb|Bagging| dựa trên việc khai thác sự độc lập của từng base learner.
+\verb|Bagging| giả định rằng, độ lỗi của mô hình có thể được giảm đáng kể bằng việc kết hợp các base learners độc lập.
+Khác với các hình dung ban đầu, từ ``bagging'' dù ám chỉ việc chia dữ liệu thành các bag nhưng không xuất phát từ từ ``bag'' mà là viết tắt của từ ``Bootstrap AGGregatING''.
+Như được đề cập trong tên gọi, \verb|Bagging| bao gồm hai ý tưởng chính là ``bootstrap'' và ``aggregation''.
+
+The main motivation of \verb|Bagging| is based on exploiting the independence of each base learner.
+\verb|Bagging| assumes that the model error can be significantly reduced by combining independent base learners.
+Contrary to the initial conception, the word ``bagging'', although referring to dividing data into bags, does not come from the word ``bag'' but is an abbreviation of the word ``Bootstrap AGGregatING''.
+As mentioned in the name, \verb|Bagging| consists of two main ideas: ``bootstrap'' and ``aggregation''.
+
+% \verb|Bagging| sử dụng bootstrap distribution để tạo ra các base learners khác nhau.
+Nói cách khác, nó áp dụng kỹ thuật bootstrap sampling để thu được các mẫu và huấn luyện các base learners trên các mẫu riêng biệt.
+Cụ thể, cho trước $n$ điểm dữ liệu huấn luyện, các mẫu chứa $n$ điểm dữ liệu được tạo ra bằng cách sampling with replacement từ $n$ điểm dữ liệu ban đầu.
+Bằng cách lặp lại quá trình này $T$ lần, $T$ mẫu được tạo thành.
+
+\verb|Bagging| uses bootstrap distribution to generate different datasets for training base learners.
+In other words, it applies bootstrap sampling technique to obtain samples and train base learners on separate samples.
+Specifically, given $n$ training data points, samples containing $n$ data points are generated by sampling with replacement from $n$ original data points.
+By repeating this process $T$ times, $T$ samples are generated.
+
+% Trong quá trình tổng hợp, \verb|Bagging| sử dụng cơ chế voting cho bài toán phân lớp và averaging cho bài toán hồi quy để tổng hợp output từ các base learners.
+Cụ thể, một \verb|Bagging| model với $T$ base learners $\{h_i, i=1\dots T\}$ trong bài toán hồi quy sẽ dự đoán instance $\mathbf{x}$ như sau:
+
+During the aggregation process, \verb|Bagging| uses voting for the classification problem and averaging for the regression problem to aggregate output from base learners.
+Specifically, a \verb|Bagging| model with $T$ base learners $\{h_i, i=1\dots T\}$ in the regression problem will predict instance $\mathbf{x}$ as follows:
+
+\begin{equation}
+    \hat{y} = \frac{1}{T}\sum_{i=1}^T{h_i(\mathbf{x})}
+\end{equation}
+
+\subsection{Boosting model}
+
+% Ý tưởng của \verb|Boosting| đến từ việc cải thiện một cách tuần tự quá trình dự đoán của mô hình trên các dữ liệu mà trước đó nó đã dự đoán sai.
+Sự cải thiện này được thực hiện thông qua quá trình điểu chỉnh lại phân phối của tập dữ liệu.
+Chúng tôi sẽ trình bày một toy example để làm rõ ý tưởng này.
+
+The idea of \verb|Boosting| comes from sequentially improving the model's prediction performance on data that it previously predicted incorrectly.
+This improvement is done by adjusting the distribution of the data set.
+We will present a toy example to illustrate this idea.
+
+% Cho tập dữ liệu $\mathcal{D}$ được rút ra từ không gian $\mathcal{X}$.
+Không gian $\mathcal{X}$ bao gồm ba phần $\mathcal{X}_1, \mathcal{X}_2$ và $\mathcal{X}_3$ chiếm tỷ lệ đều nhau.
+Giả sử mô hình $h_1$ sau khi huấn luyện trên $\mathcal{D}$ có thể dự đoán tốt các mẫu dữ liệu thuộc $\mathcal{X}_1, \mathcal{X}_2$ và có hiệu suất kém trên các mẫu dữ liệu thuộc $\mathcal{X}_3$.
+Lúc này, chúng ta có thể chỉnh sửa lỗi sai của $h_1$ bằng cách huấn luyện một mô hình $h_2$ trên tập dữ liệu $\mathcal{D}'$.
+Trong đó, $\mathcal{D}'$ là tập dữ liệu hiệu chỉnh từ $\mathcal{D}$ nhằm mục đích nhấn mạnh các mẫu $x\in\mathcal{X}_3$.
+Sau quá trình huấn luyện, $h_2$ có thể khắc phục được nhược điểm của $h_1$ trên không gian $\mathcal{X}_3$.
+Bằng việc kết hợp hai mô hình này, mô hình output được kỳ vọng là sẽ perform tốt trên cả ba sub-space của $\mathcal{X}$.
+Tuy nhiên, giả sử rằng, mô hình này chỉ perform tốt trên $\mathcal{X}_1, \mathcal{X}_3$.
+Lúc này, ta lại tiếp tục hiệu chỉnh tập dữ liệu huấn luyện để thu được $\mathcal{D}''$.
+
+Given a data set $\mathcal{D}$ drawn from the space $\mathcal{X}$.
+The space $\mathcal{X}$ consists of three parts $\mathcal{X}_1, \mathcal{X}_2$ and $\mathcal{X}_3$ with equal proportions.
+Suppose that model $h_1$ after training on $\mathcal{D}$ can predict well on data samples belonging to $\mathcal{X}_1, \mathcal{X}_2$ and has poor performance on data samples belonging to $\mathcal{X}_3$.
+At this point, we can correct the error of $h_1$ by training a model $h_2$ on the data set $\mathcal{D}'$.
+In which, $\mathcal{D}'$ is the calibrated dataset from $\mathcal{D}$ to emphasize the samples $x\in\mathcal{X}_3$.
+After training, $h_2$ can overcome the weakness of $h_1$ on the space $\mathcal{X}_3$.
+By combining these two models, the output model is expected to perform well on all three sub-spaces of $\mathcal{X}$.
+However, suppose that this model only performs well on $\mathcal{X}_1, \mathcal{X}_3$.
+At this point, we continue to calibrate the training dataset to obtain $\mathcal{D}''$.
+
+% Tóm lại, \verb|Boosting| hoạt động bằng cách huấn luyện các base learners một cách tuần tự sau đó kết hợp chúng để dự đoán. Các learner sau tập trung cải thiện lỗi của các learner trước.
+
+In a nutshell, \verb|Boosting| works by training base learners sequentially and then combining them to make predictions. Later learners focus on improving the errors of earlier learners.
+
 % cơ chế tổng hợp cùi bắp
+% Điểm hạn chế của các mô hình ensemble nằm ở quá trình tổng hợp các base learners.
+Quá trình này dựa trên hai cơ chế voting và averaging ứng với bài toán phân lớp và bài toán hồi quy.
+Không khó để nhận thấy rằng hai cơ chế này thực sự rất đơn giản và cứng nhắc.
+Điều đó khiến cho ensemble approach dù cải thiện được khả năng tổng quát hóa so với việc huấn luyện một single model nhưng khả năng tổng quá hóa này thực sự chưa cao.
+
+The limitation of ensemble models lies in the process of synthesizing base learners.
+This process is based on two mechanisms: voting and averaging, corresponding to the classification problem and the regression problem.
+\textbf{It is not difficult to see that these two mechanisms are really simple and rigid.
+That makes the ensemble approach, although improving the generalization ability compared to training a single model, the generalization ability is not really high.}
 
@@
 
+% phép tính convolution
+% Điểm khởi đầu của Convolution neural network (\verb|CNN|) bắt nguồn từ convolution operation.
+Do đó, khi nói về \verb|CNN|, không thể không nhắc đến convolution operation.
+Giả sử chúng ta sử dụng một nguồn thu phát sóng để định vị một vật thể đang chuyển động.
+Thiết bị thu phát sóng này trả về một giá trị duy nhất $x(t)$, thể hiện vị trí của vật thể tại thời điểm $t$.
+Chúng ta có thể truy xuất vị trí này tại mọi thời điểm.
+Không may, giá trị trả về của thiết bị của chúng ta không thực sự chính xác mà bị nhiễu.
+Do đó, để thu được kết quả chính xác hơn tại một thời điểm $t$, chúng ta phải tổng hợp $x(t)$ và giá trị trước đó của nó: $x(a), a=t-1, t-2, etc.$ Trên thực tế, các thời điểm $a$ càng gần $t$ thì càng ảnh hưởng đến $x(t)$.
+Do đó, một weighting function được required để đánh trọng số cho các giá trị $x(a)$.
+Lúc này, giá trị ước lượng vị trí của vật tại thời điểm $t$ được tính bằng hàm $s(t)$ như sau:
+
+The starting point of Convolution neural network (\verb|CNN|) is originated from convolution operation.
+Therefore, when talking about \verb|CNN|, it is impossible not to mention convolution operation.
+Suppose we use a sensor to locate a moving object.
+This sensor returns a single value $x(t)$, representing the position of the object at time $t$.
+We can retrieve this position at any time.
+Unfortunately, the returned value of our device is not really accurate but noisy.
+Therefore, to obtain a more accurate result at a time $t$, we have to aggregate $x(t)$ and its previous values: $x(a), a=t-1, t-2, etc.$ In fact, the closer $a$ is to $t$, the more it affects $x(t)$.
+Therefore, a weighting function is required to weight the values of $x(a)$.
+Now, the estimated value of the object's position at time $t$ is calculated using the function $s(t)$ as follows:
+
+\begin{align}
+    s(t) &= \int_{a=-\infty}^\infty{x(a) w(t-a) da}\\
+    &= (x*w)(t)
+\end{align}
+
+% Phép tính trên được gọi là \textit{convolution} operation. Trên miền rời rạc, convolution operation hoạt động như sau:
+
+The above operation is called \textit{convolution} operation. In discrete domains, the convolution operation works as follows:
+
+\begin{align}
+    (x*w)(t) = \sum_{a=-\infty}^\infty{x(a) w(t-a)}
+\end{align}
+
 % viết về conv1d, 2d, 3d...
+% Điểm mấu chốt của phép tính này là chỉ ra một quan hệ cục bộ giữa đối tượng đang xét ($x(t)$) và các đối tượng $x(a)$ với $a$ gần $t$.
+Lấy cảm hứng từ điều này, nghiên cứu \cite{lecun1989handwritten} đề xuất mạng \verb|CNN|, sử dụng hai thành tố input (ứng với $x$) và kernel (ứng với $w$) và \textit{cross-correlation} operation thay vì convolution operation truyền thống.
+Tuy nhiên, \cite{lecun1989handwritten} vẫn gọi nó là convolution neural network.
+Về cơ bản, cross-correlation operation nhìn khá giống convolution operation ngoại trừ một sự thay đổi nhỏ trong index của input và kernel để thuật tiện cho việc cài đặt.
+Cụ thể, cross-correlation operation hoạt động như sau:
+
+The key point of this operation is to show a local relationship between the object under consideration ($x(t)$) and the objects $x(a)$ with $a$ near $t$.
+Inspired by this, the research \cite{lecun1989handwritten} proposed a \verb|CNN| network, using two components input (corresponding to $x$) and kernel (corresponding to $w$) and \textit{cross-correlation} operation instead of the traditional convolution operation.
+However, \cite{lecun1989handwritten} still calls it a convolution neural network.
+Basically, the cross-correlation operation looks quite similar to the convolution operation except for a small change in the index of the input and kernel for convenience in implementation.
+Specifically, the cross-correlation operation works as follows:
+
+\begin{align}
+    (x*w)(t) = \sum_{a=-\infty}^\infty{x(t+a) w(a)}
+\end{align}
+
+% Điều đáng chú ý là \verb|CNN| không tổng hợp thông tin theo một chiều (trong ví dụ của chúng ta là chiều thời gian), mà còn có thể tổng hợp thông tin theo hai chiều (đối với ảnh RGB hoặc gray scale thông thường) hoặc ba chiều (đối với ảnh quang phổ bao gồm nhiều ảnh chụp đơn xếp chồng lên nhau).
+Tại đây, chúng tôi trình bày hai ứng dụng phổ biến của \verb|CNN| ứng với việc tổng hợp thông tin theo một chiều và hai chiều trên time-series data và dữ liệu dạng ảnh.
+Lưu ý rằng, các dữ liệu này có tính rời rạc, do đó, dấu tích phân trong phương trình gốc được thay bằng dấu tổng.
+Ngoài ra, do các thời điểm/vị trí càng xa sẽ ảnh hưởng càng ít đến thời điểm/vị trí đang xét, một giả định được sử dụng thường xuyên trong \verb|CNN| là sử ảnh hưởng là hoàn toàn cục bộ.
+Nghĩa là kernel được sử dụng sẽ có các giá trị khác 0 trong tầm ảnh hưởng và bằng 0 ở ngoài tầm ảnh hưởng.
+Có thể hình dung đơn giản rằng kernel như một window trượt dọc theo chiều của dữ liệu để tổng hợp thông tin.
+
+It is worth noting that \verb|CNN| can aggregate information not only in one dimension (in our example, the time dimension), but can also aggregate information in two dimensions (for conventional RGB or gray scale images) or three dimensions (for spectral images consisting of multiple single-shot images stacked on top of each other).
+Here, we present two common applications of \verb|CNN| for one-dimensional and two-dimensional aggregation on time-series data and image data.
+Note that these data are discrete, so the integral sign in the original equation is replaced by a summation sign.
+In addition, since farther away positions/timestamps have less influence on the current position/timestamp, a commonly used assumption in \verb|CNN| is that the influence is totally local.
+That is, the kernel has non-zero values within the influence range and zero values outside the influence range.
+It can be simply visualized the kernel as a window that slides along the direction of the data to synthesize information.
+
+% Ký hiệu $I, K$ lần lượt là input và kernel.
+Đối với time-series data, kernel chỉ có thể trượt dọc theo chiều thời gian của dữ liệu nên convolution operation được ký hiệu là \verb|Conv1D|.
+Giả sử dữ liệu đang xét là close price của cổ phiếu theo ngày của một công ty, kernel $K$ quy định mức ảnh hưởng cục bộ trong 5 ngày.
+Tại ngày thứ $i$, đặc trưng $F(i)$ được tổng hợp như sau:
+
+Denote $I, K$ as the input and kernel, respectively.
+For time-series data, the kernel can only slide along the time direction of the data, so the convolution operation is denoted as \verb|Conv1D|.
+Suppose the data under consideration is the daily close price of a company's stock, the kernel $K$ specifies the local influence over 5 days.
+At day $i$, the feature $F(i)$ is synthesized as follows:
+
+\begin{align}
+    F(i) = (I*K)(i) = \sum_{m=1}^5{I(i+m) K(m)}
+\end{align}
+
+% Đối với dữ liệu dạng ảnh thông thường, $K$ trở thành một ma trận quy định mức ảnh hưởng của các pixel cục bộ lên pixel đang xét và có thể trượt theo hai chiều từ trái sang phải và từ trên xuống dưới.
+Do đó, convolution operation được ký hiệu là \verb|Conv2D|.
+Giả sử mức ảnh hưởng này là một vùng hình chữ nhật kích thước $(M\times N)$, tại pixel $(i,j)$, đặc trưng $F(i,j)$ được tổng hợp như sau:
+
+For normal image data, $K$ becomes a matrix that determines the influence of local pixels on the pixel under consideration and can be slid in both directions from left to right and from top to bottom.
+Therefore, the convolution operation is denoted as \verb|Conv2D|.
+Suppose this influence is a rectangular region of size $(M\times N)$, at pixel $(i,j)$, the feature $F(i,j)$ is synthesized as follows:
+
+\begin{align}
+    F(i,j) = (I*K)(i,j) = \sum_m^M\sum_n^N{I(i+m, j+n)K(m,n)}
+\end{align}
+
+% Trong quá trình huấn luyện, \verb|CNN| cố gắng học một mức ảnh hưởng hợp lý từ đó có thể rút trích hiệu quả các đặc trưng không gian, thời gian của dữ liệu đầu vào.
+Sự hiệu quả của \verb|CNN| đến từ hai ý tưởng chính.
+Thứ nhất, dựa trên giả định rằng các đối tượng cục bộ mang lại nhiều ý nghĩa hơn đối với đối tượng đang xét hơn các đối tượng toàn cục, \verb|CNN| thu nhỏ không gian tìm kiếm tham số của mình từ một tập các tham số tương tác toàn diện với tất cả input thành một tập các tham số chỉ tương tác với các đối tượng cục bộ (kernel) nhưng mang lại nhiều ý nghĩa hơn.
+Giả sử mạng neural có $m$ inputs và $n$ output.
+Bằng việc sử dụng kernel kích thước $k\ll m$, độ phức tạp thời gian trong quá trình tính toán giảm từ $\mathcal{O}(mn)$ xuống còn $\mathcal{O}(kn)$ nhưng lại tăng được hiệu quả của mạng.
+Thứ hai, trong quá trình học, \verb|CNN| chỉ sử dụng duy nhất một kernel.
+Điều này làm giảm chi phí lưu trữ từ $\mathcal{O}(mn)$ trong mạng neural truyền thống xuống còn $\mathcal{O}(k)$.
+
+During training, \verb|CNN| attempts to learn a reasonable influence level to efficiently extract the spatial and temporal features of the input data.
+The effectiveness of \verb|CNN| comes from two main ideas.
+First, based on the assumption that local features are more meaningful to the object under consideration than global features, \verb|CNN| reduces its parameter search space from a set of parameters that interact comprehensively with all inputs to a set of parameters that interact only with local features (kernels) but are more meaningful.
+Suppose the neural network has $m$ inputs and $n$ outputs.
+By using a kernel of size $k\ll m$, the time complexity of the computation is reduced from $\mathcal{O}(mn)$ to $\mathcal{O}(kn)$ but the efficiency of the network is increased.
+Second, during learning, \verb|CNN| uses only one kernel.
+This reduces the space complexity from $\mathcal{O}(mn)$ in traditional neural networks to $\mathcal{O}(k)$.
+
+% Tuy nhiên, đối với time-series data nói chung và aperiodic time-series data nói riêng, tính chất dữ liệu tại một thời điểm bất kỳ không chỉ phụ thuộc vào các thời điểm gần đó mà còn phụ thuộc rất nhiều vào các thời điểm dài hạn trong quá khứ.
+Do đó, \verb|CNN| chỉ được sử dụng như một phương pháp hỗ trợ rút trích đặc trưng trong quá trình giải quyết các bài toán trên dữ liệu time-series.
+
+\textbf{However, for time-series data in general and aperiodic time-series data in particular, the data properties at any given time depend not only on recent times but also on long-term times in the past.
+Therefore, \Verb|CNN| is only used as a method to support feature extraction in the process of solving problems on time-series data.}
+
+% trong TH tệ, có thể viết thêm luồng làm việc của CNN: CNN -> pooling -> CNN -> pooling -> flatten -> clf head
 
@@
 
+Thay vì hoạt động trên một lướt pixel như \verb|CNN|, Recurrent neural network (\verb|RNN|) \cite{rumelhart1986learning} hoạt động trên một chuỗi các giá trị $\mathbf{x}^{(1)},\dots,\mathbf{x}^{(\tau)}$.
+So với mạng mạng \verb|FullyConnected|, \verb|RNN| có thể mở rộng số lượng datapoint trong một chuỗi đến vô hạn mà không cần tăng kích thước của mạng.
+Nói cách khác, \verb|RNN| có khả năng làm việc trên các chuỗi dài hiệu quả hơn nhiều so với mạng neural truyền thống.
+
+Tương tự như \verb|CNN|, \verb|RNN| cũng sử dụng một kiến trúc mạng cho phép chia sẻ tham số.
+Tuy nhiên, \verb|CNN| chia sẻ tham số trong nội bộ của một mẫu dữ liệu (cùng một kernel hoạt động trên các vùng cục bộ của một mẫu dữ liệu).
+\verb|RNN| chia sẻ tham số bằng cách sử dụng cùng một bộ tham số cho các mẫu dữ liệu khác nhau.
+Do đó, khi input một mẫu dữ liệu vào mô hình, output của mô hình chứa các thông tin của mẫu dữ liệu này và toàn bộ các mẫu dữ liệu trước đó.
+Điều này khiến cho việc mở rộng mô hình và apply mô hình lên các mẫu dữ liệu có độ dài khác nhau trở nên khả thi và hiệu quả hơn.
+
+Cụ thể, gọi hàm $f$ là mạng \verb|RNN|. Hàm $f$ sử dụng bộ tham số $\theta$ để rút trích thông tin từ trạng thái $\mathbf{s}^{(t-1)}$ để tạo ra trạng thái $\mathbf{s}^{(t)}$:
+
+\begin{align}
+    \mathbf{s}^{(t)} = f\left( \mathbf{s}^{(t-1)}; \theta \right)
+    \label{eq:recurrent}
+\end{align}
+
+Lưu ý rằng, $\mathbf{s}^{(t)}$ là một abstract variable, ám chỉ thông tin hiện tại phụ thuộc vào các thông tin trong quá khứ.
+Đối với chuỗi có độ dài $\tau$ bất kỳ, equation \ref{eq:recurrent} có thể được triển khai $\tau-1$ lần để nén các thông tin trong chuỗi vào các trạng thái $\mathbf{s}^{(t)}$.
+
+Trong các mạng \verb|RNN| thông thường, biến trạng thái $\mathbf{s}$ được thể hiện bằng thông tin rút trích được từ mạng $\mathbf{h}$ và dữ liệu $\mathbf{x}$.
+Cụ thể, trạng thái hiện tại $\mathbf{h}^{(t)}$ được rút trích dựa trên trạng thái trước đó $\mathbf{h}^{(t-1)}$ (chứa thông tin về toàn bộ chuỗi trước đó) và mẫu dữ liệu hiện tại $\mathbf{x}^{(t)}$:
+
+\begin{align}
+    \mathbf{h}^{(t)} = f\left( \mathbf{h}^{(t-1)}, \mathbf{x}^{(t)}; \theta \right)
+    \label{eq:rnn}
+\end{align}
+
+Phương trình \ref{eq:rnn} có thể được viết lại dưới dạng mô hình $g^{(t)}$ như sau:
+
+\begin{align}
+    \mathbf{h}^{(t)} = g^{(t)}\left( \mathbf{x}^{(t)}, \mathbf{x}^{(t-1)},\dots,\mathbf{x}^{(2)},\mathbf{x}^{(1)}; \theta \right)
+    \label{eq:rnn_g}
+\end{align}
+
+Từ đây, có thể thấy rằng, mô hình $g$ không cần phải cập nhật tham số khi học từng mẫu $x^{(t)}$ như mạng neural truyền thống.
+Thay vào đó, nó có thể sử dụng thông tin từ một chuỗi $\mathbf{x}^{(1)},\dots,\mathbf{x}^{(\tau)}$ để cập nhật tham số thông qua mô hình $f$.
+Ngoài ra, việc sử dụng mô hình $f$ còn cải thiện khả năng tổng quát hóa của mô hình đối với những chuỗi có độ dài bất kỳ.
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=\linewidth]{rnn.png}
+    \cprotect\caption{The computational graph of \verb|RNN| in next-word prediction problem \cite{goodfellow2016deep}.}
+    \label{fig:rnn}
+\end{figure}
+
+Dựa vào ý tưởng của recurrent function $f$ trình bày ở trên, chúng ta có thể thiết kế rất nhiều kiến trúc mạng \verb|RNN|.
+Tại đây, chúng tôi trình bày về một thiết kế đơn giản có đồ thị tính toán như hình \ref{fig:rnn}.
+Theo đó, mạng nhận vào một chuỗi các giá trị $\mathbf{x}$ với mục tiêu ánh xạ các giá trị này vào các đầu ra $\mathbf{o}$ khớp với $\mathbf{y}$.
+Để thực hiện điều này, mạng sử dụng ba bộ tham số chính: $\mathbf{U}$ để ánh xạ $\mathbf{x}$ vào không gian đặc trưng ẩn; $\mathbf{W}$ để ánh xạ đặc trưng ẩn của timestamp trước vào không gian đặc trưng ẩn của timestamp sau; $\mathbf{V}$ để ánh xạ đặc trưng ẩn vào không gian chứa các giá trị đầu ra chưa chuẩn hóa ($\mathbf{o}$).
+Giá trị dự đoán $\hat{y}$ sau đó được tính từ $\mathbf{o}$.
+Cả quá trình được thể hiện như sau:
+
+\begin{align}
+    \mathbf{a}^{(t)} &= \mathbf{b} + \mathbf{W}\mathbf{h}^{(t-1)} + \mathbf{U}\mathbf{x}^{(t)}\quad \text{($\mathbf{b}$ is bias vector)}\\
+    \mathbf{h}^{(t)} &= \tanh\left( \mathbf{a}^{(t)} \right)\\
+    \mathbf{o}^{(t)} &= \mathbf{c} + \mathbf{V}\mathbf{h}^{(t)}\quad \text{($\mathbf{c}$ is bias vector)}\\
+    \hat{y}^{(t)} &= \text{softmax}\left( \mathbf{o}^{(t)} \right)
+\end{align}
+
+Độ lỗi của mô hình được tính bằng hàm \verb|CrossEntropyLoss| (equation \ref{eq:loss_rnn}).
+Hàm lỗi này được sử dụng để cập nhật các tham số của hàm $f$ thông qua kỹ thuật back-propagation through time.
+
+\begin{align}
+    L\left( \left\{ \hat{y}^{(t)} \right\}_{t=1}^\tau, \left\{ y^{(t)} \right\}_{t=1}^\tau \right) = \sum_{t=1}^\tau{L^{(t)}\left( \hat{y}^{(t)}, y^{(t)} \right)}
+    \label{eq:loss_rnn}
+\end{align}
+
+Trong quá trình forward và backward, các bước tính toán được thực hiện tuần tự vì phải tuân theo trình tự thời gian.
+Do đó, time complexity của \verb|RNN| là $\mathcal{O}(\tau)$.
+Vì phải lưu trữ các giá trị đạo hàm sau để tính toán các giá trị đạo hàm trước đó (theo Chain rule), space complexity của thuật toán cũng là $\mathcal{O}(\tau)$.
+Do đó, đối với các chuỗi dài, chi phí tính toán có thể sẽ rất lớn.
+
 \subsection{Long Short-Term Memory}
+
+Khi làm việc trên các chuỗi dài, \verb|RNN| rất dễ mắc phải vấn đề \textit{vanishing gradient} hoặc \textit{exploding gradient}.
+Thật vậy, quá trình huấn luyện của \verb|RNN| đòi hỏi việc sử dụng Chain rule cho tất cả các trạng thái từ $t=\tau$ đến $t=1$.
+Quá trình này bao gồm rất nhiều phép nhân ma trận và vector.
+Do đó, nếu các phần tử trong ma trận hoặc vector đó nhỏ hơn 1, đạo hàm theo thời gian sẽ hội tụ về 0.
+Ngược lại, nếu các phần tử đó lớn hơn 1, theo thời gian, đạo hàm sẽ tiến đến vô cùng.
 

### 2025-01-21 doc/thesis/chapters/3.method.tex
change to maml(lstm)

@@
 Tổng quan phương pháp được minh họa trong hình \ref{fig:flow}.
-Trong phần feature extraction, chúng tôi kết hợp hai loại đặc trưng từ mạng \verb|LSTM| và \verb|LSTM+CNN|.
+Trong phần feature extraction, chúng tôi kết hợp hai loại đặc trưng từ mạng \verb|BiLSTM| và \verb|LSTM+CNN|.
 Trong phần Model synthesis, chúng tôi sử dụng \verb|MAML| để tổng hợp tham số của các mô hình.
 Với sự góp mặt của các đặc trưng từ các mạng học sâu, chúng tôi kỳ vọng sẽ rút trích được các đặc trưng ẩn trong dữ liệu aperiodic.
-Thông qua sự kết hợp giữa quá trình tổng hợp mô hình bằng thuật toán \verb|MAML| và quá trình rút trích đặc trưng bằng \verb|LSTM|/\verb|LSTM+CNN|, phương pháp đề xuất được kỳ vọng là một giải pháp thay thế hợp lý và hiệu quả cho các mô hình ensemble truyền thống trong việc tổng hợp hiệu quả external factors, giảm thiểu tác động của sự biến thiên phương sai, cũng như rút trích được các hidden long-term dependency trong quá khứ.
+Thông qua sự kết hợp giữa quá trình tổng hợp mô hình bằng thuật toán \verb|MAML| và quá trình rút trích đặc trưng bằng \verb|BiLSTM|/\verb|LSTM+CNN|, phương pháp đề xuất được kỳ vọng là một giải pháp thay thế hợp lý và hiệu quả cho các mô hình ensemble truyền thống trong việc tổng hợp hiệu quả external factors, giảm thiểu tác động của sự biến thiên phương sai, cũng như rút trích được các hidden long-term dependency trong quá khứ.
 
@@
 The overview of the method is illustrated in figure \ref{fig:flow} and the detail is presented in algorithm \ref{alg:maml}.
-In the \textit{Temporal feature extraction} section, we propose to use two types of features from \verb|LSTM| and \verb|LSTM+CNN| networks.
+In the \textit{Temporal feature extraction} section, we propose to use the features extracted from \verb|BiLSTM| network.
 In the \textit{Effective synthesis of model's parameters} section, \verb|MAML| is utilized to synthesize the parameters of the models.
 
-Due to the contribution of \verb|LSTM| and \verb|LSTM+CNN| features, we expect to effectively extract hidden features from aperiodic data.
+Due to the contribution of \verb|BiLSTM| features, we expect to effectively extract hidden temporal features from aperiodic data.
 By using \verb|MAML| in the weight aggregation process, the proposed method is expected to be a reasonable and effective alternative to traditional ensemble models in effectively synthesizing external factors, minimizing the impact of variance variation, and efficiency capturing hidden long-term dependencies in the past.
@@
                     \mathbf{x'} &= \mathbf{FullyConnected}\left( \mathbf{x} \right)\\
-                    \mathbf{h}_{LSTM} &= \mathbf{BidirectionalLSTM}\left( \mathbf{x'} \right)\\
-                    \mathbf{h}_{CNN} &= \mathbf{Convolution1D}\left( \mathbf{x'} \right)\\
-                    \hat{y} &= \begin{cases}
-                        \mathbf{FullyConnected}\left( \mathbf{h}_{LSTM} \right)\\
-                        \mathbf{FullyConnected}\left( \mathbf{Concatenate}\left( \mathbf{h}_{LSTM}, \mathbf{h}_{CNN} \right)\right)
-                    \end{cases}
+                    \mathbf{h}_{LSTM} &= \mathbf{BiLSTM}\left( \mathbf{x'} \right)\\
+                    \hat{y} &= \mathbf{FullyConnected}\left( \mathbf{h}_{LSTM} \right)
                 \end{align*}
@@
 
-% Quá trình rút trích đặc trưng được thực hiện bằng hai phương pháp khác nhau: thông qua \verb|LSTM| và \verb|LSTM+CNN|.
-Đối với phương pháp sử dụng \verb|LSTM|, từng phần tử trong ma trận $\mathbf{x}_{t-L:t}$ (viết gọn là $\mathbf{x}$) được đưa qua một lớp \verb|FullyConnected| có đầu ra lớn hơn số chiều của $\vec x_i, i\in[t-L, t]$ to present the input by a larger vector $\vec x'_i$ (equation \ref{eq:decompose}).
+% Quá trình rút trích đặc trưng được thực hiện bằng hai phương pháp khác nhau: thông qua \verb|BiLSTM| và \verb|LSTM+CNN|.
+Đối với phương pháp sử dụng \verb|BiLSTM|, từng phần tử trong ma trận $\mathbf{x}_{t-L:t}$ (viết gọn là $\mathbf{x}$) được đưa qua một lớp \verb|FullyConnected| có đầu ra lớn hơn số chiều của $\vec x_i, i\in[t-L, t]$ to present the input by a larger vector $\vec x'_i$ (equation \ref{eq:decompose}).
 Theo đó, các đặc trưng được thể hiện sâu sắc và rõ ràng hơn.
-Các đặc trưng này sau đó được truyền qua mạng \verb|LSTM| rút trích có chọn lọc các phụ thuộc thời gian dài hạn ($\mathbf{h}_{LSTM}$).
-Để có thể khai thác tối đa các ràng buộc thời gian dài hạn, chúng tôi sử dụng \verb|BidirectionalLSTM| để rút trích từ hai phía của của $\mathbf{x}$ (equation \ref{eq:feature_lstm})
+Các đặc trưng này sau đó được truyền qua mạng \verb|BiLSTM| rút trích có chọn lọc các phụ thuộc thời gian dài hạn ($\mathbf{h}_{LSTM}$).
+Để có thể khai thác tối đa các ràng buộc thời gian dài hạn, chúng tôi sử dụng \verb|BiLSTM| để rút trích từ hai phía của của $\mathbf{x}$ (equation \ref{eq:feature_lstm})
 
-Feature extraction is performed using two different methods: via \verb|LSTM| and \verb|LSTM+CNN|.
-For the method using \verb|LSTM|, each element in the matrix $\mathbf{x}_{t-L:t}$ (abbreviated as $\mathbf{x}$) is passed through a \verb|FullyConnected| layer whose output is larger than the dimension of $\vec x_i, i\in[t-L, t]$ to obtain vector $\vec x'_i$ (equation \ref{eq:decompose}).
+Feature extraction is performed using \verb|BiLSTM|.
+Specifically, each element in the matrix $\mathbf{x}_{t-L:t}$ (abbreviated as $\mathbf{x}$) is passed through a \verb|FullyConnected| layer whose output is larger than the dimension of $\vec x_i, i\in[t-L, t]$ to obtain vector $\vec x'_i$ (equation \ref{eq:decompose}).
 Accordingly, the data characteristics are expressed more deeply and clearly.
-These features are then passed through the \verb|LSTM| network to selectively extract long-term temporal dependencies ($\mathbf{h}_{LSTM}$).
-To fully exploit the long-term time constraints, we use \verb|BidirectionalLSTM| to extract from both sides of $\mathbf{x}$ (equation \ref{eq:feature_lstm})
+These features are then passed through the \verb|BiLSTM| network (equation \ref{eq:feature_lstm}) to selectively extract long-term temporal dependencies ($\mathbf{h}_{BiLSTM}$).
+At the end, $\mathbf{h}_{BiLSTM}$ is passed through a classification head of the neural network (equation \ref{eq:clf}) to predict the label of $\mathbf{x}$.
 
@@
     \mathbf{x'} &= \mathbf{FullyConnected}\left( \mathbf{x} \right) \label{eq:decompose}\\
-    \mathbf{h}_{LSTM} &= \mathbf{BidirectionalLSTM}\left( \mathbf{x'} \right) \label{eq:feature_lstm}
+    \mathbf{h}_{BiLSTM} &= \mathbf{BiLSTM}\left( \mathbf{x'} \right) \label{eq:feature_lstm}\\
+    \hat y &= \mathbf{FullyConnected}\left( \mathbf{h}_{BiLSTM} \right) \label{eq:clf}
 \end{align}
 
-% Lấy cảm hứng từ nghiên cứu \cite{vo2017multi}, chúng tôi đề xuất \verb|LSTM+CNN|, kết hợp các đặc trưng rút trích được từ mạng \verb|LSTM| và \verb|CNN|.
-Cụ thể, ngoài việc sử dụng \verb|LSTM|, $\mathbf{x'}$ còn được đưa vào \verb|CNN| để rút trích đặc trưng thời gian cục bộ ($\mathbf{h}_{CNN}$) (equation \ref{eq:feature_cnn}).
+% Lấy cảm hứng từ nghiên cứu \cite{vo2017multi}, chúng tôi đề xuất \verb|LSTM+CNN|, kết hợp các đặc trưng rút trích được từ mạng \verb|BiLSTM| và \verb|CNN|.
+Cụ thể, ngoài việc sử dụng \verb|BiLSTM|, $\mathbf{x'}$ còn được đưa vào \verb|CNN| để rút trích đặc trưng thời gian cục bộ ($\mathbf{h}_{CNN}$) (equation \ref{eq:feature_cnn}).
 Tiếp đến, $\mathbf{h}_{LSTM}$ và $\mathbf{h}_{CNN}$ được nối với nhau (equation \ref{eq:concat}) sau đó chuyển đến phần phân lớp của neural network (equation \ref{eq:clf}).
 
-Inspired by study \cite{vo2017multi}, we use \verb|LSTM+CNN| features, which combines the features extracted from \verb|LSTM| and \verb|CNN|.
-Specifically, in addition to using \verb|LSTM|, $\mathbf{x'}$ is also fed into \verb|CNN| to extract local temporal features $\mathbf{h}_{CNN}$ (equation \ref{eq:feature_cnn}).
+% Inspired by study \cite{vo2017multi}, we use \verb|LSTM+CNN| features, which combines the features extracted from \verb|BiLSTM| and \verb|CNN|.
+Specifically, in addition to using \verb|BiLSTM|, $\mathbf{x'}$ is also fed into \verb|CNN| to extract local temporal features $\mathbf{h}_{CNN}$ (equation \ref{eq:feature_cnn}).
 Next, $\mathbf{h}_{LSTM}$ and $\mathbf{h}_{CNN}$ are concatenated (equation \ref{eq:concat}) and then passed to the classification head of the neural network (equation \ref{eq:clf}).
 
-\begin{align}
-    \mathbf{h}_{CNN} &= \mathbf{Convolution1D}\left( \mathbf{x'} \right) \label{eq:feature_cnn}\\
-    \mathbf{h} &= \mathbf{Concatenate}\left( \mathbf{h}_{LSTM}, \mathbf{h}_{CNN} \right) \label{eq:concat} \\
-    \hat y &= \mathbf{FullyConnected}\left( \mathbf{h} \right) \label{eq:clf}
-\end{align}
+% \begin{align}
+%     \mathbf{h}_{CNN} &= \mathbf{Convolution1D}\left( \mathbf{x'} \right) \label{eq:feature_cnn}\\
+%     \mathbf{h} &= \mathbf{Concatenate}\left( \mathbf{h}_{LSTM}, \mathbf{h}_{CNN} \right) \label{eq:concat} \\
+%     \hat y &= \mathbf{FullyConnected}\left( \mathbf{h} \right) \label{eq:clf}
+% \end{align}
 
-% Mạng \verb|LSTM| duy trì giá trị cell-state nhằm lưu trữ có chọn lọc các phụ thuộc dài hạn.
+% Mạng \verb|BiLSTM| duy trì giá trị cell-state nhằm lưu trữ có chọn lọc các phụ thuộc dài hạn.
 Điều này rất thích hợp trong việc giải quyết các bài toán time-series data.
@@
 
-The \verb|LSTM| network maintains cell-state values to selectively store long-term dependencies.
-This is very suitable for solving time-series data problems.
+% The \verb|BiLSTM| network maintains cell-state values to selectively store long-term dependencies.
+Thus, this method is a suitable choice for solving time-series data problems.
 On the other hand, future values often depend heavily on recent historical values.
@@
 
+% The \verb|BiLSTM| network maintains cell-state values to selectively store long-term dependencies.
+Thus, this method is a suitable choice for solving time-series data problems.
+\textbf{Therefore, the proposed method not only remembers long-term features but also highlights short-term features.
+In the ideal situation, hidden constraints in aperiodic time-series data can be extracted efficiently}.
+
+% Mạng \verb|LSTM| và \verb|BiLSTM| đều duy trì các giá trị cell-state để lưu trữ một cách chọn lọc các phụ thuộc thời gian, cũng như hạn chế vấn đề vanishing gradient trong quá trình huấn luyện.
+Vì vậy, các mạng này rất phù hợp trong việc giải quyết các vấn đề time-series data.
+Tuy nhiên, đối với các chuỗi input dài, \verb|LSTM| tỏ ra khó khăn trong việc tận dụng ngữ cảnh toàn cục (toàn bộ chuỗi đầu vào) và thường đối mặt với vấn đề mất mát thông tin cục bộ (input càng xa thì càng dễ bị quên).
+Hai vấn đề này được khắc phục hiệu quả bởi \verb|BiLSTM|.
+Với khả năng rút trích và tổng hợp thông tin từ hai phía của input sequence, \textbf{\Verb|BiLSTM| có thể nắm bắt toàn bộ ngữ cảnh với lượng mất mát thông tin tối thiểu đối với các input ở xa.
+Đặt trong ngữ cảnh bài toán movement prediction of aperiodic time-series data, \Verb|BiLSTM| được kỳ vọng sẽ nắm bắt thành công sự biến đổi giá trị liên tục cũng như không bỏ sót các thông tin quan trọng tại các input ở xa}.
+Do đó, đây là một lựa chọn hợp lý hơn trong việc rút trích các ràng buộc thời gian.
+
+\verb|LSTM| and \verb|BiLSTM| both maintain cell-state values to selectively store temporal dependencies, as well as to mitigate the vanishing gradient problem during training.
+Therefore, these networks are well suited for solving time-series data problems.
+However, for long input sequences, \verb|LSTM| has difficulty in exploiting global context (the entire input sequence) and often faces the problem of local information loss (the further the input, the easier it is to forget).
+These two problems are effectively solved by \verb|BiLSTM|.
+With the ability to extract and synthesize information from both sides of the input sequence, \textbf{\Verb|BiLSTM| can capture the entire context with minimal information loss for distant inputs.
+In the context of the movement prediction of aperiodic time-series data problem, \Verb|BiLSTM| is expected to successfully capture continuous value variation as well as not miss important information at distant inputs}.
+Therefore, it is a more reasonable choice in extracting temporal constraints.
+
+\clearpage
+
 % Next, $\mathbf{h}_{LSTM}$ and $\mathbf{h}_{CNN}$ are concatenated (equation \ref{eq:concat}) and then passed to the classification part of the neural network (equation \ref{eq:clf}).
 
-\section{Effective synthesis of models' parameters}
+\section{Effective aggregation of models' parameters}
 
@@
 
-We use \verb|MAML| as presented in the algorithm \ref{alg:maml} to train and aggregate the weights of the models at the tasks.
+We use \verb|MAML| as presented in the algorithm \ref{alg:maml} to train and aggregate the weights of the models of all tasks via the \verb|CrossEntropyLoss| (equation \ref{eq:loss}).
+
+\begin{align}
+    \label{eq:loss}
+    \mathcal{L}(\mathbf{y}, \hat{\mathbf{y}}) = \frac{1}{\dim(\mathbf{y})}\sum_{i=1}^{\dim(\mathbf{y})}{(\mathbf{y}_i\log{\hat{\mathbf{y}}_i}) + (1-\mathbf{y}_i)\log{(1-\hat{\mathbf{y}}_i)}}
+\end{align}
+
 As mentioned in section \ref{sec:ml}, parameter optimization in the ML approach is to solve the two equations \ref{eq:inner_opt} and \ref{eq:outer_opt} using optimization methods on the support set and query set.
@@
     \end{cases}
-\end{align} In which, $\phi_{r-1}$ is the result of the $r-1$ outer optimization process, $\alpha$ is the inner learning rate.
+    \label{eq:inner_sol}
+\end{align} In which, $\phi_{r-1}$ is the result of the $r-1$ outer optimization process, $\alpha$ is the inner learning rate, $\mathcal{L}_t^{task}$ evaluates the use of $\theta_t^{(e-1)}$ on $\mathcal{D}_t^{support}$.
 
@@
     \label{eq:outer_sol}
-\end{align} Where, $\beta$ is the outer learning rate.
+\end{align} Where, $\beta$ is the outer learning rate, $\theta_t^*(\phi)$ is the local optimized weight obtained from equation \ref{eq:inner_sol}, $\mathcal{L}_t^{meta}$ evaluates the use of $\theta_t^*(\phi)$ on $\mathcal{D}_t^{query}$.
 

### 2025-01-21 doc/thesis/chapters/4.experiment.tex
translate into en + ablation study

@@
 
-Chương 4 mô tả các thí nghiệm để đánh giá \verb|Temporal-ML| và so sánh với \verb|NHITS| trên các tập dữ liệu lớn và phổ biến.
+% Chương 4 mô tả các thí nghiệm để đánh giá \verb|Temporal-ML| và so sánh với \verb|NHITS| trên các tập dữ liệu lớn và phổ biến.
 Các thí nghiệm được thực hiện trên aperiodic data bao gồm một dataset duy nhất và aperiodic data bao gồm nhiều datasets để kiểm chứng khả năng rút trích các thông tin về thị trường cũng như khả năng tổng hợp thông tin từ nhiều nguồn.
 Dữ liệu có chu kỳ cũng được sử dụng để tăng tính khách quan trong các đánh giá.
+Ngoài ra, chúng tôi thực hiện loại bỏ/thay thế các component của \verb|Temporal-ML| để phân tích kỹ khả năng của các components này.
+
+Chapter 4 describes experiments to evaluate \verb|Temporal-ML| and compare it with \verb|NHITS| on large and popular datasets.
+The experiments are performed on aperiodic data consisting of a single dataset and aperiodic data consisting of multiple datasets to verify the ability to extract market information as well as the ability to aggregate information from multiple sources.
+Periodic data is also used to increase the objectivity of the evaluations.
+In addition, we perform the removal/replacement of components of \verb|Temporal-ML| to analyze the capabilities of these components in detail.
 
@@
 % -> trung bình theo kiểu mean(metric), \sqrt(mean(std^2))
-Như đã đề cập trong section \ref{sec:data_prep}, để sử dụng ML, các tập dữ liệu cần phải được cấu trúc thành các task riêng biệt.
+
+% Như đã đề cập trong section \ref{sec:data_prep}, để sử dụng ML, các tập dữ liệu cần phải được cấu trúc thành các task riêng biệt.
 Trong mỗi task, tập support chiếm 20\% dữ liệu, tập query chiếm 80\% dữ liệu.
@@
 Thống kê dữ liệu cho các task theo tập dữ liệu được trình bày trong bảng \ref{tab:stat_}.
+
+As mentioned in section \ref{sec:data_prep}, to use ML, the datasets need to be structured into separate tasks.
+In each task, the support set accounts for 20\% of the data, the query set accounts for 80\% of the data.
+For \verb|multi-fx|, each currency pair is divided into 2 datasets, corresponding to 2 tasks.
+Therefore, 60 currency pairs form 120 tasks.
+For the remaining datasets, each dataset is divided into time series, each corresponding to a task.
+During the whole process, 50\% of the tasks are used in the meta-training process, 25\% of the tasks are used in the meta-validation process, and the remaining tasks are used in meta-testing.
+Statistics for tasks by dataset are presented in table \ref{tab:stat_}.
 
@@
 
-Vì các tập dữ liệu đều là time-series, dữ liệu trong quá trình huấn luyện không được tiết lộ bất cứ thông tin gì về tương lai.
+% Vì các tập dữ liệu đều là time-series, dữ liệu trong quá trình huấn luyện không được tiết lộ bất cứ thông tin gì về tương lai.
 Do đó, 120 tasks của dữ liệu \verb|multi-fx| được chia như hình \ref{fig:multi_fx_split}.
@@
 
-\begin{figure}
+Since the datasets are all time-series, the training data should not reveal any information about the future.
+Therefore, the 120 tasks of the \verb|multi-fx| dataset are divided as shown in \ref{fig:multi_fx_split}.
+For the remaining datasets, the data used in the training, testing, and validation must be in chronological order from past to future.
+
+\begin{figure}[H]
     \centering
@@
 
-    \cprotect\caption{(a): Splitting data for \verb|multi-fx|. (b): Pre-process for \verb|multi-fx| (dash line rectangle accounts for 20\% data of validation and testing tasks).}
+    \cprotect\caption{(a): Splitting data for \verb|multi-fx|. (b): Pre-process for \verb|multi-fx| (dash line rectangle accounts for all training and 20\% data of validation and testing tasks).}
     % \label{fig:ill_aperiodic}
@@
 
-Quá trình tiền xử lý bằng \verb|z-score| theo đó được thực hiện trên toàn bộ dữ liệu huấn luyện và các tập support của dữ liệu kiểm thử và kiểm tra của từng cặp tiền tệ (see figure \ref{fig:multi_fx_prep}) vì đây là các dữ liệu đã được biết trước.
+% Quá trình tiền xử lý bằng \verb|z-score| theo đó được thực hiện trên toàn bộ dữ liệu huấn luyện và các tập support của dữ liệu kiểm thử và kiểm tra của từng cặp tiền tệ (see figure \ref{fig:multi_fx_prep}) vì đây là các dữ liệu đã được biết trước.
 Cụ thể, tiền xử lý thuộc tính $X$ được thực hiện như sau:
+
+The \verb|z-score| preprocessing is then performed on the entire training data and the support sets of the test and validation data for each currency pair (see figure \ref{fig:multi_fx_prep}) since these are known data.
+Specifically, the preprocessing of attribute $X$ is performed as follows:
 
@@
 
-$\mu_{X}$ và $\sigma_{X}$ thu được từ phương trình \ref{eq:mean} và \ref{eq:std} được sử dụng để chuẩn hóa cho thuộc tính $X$ trong query set của quá trình meta-validation và meta-testing.
+% $\mu_{X}$ và $\sigma_{X}$ thu được từ phương trình \ref{eq:mean} và \ref{eq:std} được sử dụng để chuẩn hóa cho thuộc tính $X$ trong query set của quá trình meta-validation và meta-testing.
+
+$\mu_{X}$ and $\sigma_{X}$ obtained from the equations \ref{eq:mean} and \ref{eq:std} are used to normalize the attribute $X$ in the query set of the meta-validation and meta-testing processes.
 
@@
 
-Nghiên cứu sử dụng \textit{Accuracy, Precision, Recall}, và \textit{F1-score} để đánh giá các mô hình (equations \ref{eq:acc}, \ref{eq:P}, \ref{eq:R}, \ref{eq:F1}).
+% Nghiên cứu sử dụng \textit{Accuracy, Precision, Recall}, và \textit{F1-score} để đánh giá các mô hình (equations \ref{eq:acc}, \ref{eq:P}, \ref{eq:R}, \ref{eq:F1}).
 Trong đó, \textit{Accuracy} được sử dụng để tính tỷ lệ giữa số mẫu phân lớp đúng và tổng số mẫu được dự đoán, \textit{Precision} đo mức độ tin cậy của mô hình khi nó gán nhãn cho một mẫu dữ liệu, \textit{Recall} kiểm định tỷ lệ bỏ sót các mẫu dữ liệu của một phân lớp, \textit{F1-score} kiểm tra mức độ bias của mô hình trên các nhãn có số lượng lớn bằng cách lấy trung bình điều hòa trên \textit{Precision} và \textit{Recall}.
+
+The study uses \textit{Accuracy, Precision, Recall}, and \textit{F1-score} to evaluate the models (equations \ref{eq:acc}, \ref{eq:P}, \ref{eq:R}, \ref{eq:F1}).
+In which, \textit{Accuracy} is used to calculate the ratio between the number of correctly classified samples and the total number of predicted samples, \textit{Precision} measures the confidence level of the model when it labels a data sample, \textit{Recall} tests the rate of misclassified data samples of a classifier, \textit{F1-score} checks the level of bias of the model on major labels by taking the harmonic average over \textit{Precision} and \textit{Recall}.
 
@@
     F1 &= \frac{2PR}{P+R}\label{eq:F1}\\
-\end{align}Trong đó, $TP, TN, FP, FN$ lần lượt là số mẫu \textit{true\_positive, true\_negative, false\_positive, false\_negative}.
+\end{align}Where, $TP, TN, FP, FN$ are the number of \textit{true\_positive, true\_negative, false\_positive, false\_negative}, respectively.
 
-% The study uses accuracy, precision, recall, and F1 score to evaluate the models.
-Accordingly, during the inference phase, the model will run on all tasks to calculate the metrics of each one.
-Then, the average of the metrics of the tasks is calculated to obtain the final result.
-
-Giả sử thuật toán \verb|Temporal-ML| được đánh giá trên tập dữ liệu $\mathcal{D}$ gồm $a$ thuộc tính và được chia thành $n$ tasks.
+% Giả sử thuật toán \verb|Temporal-ML| được đánh giá trên tập dữ liệu $\mathcal{D}$ gồm $a$ thuộc tính và được chia thành $n$ tasks.
 Quá trình đánh giá được thực hiện bằng cách để mô hình dự đoán xu hướng của từng thuộc tính với đầu vào là tất cả các thuộc tính, sau đó thực hiện hai bước tổng hợp: (1) - Tổng hợp trên task; (2) - Tổng hợp trên thuộc tính.
@@
 
-\textbf{Tổng hợp trên task}.
+Suppose the \verb|Temporal-ML| algorithm is evaluated on a dataset $\mathcal{D}$ consisting of $a$ attributes and divided into $n$ tasks.
+The evaluation is performed by letting the model predict the movement of each attribute with all attributes as input, then performing two aggregation steps: (1) - Aggregation on tasks; (2) - Aggregation on attributes.
+This evaluation process is designed based on \cite{challu2023nhits}.
+
+% \textbf{Tổng hợp trên task}.
 Xét metric $m$ khi mô hình dự đoán thuộc tính $k$ bất kỳ ($1\leq k \leq a$).
@@
 Quá trình tổng hợp metrics $m$ của $n$ tasks được thực hiện như sau:
+
+\textbf{Aggregation on tasks}.
+Consider metric $m$ when the model predicts any attribute $k$ ($1\leq k \leq a$).
+After predicting attribute $k$, we obtain $n$ values: $\{m^{(k)}_1,\dots,m^{(k)}_n\}$.
+The process of synthesizing metrics $m$ of $n$ tasks is performed as follows:
 
@@
 
-\textbf{Tổng hợp trên thuộc tính}.
+% \textbf{Tổng hợp trên thuộc tính}.
 Mô hình thực hiện dự đoán tất cả $a$ thuộc tính và thu được $a$ metrics: $\left\{ \left( \bar{m}^{(k)}\pm s^{(k)}_m \right) \right\}_{k=1}^a$.
 Quá trình tổng hợp metric $m$ của $a$ thuộc tính diễn ra như sau:
+
+\textbf{Aggregation on attributes}.
+The model predicts all $a$ attributes and obtains $a$ metrics: $\left\{ \left( \bar{m}^{(k)}\pm s^{(k)}_m \right) \right\}_{k=1}^a$.
+The process of synthesizing metric $m$ of $a$ attributes is as follows:
 
@@
 
-Trong cài đặt của chúng tôi, chúng tôi sử dụng một lớp \verb|FullyConnected| gồm 16 units với hàm kích hoạt \verb|ReLU| để biểu diễn các đặc trưng sâu hơn và rõ ràng hơn.
-Sau đó, đặc trưng này được truyền song song đến các hai khối \verb|BidirectionalLSTM| và \verb|CNN|.
-Khối \verb|BidirectionalLSTM| bao gồm 32 hidden units, các đầu ra được nối dài để tạo thành một vector cuối.
-Khối \verb|CNN| bao gồm hai layer \verb|CNN| có số filter lần lượt là 32 và 64.
-Kernel được sử dụng trong các layer có kích thước $3\times 3$.
-Theo sau mỗi layer \verb|CNN| là một layer \verb|MaxPooling| sử dụng kernel kích thước $2\times 2$.
-Kết thúc khối \verb|CNN| là một layer \verb|Flatten|.
-Trong trường hợp chỉ sử dụng các đặc trưng dài hạn, đầu ra của khối \verb|BidirectionalLSTM| sẽ được sử dụng để phân lớp.
-Nếu sử dụng thêm các đặc trưng cục bộ, đầu ra của khối \verb|BidirectionalLSTM| và \verb|CNN| sẽ được nối lại rồi truyền qua một layer phân lớp nhị phân với hàm kích hoạt \verb|Sigmoid|.
+% Trong cài đặt của chúng tôi, chúng tôi sử dụng một lớp \verb|FullyConnected| gồm 16 units với hàm kích hoạt \verb|ReLU| để biểu diễn các đặc trưng sâu hơn và rõ ràng hơn.
+Sau đó, đặc trưng này được truyền đến khối \verb|BiLSTM|.
+Khối \verb|BiLSTM| bao gồm 32 hidden units, hai đầu ra cuối được nối dài để tạo thành một vector cuối.
+Vector này sẽ được truyền qua một layer phân lớp nhị phân với hàm kích hoạt \verb|Sigmoid|.
+
+In our implementation, we use a \verb|FullyConnected| layer of 16 units with a \verb|ReLU| activation function to represent deeper and more explicit features.
+This feature is then passed to a \verb|BiLSTM| block.
+The \verb|BiLSTM| block consists of 32 hidden units, the last two outputs of which are concatenated to form a final vector.
+This vector is then passed through a binary classification layer with a \verb|Sigmoid| activation function.
 
 % We use a \verb|FullyConnected| layer of 16 units with a \verb|ReLU| activation function to decompose the initial feature.
-This feature is then passed in parallel to the \verb|BidirectionalLSTM| and \verb|CNN| blocks.
-The \verb|BidirectionalLSTM| block consists of 32 hidden units, the outputs of which are concatenated to form a final vector.
+This feature is then passed in parallel to the \verb|BiLSTM| and \verb|CNN| blocks.
+The \verb|BiLSTM| block consists of 32 hidden units, the outputs of which are concatenated to form a final vector.
 The \verb|CNN| block consists of two \verb|CNN| layers with 32 and 64 filters, respectively.
@@
 The \verb|CNN| block ends with a \verb|Flatten| layer.
-Features of the \verb|BidirectionalLSTM| and \verb|CNN| blocks are then concatenated and passed through a binary classification layer with a \verb|Sigmoid| activation function.
+Features of the \verb|BiLSTM| and \verb|CNN| blocks are then concatenated and passed through a binary classification layer with a \verb|Sigmoid| activation function.
 
@@
     Lookback window                     & \{10, 20, 30\}                    \\
-    Inner learning rate                 & \{0.001, 0.005, 0.01, 0.05\}      \\
-    Outer learning rate                 & \{0.001, 0.005, 0.0015, 0.0055\}  \\
+    Inner learning rate                 & \{0.001, 0.00325, 0.0055, 0.00775, 0.01\}\\
+    Outer learning rate                 & \{0.001, 0.00325, 0.0055, 0.00775, 0.01\}\\
     \bottomrule
@@
 
-Quá trình fine-tune của thuật toán ML liên quan đến nhiều siêu tham số như inner batch size, outer batch size, inner training step, outer training step,...
+% Quá trình fine-tune của thuật toán ML liên quan đến nhiều siêu tham số như inner batch size, outer batch size, inner training step, outer training step,...
 Để dễ dàng fine-tune, chúng tôi cố định hầu hết các tham số và chỉ fine-tune kích thước của lookback window, inner và outer learning rate.
 Chi tiết được trình bày trong bảng \ref{tab:our_finetune}.
+
+The fine-tuning process of ML algorithms involves many hyper-parameters such as inner batch size, outer batch size, inner training step, outer training step,...
+To make fine-tuning easy, we fix most of the parameters and only fine-tune the size of lookback window, inner and outer learning rate.
+Details are presented in table \ref{tab:our_finetune}.
 
@@
 % viết về cách tổng hợp metric ở đây
-Nghiên cứu này sử dụng thuật toán \verb|NHITS| (AAAI 2023) làm baseline model.
+% Nghiên cứu này sử dụng thuật toán \verb|NHITS| (AAAI 2023) làm baseline model.
 \verb|NHITS| sử dụng toàn bộ thuộc tính để dự đoán xu hướng tiếp theo cho một thuộc tính trong một tập dữ liệu.
@@
 
-Chúng tôi tiền xử lý dữ liệu huấn luyện bằng \verb|z-score| như đã đề cập ở subsection \ref{subsec:ml_preprocess}.
+This study uses the \verb|NHITS| algorithm (AAAI 2023) as the baseline model.
+\verb|NHITS| uses the entire set of attributes to predict the next trend for an attribute in a dataset.
+The algorithm aims to decompose the frequency bands and combine them to predict the future.
+For the \verb|USD/JPY|, \verb|ETT-m2|, and \verb|WTH| datasets, the data in each set is divided into a training set, a validation set, and a test set with a ratio of 6:2:2, respectively.
+
+% Chúng tôi tiền xử lý dữ liệu huấn luyện bằng \verb|z-score| như đã đề cập ở subsection \ref{subsec:ml_preprocess}.
 Sau đó, $\mu_{X}$ và $\sigma_{X}$ thu được từ phương trình \ref{eq:mean} và \ref{eq:std} được sử dụng để chuẩn hóa cho thuộc tính $X$ trong validation set trong quá trình tuning.
@@
 
-Đối với tập dữ liệu \verb|multi-fx|, vì \verb|NHITS| không có cơ chế tổng hợp mô hình trên các tập dữ liệu khác nhau, chúng tôi lặp lại quy trình huấn luyện (huấn luyện, tuning, testing trên 60\%, 20\%, 20\% dữ liệu, respectively) cho từng cặp tiền tệ, sau đó tổng hợp metrics trên các cặp tiền tệ để thu được đánh giá cuối.
+We preprocess the training data with \verb|z-score| as mentioned in subsection \ref{subsec:ml_preprocess}.
+Then, $\mu_{X}$ and $\sigma_{X}$ obtained from \ref{eq:mean} and \ref{eq:std} equations are used to normalize the attribute $X$ in the validation set during tuning.
+Once the best hyper-parameters are selected, the training and validation data are renormalized from scratch, and then trained with the best set of hyper-parameters to predict the test set.
+
+% Đối với tập dữ liệu \verb|multi-fx|, vì \verb|NHITS| không có cơ chế tổng hợp mô hình trên các tập dữ liệu khác nhau, chúng tôi lặp lại quy trình huấn luyện (huấn luyện, tuning, testing trên 60\%, 20\%, 20\% dữ liệu, respectively) cho từng cặp tiền tệ, sau đó tổng hợp metrics trên các cặp tiền tệ để thu được đánh giá cuối.
+
+For the \verb|multi-fx| dataset, since \verb|NHITS| does not have a mechanism to aggregate models across different datasets, we repeat the training process (training, tuning, testing on 60\%, 20\%, 20\% of the data, respectively) for each currency pair, then aggregate metrics across currency pairs to obtain the final evaluation.
+
+% Ngoài ra, mô hình dự đoán ngẫu nhiên cũng được sử dụng để đối sánh.
+Mô hình ngẫu nhiên sinh ra các giá trị dự đoán 0 và 1 theo một uniform distribution.
+Quá trình tổng hợp metrics của mô hình này tương tự như \verb|NHITS|.
+
+In addition, a stochastic prediction model is also used for comparison.
+The stochastic model generates predicted values of 0 and 1 according to a uniform distribution.
+The process of synthesizing metrics of this model is similar to \verb|NHITS|.
 
@@
 
-\verb|NHITS| cũng sử dụng các metrics tượng tự \verb|Temporal-ML| và được đánh giá trên tập dữ liệu $\mathcal{D}$ với $a$ thuộc tính và phải thực hiện dự đoán trên từng thuộc tính.
+% \verb|NHITS| cũng sử dụng các metrics tượng tự \verb|Temporal-ML| và được đánh giá trên tập dữ liệu $\mathcal{D}$ với $a$ thuộc tính và phải thực hiện dự đoán trên từng thuộc tính.
 Xét metric $m$, sau quá trình dự đoán thuộc tính $k$, ta thu được $a$ giá trị: $\{m^{(k)}_1,\dots,m^{(k)}_a\}$.
 Quá trình tổng hợp metric $m$ của $a$ thuộc tính được thực hiện như sau:
+
+\verb|NHITS| also uses similar metrics to \verb|Temporal-ML| and is evaluated on the dataset $\mathcal{D}$ with $a$ attributes and must perform predictions on each attribute.
+Considering metric $m$, after predicting attribute $k$, we obtain $a$ values: $\{m^{(k)}_1,\dots,m^{(k)}_a\}$.
+The process of synthesizing metric $m$ of $a$ attributes is performed as follows:
 
@@
 
-Như đã đề cập, \verb|NHITS| không có cơ chế tổng hợp mô hình khi kiểm thử trên dữ liệu \verb|multi-fx|.
+% Như đã đề cập, \verb|NHITS| không có cơ chế tổng hợp mô hình khi kiểm thử trên dữ liệu \verb|multi-fx|.
 Do đó, chúng tôi kiểm thử \verb|NHITS| trên từng cặp tiền tệ trong \verb|multi-fx| và tổng hợp metrics giống như \verb|Temporal-ML|.
 
-\subsection{Fine-tuning}
+As mentioned, \verb|NHITS| does not have a mechanism for model aggregation when testing on \verb|multi-fx| data.
+Therefore, we test \verb|NHITS| on each currency pair in \verb|multi-fx| and aggregate metrics like \verb|Temporal-ML|.
 
-Đối với mô hình \verb|NHITS|, chúng tôi dựa trên \cite{challu2023nhits} để định nghĩa không gian tìm kiếm cho việc fine-tune tham số, cũng như kiến trúc mô hình.
-Đối với các tham số không được đề cập trong bảng, chúng tôi sử dụng giá trị mặc định của cài đặt \verb|NHITS| trong thư viện \verb|NeuralForecast| \cite{neuralforecast}.
-Kết quả tốt nhất của các lần fine-tune được chọn ra và báo cáo trong nghiên cứu này.
-
-% We rely on \cite{challu2023nhits} to define the search space (table \ref{tab:nhits_finetune}) for parameter fine-tuning, as well as to fine-tune the model architecture.
-For parameters not mentioned in the table, we use the default values of the \verb|NHITS| implementation in the \verb|NeuralForecast| library \cite{neuralforecast}.
-The best results of fine-tuning process are selected and reported in this study.
-
-\begin{table}[H]
+\begin{table}[h]
     \centering
@@
     \hline
-    Lookback window                     & \{5, 20, 30\}                                                \\
-    Pooling kernel                      & \{{[}2,2,2], [4,4,4], [8,8,8], [8,4,1], [16,8,1]\}           \\
-    Stacks' coefficients                & \{{[}168,24,1], [24,12,1], [180,60,1],[40,20,1], [64,8,1]\}  \\
+    Lookback window                     & \{10, 20, 30\}                                                \\
+    Pooling kernel                      & \{[2,2,2], [4,4,4], [8,8,8], [8,4,1], [16,8,1]\}           \\
+    Stacks' coefficients                & \{[168,24,1], [24,12,1], [180,60,1],[40,20,1], [64,8,1]\}  \\
     Number of MLP layers                & \{1,2\}                                                      \\
-    Learning rate                       & \{0.001, 0.002, 0.005, 0.01, 0.02\}                          \\
+    Learning rate                       & \{0.001, 0.00325, 0.0055, 0.00775, 0.01\}                          \\
     \bottomrule
@@
 
+\subsection{Fine-tuning}
+
+% Đối với mô hình \verb|NHITS|, chúng tôi dựa trên \cite{challu2023nhits} để định nghĩa không gian tìm kiếm cho việc fine-tune tham số (see table \ref{tab:nhits_finetune}), cũng như kiến trúc mô hình.
+Đối với các tham số không được đề cập trong bảng, chúng tôi sử dụng giá trị mặc định của cài đặt \verb|NHITS| trong thư viện \verb|NeuralForecast| \cite{neuralforecast}.
+Kết quả tốt nhất của các lần fine-tune được chọn ra và báo cáo trong nghiên cứu này.
+
+We rely on \cite{challu2023nhits} to define the search space (table \ref{tab:nhits_finetune}) for parameter fine-tuning, as well as to fine-tune the model architecture.
+For parameters not mentioned in the table, we use the default values of the \verb|NHITS| implementation in the \verb|NeuralForecast| library \cite{neuralforecast}.
+The best results of fine-tuning process are selected and reported in this study.
+
 \section{Fairness in evaluation}
 
-Gọi $m$ là số mẫu dữ liệu trong mỗi tasks, tổng số dữ liệu của quá trình huấn luyện, kiểm thử, và kiểm tra là: $mn$.
+% Gọi $m$ là số mẫu dữ liệu trong mỗi tasks, tổng số dữ liệu của quá trình huấn luyện, kiểm thử, và kiểm tra là: $mn$.
 Đối với \verb|NHITS|, mô hình được huấn luyện, kiểm thử, kiểm tra trên 60\%, 20\%, 20\% dữ liệu, respectively:
+
+Let $m$ be the number of data samples in each task, the total data of training, validating, and testing is: $mn$.
+For \verb|NHITS|, the model is trained, validated, and tested on 60\%, 20\%, 20\% of the data, respectively:
 
@@
 
-Đối với \verb|Temporal-ML|, quá trình làm giàu knowledge cho mô hình sử dụng toàn bộ dữ liệu huấn luyện và tập support của dữ liệu kiểm thử/kiểm tra.
+% Đối với \verb|Temporal-ML|, quá trình làm giàu knowledge cho mô hình sử dụng toàn bộ dữ liệu huấn luyện và tập support của dữ liệu kiểm thử/kiểm tra.
 Quá trình kiểm thử/kiểm tra được thực hiện trên tập query:
+
+For \verb|Temporal-ML|, the knowledge enrichment process for the model uses the entire training data and the support set of the testing/validation data.
+The testing/validation process is performed on the query set:
 
@@
 
-Do đó, với cách chia dữ liệu như trên, chúng tôi đạt được sự công bằng về số lượng mẫu dữ liệu trong quá trình kiểm thử.
+% Do đó, với cách chia dữ liệu như trên, chúng tôi đạt được sự công bằng về số lượng mẫu dữ liệu trong quá trình kiểm thử.
 
-% i am not sure about this
-\section{Ablation study}
+Therefore, with the above data division, we achieve fairness in the number of data samples during testing.
+
+\section{Ablation experiment}
+\label{sec:ab_ex}
 
@@
 
-Chúng tôi thực hiện đánh giá từng component trong \verb|Temporal-ML| bằng cách loại bỏ chúng khỏi thuật toán và so sánh kết quả với thuật toán gốc.
-This includes three models without \verb|MAML| (\verb|LSTM|, \verb|CNN|, \verb|LSTM+CNN|), and
+% Chúng tôi thực hiện đánh giá từng component trong \verb|Temporal-ML| cũng như sự lựa chọn phương pháp rút trích temporal feature bằng cách loại bỏ hoặc thay thế chúng khỏi thuật toán và so sánh kết quả với thuật toán gốc.
+Cụ thể, chúng tôi tiến hành thí nghiệm trên hai nhóm thuật toán: (1) - Model without \verb|BiLSTM|; (2) - Model without ML.
+Về việc đánh giá các mô hình này, các mô hình ML-based được đánh giá như \verb|Temporal-ML|, các mô hình còn lại được đánh giá giống như \verb|NHITS|.
+
+We evaluate each component in \verb|Temporal-ML| as well as the choice of temporal feature extraction method by removing or replacing them from the algorithm and compare the results with the original algorithm.
+Specifically, we conduct experiments on two groups of algorithms: (1) - Model without \verb|BiLSTM|; (2) - Model without ML.
+Regarding the evaluation of these models, ML-based models are evaluated as \verb|Temporal-ML|, the remaining models are evaluated as \verb|NHITS|.
+
+\subsection{Model without BiLSTM}
+\label{subsec:ablation_lstm}
+
+% Chúng tôi nghiên cứu khả năng rút trích temporal feature của các thuật toán bằng cách thay thế \verb|BiLSTM| trong \verb|Temporal-ML| bằng các thuật toán \verb|BiLSTM+CNN|, \verb|Transformer| \cite{vaswani2017attention}.
+\textbf{Mục tiêu của việc thay thế \Verb|BiLSTM| bằng \Verb|BiLSTM+CNN| là để thí nghiệm về khả năng rút trích đặc trưng cục bộ của extractor} bằng cách kết hợp đặc trưng của \verb|BiLSTM| và \verb|CNN| sau đó nối chúng lại và tiến hành phân lớp (equations \ref{eq:feature_lstm}, \ref{eq:feature_cnn}–\ref{eq:clf_cnn}).
+
+We study the temporal feature extraction ability of the algorithms by replacing \verb|BiLSTM| in \verb|Temporal-ML| with \verb|BiLSTM+CNN|, \verb|Transformer| algorithms \cite{vaswani2017attention}.
+\textbf{The goal of replacing \Verb|BiLSTM| with \Verb|BiLSTM+CNN| is to test the local feature extraction ability of the extractor} by combining the features of \verb|BiLSTM| and \verb|CNN| then concatenating them and performing classification (equations \ref{eq:feature_lstm}, \ref{eq:feature_cnn}–\ref{eq:clf_cnn}).
+
+\begin{align}
+    \mathbf{h}_{CNN} &= \mathbf{Convolution1D}\left( \mathbf{x'} \right) \label{eq:feature_cnn}\\
+    \mathbf{h} &= \mathbf{Concatenate}\left( \mathbf{h}_{LSTM}, \mathbf{h}_{CNN} \right) \label{eq:concat} \\
+    \hat y &= \mathbf{FullyConnected}\left( \mathbf{h} \right) \label{eq:clf_cnn}
+\end{align}
+
+% \verb|Transformer| được cấu thành từ 2 khối encoders.
+Mỗi khối encoder sử dụng 3 attention heads, số chiều của key vector là 256.
+Phần \verb|FeedForward| của khối encoder sử dụng hai \verb|Convolution1D| layers và một \verb|LayerNormalization|.
+Sau khi xếp chồng hai lớp encoders, ma trận đặc trưng được truyền qua một lớp \verb|Flatten| sau đó đưa vào classification head.
+Dựa trên nghiên cứu \cite{liu2025onsitnet}, đầu vào của \verb|Transformer| được transposed để khám phá mối quan hệ giữa các đặc trưng với nhau.
+\verb|Transformer| là một neural network có hiệu năng xuất sắc và được ứng dụng rộng rãi.
+Mục tiêu của việc sử dụng mạng này là để kiểm tra xem liệu cơ chế attention có giúp ích nhiều trên aperiodic time-series data hay không.
+
+\verb|Transformer| is composed of two encoders.
+Each encoder uses three attention heads, and the key vector dimension is 256.
+The \verb|FeedForward| part of the encoder uses two \verb|Convolution1D| layers and one \verb|LayerNormalization|.
+After stacking the two encoder layers, the feature matrix is passed through a \verb|Flatten| layer and then fed into the classification head.
+\verb|Transformer| is a widely used and excellent performing neural network.
+\textbf{The goal of using this network is to test the performance of attention mechanism on aperiodic time-series data}.
+
+\subsection{Model without ML}
+\label{subsec:ablation_ml}
+
+% Để nghiên cứu tác động của \verb|MAML| cũng như các thuật toán optimization-based ML đối với \verb|Temporal-ML|, chúng tôi loại bỏ component này và chỉ sử dụng các phương pháp rút trích đặc trưng thuần túy đã giới thiệu ở trên: \verb|BiLSTM|, \verb|BiLSTM+CNN|, \verb|Transformer|.
+Thông qua đó, khả năng tổng hợp hiệu quả mô hình của \verb|MAML| và các thuật toán optimization-based ML trong bài toán movement prediction of aperiodic time-series data sẽ được làm rõ.
+
+To study the impact of \verb|MAML| as well as optimization-based ML algorithms on \verb|Temporal-ML|, we remove this component and only use the pure feature extraction methods introduced above: \verb|BiLSTM|, \verb|BiLSTM+CNN|, \verb|Transformer|.
+Through this, \textbf{the ability to effectively synthesize the models of \Verb|MAML| and optimization-based ML algorithms in the problem of movement prediction of aperiodic time-series data will be clarified}.

### 2025-01-21 doc/thesis/chapters/5.results.tex
chems gios

--- 2024-12-31 doc/thesis/chapters/5.results.tex
@@
 
-\chapter{Results}
+\chapter{Results \& Discussions}
 \label{chap:result}
+
+% khả năng là nhập ablation study vào trong 2 phần sau main
 
@@
 
-% vẽ, mô tả cái bảng
-% bẻ nhỏ từng attribute trong fx dataset --> thành 4 bảng, 1 bảng tổng hợp là 5
-% bảng tổng hợp vẽ ở đây
-% chắc phải viết thêm cả CNN
+% vẽ, mô tả cái bảng tổng hợp
+Kết quả chính của công trình này được thể hiện trong bảng \ref{tab:mine_nhits}.
+Theo đó, phương pháp đề xuất của chúng tôi đạt hiệu quả vượt trội trên tất cả các metrics trên cả hai tập dữ liệu phi chu kỳ (\verb|USD/JPY|, \verb|multi-fx|) so với \verb|NHITS|.
+Trên các tập dữ liệu có chu kỳ (\verb|ETT-m2|, \verb|WTH|), mô hình của chúng tôi đạt hiệu quả gần như tương đương hoặc có phần tốt hơn so với baseline model.
 
-\section{Feature extraction ability}
+\begin{table}[h]
+    \centering
+    \cprotect\caption{Classification results (\%) of \verb|Temporal-ML| and \verb|NHITS|. Best results per metrics are boldfaced. (\mbox{*}): Our method.}
+    \label{tab:mine_nhits}
+    \resizebox{\columnwidth}{!}{\begin{tabular}{c|c|cccc}
+    \toprule
+    \multicolumn{1}{c}{}                &                                & $accuracy$              & $precision$             & $recall$                & $F1-score$                 \\
+    \hline
+    \multirow{2}{*}{\Verb|USD/JPY|}     & \Verb|NHITS|                   & $58.46$                 & $58.24$                 & $57.65$                 & $55.82$            \\
+                                        & \Verb|Temporal-ML|\mbox{*}     & $\mathbf{70.33\pm1.69}$ & $\mathbf{70.73\pm1.82}$ & $\mathbf{70.26\pm1.93}$ & $\mathbf{69.28\pm2.50}$    \\
+                                        % & Ours(\Verb|BiLSTM+CNN|)  & $68.75\pm1.78$          & $69.15\pm1.55$          & $68.73\pm1.77$          & $69.14\pm2.17$             \\
+    \hline
+    \multirow{2}{*}{\Verb|multi-fx|}    & \Verb|NHITS|                   & $53.51\pm5.02$          & $53.90\pm7.68$          & $53.64\pm4.05$          & $50.20\pm7.35$              \\
+                                        & \Verb|Temporal-ML|\mbox{*}     & $\mathbf{66.26\pm7.45}$ & $\mathbf{67.08\pm8.97}$ & $\mathbf{65.76\pm7.79}$ & $\mathbf{64.06\pm10.00}$   \\
+                                        % & Ours(\Verb|BiLSTM+CNN|)  & $66.08\pm8.19$          & $66.56\pm9.23$          & $65.41\pm8.66$          & $63.78\pm10.83$            \\
+    \hline
+    \multirow{2}{*}{\Verb|ETT-m2|}      & \Verb|NHITS|                   & $\mathbf{71.88}$        & $\mathbf{66.86}$        & $\mathbf{62.49}$        & $\mathbf{62.69}$   \\
+                                        & \Verb|Temporal-ML|\mbox{*}     & $71.14\pm9.13$          & $62.21\pm7.51$          & $58.75\pm5.30$          & $57.61\pm6.74$             \\
+                                        % & Ours(\Verb|BiLSTM+CNN|)  & $71.40\pm9.10$          & $63.25\pm8.82$          & $60.22\pm8.42$          & $59.43\pm9.24$             \\
+    \hline
+    \multirow{2}{*}{\Verb|WTH|}         & \Verb|NHITS|                   & $74.18$                 & $68.05$                 & $65.04$                 & $65.40$             \\
+                                        & \Verb|Temporal-ML|\mbox{*}     & $\mathbf{74.97\pm2.63}$ & $\mathbf{69.13\pm3.44}$ & $\mathbf{65.98\pm3.96}$ & $\mathbf{66.00\pm3.80}$    \\
+                                        % & Ours(\Verb|BiLSTM+CNN|)  & $74.41\pm2.71$          & $68.17\pm4.60$          & $65.50\pm4.82$          & $65.63\pm4.47$             \\
+    \bottomrule
+    \end{tabular}}
+\end{table}
 
-% 4 bảng phụ vẽ ở đây
-% kiểu: sử dụng mô hình để dự đoán từng attribute, ta có bảng kq sau
+% \begin{table}[h]
+%     \centering
+%     \cprotect\caption{Classification results (\%) of \verb|Temporal-ML| and \verb|NHITS|. Best results per metrics are boldfaced. (\mbox{*}): Our method.}
+%     \label{tab:mine_nhits}
+%     \resizebox{\columnwidth}{!}{\begin{tabular}{c|c|cccc}
+%     \toprule
+%     \multicolumn{1}{c}{}                &                                & $accuracy$              & $precision$             & $recall$                & $F1-score$                 \\
+%     \hline
+%     \multirow{2}{*}{\Verb|USD/JPY|}     & \Verb|NHITS|                   & $58.46\pm9.60$          & $58.24\pm9.89$          & $57.65\pm10.08$         & $55.82\pm11.22$            \\
+%                                         & \Verb|Temporal-ML|\mbox{*}     & $\mathbf{70.33\pm1.69}$ & $\mathbf{70.73\pm1.82}$ & $\mathbf{70.26\pm1.93}$ & $\mathbf{69.28\pm2.50}$    \\
+%                                         % & Ours(\Verb|BiLSTM+CNN|)  & $68.75\pm1.78$          & $69.15\pm1.55$          & $68.73\pm1.77$          & $69.14\pm2.17$             \\
+%     \hline
+%     \multirow{2}{*}{\Verb|multi-fx|}    & \Verb|NHITS|                   & $53.51\pm5.02$          & $53.90\pm7.68$          & $53.64\pm4.05$          & $50.20\pm7.35$              \\
+%                                         & \Verb|Temporal-ML|\mbox{*}     & $\mathbf{66.26\pm7.45}$ & $\mathbf{67.08\pm8.97}$ & $\mathbf{65.76\pm7.79}$ & $\mathbf{64.06\pm10.00}$   \\
+%                                         % & Ours(\Verb|BiLSTM+CNN|)  & $66.08\pm8.19$          & $66.56\pm9.23$          & $65.41\pm8.66$          & $63.78\pm10.83$            \\
+%     \hline
+%     \multirow{2}{*}{\Verb|ETT-m2|}      & \Verb|NHITS|                   & $\mathbf{71.88\pm14.06}$& $\mathbf{66.86\pm10.21}$& $\mathbf{62.49\pm10.40}$& $\mathbf{62.69\pm10.72}$   \\
+%                                         & \Verb|Temporal-ML|\mbox{*}     & $71.14\pm9.13$          & $62.21\pm7.51$          & $58.75\pm5.30$          & $57.61\pm6.74$             \\
+%                                         % & Ours(\Verb|BiLSTM+CNN|)  & $71.40\pm9.10$          & $63.25\pm8.82$          & $60.22\pm8.42$          & $59.43\pm9.24$             \\
+%     \hline
+%     \multirow{2}{*}{\Verb|WTH|}         & \Verb|NHITS|                   & $74.18\pm8.11$          & $68.05\pm6.65$          & $65.04\pm7.67$          & $65.40\pm7.97$             \\
+%                                         & \Verb|Temporal-ML|\mbox{*}     & $\mathbf{74.97\pm2.63}$ & $\mathbf{69.13\pm3.44}$ & $\mathbf{65.98\pm3.96}$ & $\mathbf{66.00\pm3.80}$    \\
+%                                         % & Ours(\Verb|BiLSTM+CNN|)  & $74.41\pm2.71$          & $68.17\pm4.60$          & $65.50\pm4.82$          & $65.63\pm4.47$             \\
+%     \bottomrule
+%     \end{tabular}}
+% \end{table}
+
+Cụ thể, trên hai tập dữ liệu phi chu kỳ \verb|USD/JPY| và \verb|multi-fx|, \verb|Temporal-ML| cải thiện từ 12\% đến 14\% trên các metrics so với baseline model.
+Mặt khác, đối với dữ liệu có chu kỳ, độ chính xác của \verb|Temporal-ML| thấp hơn chưa đến 1\%, các metrics còn lại thấp hơn từ 4-5\% so với \verb|NHITS| trên dữ liệu \verb|ETT-m2|.
+Đối với dữ liệu có tính chu kỳ mạnh như \verb|WTH|, thuật toán của chúng tôi đạt hiệu suất tốt hơn \verb|NHITS| khoảng 1\%.
+Ngoài ra, \verb|Temporal-ML| cũng đạt được mức phân tán trên các metrics thấp hơn \verb|NHITS| từ 1-3\% khi hoạt động trong ngữ cảnh dữ liệu đa nguồn.
+Quan sát tổng quan, sự chênh lệch của các metrics của cùng một phương pháp là không quá lớn nên không tồn tại vấn đề dữ liệu bị lệch và có thể đánh giá mô hình bằng cách chỉ dựa trên độ chính xác.
+
+Trong bảng \ref{tab:mine_nhits}, kết quả được tổng hợp dưới dạng trung bình cộng của tất cả các đặc trưng nên có tính đại diện cao.
+Tuy nhiên, các kết quả này thiếu đi sự chi tiết trong đánh giá.
+Do đó, chúng tôi tiến hành phân tích kết quả của các thuật toán trên từng attribute của aperiodic datasets trong bảng \ref{tab:mine_nhits_att}.
+Theo đó, khi \verb|NHITS| dự đoán các thuộc tính \textit{hight, low, close} price trên cả hai tập dữ liệu, kết quả đều chỉ đạt ở mức rất gần mô hình dự đoán ngẫu nhiên.
+Điều này cho thấy việc dự đoán trên aperiodic data thực sự rất khó khăn.
+Tuy nhiên, trên cả hai tập dữ liệu, độ chính xác của \verb|NHITS| thường đạt giá trị lớn nhất khi dự đoán \textit{open} price và lớn hơn đáng kể so với kết quả dự đoán các attribute còn lại.
+Do đó, khi lấy trung bình cộng, kết quả của \verb|NHITS| có vẻ khá cao.
+
+\begin{table}[h]
+    \centering
+    \cprotect\caption{Accuracy (\%) of \verb|Temporal-ML| and \verb|NHITS| on each attribute of \verb|USD/JPY| and \verb|multi-fx|. Best results per metrics are boldfaced. (\mbox{*}): Our method.}
+    \label{tab:mine_nhits_att}
+    \resizebox{\columnwidth}{!}{\begin{tabular}{c|c|cccc}
+    \toprule
+    \multicolumn{1}{c}{}                &                                & $Open$                  & $High$                  & $Low$                   & $Close$                     \\
+    \hline
+    \multirow{3}{*}{\Verb|USD/JPY|}     & Random                         & $49.91$                 & $49.91$                 & $50.33$                 & $50.66$                     \\
+                                        & \Verb|NHITS|                   & $75.03$                 & $53.74$                 & $51.57$                 & $53.50$                     \\
+                                        & \Verb|Temporal-ML|\mbox{*}     & $\mathbf{93.06\pm1.36}$ & $\mathbf{68.87\pm1.53}$ & $\mathbf{51.01\pm1.58}$ & $\mathbf{68.37\pm2.16}$     \\
+    \hline
+    \multirow{3}{*}{\Verb|multi-fx|}    & Random                         & $49.92\pm2.19$          & $49.93\pm2.22$          & $49.94\pm2.17$          & $49.85\pm2.13$              \\
+                                        & \Verb|NHITS|                   & $58.36\pm7.13$          & $50.66\pm4.28$          & $51.17\pm3.21$          & $53.84\pm4.60$              \\
+                                        & \Verb|Temporal-ML|\mbox{*}     & $\mathbf{78.13\pm6.33}$ & $\mathbf{62.74\pm3.30}$ & $\mathbf{58.28\pm3.86}$ & $\mathbf{65.88\pm12.49}$    \\
+    \bottomrule
+    \end{tabular}}
+\end{table}
+
+Mặt khác, \verb|Temporal-ML| chỉ thất bại trong việc dự đoán \textit{low} price của \verb|USD/JPY| (độ chính xác tiệm cận mức dự đoán ngẫu nhiên).
+Tất cả các thuộc tính còn lại của cả hai tập dữ liệu đều đạt mức hội tụ cao và cao hơn \verb|NHITS| khá nhiều.
+Điều này chứng tỏ khả năng vượt trội của thuật toán đề xuất trên dữ liệu phi chu kỳ.
+
+Để làm rõ khả năng của \verb|Temporal-ML|, dựa trên kết quả thu được từ các ablation experiments (section \ref{sec:ab_ex}) chúng tôi tiến hành phân tích thuật toán đề xuất trên hai khía cạnh: (1) - Temporal feature extraction ability; (2) - Models' parameters aggregation ability.
+
+\section{Temporal feature extraction ability}
+
+% đấm nhits
+Từ bảng \ref{tab:mine_nhits} và \ref{tab:mine_nhits_att}, không khó để nhận ra khả năng rút trích temporal feature của \verb|Temporal-ML| tốt hơn nhiều so với \verb|NHITS|.
+Thật vậy, quá trình rút trích và tổng hợp đặc trưng của \verb|NHITS| thể hiện rằng phương pháp này đang cố gắng mô phỏng lại quá trình phân giải tần số của Fourier transform.
+Đặc trưng về các giải tần là thông tin tối quan trọng đối với dữ liệu có chu kỳ.
+Do đó, các dự đoán của \verb|NHITS| có thể dễ dàng đạt được độ chính xác cao trên loại dữ liệu này.
+Tuy nhiên, rất khó để có thể phân giải tần số trên dữ liệu phi chu kỳ.
+Hơn nữa, time-series data, dù được lọc qua nhiều stack để rút trích các dải tần từ thấp đến cao, nhưng các kỹ thuật lọc được sử dụng trong mạng là vô cùng đơn giản (chỉ sử dụng \verb|FullyConnected| với hàm kích hoạt phi tuyến và \verb|Pooling| layers).
+Chính những điều này đã khiến cho \verb|NHITS| gặp trở ngại lớn trên \verb|multi-fx|, \verb|USD/JPY| datasets và các dữ liệu phi chu kỳ nói chung.
+
+Mặt khác, phương pháp đề xuất của chúng tôi sử dụng các mạng deep learning để rút trích đặc trưng.
+\verb|BiLSTM| và \verb|RNN| nói chung đã chúng tỏ được khả năng của chúng qua nhiều bài toán có liên quan đến các ràng buộc thời gian, rằng chúng có thể dễ dàng thu được các deep hidden feature.
+Do đó, trên cả periodic data lẫn aperiodic data, phương pháp của chúng tôi đều đạt được kết quả xấp xỉ hoặc cao hơn \verb|NHITS|.
+
 % phân tích bảng kết quả
+Để nắm bắt rõ hơn khả năng của \verb|BiLSTM|, chúng tôi thực hiện ablation study trên component này của \verb|Temporal-ML| như đã đề cập trong subsection \ref{subsec:ablation_lstm} và trình bày kết quả trong bảng \ref{tab:ab_lstm}.
+Theo đó, khi \verb|BiLSTM| được thay thế bởi các component khác (\verb|BiLSTM+CNN|, \verb|Transformer|) để phục vụ quá trình rút trích đặc trưng thời gian, kết quả  trên cả bốn metrics của hầu hết các tập dữ liệu đều thấp hơn kết quả thu được từ \verb|Temporal-ML|.
+Cụ thể, trên hai tập dữ liệu phi chu kỳ (\verb|USD/JPY| and \verb|multi-fx|), các thuật toán sử dụng \verb|BiLSTM+CNN| và \verb|Transformer| để rút trích đặc trưng cho khả năng dự đoán thấp hơn từ 2-10\% so với \verb|Temporal-ML|.
+Trên dữ liệu có chu kỳ, \verb|MAML(Transformer)| vẫn cho kết quả thấp hơn nhiều so với \verb|Temporal-ML|.
+Tuy nhiên \verb|MAML(BiLSTM+CNN)| cho kết quả gần như tiệm cận với thuật toán đề xuất.
+Trên cả bốn datasets, độ lệch chuẩn của các metrics của thuật toán đề xuất nhìn chung nhỏ hơn từ 1-5\% so với các thuật toán còn lại.
+
+\begin{table}[h]
+    \centering
+    \cprotect\caption{Classification results (\%) of \verb|Temporal-ML| and models without \verb|BiLSTM|. Best results per metrics are boldfaced. (\mbox{*}): Our method.}
+    \label{tab:ab_lstm}
+    \resizebox{\columnwidth}{!}{\begin{tabular}{c|c|cccc} 
+    \toprule
+    \multicolumn{1}{c}{}      &                                          & $accuracy$               & $precision$              & $recall$                 & $F1-score$                \\ 
+    \hline
+    \multirow{3}{*}{\Verb|USD/JPY|} & \Verb|Temporal-ML|\mbox{*}         & $\mathbf{70.33\pm1.69}$  & $\mathbf{70.73\pm1.82}$  & $\mathbf{70.26\pm1.93}$  & $\mathbf{69.28\pm2.50}$   \\
+                                    & \Verb|MAML|(\Verb|BiLSTM+CNN|)     & $68.75\pm1.78$           & $69.15\pm1.55$           & $68.73\pm1.77$           & $69.14\pm2.17$            \\
+                                    & \Verb|MAML|(\Verb|Transformers|)   & $60.67\pm4.09$           & $59.74\pm6.24$           & $59.32\pm3.79$           & $53.15\pm6.19$            \\ 
+    \hline
+    \multirow{3}{*}{\Verb|multi-fx|}& \Verb|Temporal-ML|\mbox{*}         & $\mathbf{66.26\pm7.45}$  & $\mathbf{67.08\pm8.97}$  & $\mathbf{65.76\pm7.79}$  & $\mathbf{64.06\pm10.00}$  \\
+                                    & \Verb|MAML|(\Verb|BiLSTM+CNN|)     & $66.08\pm8.19$           & $66.56\pm9.23$           & $65.41\pm8.66$           & $63.78\pm10.83$           \\
+                                    & \Verb|MAML|(\Verb|Transformers|)   & $56.03\pm6.08$           & $56.32\pm6.01$           & $55.63\pm5.01$           & $52.33\pm7.08$            \\ 
+    \hline
+    \multirow{3}{*}{\Verb|ETT-m2|}  & \Verb|Temporal-ML|\mbox{*}         & $71.14\pm9.13$           & $62.21\pm7.51$           & $58.75\pm5.30$           & $57.61\pm6.74$            \\
+                                    & \Verb|MAML|(\Verb|BiLSTM+CNN|)     & $\mathbf{71.40\pm9.10}$  & $\mathbf{63.25\pm8.82}$  & $\mathbf{60.22\pm8.42}$  & $\mathbf{59.43\pm9.24}$   \\
+                                    & \Verb|MAML|(\Verb|Transformers|)   & $68.69\pm8.85$           & $58.54\pm4.38$           & $57.85\pm4.71$           & $57.07\pm5.14$            \\ 
+    \hline
+    \multirow{3}{*}{\Verb|WTH|}     & \Verb|Temporal-ML|\mbox{*}         & $\mathbf{74.97\pm2.63}$  & $\mathbf{69.13\pm3.44}$  & $\mathbf{65.98\pm3.96}$  & $\mathbf{66.00\pm3.80}$   \\
+                                    & \Verb|MAML|(\Verb|BiLSTM+CNN|)     & $74.41\pm2.71$           & $68.17\pm4.60$           & $65.50\pm4.82$           & $65.63\pm4.47$            \\
+                                    & \Verb|MAML|(\Verb|Transformers|)   & $72.28\pm3.06$           & $62.35\pm8.73$           & $60.90\pm4.97$           & $58.65\pm4.72$            \\
+    \bottomrule
+\end{tabular}}
+\end{table}
+
 % giải thích lý do tại sao LSTM lại ngon hơn LSTM+CNN
+Thông qua kết quả này, có thể nhận thấy việc sử dụng \verb|BiLSTM| mang lại kết quả cao và ổn định trong quá trình dự đoán.
+Việc sử dụng các mạng như \verb|CNN|, \verb|Transformer| không những không tạo ra các cải thiện hiệu quả mà còn làm nhiễu quá trình học hoặc giảm hiệu quả của cả mô hình.
+Đối với việc sử dụng \verb|BiLSTM+CNN|, quá trình học bị nhiễu do bản thân \verb|BiLSTM| đã chọn lọc được các ràng buộc thời gian hiệu quả trong quá trình rút trích trong khi \verb|CNN| chỉ rút trích đặc trưng chứ không hề chọn lọc.
+Điều này dẫn đến việc đặc trưng của \verb|BiLSTM+CNN| không ổn định.
+Nếu \verb|CNN| rút trích được đặc trưng tốt, mạng sẽ hoạt động hiệu quả (e.g., metrics on \verb|ETT-m2|).
+Ngược lại, mạng sẽ hoạt động kém hiệu quả hơn so với việc chỉ sử dụng \verb|BiLSTM| (e.g., metrics on \verb|USD/JPY|, \verb|multi-fx|, \verb|WTH|).
 
-\section{Sub-model aggregation ability}
+% giải thích lý do tại sao LSTM lại ngon hơn Transformer
+Ngoài ra, chúng tôi thực hiện phân tích quá trình hội tụ của \verb|Temporal-ML| và các thuật toán nêu trên trên hai tập dữ liệu phi chu kỳ.
+Kết quả được trình bày trong hình \ref{fig:ablation_att_usd} và \ref{fig:ablation_att_multi}.
+Nhìn chung, trong quá trình dự đoán các attribute của \verb|USD/JPY| và \verb|multi-fx|, \verb|Temporal-ML| đều cho độ chính xác cao và quá trình hội tụ nhanh.
+Riêng dự đoán \textit{close} price trên \verb|multi-fx|, \verb|MAML(BiLSTM+CNN)| cho độ chính xác lẫn quá trình hội tụ tốt hơn thuật toán đề xuất.
+Điều đáng ngạc nhiên là \verb|MAML(Transformer)| chỉ hội tụ khi dự đoán \textit{open} price của cả hai tập dữ liệu foreign exchange.
+Chúng tôi cho rằng, để có thể sử dụng một mô hình phức tạp như \verb|Transformer|, cần sử dụng một lượng lớn dữ liệu với số lượng inner epochs lớn.
+Điều này khiến cho \verb|Transformer| không phải là một lựa chọn tốt khi kết hợp với các thuật toán ML.
 
-% vẽ về quá trình hội tụ
-% mô tả về quá trình hội tụ của các thuật toán
-% múa
+Với các ý phân tích nêu trên, chúng tôi kết luận rằng, \verb|BiLSTM| không chỉ là lựa chọn phù hợp trong việc khám phá sự biến thiên của aperiodic time-series data mà còn là thuật toán thích hợp để kết hợp với các thuật toán ML khi làm việc trên time-series data nói chung.
+\textbf{Bằng việc sử dụng \Verb|BiLSTM|, chúng tôi đã rút trích được sự biến động của thị trường cũng như dự đoán tương lai một cách hiệu quả}.
+
+\begin{sidewaysfigure}[h]
+    \centering
+    \begin{subfigure}[b]{\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{ab_att_usd.pdf}
+        \cprotect\caption{\verb|Temporal-ML| vs. Models without \verb|BiLSTM| on \verb|USD/JPY|.}
+        \label{fig:ablation_att_usd}
+    \end{subfigure}
+    ~
+    \begin{subfigure}[b]{\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{ab_att_multi.pdf}
+        \cprotect\caption{\verb|Temporal-ML| vs. Models without \verb|BiLSTM| on \verb|multi-fx|.}
+        \label{fig:ablation_att_multi}
+    \end{subfigure}
+    ~
+    \begin{subfigure}[b]{\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{ab_ml_multi.pdf}
+        \cprotect\caption{\verb|Temporal-ML| vs. Models without \verb|MAML| on \verb|multi-fx|.}
+        \label{fig:ablation_ml_multi}
+    \end{subfigure}
+    \cprotect\caption{Convergence process of algorithms in ablation study on all attribute of foreign exchange datasets.}
+    \label{fig:ablation}
+\end{sidewaysfigure}
+
+\section{Models' parameters aggregation ability}
+
+% đấm nhits
+Trong quá trình tổng hợp các đặc trưng để đưa ra dự đoán cuối, \verb|NHITS| chỉ sử dụng các phép cộng trừ đơn giản.
+Điều này đặt trong ngữ cảnh của việc tổng hợp thông tin dựa trên các dải tần phân tách được từ dữ liệu đầu vào là hợp lý.
+Thật vậy, các dữ liệu liên quan đến thời tiết, nhu cầu sử dụng máy bay hoặc điện thể hiện tính chu kỳ rất mạnh vì thời tiết và hành vi tiêu dùng của con người, dù có thể chênh lệch theo thời gian nhưng vẫn thể hiện rất rõ tính mùa vụ (e.g., trời lạnh vào mùa đông, nhu cầu sử dụng máy bay tăng cao vào ngày lễ).
+Do đó, tần suất của loại dữ liệu này là một đặc trưng tốt và dễ dàng rút trích.
+Sau khi phân rã dữ liệu đầu vào thành các dải tần, chỉ cần tổng hợp chúng để có thể dự đoán tương lai.
+Ngoài ra, việc tổng hợp các dải tần bằng phép cộng đơn thuần có thể coi là rất đơn giản so với các thuật toán học sâu hiện tại, vốn làm việc trên không gian đặc trưng ẩn của dữ liệu.
+Do đó, quá trình tổng hợp của \verb|NHITS| trên dữ liệu phi chu kỳ gặp rất nhiều khó khăn vì thông tin đầu vào cho quá trình tổng hợp không tốt và phương pháp tổng hợp không hiệu quả.
+Từ đây, có thể thấy rằng, aperiodic time-series data trở thành điểm yếu chí mạng cho \verb|NHITS| và các phương pháp dựa trên việc phân rã tần số nói chung.
+
+% nâng bi bản thân
+Đối diện với sự phức tạp trong dữ liệu phi chu kỳ đòi hỏi một cấu trúc phức tạp để có thể tổng hợp các đặc trưng một cách hợp hiệu quả.
+Như đã đề cập trong phần Motivation (section \ref{sec:moti}), việc phân tích và tổng hợp thông tin từ nhiều nguồn dữ liệu trong cùng một domain có thể cung cấp các thông tin hữu ích trong việc dự đoán các chỉ số quan tâm.
+\verb|Temporal-ML| với khả năng tổng hợp thông tin dựa trên quá trình tối ưu của các thuật toán ML, cho phép mô hình nắm bắt thông tin đa chiều từ các nguồn dữ liệu khác nhau.
+Từ đó đưa ra dự đoán hiệu quả.
+
+% phân tích bảng kq
+Để nắm bắt rõ khả năng của các thuật toán ML trong \verb|Temporal-ML|, chúng tôi loại bỏ component này khỏi thuật toán đề xuất sau đó chạy thực nghiệm để so sánh với kết quả ban đầu (subsection \ref{subsec:ablation_ml}).
+Chúng tôi sử dụng các thuật toán ML nhằm mục đích tổng hợp thông tin đa nguồn.
+Do đó, các so sánh chỉ được thực hiện trên tập dữ liệu \verb|multi-fx| vì chỉ tập dữ liệu này chứa nhiều tập dữ liệu con từ nhiều nguồn khác nhau.
+Kết quả được tổng hợp trong bảng \ref{tab:ab_ml}.
+Theo đó, với khả năng tổng hợp thông tin của mình, \verb|Temporal-ML| cho kết quả cao hơn từ 2-12\% so với các thuật toán không sử dụng ML trong quá trình huấn luyện.
+Ngoài ra, mức phân tán của các metrics cũng thấp hơn từ 1-7\%, chứng tỏ rằng mức chênh lệch trong quá trình dự đoán các task tốt hơn nhiều so với các thuật toán còn lại.
+
+Việc đạt được kết quả tốt hơn trên \verb|multi-fx| của \verb|Temporal-ML| là hoàn toàn dễ hiểu vì tất cả phương pháp còn lại trong bảng \ref{tab:ab_ml} đều không có khả năng tổng hợp thông tin như thuật toán đề xuất mà chỉ dựa hoàn toàn vào dữ liệu quá khứ để dự đoán tương lai.
+Do đó, không chỉ \verb|BiLSTM|, \verb|BiLSTM+CNN| mà ngay cả \verb|Transformer| cũng không thể tận dụng được mối tương quan giữa các data sources.
+
+\begin{table}[H]
+    \centering
+    \cprotect\caption{Classification results (\%) of \verb|Temporal-ML| and models without \verb|MAML| on \verb|multi-fx|. Best results per metrics are boldfaced. (\mbox{*}): Our method.}
+    \label{tab:ab_ml}
+    \begin{tabular}{c|cccc} 
+    \toprule
+                                & $accuracy$              & $precision$             & $recall$                & $F1-score$                \\ 
+    \hline
+    \Verb|Temporal-ML|\mbox{*}  & $\mathbf{66.26\pm7.45}$ & $\mathbf{67.08\pm8.97}$ & $\mathbf{65.76\pm7.79}$ & $\mathbf{64.06\pm10.00}$  \\
+    \Verb|BiLSTM|               & $63.88\pm9.00$          & $64.18\pm13.39$         & $63.66\pm9.13$          & $60.07\pm13.09$           \\
+    \Verb|BiLSTM+CNN|           & $62.23\pm8.63$          & $62.89\pm9.76$          & $62.20\pm8.34$          & $59.91\pm10.97$           \\
+    \Verb|Transformers|         & $58.44\pm9.15$          & $56.71\pm15.10$         & $58.22\pm8.84$          & $52.22\pm13.12$           \\
+    \bottomrule
+    \end{tabular}
+\end{table}
+
+% phân tích hội tụ
+Quá trình hội tụ của \verb|Temporal-ML| và các model without ML được trình bày trong hình \ref{fig:ablation_ml_multi}.
+Trong đó, khi dự đoán \textit{open} price, thuật toán đề xuất cho khả năng hội tụ vượt trội so với các thuật toán còn lại.
+Đối với các thuộc tính \textit{high} price và \textit{close} price, \verb|Temporal-ML| tỏ ra đuối sức trong giai đoạn đầu của quá trình huấn luyện.
+Tuy nhiên, khi về cuối, mô hình đề xuất cho thấy khả năng vượt lên nhanh chóng và tiềm năng cải thiện độ chính xác hơn nữa so với các thuật toán khác, vốn đã hội tụ.
+
+Do đó, có thể thấy rằng, \textbf{việc sử dụng các thuật toán optimization-based ML trong việc tổng hợp thông tin là hoàn toàn hợp lý vì đã tận dụng thành công mối tương quan giữa các tập dữ liệu và đạt được độ chính xác cao mà không đòi hỏi quá nhiều dữ liệu như trong ngữ cảnh đơn nguồn}.

### 2025-01-21 doc/thesis/chapters/6.conclusion.tex
chems gios

@@
 
-\chapter{Conclusion \& Future works}
+\chapter{Conclusions \& Future works}
 \label{chap:conclusion}
 
-\section{Conclusion}
+\section{Conclusions}
 
-% viết về contribution
+Các bài toán trên aperiodic time-series data đặt ra một thử thách lớn cho các mô hình học máy hiện nay vì dữ liệu này phụ thuộc vào cả lịch sử giao dịch lẫn các external factor như tình hình kinh tế, chính trị; variance của dữ liệu thay đổi liên tục; tính phi chu kỳ của dữ liệu thể hiện rất mạng.
+Trong khi các thách thức này trở nên rất đặc thù cho dữ liệu và yêu cầu một phương pháp được thiết kế riêng, các thuật toán hiện tại hầu như không tập trung giải quyết các thách thức này mà chỉ hướng đến time-series nói chung.
+Thật vậy, các phương pháp học sâu như \verb|LSTM|, \verb|CNN|, \verb|Transformer| được thiết kế để phân tích dữ liệu dạng chuỗi nói chung.
+Đối với dạng dữ liệu đặc thù như aperiodic time-series data, chúng trở nên kém hiệu quả do hạn chế trong việc lưu trữ các ràng buộc thời gian.
+Các phương pháp dựa trên phân rã tần số như \verb|NHITS| và \verb|bla_bla| hướng đến phân rã tín hiệu đầu vào thành các dải tần nhưng tính chu kỳ hoàn toàn không tồn tại trong aperiodic time-series data.
+
+Ý thức được các thách thức nêu trên cùng với việc tận dụng khả năng rút trích ràng buộc thời gian của \verb|BiLSTM| và khả năng tổng hợp tham số hiệu quả của \verb|MAML|, chúng tôi đề xuất thuật toán \verb|Temporal-ML| với khả năng rút trích thời gian đặc trưng ẩn cũng như tổng hợp thông tin từ dữ liệu đa nguồn.
+Dựa trên thực nghiệm, trong quá trình giải quyết bài toán dự đoán xu hướng tiếp theo, phương pháp đề xuất đạt hiệu quả cao trên hai tập dữ liệu phi chu kỳ (\verb|USD/JPY| và \verb|multi-fx|) và hiệu quả tương đương trên các tập dữ liệu có chu kỳ (\verb|ETT-m2| và \verb|WTH|) so với \verb|NHITS| (SOTA model in AAAI 2023).
+Ngoài ra, bằng việc thực hiện ablation study, chúng tôi chứng minh được sự hiệu quả của từng component trong \verb|Temporal-ML|.
+Cụ thể, khả năng rút trích hiệu quả các ràng buộc thời gian đến từ \verb|BiLSTM| và khả năng tổng hợp thông tin đa nguồn đến từ \verb|MAML|.
+Theo đó, thuật toán của chúng tôi đã giải quyết hiệu quả vấn đề dữ liệu phụ thuộc vào external factor, vấn đề variance bất định và vấn đề phi chu kỳ của dữ liệu.
+
+% Bằng việc kết hợp các đặc trưng cục bộ và các phụ thuộc dài hạn cũng như khai thác các phụ thuộc ẩn trong từng khoảng thời gian quá khứ, chúng tôi đã cải khả năng của mô hình học máy trên các classification metrics trong bài toán dự đoán xu hướng giá của ngày tiếp theo trên dữ liệu FX.
+Thuật toán đề xuất đã chứng minh được khả năng vượt trội của mình khi so sánh với mô hình \verb|NHITS|.
+
+% Aperiodic time-series data là một dạng dữ liệu phổ biến.
+Việc khai thác các insight trong loại dữ liệu này mang lại nhiều thông tin hữu ích.
+Tuy nhiên các phương pháp hiện tại gặp khó khăn rất lớn trong việc phân tích và dự đoán loại dữ liệu này.
 
 \section{Future works}
+
+In the future, we propose two main directions of development related to the architecture and personalization of the model
+
+\textbf{Model architecture}.
+Phương pháp của chúng tôi được phát triển dưới dạng component với hai components chính hoạt động song song: Temporal feature extraction and Models' parameter synthesis.
+Điều này cung chấp cho phương pháp của chúng tôi một khả năng nâng cấp linh hoạt.
+Thí nghiệm của chúng tôi chỉ minh họa một trường hợp điển hình trong tổng hợp hiệu quả các đặc trưng.
+Bằng việc sử dụng các thuật toán ML khác nhau như \verb|Reptile|, \verb|Meta-SGD|, \verb|iMAML|, hoàn toàn có thể tạo ra mô hình mới với độ chính xác cao hơn.
+
+\textbf{Long-horizon problem}.
+Hoàn toàn có thể mở rộng phương pháp này để giải các bài toán về long-horizon prediction.
+Thật vậy, bằng việc thay đổi đầu ra và độ lỗi của mô hình, có thể tiến hành giải các bài toán này.
+Tuy nhiên, kiến trúc của các sub-model cần được nghiên cứu lại để phù hợp hơn với bài toán mới.

### 2025-01-21 doc/thesis/pre_info/acknowledge.tex
ack, title, ref stuff

@@
 
-I would like to thank everyone in Maezono and Hongo laboratory, fellow students and the laboratory staff, for helping me out whenever I run into trouble with the administrative documents or health.
+In addition, I would like to thank everyone in Maezono and Hongo laboratory, fellow students and the laboratory staff, for helping me out whenever I run into trouble with the administrative documents or health.
 

### 2025-01-30 doc/thesis/chapters/2.related_works.tex
add thesis, create research log

@@
 
-Thay vì hoạt động trên một lướt pixel như \verb|CNN|, Recurrent neural network (\verb|RNN|) \cite{rumelhart1986learning} hoạt động trên một chuỗi các giá trị $\mathbf{x}^{(1)},\dots,\mathbf{x}^{(\tau)}$.
-So với mạng mạng \verb|FullyConnected|, \verb|RNN| có thể mở rộng số lượng datapoint trong một chuỗi đến vô hạn mà không cần tăng kích thước của mạng.
+% Thay vì hoạt động trên một lưới pixel như \verb|CNN|, Recurrent neural network (\verb|RNN|) \cite{rumelhart1986learning} hoạt động trên một chuỗi các giá trị $\mathbf{x}^{(1)},\dots,\mathbf{x}^{(\tau)}$.
+So với mạng mạng \verb|FullyConnected|, \verb|RNN| có thể hoạt động trên chuỗi có độ dài bất kỳ mà không cần tăng kích thước của mạng.
 Nói cách khác, \verb|RNN| có khả năng làm việc trên các chuỗi dài hiệu quả hơn nhiều so với mạng neural truyền thống.
 
-Tương tự như \verb|CNN|, \verb|RNN| cũng sử dụng một kiến trúc mạng cho phép chia sẻ tham số.
+Instead of operating on a pixel grid like \verb|CNN|, Recurrent neural network (\verb|RNN|) \cite{rumelhart1986learning} operates on a sequence of values $\mathbf{x}^{(1)},\dots,\mathbf{x}^{(\tau)}$.
+Compared to \verb|FullyConnected|, \verb|RNN| can operate on sequences of any length without increasing the size of the network.
+In other words, \verb|RNN| is able to work on long sequences much more efficiently than traditional neural networks.
+
+% Tương tự như \verb|CNN|, \verb|RNN| cũng sử dụng một kiến trúc mạng cho phép chia sẻ tham số.
 Tuy nhiên, \verb|CNN| chia sẻ tham số trong nội bộ của một mẫu dữ liệu (cùng một kernel hoạt động trên các vùng cục bộ của một mẫu dữ liệu).
@@
 
-Cụ thể, gọi hàm $f$ là mạng \verb|RNN|. Hàm $f$ sử dụng bộ tham số $\theta$ để rút trích thông tin từ trạng thái $\mathbf{s}^{(t-1)}$ để tạo ra trạng thái $\mathbf{s}^{(t)}$:
+Similar to \verb|CNN|, \verb|RNN| also uses a network architecture that allows parameter sharing.
+However, \verb|CNN| shares parameters within a data sample (the same kernel operates on local regions of a data sample).
+\verb|RNN| shares parameters by using the same set of parameters for different data samples.
+Therefore, when a data sample is input to the model, the model output contains information about this data sample and all previous data samples.
+This makes it possible and efficient to extend the model and apply it to data samples of different lengths.
+
+% Cụ thể, gọi hàm $f$ là mạng \verb|RNN|. Hàm $f$ sử dụng bộ tham số $\theta$ để rút trích thông tin từ trạng thái $\mathbf{s}^{(t-1)}$ để tạo ra trạng thái $\mathbf{s}^{(t)}$:
+
+Specifically, let the function $f$ be the \verb|RNN| network.
+The function $f$ uses the parameter set $\theta$ to extract information from state $\mathbf{s}^{(t-1)}$ to generate state $\mathbf{s}^{(t)}$ as follows:
 
@@
 
-Lưu ý rằng, $\mathbf{s}^{(t)}$ là một abstract variable, ám chỉ thông tin hiện tại phụ thuộc vào các thông tin trong quá khứ.
+% Lưu ý rằng, $\mathbf{s}^{(t)}$ là một abstract variable, ám chỉ thông tin hiện tại phụ thuộc vào các thông tin trong quá khứ.
 Đối với chuỗi có độ dài $\tau$ bất kỳ, equation \ref{eq:recurrent} có thể được triển khai $\tau-1$ lần để nén các thông tin trong chuỗi vào các trạng thái $\mathbf{s}^{(t)}$.
 
-Trong các mạng \verb|RNN| thông thường, biến trạng thái $\mathbf{s}$ được thể hiện bằng thông tin rút trích được từ mạng $\mathbf{h}$ và dữ liệu $\mathbf{x}$.
+Note that, $\mathbf{s}^{(t)}$ is an abstract variable, implying that current information depends on past information.
+For any sequence of length $\tau$, the equation \ref{eq:recurrent} can be implemented $\tau-1$ times to compress the information in the sequence into $\mathbf{s}^{(t)}$ states.
+
+% Trong các mạng \verb|RNN| thông thường, biến trạng thái $\mathbf{s}$ được thể hiện bằng thông tin rút trích được từ mạng $\mathbf{h}$ và dữ liệu $\mathbf{x}$.
 Cụ thể, trạng thái hiện tại $\mathbf{h}^{(t)}$ được rút trích dựa trên trạng thái trước đó $\mathbf{h}^{(t-1)}$ (chứa thông tin về toàn bộ chuỗi trước đó) và mẫu dữ liệu hiện tại $\mathbf{x}^{(t)}$:
+
+In typical \verb|RNNs|, the state variable $\mathbf{s}$ is represented by information extracted from the network $\mathbf{h}$ and the data $\mathbf{x}$.
+Specifically, the current state $\mathbf{h}^{(t)}$ is extracted based on the previous state $\mathbf{h}^{(t-1)}$ (which contains information about the entire previous sequence) and the current data sample $\mathbf{x}^{(t)}$:
 
@@
 
-Phương trình \ref{eq:rnn} có thể được viết lại dưới dạng mô hình $g^{(t)}$ như sau:
+% Phương trình \ref{eq:rnn} có thể được viết lại dưới dạng mô hình $g^{(t)}$ như sau:
+
+The equation \ref{eq:rnn} can be rewritten in the form of the $g^{(t)}$ model as follows:
 
@@
 
-Từ đây, có thể thấy rằng, mô hình $g$ không cần phải cập nhật tham số khi học từng mẫu $x^{(t)}$ như mạng neural truyền thống.
+% Từ đây, có thể thấy rằng, mô hình $g$ không cần phải cập nhật tham số khi học từng mẫu $x^{(t)}$ như mạng neural truyền thống.
 Thay vào đó, nó có thể sử dụng thông tin từ một chuỗi $\mathbf{x}^{(1)},\dots,\mathbf{x}^{(\tau)}$ để cập nhật tham số thông qua mô hình $f$.
-Ngoài ra, việc sử dụng mô hình $f$ còn cải thiện khả năng tổng quát hóa của mô hình đối với những chuỗi có độ dài bất kỳ.
+Do đó, việc sử dụng mô hình $f$ còn cải thiện khả năng tổng quát hóa của mô hình đối với những chuỗi có độ dài bất kỳ.
+
+As a result, it can be seen that the model $g$ does not need to update its parameters when learning each sample $x^{(t)}$ like a traditional neural network.
+Instead, it can use information from a sequence $\mathbf{x}^{(1)},\dots,\mathbf{x}^{(\tau)}$ to update its parameters through the model $f$.
+Thus, using the model $f$ also improves the model's generalization ability to sequences of any length.
 
@@
     \includegraphics[width=\linewidth]{rnn.png}
-    \cprotect\caption{The computational graph of \verb|RNN| in next-word prediction problem \cite{goodfellow2016deep}.}
+    \cprotect\caption{An example of computational graph of \verb|RNN| \cite{goodfellow2016deep}.}
     \label{fig:rnn}
@@
 
-Dựa vào ý tưởng của recurrent function $f$ trình bày ở trên, chúng ta có thể thiết kế rất nhiều kiến trúc mạng \verb|RNN|.
+% Dựa vào ý tưởng của recurrent function $f$ trình bày ở trên, chúng ta có thể thiết kế rất nhiều kiến trúc mạng \verb|RNN|.
 Tại đây, chúng tôi trình bày về một thiết kế đơn giản có đồ thị tính toán như hình \ref{fig:rnn}.
-Theo đó, mạng nhận vào một chuỗi các giá trị $\mathbf{x}$ với mục tiêu ánh xạ các giá trị này vào các đầu ra $\mathbf{o}$ khớp với $\mathbf{y}$.
-Để thực hiện điều này, mạng sử dụng ba bộ tham số chính: $\mathbf{U}$ để ánh xạ $\mathbf{x}$ vào không gian đặc trưng ẩn; $\mathbf{W}$ để ánh xạ đặc trưng ẩn của timestamp trước vào không gian đặc trưng ẩn của timestamp sau; $\mathbf{V}$ để ánh xạ đặc trưng ẩn vào không gian chứa các giá trị đầu ra chưa chuẩn hóa ($\mathbf{o}$).
-Giá trị dự đoán $\hat{y}$ sau đó được tính từ $\mathbf{o}$.
+Theo đó, mạng nhận vào một chuỗi các giá trị $\mathbf{x}$ với mục tiêu ánh xạ các giá trị này vào các đầu ra $o$ khớp với $\mathbf{y}$.
+Để thực hiện điều này, mạng sử dụng ba bộ tham số chính: $U$ để ánh xạ $\mathbf{x}$ vào không gian đặc trưng ẩn; $W$ để ánh xạ đặc trưng ẩn của timestamp trước vào không gian đặc trưng ẩn của timestamp sau; $V$ để ánh xạ đặc trưng ẩn vào không gian chứa các giá trị đầu ra chưa chuẩn hóa.
+Giá trị dự đoán $o$ thu được sau quá trình chuẩn hóa.
 Cả quá trình được thể hiện như sau:
 
+Based on the idea of recurrent function $f$ presented above, we can design many \verb|RNN| network architectures.
+Here, we present a simple design with a computational graph as shown in \ref{fig:rnn}.
+Accordingly, the network takes in a sequence of values $\mathbf{x}$ with the goal of mapping these values to outputs $o$ matching $\mathbf{y}$.
+To do this, the network uses three main sets of parameters: $U$ to map $\mathbf{x}$ to the hidden feature space; $W$ to map the hidden feature of the previous timestamp to the hidden feature space of the current timestamp; $V$ to map the hidden feature to the space containing unnormalized output values.
+The predicted value $o$ is obtained after the normalization process.
+The whole process is shown as follows:
+
 \begin{align}
-    \mathbf{a}^{(t)} &= \mathbf{b} + \mathbf{W}\mathbf{h}^{(t-1)} + \mathbf{U}\mathbf{x}^{(t)}\quad \text{($\mathbf{b}$ is bias vector)}\\
+    \mathbf{a}^{(t)} &= \mathbf{b} + W\mathbf{h}^{(t-1)} + U\mathbf{x}^{(t)} \quad \text{($\mathbf{b}$ is bias vector)}\\
     \mathbf{h}^{(t)} &= \tanh\left( \mathbf{a}^{(t)} \right)\\
-    \mathbf{o}^{(t)} &= \mathbf{c} + \mathbf{V}\mathbf{h}^{(t)}\quad \text{($\mathbf{c}$ is bias vector)}\\
-    \hat{y}^{(t)} &= \text{softmax}\left( \mathbf{o}^{(t)} \right)
+    o^{(t)} &= \text{softmax}\left( \mathbf{c} + V\mathbf{h}^{(t)} \right) \quad \text{($\mathbf{c}$ is bias vector)}
 \end{align}
 
-Độ lỗi của mô hình được tính bằng hàm \verb|CrossEntropyLoss| (equation \ref{eq:loss_rnn}).
-Hàm lỗi này được sử dụng để cập nhật các tham số của hàm $f$ thông qua kỹ thuật back-propagation through time.
+% Độ lỗi của mô hình được tính bằng hàm \verb|CrossEntropyLoss| (equation \ref{eq:loss_rnn}).
+Hàm lỗi này được sử dụng để cập nhật các tham số của hàm $f$ thông qua kỹ thuật back-propagation through time (BPTT).
+
+The model loss is calculated using the \verb|CrossEntropyLoss| function (equation \ref{eq:loss_rnn}).
+This error function is used to update the parameters of the function $f$ via back-propagation through time (BPTT).
 
 \begin{align}
-    L\left( \left\{ \hat{y}^{(t)} \right\}_{t=1}^\tau, \left\{ y^{(t)} \right\}_{t=1}^\tau \right) = \sum_{t=1}^\tau{L^{(t)}\left( \hat{y}^{(t)}, y^{(t)} \right)}
+    L\left( \left\{ o^{(t)} \right\}_{t=1}^\tau, \left\{ y^{(t)} \right\}_{t=1}^\tau \right) = \sum_{t=1}^\tau{L^{(t)}\left( o^{(t)}, y^{(t)} \right)}
     \label{eq:loss_rnn}
@@
 
-Trong quá trình forward và backward, các bước tính toán được thực hiện tuần tự vì phải tuân theo trình tự thời gian.
+% Kỹ thuật BPTT được thực hiện bằng cách lấy đạo hàm của $L$ theo các trọng số xuyên suốt các timestamp từ $t=\tau$ đến $t=1$.
+Để thuận tiện trong việc trình bày, chúng tôi viết lại quá trình forward như sau:
+
+The BPTT technique is implemented by taking the derivative of $L$ with respect to the weights across timestamps from $t=\tau$ to $t=1$.
+For convenience in presentation, we rewrite the forward process as follows:
+
+\begin{align}
+    \mathbf{h}^{(t)} &= f\left( \mathbf{x}^{(t)}, \mathbf{h}^{(t-1)}, W_h \right)\\
+    o^{(t)} &= g\left( \mathbf{h}^{(t)}, W_o \right)
+\end{align}
+
+% Theo đó, đạo hàm của hàm lỗi với $W_o$ và $W_h$ được tính như sau:
+Accordingly, the derivative of the loss function with respect to $W_o$ and $W_h$ is calculated as follows:
+
+\begin{align}
+    \frac{\partial L}{\partial W_o} &= \sum_{t=1}^\tau{\frac{\partial L}{\partial L^{(t)}} \frac{\partial L^{(t)}}{\partial o^{(t)}} \frac{\partial o^{(t)}}{\partial W_o}}\\
+    &= \sum_{t=1}^\tau{\frac{\partial L^{(t)}}{\partial o^{(t)}} \frac{\partial o^{(t)}}{\partial W_o}}
+    \label{eq:grad_wo}
+\end{align}
+
+\begin{align}
+    \frac{\partial L}{\partial W_h} &= \sum_{t=1}^\tau{\frac{\partial L^{(t)}}{\partial W_h}}\\
+    &= \sum_{t=1}^\tau{\frac{\partial L^{(t)}}{\partial o^{(t)}} \frac{\partial o^{(t)}}{\partial \mathbf{h}^{(t)}} \frac{\partial \mathbf{h}^{(t)}}{\partial W_h}}
+\end{align} In which,
+
+\begin{align}
+    \frac{\partial \mathbf{h}^{(t)}}{\partial W_h} &= \frac{\partial f\left( \mathbf{x}^{(t)}, \mathbf{h}^{(t-1)}, W_h \right)}{\partial W_h} + \frac{\partial f\left( \mathbf{x}^{(t)}, \mathbf{h}^{(t-1)}, W_h \right)}{\partial \mathbf{h}^{(t-1)}} \frac{\partial \mathbf{h}^{(t-1)}}{\partial W_h}\\
+    &= \frac{\partial f\left( \mathbf{x}^{(t)}, \mathbf{h}^{(t-1)}, W_h \right)}{\partial W_h} + \sum_{i=1}^{t-1}\left( \prod_{j=i+1}^t {\frac{\partial f\left( \mathbf{x}^{(j)}, \mathbf{h}^{(j-1)}, W_h \right)}{\partial \mathbf{h}^{(j-1)}}} \right) \frac{\partial f\left( \mathbf{x}^{(i)}, \mathbf{h}^{(i-1)}, W_h \right)}{\partial W_h}
+    \label{eq:grad_wh}
+\end{align}
+
+% Sau khi thu được các đạo hàm như phương trình \ref{eq:grad_wo} và \ref{eq:grad_wh}, chúng ta có thể sử dụng các gradient-based optimization strategies để cập nhật tham số trong mạng.
+
+Once we obtain the derivatives as in equation \ref{eq:grad_wo} and \ref{eq:grad_wh}, we can use gradient-based optimization strategies to update the parameters in the network.
+
+% Trong quá trình forward và backward, các bước tính toán được thực hiện tuần tự vì phải tuân theo trình tự thời gian.
 Do đó, time complexity của \verb|RNN| là $\mathcal{O}(\tau)$.
@@
 
+In the forward and backward processes, the computation steps are performed sequentially because they must follow the input order.
+Therefore, the time complexity of \verb|RNN| is $\mathcal{O}(\tau)$.
+Since the later derivative values must be stored to compute the previous derivative values (according to the Chain rule), the space complexity of the algorithm is also $\mathcal{O}(\tau)$.
+Therefore, for long sequences, the computation can be extremely expensive.
+
 \subsection{Long Short-Term Memory}
 
-Khi làm việc trên các chuỗi dài, \verb|RNN| rất dễ mắc phải vấn đề \textit{vanishing gradient} hoặc \textit{exploding gradient}.
+% Giả sử các dữ liệu thuộc không gian $\mathbb{R}$ và hàm $f$ trong equation \ref{eq:grad_wh} có dạng $\sigma(w(h+x))$. Lúc này, the product có dạng:
+
+Suppose the data are in 1-dimensional space $\mathbb{R}$ and the function $f$ in equation \ref{eq:grad_wh} is of the form $\sigma(w(h+x))$. Then, the product becomes:
+
+\begin{align}
+    \prod_{j=i+1}^t {\frac{\partial \sigma(w(h+x))}{\partial h}} &= \prod_{j=i+1}^t {w \sigma'(w(h+x))}\\
+    &= w^{t-i} \prod_{j=i+1}^t {\sigma'(w(h+x))}
+    \label{eq:vanish}
+\end{align}
+
+% Trong trường hợp $t$ lớn và $w \ne 1$, this product sẽ giảm về 0 hoặc tăng đến vô cùng theo hàm số mũ.
+Do đó, khi làm việc trên các chuỗi dài, \verb|RNN| rất dễ mắc phải vấn đề \textit{vanishing gradient} hoặc \textit{exploding gradient}.
 Thật vậy, quá trình huấn luyện của \verb|RNN| đòi hỏi việc sử dụng Chain rule cho tất cả các trạng thái từ $t=\tau$ đến $t=1$.
@@
 
-% viết 1 chút về bidirectional lstm
+In the case of large $t$ and $w \ne 1$, this product will either decrease to 0 or increase to infinity exponentially.
+Therefore, when working on long sequences, \verb|RNN| is susceptible to the \textit{vanishing gradient} or \textit{exploding gradient} problem.
+Indeed, the training process of \verb|RNN| requires the application of the Chain rule for all states from $t=\tau$ to $t=1$.
+This process involves a lot of matrix and vector multiplication as illustrated in equation \ref{eq:vanish}.
+Therefore, if the elements in that matrix or vector are less than 1, the derivative with respect to time will converge to 0.
+Conversely, if the elements are greater than 1, the derivative will approach infinity over time.
+
+% Để giải quyết vấn đề trên, nghiên cứu \cite{hochreiter1997long} đề xuất mạng Long Short-Term Memory (\verb|LSTM|), một biến thể của \verb|RNN|.
+\verb|LSTM| được phát triển dựa trên ba ý tưởng chính: Sử dụng cell state cho mỗi timestamp; Sử dụng các cổng để điều chỉnh dòng thông tin; Kỹ thuật dòng đạo hàm không gián đoạn.
+
+To solve the above problem, \cite{hochreiter1997long} proposed Long Short-Term Memory (\verb|LSTM|).
+\verb|LSTM| is developed based on three main ideas: Using cell state for each timestamp; Using gates to regulate information flow; Uninterrupted derivative flow technique.
+
+% Thứ nhất, mỗi cell của mạng duy trì một trạng thái riêng biệt $\mathbf{c}^{(t)}$.
+Trạng thái này chủ động tổng hợp các thông tin thu được từ đầu vào $\mathbf{x}^{(t)}$ và thông tin từ các trạng thái trước đó.
+Quá trình này lưu lại những thông tin quan trọng và loại bỏ các thông tin không quan trọng.
+Từ đó, mạng \verb|LSTM| có thể nén được nhiều thông tin có giá trị khi học qua một chuỗi thay vì cố gắng ghi nhớ tất cả các thông tin như \verb|RNN| truyền thống.
+
+First, each cell of the network maintains a separate state $\mathbf{c}^{(t)}$.
+This cell state actively synthesizes information obtained from the input $\mathbf{x}^{(t)}$ and information from previous states.
+This process retains important information and discards unimportant information.
+As a result, the \verb|LSTM| network can compress more valuable information when learning through a sequence instead of trying to remember all the information like traditional \verb|RNN|.
+
+% Thứ hai, \verb|LSTM| sử dụng ba cổng forget, input và output gate và giá trị input node để điều chỉnh các luồng thông tin.
+Tại timestamp $t$, các toán tử này nhận vào một vector nối dài bao gồm hidden state của trạng thái trước đó $\mathbf{h}^{(t-1)}$ và đầu vào hiện tại $\mathbf{x}^{(t)}$.
+Cụ thể, chúng hoạt động như sau:
+
+Second, \verb|LSTM| uses three forget gates, input and output gates, and input node values to regulate the information flow.
+At timestamp $t$, these operators take in an extended vector consisting of the hidden state of the previous state $\mathbf{h}^{(t-1)}$ and the current input $\mathbf{x}^{(t)}$.
+Specifically, they work as follows:
+
+\begin{align}
+    \mathbf{f}^{(t)} &= \sigma\left(W_f \left[ \mathbf{x}^{(t)}, \mathbf{h}^{(t-1)} \right]\right)\\
+    \mathbf{i}^{(t)} &= \sigma\left(W_i \left[ \mathbf{x}^{(t)}, \mathbf{h}^{(t-1)} \right]\right)\\
+    \hat{\mathbf{c}}^{(t)} &= \tanh\left(W_c \left[ \mathbf{x}^{(t)}, \mathbf{h}^{(t-1)} \right]\right)\\
+    \mathbf{o}^{(t)} &= \sigma\left(W_o \left[ \mathbf{x}^{(t)}, \mathbf{h}^{(t-1)} \right]\right)
+\end{align}
+
+% Dễ thấy, giá trị của ba cổng này đều nằm trong đoạn $[0,1]$.
+Do đó, chúng có tác dụng gia giảm mức độ thông tin trong trạng thái bộ nhớ.
+Cụ thể, $\mathbf{f}^{(t)}$ chọn lọc và làm suy yếu các thông tin kém hiệu quả của cell state trước đó $\mathbf{c}^{(t-1)}$ bằng cách thực hiện element wise multiplication.
+$\mathbf{i}^{(t)}$ chọn lọc các thông tin đầu vào hiện tại $\hat{\mathbf{c}}^{(t)}$ theo cách tương tự sau đó cập nhật vào cell state hiện tại $\mathbf{c}^{(t)}$ bằng các thực hiện element wise addition với $\mathbf{c}^{(t-1)}$ sau khi đã loại bỏ bớt thông tin dư thừa.
+Quá trình này diễn ra trong phương trình sau:
+
+It is easy to see that the values of these three gates are all in the range $[0,1]$.
+Therefore, they have the effect of decreasing or maintaining the level of information in the memory state.
+Specifically, $\mathbf{f}^{(t)}$ selects and weakens the ineffective information of the previous cell state $\mathbf{c}^{(t-1)}$ by performing element wise multiplication.
+$\mathbf{i}^{(t)}$ selects the current input information $\hat{\mathbf{c}}^{(t)}$ in a similar way and then updates the current cell state $\mathbf{c}^{(t)}$ by performing element wise addition with $\mathbf{c}^{(t-1)}$ (after removing redundant information).
+This process takes place in the following equation:
+
+\begin{align}
+    \mathbf{c}^{(t)} = \mathbf{f}^{(t)} \otimes \mathbf{c}^{(t-1)} + \mathbf{i}^{(t)} \otimes \hat{\mathbf{c}}^{(t)}
+\end{align}
+
+% Quá trình dự đoán sử dụng hidden state $\mathbf{h}^{(t)}$. $\mathbf{h}^{(t)}$ được tạo ra bằng cách điều chỉnh giá trị cell state hiện tại $\mathbf{c}^{(t)}$ bằng giá trị của output gate:
+
+The prediction process uses the hidden state $\mathbf{h}^{(t)}$. $\mathbf{h}^{(t)}$ is generated by adjusting the current cell state value $\mathbf{c}^{(t)}$ based on the value of the output gate:
+
+\begin{align}
+    \mathbf{h}^{(t)} = \mathbf{o}^{(t)} \otimes \tanh\left(\mathbf{c}^{(t-1)}\right)
+\end{align}
+
+% Thứ ba, trong quá trình BPTT, thay vì theo các timestamp thông qua $\mathbf{h}$ như \verb|RNN|, \verb|LSTM| đạo hàm thông qua $\mathbf{c}$.
+Các phép toán thực hiện trên $\mathbf{c}$ không có sự tham gia của bất kỳ trọng số nào.
+Do đó, chúng không bị dư một lượng trọng số như phương trình \ref{eq:vanish}.
+Bằng cách này, \verb|LSTM| đã hạn chế thành công vấn đề vanishing gradient của \verb|RNN|.
+
+Third, in the BPTT process, instead of following the timestamps through $\mathbf{h}$ like \verb|RNN|, \verb|LSTM| derives the derivatives through $\mathbf{c}$.
+The operations performed on $\mathbf{c}$ do not involve any weights.
+Therefore, they do not have a residual amount of weights like the \ref{eq:vanish} equation.
+In this way, \textbf{\Verb|LSTM| successfully mitigates the vanishing gradient problem of \Verb|RNN|}.
 
@@
 
-\subsection{A transformer-based method}
+% \subsection{A transformer-based method}
 
@@
 
-\subsection{NHITS}
+% \subsection{NHITS}
 
@@
 
-\subsection{Model-Agnostic Meta-Learning (MAML)}
+\subsection{Model-Agnostic Meta-Learning}
 
@@
 
-For the optimization-based approach, a basic ML algorithm, typically \verb|MAML|, is learned on multiple tasks $t$ drawn from the same task distribution $\mathcal{T}$ \cite{hospedales2021meta}.
+For the optimization-based approach, a basic ML algorithm, typically Model-Agnostic Meta-Learning (\verb|MAML|), is learned on multiple tasks $t$ drawn from the same task distribution $\mathcal{T}$ \cite{hospedales2021meta}.
 The data for each task is divided into a support set $\mathcal{D}_t^{support}$ (usually small, around 20\%) and a query set $\mathcal{D}_t^{query}$.

### 2025-01-30 doc/thesis/chapters/6.conclusion.tex
add thesis, create research log

--- 2024-12-31 doc/thesis/chapters/6.conclusion.tex
@@
 
-\chapter{Conclusion \& Future works}
+\chapter{Conclusions \& Future works}
 \label{chap:conclusion}
 
-\section{Conclusion}
+\section{Conclusions}
 
-% viết về contribution
+% Các bài toán trên aperiodic time-series data đặt ra một thử thách lớn cho các mô hình học máy hiện nay vì dữ liệu này phụ thuộc vào cả lịch sử giao dịch lẫn các external factor như tình hình kinh tế, chính trị; variance của dữ liệu thay đổi liên tục; tính phi chu kỳ của dữ liệu thể hiện rất mạng.
+Trong khi các thách thức này trở nên rất đặc thù cho dữ liệu và yêu cầu một phương pháp được thiết kế riêng, các thuật toán hiện tại hầu như không tập trung giải quyết các thách thức này mà chỉ hướng đến time-series nói chung.
+Thật vậy, các phương pháp học sâu như \verb|LSTM|, \verb|CNN|, \verb|Transformer| được thiết kế để phân tích dữ liệu dạng chuỗi nói chung.
+Đối với dạng dữ liệu đặc thù như aperiodic time-series data, chúng trở nên kém hiệu quả do hạn chế trong việc lưu trữ các ràng buộc thời gian.
+Các phương pháp dựa trên phân rã tần số như \verb|NHITS| và \verb|bla_bla| hướng đến phân rã tín hiệu đầu vào thành các dải tần nhưng tính chu kỳ hoàn toàn không tồn tại trong aperiodic time-series data.
+
+Problems on aperiodic time-series data pose a great challenge to current machine learning models because this data depends on both transaction history and external factors such as economic and political situations; the variance of the data changes continuously; the aperiodicity of the data is very strong.
+While these challenges are very data-specific and require a tailored approach, most of the current algorithms do not focused on solving these challenges but are only aimed at time-series in general.
+Indeed, deep learning methods such as \verb|LSTM|, \verb|CNN|, \verb|Transformer| are designed to analyze general serial data.
+For special data types such as aperiodic time-series data, they become less effective due to the limitation in storing temporal constraints.
+Frequency decomposition-based methods such as \verb|NHITS| and \verb|bla_bla| aims to decompose the input signal into frequency bands but periodicity does not exist at all in aperiodic time-series data.
+
+% Ý thức được các thách thức nêu trên cùng với việc tận dụng khả năng rút trích ràng buộc thời gian của \verb|BiLSTM| và khả năng tổng hợp tham số hiệu quả của \verb|MAML|, chúng tôi đề xuất thuật toán \verb|Temporal-ML| với khả năng rút trích thời gian đặc trưng ẩn cũng như tổng hợp thông tin từ dữ liệu đa nguồn.
+Dựa trên thực nghiệm, trong quá trình giải quyết bài toán dự đoán xu hướng tiếp theo, phương pháp đề xuất đạt hiệu quả cao trên hai tập dữ liệu phi chu kỳ (\verb|USD/JPY| và \verb|multi-fx|) và hiệu quả tương đương trên các tập dữ liệu có chu kỳ (\verb|ETT-m2| và \verb|WTH|) so với \verb|NHITS| (SOTA model in AAAI 2023).
+Ngoài ra, bằng việc thực hiện ablation study, chúng tôi chứng minh được sự hiệu quả của từng component trong \verb|Temporal-ML|.
+Cụ thể, khả năng rút trích hiệu quả các ràng buộc thời gian đến từ \verb|BiLSTM| và khả năng tổng hợp thông tin đa nguồn đến từ \verb|MAML|.
+Theo đó, thuật toán của chúng tôi đã giải quyết hiệu quả vấn đề dữ liệu phụ thuộc vào external factor, vấn đề variance bất định và vấn đề phi chu kỳ của dữ liệu.
+
+Keeping in mind the above challenges, by utilizing the time constraint extraction capability of \verb|BiLSTM| and the efficient parameter synthesis capability of \verb|MAML|, we propose the \verb|Temporal-ML| algorithm with the ability to extract hidden time features as well as synthesize information from multi-source data.
+Based on experiments, in the process of solving the next trend prediction problem, the proposed method achieves high efficiency on two aperiodic data sets (\verb|USD/JPY| and \verb|multi-fx|) and equivalent efficiency on periodic data sets (\verb|ETT-m2| and \verb|WTH|) compared to \verb|NHITS| (SOTA model in AAAI 2023).
+In addition, by performing an ablation study, we demonstrate the effectiveness of each component in \verb|Temporal-ML|.
+Specifically, the ability to efficiently extract temporal constraints comes from \verb|BiLSTM| and the ability to synthesize multi-source information comes from \verb|MAML|.
+Accordingly, our algorithm effectively solves the problem of data dependence on external factors, the problem of unstable variance, and the problem of data aperiodicity.
 
 \section{Future works}
+
+In the future, we propose two main directions of development related to the architecture and personalization of the model
+
+% \textbf{Model architecture}.
+Phương pháp của chúng tôi được phát triển dưới dạng component với hai components chính hoạt động song song: Temporal feature extraction and Models' parameter synthesis.
+Điều này cung chấp cho phương pháp của chúng tôi một khả năng nâng cấp linh hoạt.
+Thí nghiệm của chúng tôi chỉ minh họa một trường hợp điển hình trong tổng hợp hiệu quả các mô hình.
+Bằng việc sử dụng các thuật toán ML khác nhau như \verb|Reptile|, \verb|Meta-SGD|, \verb|iMAML|, hoàn toàn có thể tạo ra mô hình mới với độ chính xác cao hơn.
+
+\textbf{Model architecture}.
+Our method is developed as a component with two main components working in parallel: Temporal feature extraction and Models' parameter synthesis.
+This provides our method with a flexible upgradability.
+Our experiment only illustrates a typical case of efficient model synthesis.
+By using different ML algorithms such as \verb|Reptile|, \verb|Meta-SGD|, \verb|iMAML|, it is possible to generate new models with higher accuracy.
+
+% \textbf{Long-horizon problem}.
+Hoàn toàn có thể mở rộng phương pháp này để giải các bài toán về long-horizon prediction.
+Thật vậy, bằng việc thay đổi đầu ra và độ lỗi của mô hình, có thể tiến hành giải các bài toán này.
+Tuy nhiên, kiến trúc của các sub-model cần được nghiên cứu lại để phù hợp hơn với bài toán mới.
+
+\textbf{Long-horizon problem}.
+It is possible to extend this method to solve long-horizon prediction problems.
+Indeed, by changing the output and error of the model, it is possible to solve these problems.
+However, the architecture of the sub-models needs to be re-examined to better suit the new problem.

### 2025-01-30 doc/thesis/chapters/1.intro.tex
contributions undone

@@
 
-Time-series data is gaining popularity alongside numerical, categorical, and text data not only in terms of data sources (like finance, energy, and transportation), but also in terms of methods for analyzing and forecasting them (like frequency decomposition and methods based on historical data memorization).
+Time-series data is gaining popularity alongside numerical, categorical, and text data not only in terms of data sources (e.g., finance, energy, and transportation), but also in terms of methods for analyzing and forecasting them (e.g., frequency decomposition and methods based on historical data memorization).
 This makes sense as planning, organizing, and developing business strategies are greatly aided by the analysis and prediction of time-series data.
@@
 
-The two main approaches used in analyzing and predicting time-series data are fundamental analysis and technical analysis \cite{ayitey2023forex}.
-While fundamental analysis concentrates on examining external factors that are hard to capture from past price changes, such as the economic strategies and policies of nations and businesses, technical analysis relies entirely on historical price changes to forecast future trends
+Two main approaches used in analyzing and predicting time-series data are fundamental analysis and technical analysis \cite{ayitey2023forex}.
+While fundamental analysis concentrates on examining external factors that are hard to capture from past price changes, such as the economic strategies and policies of nations and businesses, technical analysis relies entirely on historical price changes to forecast future trends.
 
@@
 
-It is challenging to achieve efficient automation of basic analytical methods due to the unstructured nature of news data.
-As a result, academics frequently concentrate on creating techniques for technical analysis.
+It is challenging to achieve efficient automation of basic analytical methods due to the unstructured of news data.
+As a result, academics usually concentrate on creating techniques for technical analysis.
 The forecasting process can now be intelligently automated with the help of machine learning and deep learning techniques, such as Auto-regressive model (1951) \cite{moran1951hypothesis}, Moving average model (2000) \cite{rosenblatt2000gaussian}, Long short-term memory model - \verb|LSTM| (1997) \cite{hochreiter1997long}, and Transformers (2017) \cite{vaswani2017attention}.
@@
 
-The aforementioned techniques, together with their variations, are mostly applicable to periodic time-series data, such as traffic, weather, etc.
+The aforementioned techniques, are mostly applicable to periodic time-series data, such as traffic, weather, etc.
 Over time, this kind of data clearly shows a seasonal or cyclical character.
-As a result, analysis and prediction may be done quickly and accurately.
-This is especially true for existing techniques that use deep neural networks to try to decompose periods \cite{liu2022scinet,chen2021autoformer,zhou2022fedformer}.
+Consequently, analysis and prediction can be performed quickly and accurately.
+This is especially true for existing techniques that use deep neural networks to decompose periods \cite{liu2022scinet,chen2021autoformer,zhou2022fedformer}.
 Nevertheless, \textbf{aperiodic time-series data} (e.g., stock prices, foreign exchange rates, etc.) has been the subject of very few investigations.
@@
 
-As can be seen in figure \ref{fig:ill_aperiodic}, the primary distinction between aperiodic and periodic time-series data is their lack of a distinct periodicity.
-In fact, because of seasonal fluctuations in electricity demand, the \textit{Temperature oil} (target variable) attribute of the Electricity Transformer Temperature dataset (\verb|ETT-m2|) \cite{zhou2021informer} shows a very noticeable periodicity over time in figure \ref{fig:periodic}.
+As can be seen in figure \ref{fig:ill_aperiodic}, the primary distinction between aperiodic and periodic time-series data is the lack of a distinct periodicity.
+In fact, because of seasonal fluctuations in electricity demand, the \textit{Temperature oil} (target variable) attribute of the Electricity Transformer Temperature dataset (\verb|ETT-m2|) \cite{zhou2021informer} shows a noticeable periodicity over time in figure \ref{fig:periodic}.
 In the meanwhile, a number of external factors influence the \textit{Close price} between the US dollar and the Japanese yen (figure \ref{fig:aperiodic}).
 This exchange rate can be significantly impacted by recent economic news.
-As a result, there is no periodicity in this exchange rate.
+Thus, there is no periodicity in this exchange rate.
 
@@
 Aperiodic time-series data's primary feature is that it is heavily impacted by external factors, including production strategy and the political and economic policies of nations or businesses.
-This has resulted in notable modifications to the data's characteristics, which are also referred to as concept drift scenarios \cite{liu2025onsitnet}.
-Mathematically, this results in aperiodic time-series data with a continuous and incredibly wide variance.
+This results in notable modifications to the data's characteristics, which are also referred to as concept drift scenarios \cite{liu2025onsitnet}.
+Mathematically, this results in aperiodic time-series data with a continuous and unstable variance.
 Additionally, because the data no longer exhibits periodicity, it becomes extremely challenging to extract characteristics from it using a sliding window.
@@
 
-In conclusion, there are three primary obstacles to overcome when dealing with aperiodic time-series data: (1) External influences including politics, economy, and society have a significant impact on the data; (2) Data has a massive and non-stationary variance; and (3) Data's strong non-periodicity makes feature extraction challenging.
+In conclusion, there are three primary obstacles in dealing with aperiodic time-series data: (1) External influences including politics, economy, and society have a significant impact on the data; (2) Data has a massive and non-stationary variance; and (3) Data's strong non-periodicity makes feature extraction challenging.
 \textbf{In this study, we focus on solving the three challenges mentioned above in the problem of trend prediction (upward or downward) on aperiodic time-series data.}
@@
 \section{Motivation}
-% cần paraphrase lại
+\label{sec:moti}
 
@@
 
-In three aforementioned challenges, the first challenge is the most difficult to solve based on technical analysis methods because political and social information is difficult to extract through historical data of a data set.
-However, according to the efficient market hypothesis \cite{fama1970efficient}, historical data itself clearly reflects information about market fluctuations.
-That means, \textbf{by carefully examining the transaction history, one can completely grasp market fluctuations and forecast the future} without using political and social information.
-In addition, \cite{overreactioncontrarian, mech1993portfolio} studies also point out the dependence between the financial indicators of a certain company and the indicators of other companies.
+The first of the three issues listed above is the most challenging to resolve using technical analysis techniques as it is challenging to extract political and social information from historical data in a data collection.
+However, according to the efficient market hypothesis \cite{fama1970efficient}, historical data itself clearly reflects information about market movements.
+This implies that, \textbf{by carefully examining the transaction history, one can completely grasp market fluctuations and forecast the future} without aggregating political and social information.
+In addition, studies \cite{overreactioncontrarian, mech1993portfolio} indicate the dependence between the financial indicators of a certain company and the indicators of other companies.
 In other words, \textbf{integrating analysis of multiple sources of information within the same domain can yield valuable analytical insights for the indicator of interest.}
@@
 
-When faced with the second challenge, ensemble learning is often used to mitigate the effects of variance \cite{ali2020complete, zafeiriou2020intraday, sadeghi2021combined}.
-Basically, ensemble learning works by splitting the data into multiple parts and assuming a fixed level of variance across these parts, then simulating that variation.
-This approach is reasonable because it is difficult for a single model to capture all the variability in the data over a long period of time.
+Ensemble learning is frequently utilized to mitigate the impacts of variation while dealing with the second challenge \cite{ali2020complete, zafeiriou2020intraday, sadeghi2021combined}.
+This type of learning involves dividing the data into several parts, assuming a certain amount of variation in each part, and then learn that variation.
+This approach is reasonable because it is challenging for a single model to capture all of the variability in the data over an extended period of time.
 \textbf{By analyzing and synthesizing information from multiple sub-models, ensemble learning provides a multidimensional view of the data, which helps the overall model adapt well to strong variance changes.}
@@
 
+% cần paraphrase lại
 Up to now, many models have been proposed to be able to extract features on data in general and time-series data in particular.

### 2025-01-30 doc/thesis/chapters/5.results.tex

--- 2024-12-31 doc/thesis/chapters/5.results.tex
@@
 
-% vẽ, mô tả cái bảng
-% bẻ nhỏ từng attribute trong fx dataset --> thành 4 bảng, 1 bảng tổng hợp là 5
-% bảng tổng hợp vẽ ở đây
-% chắc phải viết thêm cả CNN
+% vẽ, mô tả cái bảng tổng hợp
+% Kết quả chính của công trình này được thể hiện trong bảng \ref{tab:mine_nhits}.
+Theo đó, phương pháp đề xuất của chúng tôi đạt hiệu quả vượt trội trên tất cả các metrics trên cả hai tập dữ liệu phi chu kỳ (\verb|USD/JPY|, \verb|multi-fx|) so với \verb|NHITS|.
+Trên các tập dữ liệu có chu kỳ (\verb|ETT-m2|, \verb|WTH|), mô hình của chúng tôi đạt hiệu quả gần như tương đương hoặc có phần tốt hơn so với baseline model.
 
-\section{Feature extraction ability}
+The main results of this work are shown in table \ref{tab:mine_nhits}.
+Accordingly, our proposed method outperforms \verb|NHITS| on all metrics on both aperiodic datasets (\verb|USD/JPY|, \verb|multi-fx|).
+On periodic datasets (\verb|ETT-m2|, \verb|WTH|), our model performs almost the same or better than the baseline model in some cases.
 
-% 4 bảng phụ vẽ ở đây
-% kiểu: sử dụng mô hình để dự đoán từng attribute, ta có bảng kq sau
+\begin{table}[h]
+    \centering
+    \cprotect\caption{Classification results (\%) of \verb|Temporal-ML| and \verb|NHITS|. Best results per metrics are boldfaced. (\mbox{*}): Our method.}
+    \label{tab:mine_nhits}
+    \resizebox{\columnwidth}{!}{\begin{tabular}{c|c|cccc}
+    \toprule
+    \multicolumn{1}{c}{}                &                                & $accuracy$              & $precision$             & $recall$                & $F1-score$                 \\
+    \hline
+    \multirow{2}{*}{\Verb|USD/JPY|}     & \Verb|NHITS|                   & $58.46$                 & $58.24$                 & $57.65$                 & $55.82$            \\
+                                        & \Verb|Temporal-ML|\mbox{*}     & $\mathbf{70.33\pm1.69}$ & $\mathbf{70.73\pm1.82}$ & $\mathbf{70.26\pm1.93}$ & $\mathbf{69.28\pm2.50}$    \\
+                                        % & Ours(\Verb|BiLSTM+CNN|)  & $68.75\pm1.78$          & $69.15\pm1.55$          & $68.73\pm1.77$          & $69.14\pm2.17$             \\
+    \hline
+    \multirow{2}{*}{\Verb|multi-fx|}    & \Verb|NHITS|                   & $53.51\pm5.02$          & $53.90\pm7.68$          & $53.64\pm4.05$          & $50.20\pm7.35$              \\
+                                        & \Verb|Temporal-ML|\mbox{*}     & $\mathbf{66.26\pm7.45}$ & $\mathbf{67.08\pm8.97}$ & $\mathbf{65.76\pm7.79}$ & $\mathbf{64.06\pm10.00}$   \\
+                                        % & Ours(\Verb|BiLSTM+CNN|)  & $66.08\pm8.19$          & $66.56\pm9.23$          & $65.41\pm8.66$          & $63.78\pm10.83$            \\
+    \hline
+    \multirow{2}{*}{\Verb|ETT-m2|}      & \Verb|NHITS|                   & $\mathbf{71.88}$        & $\mathbf{66.86}$        & $\mathbf{62.49}$        & $\mathbf{62.69}$   \\
+                                        & \Verb|Temporal-ML|\mbox{*}     & $71.14\pm9.13$          & $62.21\pm7.51$          & $58.75\pm5.30$          & $57.61\pm6.74$             \\
+                                        % & Ours(\Verb|BiLSTM+CNN|)  & $71.40\pm9.10$          & $63.25\pm8.82$          & $60.22\pm8.42$          & $59.43\pm9.24$             \\
+    \hline
+    \multirow{2}{*}{\Verb|WTH|}         & \Verb|NHITS|                   & $74.18$                 & $68.05$                 & $65.04$                 & $65.40$             \\
+                                        & \Verb|Temporal-ML|\mbox{*}     & $\mathbf{74.97\pm2.63}$ & $\mathbf{69.13\pm3.44}$ & $\mathbf{65.98\pm3.96}$ & $\mathbf{66.00\pm3.80}$    \\
+                                        % & Ours(\Verb|BiLSTM+CNN|)  & $74.41\pm2.71$          & $68.17\pm4.60$          & $65.50\pm4.82$          & $65.63\pm4.47$             \\
+    \bottomrule
+    \end{tabular}}
+\end{table}
+
+% \begin{table}[h]
+%     \centering
+%     \cprotect\caption{Classification results (\%) of \verb|Temporal-ML| and \verb|NHITS|. Best results per metrics are boldfaced. (\mbox{*}): Our method.}
+%     \label{tab:mine_nhits}
+%     \resizebox{\columnwidth}{!}{\begin{tabular}{c|c|cccc}
+%     \toprule
+%     \multicolumn{1}{c}{}                &                                & $accuracy$              & $precision$             & $recall$                & $F1-score$                 \\
+%     \hline
+%     \multirow{2}{*}{\Verb|USD/JPY|}     & \Verb|NHITS|                   & $58.46\pm9.60$          & $58.24\pm9.89$          & $57.65\pm10.08$         & $55.82\pm11.22$            \\
+%                                         & \Verb|Temporal-ML|\mbox{*}     & $\mathbf{70.33\pm1.69}$ & $\mathbf{70.73\pm1.82}$ & $\mathbf{70.26\pm1.93}$ & $\mathbf{69.28\pm2.50}$    \\
+%                                         % & Ours(\Verb|BiLSTM+CNN|)  & $68.75\pm1.78$          & $69.15\pm1.55$          & $68.73\pm1.77$          & $69.14\pm2.17$             \\
+%     \hline
+%     \multirow{2}{*}{\Verb|multi-fx|}    & \Verb|NHITS|                   & $53.51\pm5.02$          & $53.90\pm7.68$          & $53.64\pm4.05$          & $50.20\pm7.35$              \\
+%                                         & \Verb|Temporal-ML|\mbox{*}     & $\mathbf{66.26\pm7.45}$ & $\mathbf{67.08\pm8.97}$ & $\mathbf{65.76\pm7.79}$ & $\mathbf{64.06\pm10.00}$   \\
+%                                         % & Ours(\Verb|BiLSTM+CNN|)  & $66.08\pm8.19$          & $66.56\pm9.23$          & $65.41\pm8.66$          & $63.78\pm10.83$            \\
+%     \hline
+%     \multirow{2}{*}{\Verb|ETT-m2|}      & \Verb|NHITS|                   & $\mathbf{71.88\pm14.06}$& $\mathbf{66.86\pm10.21}$& $\mathbf{62.49\pm10.40}$& $\mathbf{62.69\pm10.72}$   \\
+%                                         & \Verb|Temporal-ML|\mbox{*}     & $71.14\pm9.13$          & $62.21\pm7.51$          & $58.75\pm5.30$          & $57.61\pm6.74$             \\
+%                                         % & Ours(\Verb|BiLSTM+CNN|)  & $71.40\pm9.10$          & $63.25\pm8.82$          & $60.22\pm8.42$          & $59.43\pm9.24$             \\
+%     \hline
+%     \multirow{2}{*}{\Verb|WTH|}         & \Verb|NHITS|                   & $74.18\pm8.11$          & $68.05\pm6.65$          & $65.04\pm7.67$          & $65.40\pm7.97$             \\
+%                                         & \Verb|Temporal-ML|\mbox{*}     & $\mathbf{74.97\pm2.63}$ & $\mathbf{69.13\pm3.44}$ & $\mathbf{65.98\pm3.96}$ & $\mathbf{66.00\pm3.80}$    \\
+%                                         % & Ours(\Verb|BiLSTM+CNN|)  & $74.41\pm2.71$          & $68.17\pm4.60$          & $65.50\pm4.82$          & $65.63\pm4.47$             \\
+%     \bottomrule
+%     \end{tabular}}
+% \end{table}
+
+% Cụ thể, trên hai tập dữ liệu phi chu kỳ \verb|USD/JPY| và \verb|multi-fx|, \verb|Temporal-ML| cải thiện từ 12\% đến 14\% trên các metrics so với baseline model.
+Mặt khác, đối với dữ liệu có chu kỳ, độ chính xác của \verb|Temporal-ML| thấp hơn chưa đến 1\%, các metrics còn lại thấp hơn từ 4-5\% so với \verb|NHITS| trên dữ liệu \verb|ETT-m2|.
+Đối với dữ liệu có tính chu kỳ mạnh như \verb|WTH|, thuật toán của chúng tôi đạt hiệu suất tốt hơn \verb|NHITS| khoảng 1\%.
+Ngoài ra, \verb|Temporal-ML| cũng đạt được mức phân tán trên các metrics thấp hơn \verb|NHITS| từ 1-3\% khi hoạt động trong ngữ cảnh dữ liệu đa nguồn.
+Quan sát tổng quan, sự chênh lệch của các metrics của cùng một phương pháp là không quá lớn nên không tồn tại vấn đề dữ liệu bị lệch và có thể đánh giá mô hình bằng cách chỉ dựa trên độ chính xác.
+
+Specifically, on the two aperiodic datasets \verb|USD/JPY| and \verb|multi-fx|, \verb|Temporal-ML| improves by 12\% to 14\% on metrics compared to the baseline model.
+On the other hand, for periodic data, the accuracy of \verb|Temporal-ML| is less than 1\% lower, and the remaining metrics are 4-5\% lower than \verb|NHITS| on \verb|ETT-m2| data.
+For strongly cyclical data such as \verb|WTH|, our algorithm outperforms \verb|NHITS| by about 1\%.
+In addition, \verb|Temporal-ML| also achieves a 1-3\% lower dispersion on metrics than \verb|NHITS| when operating in a multi-source data context.
+Overall, the difference between metrics of the same method is not too large, so there is no problem of data bias and the model can be evaluated based on accuracy alone.
+
+% Trong bảng \ref{tab:mine_nhits}, kết quả được tổng hợp dưới dạng trung bình cộng của tất cả các đặc trưng nên có tính đại diện cao.
+Tuy nhiên, các kết quả này thiếu đi sự chi tiết trong đánh giá.
+Do đó, chúng tôi tiến hành phân tích kết quả của các thuật toán trên từng attribute của aperiodic datasets trong bảng \ref{tab:mine_nhits_att}.
+Theo đó, khi \verb|NHITS| dự đoán các thuộc tính \textit{hight, low, close price} trên cả hai tập dữ liệu, kết quả đều chỉ đạt ở mức rất gần mô hình dự đoán ngẫu nhiên.
+Điều này cho thấy việc dự đoán trên aperiodic data thực sự rất khó khăn.
+Tuy nhiên, trên cả hai tập dữ liệu, độ chính xác của \verb|NHITS| thường đạt giá trị lớn nhất khi dự đoán \textit{open price} và lớn hơn đáng kể so với kết quả dự đoán các attribute còn lại.
+Do đó, khi lấy trung bình cộng, kết quả của \verb|NHITS| có vẻ khá cao.
+
+In table \ref{tab:mine_nhits}, the results are aggregated as the average of all features, so they are highly representative.
+However, these results lack detail in the evaluation.
+Therefore, we analyze the results of the algorithms on each attribute of the aperiodic datasets in table \ref{tab:mine_nhits_att}.
+Accordingly, when \verb|NHITS| predicts the attributes \textit{hight, low, close price} on both datasets, the results are only very close to the random prediction model.
+This shows that prediction on aperiodic data is really difficult.
+However, on both datasets, the accuracy of \verb|NHITS| often reaches the largest value when predicting \textit{open price} and is significantly larger than the prediction results of the remaining attributes.
+Therefore, when taking the average, the results of \verb|NHITS| seem to be quite high.
+
+\begin{table}[h]
+    \centering
+    \cprotect\caption{Accuracy (\%) of \verb|Temporal-ML| and \verb|NHITS| on each attribute of \verb|USD/JPY| and \verb|multi-fx|. Best results per metrics are boldfaced. (\mbox{*}): Our method.}
+    \label{tab:mine_nhits_att}
+    \resizebox{\columnwidth}{!}{\begin{tabular}{c|c|cccc}
+    \toprule
+    \multicolumn{1}{c}{}                &                                & $Open$                  & $High$                  & $Low$                   & $Close$                     \\
+    \hline
+    \multirow{3}{*}{\Verb|USD/JPY|}     & Random                         & $49.91$                 & $49.91$                 & $50.33$                 & $50.66$                     \\
+                                        & \Verb|NHITS|                   & $75.03$                 & $53.74$                 & $51.57$                 & $53.50$                     \\
+                                        & \Verb|Temporal-ML|\mbox{*}     & $\mathbf{93.06\pm1.36}$ & $\mathbf{68.87\pm1.53}$ & $\mathbf{51.01\pm1.58}$ & $\mathbf{68.37\pm2.16}$     \\
+    \hline
+    \multirow{3}{*}{\Verb|multi-fx|}    & Random                         & $49.92\pm2.19$          & $49.93\pm2.22$          & $49.94\pm2.17$          & $49.85\pm2.13$              \\
+                                        & \Verb|NHITS|                   & $58.36\pm7.13$          & $50.66\pm4.28$          & $51.17\pm3.21$          & $53.84\pm4.60$              \\
+                                        & \Verb|Temporal-ML|\mbox{*}     & $\mathbf{78.13\pm6.33}$ & $\mathbf{62.74\pm3.30}$ & $\mathbf{58.28\pm3.86}$ & $\mathbf{65.88\pm12.49}$    \\
+    \bottomrule
+    \end{tabular}}
+\end{table}
+
+% Mặt khác, \verb|Temporal-ML| chỉ thất bại trong việc dự đoán \textit{low price} của \verb|USD/JPY| (độ chính xác tiệm cận mức dự đoán ngẫu nhiên).
+Tất cả các thuộc tính còn lại của cả hai tập dữ liệu đều đạt mức hội tụ cao và cao hơn \verb|NHITS| khá nhiều.
+Điều này chứng tỏ khả năng vượt trội của thuật toán đề xuất trên dữ liệu phi chu kỳ.
+
+On the other hand, \verb|Temporal-ML| only fails to predict \textit{low price} of \verb|USD/JPY| (accuracy approaching random prediction level).
+All other features of both datasets achieve high convergence and are significantly higher than \verb|NHITS|.
+This demonstrates the superiority of the proposed algorithm on aperiodic data.
+
+% Để làm rõ khả năng của \verb|Temporal-ML|, dựa trên kết quả thu được từ các ablation experiments (section \ref{sec:ab_ex}) chúng tôi tiến hành phân tích thuật toán đề xuất trên hai khía cạnh: (1) - Temporal feature extraction ability; (2) - Models' parameters aggregation ability.
+
+To clarify the capabilities of \verb|Temporal-ML|, based on the results obtained from ablation experiments (section \ref{sec:ab_ex}), we analyze the proposed algorithm on two aspects: (1) - Temporal feature extraction ability; (2) - Models' parameters aggregation ability.
+
+\section{Temporal feature extraction ability}
+
+% đấm nhits
+% Từ bảng \ref{tab:mine_nhits} và \ref{tab:mine_nhits_att}, không khó để nhận ra khả năng rút trích temporal feature của \verb|Temporal-ML| tốt hơn nhiều so với \verb|NHITS|.
+Thật vậy, quá trình rút trích và tổng hợp đặc trưng của \verb|NHITS| thể hiện rằng phương pháp này đang cố gắng mô phỏng lại quá trình phân giải tần số của Fourier transform.
+Đặc trưng về các giải tần là thông tin tối quan trọng đối với dữ liệu có chu kỳ.
+Do đó, các dự đoán của \verb|NHITS| có thể dễ dàng đạt được độ chính xác cao trên loại dữ liệu này.
+Tuy nhiên, rất khó để có thể phân giải tần số trên dữ liệu phi chu kỳ.
+Hơn nữa, time-series data, dù được lọc qua nhiều stack để rút trích các dải tần từ thấp đến cao, nhưng các kỹ thuật lọc được sử dụng trong mạng là vô cùng đơn giản (chỉ sử dụng \verb|FullyConnected| với hàm kích hoạt phi tuyến và \verb|Pooling| layers).
+Chính những điều này đã khiến cho \verb|NHITS| gặp trở ngại lớn trên \verb|multi-fx|, \verb|USD/JPY| datasets và các dữ liệu phi chu kỳ nói chung.
+
+From the tables \ref{tab:mine_nhits} and \ref{tab:mine_nhits_att}, it is not difficult to see that the temporal feature extraction ability of \verb|Temporal-ML| is much better than that of \verb|NHITS|.
+Indeed, the feature extraction and synthesis process of \verb|NHITS| shows that this method is trying to simulate the frequency resolution process of Fourier transform.
+The frequency band features are the most important information for periodic data.
+Therefore, the predictions of \verb|NHITS| can easily achieve high accuracy on this type of data.
+However, it is very difficult to achieve frequency resolution on aperiodic data.
+Furthermore, the time-series data, although filtered through multiple stacks to extract low-to-high frequency bands, the filtering techniques used in the network are extremely simple (using only \verb|FullyConnected| with nonlinear activation functions and \verb|Pooling| layers).
+This is what makes \verb|NHITS| a major obstacle on \verb|multi-fx|, \verb|USD/JPY| datasets and aperiodic data in general.
+
+% Mặt khác, phương pháp đề xuất của chúng tôi sử dụng các mạng deep learning để rút trích đặc trưng.
+\verb|BiLSTM| và \verb|RNN| nói chung đã chúng tỏ được khả năng của chúng qua nhiều bài toán có liên quan đến các ràng buộc thời gian, rằng chúng có thể dễ dàng thu được các deep hidden feature.
+Do đó, trên cả periodic data lẫn aperiodic data, phương pháp của chúng tôi đều đạt được kết quả xấp xỉ hoặc cao hơn \verb|NHITS|.
+
+On the other hand, our proposed method uses deep neural networks for feature extraction.
+\verb|BiLSTM| and \verb|RNN| have generally demonstrated their capabilities in many problems involving temporal constraints, that they can easily obtain deep hidden features.
+Therefore, on both periodic and aperiodic data, our method achieves results approximately equal to or better than \verb|NHITS|.
+
 % phân tích bảng kết quả
+% Để nắm bắt rõ hơn khả năng của \verb|BiLSTM|, chúng tôi thực hiện ablation study trên component này của \verb|Temporal-ML| như đã đề cập trong subsection \ref{subsec:ablation_lstm} và trình bày kết quả trong bảng \ref{tab:ab_lstm}.
+Theo đó, khi \verb|BiLSTM| được thay thế bởi các component khác (\verb|BiLSTM+CNN|, \verb|Transformer|) để phục vụ quá trình rút trích đặc trưng thời gian, kết quả  trên cả bốn metrics của hầu hết các tập dữ liệu đều thấp hơn kết quả thu được từ \verb|Temporal-ML|.
+Cụ thể, trên hai tập dữ liệu phi chu kỳ (\verb|USD/JPY| and \verb|multi-fx|), các thuật toán sử dụng \verb|BiLSTM+CNN| và \verb|Transformer| để rút trích đặc trưng cho khả năng dự đoán thấp hơn từ 2-10\% so với \verb|Temporal-ML|.
+Trên dữ liệu có chu kỳ, \verb|MAML(Transformer)| vẫn cho kết quả thấp hơn nhiều so với \verb|Temporal-ML|.
+Tuy nhiên \verb|MAML(BiLSTM+CNN)| cho kết quả gần như tiệm cận với thuật toán đề xuất.
+Trên cả bốn datasets, độ lệch chuẩn của các metrics của thuật toán đề xuất nhìn chung nhỏ hơn từ 1-5\% so với các thuật toán còn lại.
+
+To better understand the capabilities of \verb|BiLSTM|, we performed an ablation study on this component of \verb|Temporal-ML| as mentioned in the subsection \ref{subsec:ablation_lstm} and presented the results in table \ref{tab:ab_lstm}.
+Accordingly, when \verb|BiLSTM| is replaced by other components (\verb|BiLSTM+CNN|, \verb|Transformer|) for temporal feature extraction, the results on all four metrics of most datasets are lower than those obtained by \verb|Temporal-ML|.
+Specifically, on two aperiodic datasets (\verb|USD/JPY| and \verb|multi-fx|), the algorithms using \verb|BiLSTM+CNN| and \verb|Transformer| to extract features give 2-10\% lower in prediction ability than \verb|Temporal-ML|.
+On periodic data, \verb|MAML(Transformer)| still gives much lower results than \verb|Temporal-ML|.
+However, \verb|MAML(BiLSTM+CNN)| gives results that are almost asymptotically close to the proposed algorithm.
+On all four datasets, the standard deviation of the metrics of the proposed algorithm is generally 1-5\% smaller than the remaining algorithms.
+
+\begin{table}[h]
+    \centering
+    \cprotect\caption{Classification results (\%) of \verb|Temporal-ML| and models without \verb|BiLSTM|. Best results per metrics are boldfaced. (\mbox{*}): Our method.}
+    \label{tab:ab_lstm}
+    \resizebox{\columnwidth}{!}{\begin{tabular}{c|c|cccc} 
+    \toprule
+    \multicolumn{1}{c}{}      &                                          & $accuracy$               & $precision$              & $recall$                 & $F1-score$                \\ 
+    \hline
+    \multirow{3}{*}{\Verb|USD/JPY|} & \Verb|Temporal-ML|\mbox{*}         & $\mathbf{70.33\pm1.69}$  & $\mathbf{70.73\pm1.82}$  & $\mathbf{70.26\pm1.93}$  & $\mathbf{69.28\pm2.50}$   \\
+                                    & \Verb|MAML|(\Verb|BiLSTM+CNN|)     & $68.75\pm1.78$           & $69.15\pm1.55$           & $68.73\pm1.77$           & $69.14\pm2.17$            \\
+                                    & \Verb|MAML|(\Verb|Transformers|)   & $60.67\pm4.09$           & $59.74\pm6.24$           & $59.32\pm3.79$           & $53.15\pm6.19$            \\ 
+    \hline
+    \multirow{3}{*}{\Verb|multi-fx|}& \Verb|Temporal-ML|\mbox{*}         & $\mathbf{66.26\pm7.45}$  & $\mathbf{67.08\pm8.97}$  & $\mathbf{65.76\pm7.79}$  & $\mathbf{64.06\pm10.00}$  \\
+                                    & \Verb|MAML|(\Verb|BiLSTM+CNN|)     & $66.08\pm8.19$           & $66.56\pm9.23$           & $65.41\pm8.66$           & $63.78\pm10.83$           \\
+                                    & \Verb|MAML|(\Verb|Transformers|)   & $56.03\pm6.08$           & $56.32\pm6.01$           & $55.63\pm5.01$           & $52.33\pm7.08$            \\ 
+    \hline
+    \multirow{3}{*}{\Verb|ETT-m2|}  & \Verb|Temporal-ML|\mbox{*}         & $71.14\pm9.13$           & $62.21\pm7.51$           & $58.75\pm5.30$           & $57.61\pm6.74$            \\
+                                    & \Verb|MAML|(\Verb|BiLSTM+CNN|)     & $\mathbf{71.40\pm9.10}$  & $\mathbf{63.25\pm8.82}$  & $\mathbf{60.22\pm8.42}$  & $\mathbf{59.43\pm9.24}$   \\
+                                    & \Verb|MAML|(\Verb|Transformers|)   & $68.69\pm8.85$           & $58.54\pm4.38$           & $57.85\pm4.71$           & $57.07\pm5.14$            \\ 
+    \hline
+    \multirow{3}{*}{\Verb|WTH|}     & \Verb|Temporal-ML|\mbox{*}         & $\mathbf{74.97\pm2.63}$  & $\mathbf{69.13\pm3.44}$  & $\mathbf{65.98\pm3.96}$  & $\mathbf{66.00\pm3.80}$   \\
+                                    & \Verb|MAML|(\Verb|BiLSTM+CNN|)     & $74.41\pm2.71$           & $68.17\pm4.60$           & $65.50\pm4.82$           & $65.63\pm4.47$            \\
+                                    & \Verb|MAML|(\Verb|Transformers|)   & $72.28\pm3.06$           & $62.35\pm8.73$           & $60.90\pm4.97$           & $58.65\pm4.72$            \\
+    \bottomrule
+\end{tabular}}
+\end{table}
+
 % giải thích lý do tại sao LSTM lại ngon hơn LSTM+CNN
+% Thông qua kết quả này, có thể nhận thấy việc sử dụng \verb|BiLSTM| mang lại kết quả cao và ổn định trong quá trình dự đoán.
+Việc sử dụng các mạng như \verb|CNN|, \verb|Transformer| không những không tạo ra các cải thiện hiệu quả mà còn làm nhiễu quá trình học hoặc giảm hiệu quả của cả mô hình.
+Đối với việc sử dụng \verb|BiLSTM+CNN|, quá trình học bị nhiễu do bản thân \verb|BiLSTM| đã chọn lọc được các ràng buộc thời gian hiệu quả trong quá trình rút trích trong khi \verb|CNN| chỉ rút trích đặc trưng chứ không hề chọn lọc.
+Điều này dẫn đến việc đặc trưng của \verb|BiLSTM+CNN| không ổn định.
+Nếu \verb|CNN| rút trích được đặc trưng tốt, mạng sẽ hoạt động hiệu quả (e.g., metrics on \verb|ETT-m2|).
+Ngược lại, mạng sẽ hoạt động kém hiệu quả hơn so với việc chỉ sử dụng \verb|BiLSTM| (e.g., metrics on \verb|USD/JPY|, \verb|multi-fx|, \verb|WTH|).
 
-\section{Sub-model aggregation ability}
+According to this result, it can be seen that using \verb|BiLSTM| brings high and stable results in the prediction process.
+Using networks such as \verb|CNN|, \verb|Transformer| not only create negative improvements but also disturbs the learning process or reduces the efficiency of the whole model.
+For using \verb|BiLSTM+CNN|, the learning process is disturbed because \verb|BiLSTM| itself has selected effective time dependencies during the extraction process while \verb|CNN| only extracts features but does not select.
+This leads to the instability of \verb|BiLSTM+CNN|'s features.
+If \verb|CNN| extracts good features, the network will perform well (e.g., metrics on \verb|ETT-m2|).
+On the contrary, the network will perform worse than using only \verb|BiLSTM| (e.g., metrics on \verb|USD/JPY|, \verb|multi-fx|, \verb|WTH|).
 
-% vẽ về quá trình hội tụ
-% mô tả về quá trình hội tụ của các thuật toán
-% múa
+% giải thích lý do tại sao LSTM lại ngon hơn Transformer
+% Ngoài ra, chúng tôi thực hiện phân tích quá trình hội tụ của \verb|Temporal-ML| và các thuật toán nêu trên trên hai tập dữ liệu phi chu kỳ.
+Kết quả được trình bày trong hình \ref{fig:ablation_att_usd} và \ref{fig:ablation_att_multi}.
+Nhìn chung, trong quá trình dự đoán các attribute của \verb|USD/JPY| và \verb|multi-fx|, \verb|Temporal-ML| đều cho độ chính xác cao và quá trình hội tụ nhanh.
+Riêng dự đoán \textit{close price} trên \verb|multi-fx|, \verb|MAML(BiLSTM+CNN)| cho độ chính xác lẫn quá trình hội tụ tốt hơn thuật toán đề xuất.
+Điều đáng ngạc nhiên là \verb|MAML(Transformer)| chỉ hội tụ khi dự đoán \textit{open price} của cả hai tập dữ liệu foreign exchange.
+Chúng tôi cho rằng, để có thể sử dụng một mô hình phức tạp như \verb|Transformer|, cần sử dụng một lượng lớn dữ liệu với số lượng inner epochs lớn.
+Điều này khiến cho \verb|Transformer| không phải là một lựa chọn tốt khi kết hợp với các thuật toán ML.
+
+In addition, we analyze the convergence of \verb|Temporal-ML| and the above algorithms on two aperiodic datasets.
+The results are presented in figures \ref{fig:ablation_att_usd} and \ref{fig:ablation_att_multi}.
+In general, in the process of predicting attributes of \verb|USD/JPY| and \verb|multi-fx|, \verb|Temporal-ML| both give high accuracy and fast convergence.
+Particularly for predicting \textit{close price} on \verb|multi-fx|, \verb|MAML(BiLSTM+CNN)| gives better accuracy and convergence than the proposed algorithm.
+Surprisingly, \verb|MAML(Transformer)| only converges when predicting \textit{open price} of both foreign exchange datasets.
+We believe that in order to use a complex model like \verb|Transformer|, a large amount of data with a large number of inner epochs is required.
+This makes \verb|Transformer| not a good choice when combined with ML algorithms.
+
+% Với các ý phân tích nêu trên, chúng tôi kết luận rằng, \verb|BiLSTM| không chỉ là lựa chọn phù hợp trong việc khám phá sự biến thiên của aperiodic time-series data mà còn là thuật toán thích hợp để kết hợp với các thuật toán ML khi làm việc trên time-series data nói chung.
+\textbf{Bằng việc sử dụng \Verb|BiLSTM|, chúng tôi đã rút trích được sự biến động của thị trường cũng như dự đoán tương lai một cách hiệu quả}.
+
+With the above analysis, we conclude that \verb|BiLSTM| is not only a suitable choice for exploring the variability of aperiodic time-series data but also a suitable algorithm to combine with ML algorithms when working on time-series data in general.
+\textbf{By using \Verb|BiLSTM|, we have effectively extracted market fluctuations and predicted the future}.
+
+\begin{sidewaysfigure}[h]
+    \centering
+    \begin{subfigure}[b]{\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{ab_att_usd.pdf}
+        \cprotect\caption{\verb|Temporal-ML| vs. Models without \verb|BiLSTM| on \verb|USD/JPY|.}
+        \label{fig:ablation_att_usd}
+    \end{subfigure}
+    ~
+    \begin{subfigure}[b]{\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{ab_att_multi.pdf}
+        \cprotect\caption{\verb|Temporal-ML| vs. Models without \verb|BiLSTM| on \verb|multi-fx|.}
+        \label{fig:ablation_att_multi}
+    \end{subfigure}
+    ~
+    \begin{subfigure}[b]{\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{ab_ml_multi.pdf}
+        \cprotect\caption{\verb|Temporal-ML| vs. Models without \verb|MAML| on \verb|multi-fx|.}
+        \label{fig:ablation_ml_multi}
+    \end{subfigure}
+    \cprotect\caption{Convergence process of algorithms in ablation study on all attribute of foreign exchange datasets.}
+    \label{fig:ablation}
+\end{sidewaysfigure}
+
+\section{Models' parameters aggregation ability}
+
+% đấm nhits
+% Trong quá trình tổng hợp các đặc trưng để đưa ra dự đoán cuối, \verb|NHITS| chỉ sử dụng các phép cộng trừ đơn giản.
+Điều này đặt trong ngữ cảnh của việc tổng hợp thông tin dựa trên các dải tần phân tách được từ dữ liệu đầu vào là hợp lý.
+Thật vậy, các dữ liệu liên quan đến thời tiết, nhu cầu sử dụng máy bay hoặc điện thể hiện tính chu kỳ rất mạnh vì thời tiết và hành vi tiêu dùng của con người, dù có thể chênh lệch theo thời gian nhưng vẫn thể hiện rất rõ tính mùa vụ (e.g., trời lạnh vào mùa đông, nhu cầu sử dụng máy bay tăng cao vào ngày lễ).
+Do đó, tần suất của loại dữ liệu này là một đặc trưng tốt và dễ dàng rút trích.
+Sau khi phân rã dữ liệu đầu vào thành các dải tần, chỉ cần tổng hợp chúng để có thể dự đoán tương lai.
+Ngoài ra, việc tổng hợp các dải tần bằng phép cộng đơn thuần có thể coi là rất đơn giản so với các thuật toán học sâu hiện tại, vốn làm việc trên không gian đặc trưng ẩn của dữ liệu.
+Do đó, quá trình tổng hợp của \verb|NHITS| trên dữ liệu phi chu kỳ gặp rất nhiều khó khăn vì thông tin đầu vào cho quá trình tổng hợp không tốt và phương pháp tổng hợp không hiệu quả.
+Từ đây, có thể thấy rằng, aperiodic time-series data trở thành điểm yếu chí mạng cho \verb|NHITS| và các phương pháp dựa trên việc phân rã tần số nói chung.
+
+In the process of synthesizing features to make the final prediction, \verb|NHITS| only uses simple addition and subtraction.
+This makes sense in the context of synthesizing information based on the frequency bands extracted from the input data.
+Indeed, data related to weather, demand for airplanes or electricity show very strong periodic characteristics because weather and human consumption behavior, although they may vary over time, still show a very clear seasonality (e.g., cold weather in winter, high demand for airplanes during holidays).
+Therefore, the frequency of this type of data is a good feature and can be easily extracted.
+After decomposing the input data into frequency bands, we can simply synthesize them to be able to predict the future.
+In addition, synthesizing the frequency bands by an addition can be considered very simple compared to current deep learning algorithms, which work on the hidden feature space of the data.
+Therefore, the synthesis process of \verb|NHITS| on aperiodic data encounters many difficulties because the input information for the process is not good and the synthesis method is ineffective.
+Consequently, it can be seen that aperiodic time-series data becomes a fatal weakness for \verb|NHITS| and frequency decomposition-based methods in general.
+
+% nâng bi bản thân
+% Đối diện với sự phức tạp trong dữ liệu phi chu kỳ đòi hỏi một cấu trúc phức tạp để có thể tổng hợp các đặc trưng một cách hợp hiệu quả.
+Như đã đề cập trong phần Motivation (section \ref{sec:moti}), việc phân tích và tổng hợp thông tin từ nhiều nguồn dữ liệu trong cùng một domain có thể cung cấp các thông tin hữu ích trong việc dự đoán các chỉ số quan tâm.
+\verb|Temporal-ML| với khả năng tổng hợp thông tin dựa trên quá trình tối ưu của các thuật toán ML, cho phép mô hình nắm bắt thông tin đa chiều từ các nguồn dữ liệu khác nhau.
+Từ đó đưa ra dự đoán hiệu quả.
+
+Dealing with the complexity of non-periodic data requires a complex structure that can efficiently aggregate features.
+As mentioned in the Motivation section (section \ref{sec:moti}), analyzing and synthesizing information from multiple data sources in the same domain can provide useful information for predicting metrics of interest.
+\verb|Temporal-ML| with the ability to synthesize information based on the optimization process of ML algorithms, allows the model to capture multi-dimensional information from different data sources.
+which helps produce effective predictions.
+
+% phân tích bảng kq
+% Để nắm bắt rõ khả năng của các thuật toán ML trong \verb|Temporal-ML|, chúng tôi loại bỏ component này khỏi thuật toán đề xuất sau đó chạy thực nghiệm để so sánh với kết quả ban đầu (subsection \ref{subsec:ablation_ml}).
+Chúng tôi sử dụng các thuật toán ML nhằm mục đích tổng hợp thông tin đa nguồn.
+Do đó, các so sánh chỉ được thực hiện trên tập dữ liệu \verb|multi-fx| vì chỉ tập dữ liệu này chứa nhiều tập dữ liệu con từ nhiều nguồn khác nhau.
+Kết quả được tổng hợp trong bảng \ref{tab:ab_ml}.
+Theo đó, với khả năng tổng hợp thông tin của mình, \verb|Temporal-ML| cho kết quả cao hơn từ 2-12\% so với các thuật toán không sử dụng ML trong quá trình huấn luyện.
+Ngoài ra, mức phân tán của các metrics cũng thấp hơn từ 1-7\%, chứng tỏ rằng mức chênh lệch trong quá trình dự đoán các task tốt hơn nhiều so với các thuật toán còn lại.
+
+To evaluate the capabilities of ML algorithms in \verb|Temporal-ML|, we remove this component from the proposed algorithm and then run experiments to compare with the original results (subsection \ref{subsec:ablation_ml}).
+We use ML algorithms for the purpose of synthesizing multi-source information.
+Therefore, comparisons are only performed on the \verb|multi-fx| dataset because this dataset contains many sub-datasets from many different sources, which helps to accurately evaluate the information synthesis capability.
+The results are summarized in table \ref{tab:ab_ml}.
+Accordingly, with the information synthesis capability, \verb|Temporal-ML| achieves 2-12\% higher in all metrics than algorithms that do not use ML in testing process.
+In addition, the dispersion of the metrics is also lower from 1-7\%, proving that the difference in the task prediction process is much better than the remaining algorithms.
+
+% Việc đạt được kết quả tốt hơn trên \verb|multi-fx| của \verb|Temporal-ML| là hoàn toàn dễ hiểu vì tất cả phương pháp còn lại trong bảng \ref{tab:ab_ml} đều không có khả năng tổng hợp thông tin như thuật toán đề xuất mà chỉ dựa hoàn toàn vào dữ liệu quá khứ để dự đoán tương lai.
+Do đó, không chỉ \verb|BiLSTM|, \verb|BiLSTM+CNN| mà ngay cả \verb|Transformer| cũng không thể tận dụng được mối tương quan giữa các data sources.
+
+The better performance of \verb|Temporal-ML| on \verb|multi-fx| is understandable because all the remaining methods in the \ref{tab:ab_ml} table do not have the ability to synthesize information like the proposed algorithm, but rely entirely on past data to predict the future.
+Therefore, not only \verb|BiLSTM|, \verb|BiLSTM+CNN| but also \verb|Transformer| cannot take advantage of the correlation between data sources.
+
+\begin{table}[H]
+    \centering
+    \cprotect\caption{Classification results (\%) of \verb|Temporal-ML| and models without \verb|MAML| on \verb|multi-fx|. Best results per metrics are boldfaced. (\mbox{*}): Our method.}
+    \label{tab:ab_ml}
+    \begin{tabular}{c|cccc} 
+    \toprule
+                                & $accuracy$              & $precision$             & $recall$                & $F1-score$                \\ 
+    \hline
+    \Verb|Temporal-ML|\mbox{*}  & $\mathbf{66.26\pm7.45}$ & $\mathbf{67.08\pm8.97}$ & $\mathbf{65.76\pm7.79}$ & $\mathbf{64.06\pm10.00}$  \\
+    \Verb|BiLSTM|               & $63.88\pm9.00$          & $64.18\pm13.39$         & $63.66\pm9.13$          & $60.07\pm13.09$           \\
+    \Verb|BiLSTM+CNN|           & $62.23\pm8.63$          & $62.89\pm9.76$          & $62.20\pm8.34$          & $59.91\pm10.97$           \\
+    \Verb|Transformers|         & $58.44\pm9.15$          & $56.71\pm15.10$         & $58.22\pm8.84$          & $52.22\pm13.12$           \\
+    \bottomrule
+    \end{tabular}
+\end{table}
+
+% phân tích hội tụ
+% Quá trình hội tụ của \verb|Temporal-ML| và các model without ML được trình bày trong hình \ref{fig:ablation_ml_multi}.
+Trong đó, khi dự đoán \textit{open price}, thuật toán đề xuất cho khả năng hội tụ vượt trội so với các thuật toán còn lại.
+Đối với các thuộc tính \textit{high price} và \textit{close price}, \verb|Temporal-ML| tỏ ra đuối sức trong giai đoạn đầu của quá trình huấn luyện.
+Tuy nhiên, khi về cuối, mô hình đề xuất cho thấy khả năng vượt lên nhanh chóng và tiềm năng cải thiện độ chính xác hơn nữa so với các thuật toán khác, vốn đã hội tụ.
+
+The convergence of \verb|Temporal-ML| and the models without ML is shown in Fig.
+\ref{fig:ablation_ml_multi}.
+In particular, when predicting \textit{open price}, the proposed algorithm shows superior convergence compared to the remaining algorithms.
+For the attributes \textit{high price} and \textit{close price}, \verb|Temporal-ML| shows signs of slowing down in the early stages of training.
+However, towards the end, the proposed model shows the ability to quickly surpass and has the potential to further improve the accuracy compared to the other algorithms, which have already converged.
+
+% Do đó, có thể thấy rằng, \textbf{việc sử dụng các thuật toán optimization-based ML trong việc tổng hợp thông tin là hoàn toàn hợp lý vì đã tận dụng thành công mối tương quan giữa các tập dữ liệu và đạt được độ chính xác cao mà không đòi hỏi quá nhiều dữ liệu như trong ngữ cảnh đơn nguồn}.
+
+Therefore, it can be seen that, \textbf{the use of optimization-based ML algorithms in information synthesis is completely reasonable because it successfully takes advantage of the correlation between data sets and achieves high accuracy without requiring too much data as in the single-source context}.

### 2025-01-30 doc/thesis/pre_info/abstract.tex

@@
 
+% context
+% Phân tích và dự báo trên time-series data từ lâu đã nhận được sự quan tâm rất lớn của cả cộng đồng nghiên cứu và doanh nghiệp bởi sự phổ biến cũng như nguồn lợi lớn mà nó mang lại về mặt kinh tế lẫn học thuật.
+Tuy nhiên, trong khi các nghiên cứu trên periodic time-series data được mở rộng và đạt được nhiều kết quả tích cực, aperiodic time-series data như foreign exchange, stock price chưa được nghiên cứu sâu.
+So với periodic time-series data, dạng dữ liệu này có nhiều tính chất phức tạp hơn, tạo ra những khó khăn riêng và cần được giải quyết bằng các phương pháp được thiết kế riêng.
+
+Analysis and forecasting on time-series data have received great attention from both the research community and businesses due to its popularity and the great benefits it brings in terms of economics and academia.
+However, while research on periodic time-series data has been expanded and achieved many positive results, aperiodic time-series data such as foreign exchange, stock price has not been studied in depth.
+Compared to periodic time-series data, this type of data has more complex properties, creating its own difficulties and needs to be solved by specially designed methods.
+
+% proposed method
+% Công trình này đề xuất \verb|Temporal-ML|, một hướng tiếp cận mới, kết hợp giữa thuật toán Model-Agnostic Meta-Learning (\verb|MAML|) và Bidirectional Long Shot-Term Memory (\verb|BiLSTM|), trong việc giải quyết bài toán dự đoán xu hướng (tăng, giảm) của aperiodic time-series data.
+Mục tiêu của thuật toán là kết hợp khả năng rút trích có chọn lọc các ràng buộc thời gian của \verb|BiLSTM| và khả năng tổng hợp mô hình với tính tổng quát hóa cao của \verb|MAML|.
+\verb|Temporal-ML| theo đó không chỉ rút trích hiệu quả các ràng buộc thời gian mà còn có thể rút trích được mối tương quan giữa các tập dữ liệu khác nhau trong ngữ cảnh dữ liệu đa nguồn.
+
+This paper proposes \verb|Temporal-ML|, a new approach that combines Model-Agnostic Meta-Learning (\verb|MAML|) and Bidirectional Long Shot-Term Memory (\verb|BiLSTM|) to solve the problem of movement prediction (upward, downward) of aperiodic time-series data.
+The goal of the algorithm is to combine the ability to selectively extract temporal dependencies of \verb|BiLSTM| and the ability to synthesize models with high generalization of \verb|MAML|.
+Accordingly, \verb|Temporal-ML| not only efficiently extracts temporal dependencies but also extract correlations between different datasets in the context of multi-source data.
+
+% compare to nhits
+% Trong qua quá trình thực nghiệm, chúng tôi so sánh \verb|Temporal-ML| và \verb|NHITS| (state-of-the-art (SOTA) model in 2023) trên hai tập dữ liệu phi chu kỳ \verb|USD/JPY| (foreign exchange rate giữa US dollar và Japanese yen, được sample theo giờ từ năm 2000 đến năm 2024) và \verb|multi-fx| (foreign exchange rate của 60 cặp tiền tệ giữa 18 quốc gia, được sample theo ngày từ năm 2014 đến 2024).
+Kết quả trên hai tập dữ liệu phi chu kỳ cho thấy sự vượt trội của thuật toán đề xuất so với \verb|NHITS| trên tất cả các metrics phân lớp.
+Trên dữ liệu có chu kỳ, \verb|Temporal-ML| đạt kết quả tương đương mô hình baseline.
+Điều này chứng tỏ khả năng vượt trội của thuật toán đề xuất trên dữ liệu phi chu kỳ cũng như khả năng tương đương mô hình SOTA trên dữ liệu có chu kỳ.
+
+In our experiments, we compare \verb|Temporal-ML| and \verb|NHITS| (state-of-the-art (SOTA) model in 2023) on two aperiodic time-series datasets \verb|USD/JPY| (foreign exchange rate between US dollar and Japanese yen, sampled hourly from 2000 to 2024) and \verb|multi-fx| (foreign exchange rate of 60 currency pairs between 18 countries, sampled daily from 2014 to 2024).
+The results on the two aperiodic datasets show the superiority of the proposed algorithm over \verb|NHITS| on all classification metrics.
+On the periodic data, \verb|Temporal-ML| achieves results equivalent to the baseline model.
+These results demonstrate the superiority of the proposed algorithm on aperiodic data as well as the equivalence of the SOTA model on periodic data.
+
+% ablation study
+% Ngoài ra, công trình này thực hiện ablation study để phân tích sâu sự ảnh hưởng của từng component trong \verb|Temporal-ML|.
+Dựa trên kết quả thu được, khóa luận chứng minh được vai trò của từng component cũng như sự hợp lý trong việc lựa chọn các thuật toán trong quá trình kết hợp.
+
+Additionally, this work conducts an ablation study to deeply analyze the influence of each component in \verb|Temporal-ML|.
+Based on the obtained results, the thesis proves the role of each component as well as the rationality in choosing algorithms in the combination process.
+
+% Tóm lại, công trình này nhấn mạnh tầm quan trọng của việc thiết kế một phương pháp riêng cho aperiodic time-series data.
+Các khám phá trong khóa luận không chỉ giúp tháo gỡ các khó khăn trong quá trình phân tích và dự đoán aperiodic time-series data, mà còn trực tiếp thúc đẩy cộng đồng nghiên cứu trong việc tìm ra các giải pháp hiệu quả trên dạng dữ liệu này.
+Từ đó các nghiên cứu không chỉ hướng đến dự đoán xu hướng mà còn dự đoán được giá trị cụ thể.
+
+In summary, this work emphasizes the importance of designing a specific method for aperiodic time-series data.
+The discoveries in this thesis not only help to solve the difficulties in the process of analyzing and predicting aperiodic time-series data, but also directly promote the research community in finding effective solutions on this type of data.
+Consequently, the research can move from movement prediction to value-specified prediction.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \noindent
-\textbf{Keywords:} Meta-learning, \verb|LSTM|, \verb|CNN|, Aperiodic time-series data, Foreign exchange
+\textbf{Keywords:} Aperiodic time-series data, Meta-learning, \verb|BiLSTM|, \verb|BiLSTM+CNN|, \verb|Transformer|, Foreign exchange

### 2025-01-30 doc/thesis/pre_info/bio.tex

\begin{center}
{\LARGE\bf Biography}
\end{center}
\vspace{20pt}

\begin{table}[h]
\setlength{\tabcolsep}{0.7ex}
\begin{tabular}{lll}
\textbf{Name} && Rohit Sanjay Dahule \\[3ex]
\textbf{Date of Birth} && 23 May 1994 \\[3ex]
\textbf{Educational Attainment} && 2020: M.Tech. in Materials Science \& Engineering \\[0.5ex] 
    && Indian Institute of Technology Gandhinagar, \\[0.5ex]
    && Palaj, Gujarat, India \\[2ex]
    && 2016: B.E. in Mechanical Engineering\\[0.5ex] 
    && Rajiv Gandhi College of Engineering, Research and Technology \\[0.5ex]
    && Chandrapur, Maharashtra, India \\[2ex]

\textbf{Scholarship} && 2021-2025: Monbukagakusho (MEXT) Scholarship \\[2ex]
\textbf{Work Experience} && 2020-2021: Junior Research Fellow \\[0.5ex]
    && Indian Institute of Technology Gandhinagar, \\[0.5ex] 
    && Palaj, Gujarat, India \\[2ex]
\end{tabular}
\end{table}
\vspace{-10pt} 
\noindent
\textbf{Publications}
\begin{enumerate}[label={[\arabic*]}]
    \item \underline{Dahule R.}, Raghav A., Hanindriyo A. T., Hongo K., Maezono R.,
        \& Panda, E. (2021). Surface study of Cu$_2$SnS$_3$ using first‐principles
        density functional theory,  Adv. The. Simu., 4(6), 2000315.
    \item \underline{Dahule R.}, Singh C. C., Hongo K., Maezono R., \& Panda E.
        (2022). Anomalies in the bulk and surface electronic properties of SnS:
        effects of native defects,  J. Mate. Chem. C, 10(14), 5514-5525.
    \item Song P., Durajski A, Hou Z., Ghaffar A, \underline{Dahule R.}, Szczesniak R.,Hongo K.,
        \& Maezono R., (2024) (La,Th)H$_{10}$: Potential high-T$_c$ (242 K) 
        superconductors stabilized thermodynamically below 200 GP. 
        J. Phys. Chem. C 2024, 128, 6, 2656–2665.
    \item Bandi H., Kakarla A., \underline{Dahule R.}, Maezono R., Narsimulu D.,
         Shanthappa R. \& Yu J. S.,
         Nickel vanadate cathode induced in-situ phase transition for improved zinc storage
         by low migration barrier and Zn$^{2+}$/H$^+$ co-insertion mechanism.
         small 2024, 2408568.
    \item \underline{Dahule R.}, Gao, B., Hongo, K., Panda E., Ma, Y. \& Maezono R., 
        Stability and Electronic Properties of SnS/ZnS Interfaces: 
        A First-Principles Investigation, 
        (under review for J. Phys. Chem. C)
    \item \underline{Dahule R.}, Oqumhula K., Maezono, R., \& Hongo, K., 
        Physics-informed data-driven discovery of high thermal conductive polymer crystals. 
        (under review for ACS Appl. Polym. Mater.)
    \item Bandi H., \underline{Dahule R.}, Kakarla A., Maezono R., Narsimulu D., 
        Akram S. W.,  Shanthappa R. \& Yu J. S., 
        New insights into the green synthesis of CoMoO$_{4}$ nanorods decorated with 
        silver nanoparticles for high-performance Li$^+$ storage anode 
        (under review for J. Mate. Chem. A)
\end{enumerate}

### 2025-01-30 execute/ablation.py
trash code, please read the comment

@@ -11,16 +11,20 @@ import keras
 import argparse
 import json
 
+from neuralforecast import NeuralForecast
+from neuralforecast.models import NHITS
+from neuralforecast.losses.pytorch import DistributionLoss
+
 from model.based_model import get_model
 from data.dataloader import DataLoader
 from common.constants import *
 from common.common import compute_metrics
 
 class DataLoaderAblation(DataLoader):
-    def __init__(self, dataset, look_back = None, batch_size = -1, test_size = 0.2, mode = CLF, is_multi = False):
+    def __init__(self, dataset, look_back = None, batch_size = -1, test_size = 0.4, mode = CLF, is_multi = False):
         super().__init__(dataset, look_back, batch_size, test_size, mode, is_multi)
 
-    def get_ablation_data(self, label:int=-1):
+    def get_ablation_data(self, label:int=-1, nhits:bool=False):
         data_dir = self.data_dir if self.is_multi else os.path.join(self.data_dir, 'all')
         data_dict = {}
 
@@ -31,18 +35,29 @@ class DataLoaderAblation(DataLoader):
                 df = pd.read_csv(os.path.join(data_dir, file_name)).to_numpy()[:,1:]
 
                 # normalize data
-                transformed_df, mean_, std_ = self._normalize_data(df=df, label=label, test_size=0.2)
+                transformed_df, mean_, std_ = self._normalize_data(df=df, label=label, test_size=0.4)
 
-                # create dataset with lookback window
-                X, y = self._create_dataset(df=transformed_df, label=label)
+                if nhits:
+                    # create label
+                    y = self._create_label(transformed_df, label)
 
-                # add to data_dict
-                tmp_dict = self.split_support_query(X, y, mean_, std_)
-                num_train = tmp_dict['support_X'].shape[0]
-                # only use 60% data for training
-                tmp_dict['support_X'] = tmp_dict['support_X'][:int(num_train*0.75)]
-                tmp_dict['support_y'] = tmp_dict['support_y'][:int(num_train*0.75)]
-                data_dict[cid] = tmp_dict
+                    # concat transformed_df, y
+                    transformed_df = np.concatenate([transformed_df, y.reshape(-1,1)], axis=1)
+                    transformed_df = pd.DataFrame(transformed_df, columns=self.columns+['y'])
+                    transformed_df['unique_id'] = 0
[…]

### 2025-01-30 model/based_model.py
minor change

@@ -49,18 +49,19 @@ def transformer_encoder(inputs, head_size, num_heads, n_filters, dropout=0):
         key_dim=head_size, num_heads=num_heads, dropout=dropout
     )(inputs, inputs)
     x = layers.Dropout(dropout)(x)
-    x = layers.LayerNormalization(epsilon=1e-6)(x)
+    x = layers.LayerNormalization(epsilon=0.001)(x)
     res = x + inputs
 
     # Feed Forward Part
     x = layers.Conv1D(filters=n_filters, kernel_size=1, activation="relu")(res)
     x = layers.Dropout(dropout)(x)
     x = layers.Conv1D(filters=inputs.shape[-1], kernel_size=1)(x)
-    x = layers.LayerNormalization(epsilon=1e-6)(x)
+    x = layers.LayerNormalization(epsilon=0.001)(x)
     return x + res
 
 def _get_attention(input_layer, head_size=256, num_heads=3, n_filters=4, num_transformer_blocks=2, mlp_units=[64], dropout=0, mlp_dropout=0):
-    x = input_layer
+    # x = input_layer
+    x = layers.Permute((2,1))(input_layer)
     for _ in range(num_transformer_blocks):
         x = transformer_encoder(x, head_size, num_heads, n_filters, dropout)
 

### 2025-01-30 model/baseline_models.py

@@ -81,10 +81,10 @@ if __name__ == '__main__':
         os.mkdir(dataset_dir)
 
     # hyper-param search space
-    pooling_sizes = [[2,2,2], [4,4,4], [8,8,8], [8,4,1], [16,8,1]]
-    freqs = [[168,24,1], [24,12,1], [180,60,1], [40,20,1], [64,8,1]]
+    pooling_sizes = [[8,4,1]]
+    freqs = [[168,24,1]]
     lrs = [0.001]
-    input_sizes = [5,20,30]
+    input_sizes = [30]
     loss_fn = DistributionLoss('Bernoulli') if mode==CLF else MSE() if mode==REG else None
 
     # predict all columns (1 by 1) in the dataset

### 2025-03-03 doc/paper/els-cas-templates/cas-sc-sample.tex
paper + thesis

%% 
%% Copyright 2019-2024 Elsevier Ltd
%% 
%% Version 2.4
%% 
%% This file is part of the 'CAS Bundle'.
%% --------------------------------------
%% 
%% It may be distributed under the conditions of the LaTeX Project Public
%% License, either version 1.2 of this license or (at your option) any
%% later version.  The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.2 or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%% 
%% The list of all files belonging to the 'CAS Bundle' is
%% given in the file `manifest.txt'.
%% 
%% Template article for cas-sc documentclass for 
%% single column output.

%\documentclass[a4paper,fleqn,longmktitle]{cas-sc}
\documentclass[a4paper,fleqn]{cas-sc}

\usepackage[numbers]{natbib}
%\usepackage[authoryear]{natbib}
% \usepackage[authoryear,longnamesfirst]{natbib}
\usepackage{fvextra}
\usepackage{tabularray}
\usepackage{float}

% % remove left footer
% \usepackage{fancyhdr} % Ensure this is included (usually already in Elsevier templates)
% \pagestyle{fancy}     % Use the fancy page style
% \lfoot{}              % Clear the left footer

%%%Author macros
\def\tsc#1{\csdef{#1}{\textsc{\lowercase{#1}}\xspace}}
\tsc{WGM}
\tsc{QE}
\tsc{EP}
% \tscawew{PMS}
\tsc{BEC}
\tsc{DE}
%%%

\begin{document}
\let\WriteBookmarks\relax
\def\floatpagepagefraction{1}
\def\textpagefraction{.001}

\shorttitle{A Meta-learning approach in movement prediction of aperiodic time-series data}
\shortauthors{B.L. Nguyen et~al.}

%\begin{frontmatter}

\title [mode = title]{A Meta-learning approach in movement prediction of aperiodic time-series data}
\tnotemark[1]

\tnotetext[1]{This research did not receive any specific grant from funding agencies in the public, commercial, or not-for-profit sectors.}

% % \tnotetext[2]{The second title footnote which is a longer text matter to fill through the whole text width and overflow into another line in the footnotes area of the first page.}

\author[1]{Bao Long Nguyen}[
    % type=editor,
    % auid=000,bioid=1,
    % prefix=Sir,
    % role=Researcher,
    orcid=0000-0002-6411-8943
]
\cormark[2]
% \fnmark[1]
% \ead{mwklng2309@icloud.com}
\ead{s2210434@jaist.ac.jp}
\credit{Data curation, Methodology, Programming, Writing - Original draft preparation}

\affiliation[1]{
    organization={School of Information Science, Japan Advanced Institute of Science and Technology (JAIST)},
    city={Nomi},
    state={Ishikawa},
    postcode={923-1292}, 
    country={Japan}
}

\author[2]{Kenta Hongo}[
    orcid=0000-0002-2580-0907
]
\ead{hongo@jaist.ac.jp}
\credit{Literature review}
\affiliation[2]{
    organization={Research Center for Advanced Computing Infrastructure, JAIST},
    city={Nomi},
    state={Ishikawa},
    postcode={923-1292}, 
    country={Japan}
}

\author[1]{Ryo Maezono}[
    orcid=0000-0002-5875-971X
]
\ead{rmaezono@jaist.ac.jp}
\credit{Conceptialization}

\author[3]{Bac Le}[
    % type=editor,
    % auid=000,bioid=1,
    % prefix=Sir,
    % role=Researcher,
    orcid=0000-0002-4306-6945
]
\ead{lhbac@fit.hcmus.edu.vn}
\credit{Review paper}
\affiliation[3]{
    organization={Deparment of Computer Science, University of Science, Vietnam National University},
    city={Ho Chi Minh City},
    postcode={700000},
    country={Vietnam}
}

\author[1]{Tom Ichibha}[
    orcid=0000-0002-7455-4968
]
\cormark[1]
\ead{ichibha@icloud.com}
\credit{Supervision}

\cortext[cor1]{Corresponding author}
\cortext[cor2]{Principal corresponding author}

\begin{abstract}
    Predicting aperiodic time-series data (e.g. stock price, foreign exchange, Bitcoin price, etc.) is a difficult task for machine learning models because eternal factors including politics, economy, and society have a significant	impact on the data; The data has a massive and non-stationary variance; Data's strong non-periodicity makes feature extraction challenging. These challenges make the data unstable and non-cyclical. To overcome the above challenges, we proposed \verb|Temporal-ML|, which effectively extracting temporal feature of the data and synthesizing sub-models across multi-source. Experiments on aperiodic data (foreign exchange data of 60 currency pairs over 24 years from 2000 to 2024) show that the proposed method performs well and has 12-14\% higher accuracy compared to the \verb|NHITS| - the state of the art (SOTA) model in 2023 on time-series data, in the task of predicting the trend (upward or downward) of the next trading day. Meanwhile, the results obtained on periodic data (Electricity Transformer Temperature, Weather) are approximately \verb|NHITS|. Ablation study is also conducted, which analyze the contribution of each component in the proposed method. Based on the results, this work proves the role of each component as well as the rationality in choosing algorithms in the method designation.
\end{abstract}

% \begin{graphicalabstract}
% \includegraphics{figs/cas-grabs.pdf}
% \end{graphicalabstract}

\begin{highlights}
    \item Aggregating information from multi-source, multi-period of data improves the model.
    \item Meta-learning efficiently aggregates information from multi-source, multi-period of time.
    \item Achieve better results than frequency decomposition method on aperiodic datasets.
    \item Ablation study shows the superior of Temporal-ML compared to Transformers.
\end{highlights}

\begin{keywords}
    Aperiodic time-series data \sep Meta-learning \sep Foreign exchange \sep \verb|Temporal-ML|
\end{keywords}

\maketitle

\section{Introduction}

Time-series data is becoming increasingly diverse not only in terms of data sources (e.g., finance, energy, transportation, etc.), but also in terms of methods for analyzing and forecasting them (e.g., frequency decomposition and methods based on historical data memorization). This makes sense as planning, organizing, and developing business strategies are greatly aided by the analysis and prediction of time-series data.

\vspace{1mm}

Two main approaches used in analyzing and predicting time-series data are fundamental analysis and technical analysis \cite{ayitey2023forex}. While fundamental analysis concentrates on examining external factors that are hard to capture from past price changes, such as the economic strategies and policies of nations and businesses, technical analysis relies entirely on historical price changes to forecast future trends.

\vspace{1mm}

% It is challenging to achieve efficient automation of fundamental analysis due to the unstructured of news data. As a result, academics usually concentrate on creating techniques for technical analysis. The forecasting process can now be intelligently automated with the help of machine learning and deep learning techniques, such as Auto-regressive model \cite{moran1951hypothesis}, Moving average model \cite{rosenblatt2000gaussian}, Convolutional neural network - \verb|CNN| \cite{lecun1989handwritten}, Long short-term memory model - \verb|LSTM| \cite{hochreiter1997long}, and methods based on \verb|Transformers| \cite{chen2021autoformer, zhou2022fedformer, liu2025onsitnet}.

% \vspace{1mm}

% The aforementioned techniques, are mostly applicable to periodic time-series data, such as traffic, weather. Over time, this kind of data clearly shows a seasonal or cyclical character. Consequently, analysis and prediction can be performed quickly and accurately. This is especially true for existing techniques that use deep neural networks to decompose periods \cite{chen2021autoformer,liu2022scinet,zhou2022fedformer}. Nevertheless, \textbf{aperiodic time-series data} (e.g., stock prices, foreign exchange rates, etc.) has been the subject of very few investigations.

% \vspace{1mm}

Unfortunately, these methods mainly focus on data with strong periodicity. \textbf{Aperiodic time-series data} such as stock prices, foreign exchange rates, has been the subject of very few investigations. The primary characteristic of aperiodic data is that it is \textbf{heavily impacted by external influences}, including production strategy and the political and economic policies of nations or businesses. This results in notable modifications to the data's characteristics, which are also referred to as concept drift scenarios \cite{liu2025onsitnet}. Mathematically, this implies that aperiodic time-series data has a \textbf{continuous and unstable variance}. Additionally, because the data no longer exhibits periodicity, it becomes extremely \textbf{challenging to extract characteristics} from it using a sliding window.

\vspace{1mm}

% Consequently, there are three primary obstacles in dealing with aperiodic time-series data: (1) External influences including politics, economy, and society have a significant impact on the data; (2) Data has a massive and non-stationary variance; and (3) Data's strong non-periodicity makes feature extraction challenging.

% \vspace{1mm}

% handle challenge 1
External influences is the most challenging to resolve using technical analysis techniques as it is difficult to extract political and social information from historical data. However, according to the efficient market hypothesis \cite{fama1970efficient}, historical data itself clearly reflects information about market movements. This implies that, by carefully examining the transaction history, one can completely grasp market fluctuations and forecast the future without aggregating external news. In addition, studies \cite{overreactioncontrarian, mech1993portfolio} indicate the correlations between the financial indicators of a certain company and the indicators of other companies. In other words, integrating analysis of multiple data sources within the same domain can yield valuable insights for the indicator of interest. However, recent methods concentrate on exploiting insight from single source rather than mining correlations between multiple sources.

\vspace{1mm}

% handle challenge 2
Ensemble learning is frequently utilized to mitigate the impacts of variation \cite{ali2020complete, sadeghi2021combined, zafeiriou2020intraday}. This type of learning involves dividing the data into several parts, assuming and learning a certain amount of variation in each part. This approach is reasonable because it is challenging for using only one single model to capture all of the variability in the data over long time. Even so, the synthesis mechanism of ensemble learning methods is still simple and rigid, which may not be able to capture the exploited information in sub-models.

\vspace{1mm}

% handle challenge 3
So far, many models have been proposed to be able to extract features on data in general and time-series data in particular. Typically, Convolutional neural network (\verb|CNN|) \cite{lecun1989handwritten}, Long short-term memory (\verb|LSTM|) \cite{hochreiter1997long}, and attention mechanism \cite{bahdanau2014neural, luong2015effective, vaswani2017attention} have been proposed to extract local features, selectively remember long, short-term features, and emphasize vectors in the feature matrix, respectively. However, these methods still have the problem of information forgetting, which limits the ability to learn from long sequences. In addition, recent approaches propose to decompose input signal into frequency bands \cite{chen2021autoformer, liu2022scinet, zhou2022fedformer}. These methods are reported to achieve an excellent performance but most of experiments are conducted on highly periodic data. Our data exhibits strong non-periodicity, which makes it a huge challenge for recent methods.

\vspace{1mm}

On the other hand, optimization-based meta-learning algorithm (ML) algorithms are known for their ability to increase the generalization and adaptability of a model on limited data \cite{hospedales2021meta, vettoruzzo2024advances}. During training, the optimization process of ML can be viewed as an efficient synthesis of models across sub-tasks (sub-datasets). Based on this ability, a machine learning models trained by ML algorithms are expected to efficiently synthesize information from multiple data sources/time periods as well as adapt well to the continuous change of variance.

\vspace{1mm}

% viết 1 đoạn kiểu: method của chúng toi overcomes 3 challenges trên
We approach the above challenges through the combination of temporal feature extraction and model synthesis. Specifically, by using \verb|MAML| \cite{finn2017model}, an ML algorithm, we efficiently synthesize the parameters of local models, which significantly reduces variance loss and increases the ability to aggregate information from multiple data sources as well as provides the generalization to the aggregated model. To effectively learn the temporal feature, we use Bidirectional LSTM (\verb|BiLSTM|) \cite{cui2018deep}, which is an old but efficient extraction method in dealing with sequenced data. Deep learning approaches such as \verb|BiLSTM| encounter obstacles in learning long sequences because long-term features will be forgotten over time. We overcome this problem by dividing dataset into multiple sequences as approached in ensemble model and synthesize the learned parameters effectively by ML. As a result, these challenges can be handled well.

\vspace{1mm}

We demonstrate the superiority of the proposed algorithm by solving the problem of predicting the trend (upward or downward) of time-series data and comparing to \verb|NHITS| \cite{challu2023nhits} (SOTA model in 2023) on two types of data: (1) - Aperiodic data, which contains USD/JPY exchange rate data (2000-2024) and the exchange rate data of 60 currency pairs made of 18 countries (2014-2024); (2) - Periodic data which contains Electricity Transformer Temperature \cite{zhou2021informer} and Weather \cite{Kolle} dataset. These datasets are publicly available on the Internet and can be downloaded easily. Moreover, we conduct an ablation study to analyze the contribution of \verb|MAML| and \verb|BiLSTM| in our method. The official implementation can be found at \url{https://github.com/baolongnguyenmac/multi_fx}.

\vspace{1mm}

% main contributions
In summary, our main contributions are as follows:

\begin{itemize}
    \item We propose \verb|Temporal-ML|, a method that effectively extracts temporal features of aperiodic time-series data and synthesizes sub-models across multi-source data.
    \item We conduct experiments on aperiodic and periodic data. The result shows that the proposed method outperforms \verb|NHITS| on aperiodic data and achieves competitive performance on periodic data, which indicates that ML algorithms are efficient in aggregating information from multiple sources.
    \item Ablation study is conducted to evaluate the contribution of components in proposed method as well as the superior of \verb|Temporal-ML| compared to \verb|Transformers| and other models.
\end{itemize}

\section{Related work}

\subsection{Traditional approach}

% boosting, bagging, stacking
% By analyzing and synthesizing information from multiple sub-models, we can obtain a multidimensional view of the data and reduce the impact of unstable variance, which helps the overall model adapt well to strong variance changes.

Two main directions in tackling the problems of time-series data are deep learning methods and ensemble methods. Meanwhile, the former approach tends to extract the deep hidden feature from data; the latter tries to integrate the strength of several sub-models to enhance the predictive ability.

\vspace{1mm}

% deep learning stuff
% For deep learning methods, \verb|CNN|-based \cite{lecun1989handwritten, samarawickrama2019multi, zhao2020wavelet} and \verb|RNN|-based methods \cite{lai2018modeling, rumelhart1986learning, shih2019temporal} are commonly utilized.

\verb|CNN| and its variants \cite{lecun1989handwritten, samarawickrama2019multi, zhao2020wavelet} use a sliding window (kernel) to slide along the time direction of data, which effectively extract the local relationship between the timestamp under consideration and the nearby timestamps. During the training process, \verb|CNN| tries to learn a suitable kernel to extract the temporal features, making it perform faster compare to conventional neural network. However, for time-series data in general and aperiodic time-series data in particular, the data properties at any given time depend not only on recent times but also on long-term times in the past. Therefore, \verb|CNN| should only be used as a method to support feature extraction in the process of solving problems on time-series data.

\vspace{1mm}

Instead of focusing on extracting local feature, \verb|RNN|-based methods \cite{lai2018modeling, rumelhart1986learning, shih2019temporal} concatenate on creating a hidden state which selectively presents a long period of time from input data. Indeed, these algorithms perform on an input sequence of arbitrary size. At each step, the hidden state is updated by extracting information from the current input element and combining with the previous state. As a result, \verb|RNN|-based methods easily outperform \verb|CNN| since they can effectively capture the historical data. Nevertheless, \verb|RNN| and its variations face a big problem of forgetting data in tackling long sequences.

\vspace{1mm}

\verb|Transformers|-based methods \cite{chen2021autoformer, zhou2022fedformer} utilizes the attention mechanism to extract the relationships between each element in the input sequence. Accordingly, we can predict the future value based on its dependence on each individual element in the past. In addition, \verb|Transformers| can also be used to extract relationships between dimensions in data \cite{liu2025onsitnet, zhang2023crossformer}. Although it has been proven effective on language processing and time-series data prediction tasks, \verb|Transformers| are only effective if the elements in the input sequence contain a lot of information to extract (high-dimensional vectors). For simple data types such as financial indicators, using \verb|Transformers| will encounter many difficulties.

\vspace{1mm}

On the other hand, aperiodic data is greatly influenced by external factors. These factors are directly reflected in the data of not only the indicator of interest but also other indicators in the same domain. Therefore, exploiting multi-source data can bring valuable information. However, the aforementioned methods have absolutely no mechanism to extract or synthesize this type of information. At this point, ensemble methods (bagging, boosting, stacking models) can be considered an obvious solution since it has a specific mechanism to learn and synthesize information from multiple data sources. However, the simplicity and rigidity in information aggregation of these methods have limited the idea of combining multi-source information. As a result, ensemble models have been gradually forgotten in recent studies.

\subsection{Frequency decomposition-based approach}

Fourier series has been widely used in the analysis and prediction of time-series data. It was shown by Jean-Baptiste Fourier to be able to approximate any periodic function with sine and cosine basis functions. With this ability, Fourier series can easily identify seasonal patterns such as weekly, monthly seasonality. STL decomposition \cite{rb1990stl} has similar capabilities but is more powerful because it allows users to change seasonal parameters and handles outlier data well (i.e., outlier data does not affect the trend and seasonal approximation). However, these methods are often inflexible because they depend on fixed parameters and cannot be tuned automatically. As a result, they become less flexible when dealing with complex data such as aperiodic time-series data.

\vspace{1mm}

To overcome this drawback, deep learning methods are used to exploit or simulate frequency features. With the flexibility in network architecture and feature extraction capabilities, deep learning methods quickly integrate the frequency features of data into the prediction process. For example, Fractional Fourier Transform \cite{kocc2022fractional} uses Fourier transform as a feature extraction step with the goal of transforming the original data to the spectral domain. The features in the spectral domain will then be fed into recurrent networks networks (e.g., \verb|GRU|, \verb|RNN|) for further processing.

\vspace{1mm}

% nbeats + nhits
Beyond incorporating frequency features into the learning process, deep learning networks can also simulate them. For example, \verb|NBEATS| \cite{oreshkin2019n} decomposes the input data into frequency bands using multiple stacked blocks, each of which consists of \verb|FullyConnected| (with or without non-linear activation function) layers. Each block is trained to predict future values and approximate its input values. The next block learns from the difference between the actual input values and the previous block's input data estimates. The prediction of each block is thus an approximation of a data's frequency band. Summing these values, we obtain the final predicted value. \verb|NHITS| uses a similar architecture but adds a \verb|MaxPooling| layer at the top of each block. The kernel size of these layers decreases to sample data at multiple rates. As a result, the frequency bands can be extracted in decreasing order of wavelength. In addition, during the synthesis process, \verb|NHITS| uses an interpolation function to reduce the computation time of the network compared to \verb|NBEATS|.

\vspace{1mm}

% 1 con transformer
% Mặc dù có thể phân tách được các giải tần trong dữ liệu, \verb|NBEATS| và \verb|NHITS| chỉ sử dụng các lớp tuyến tính đơn giản trong quá trình rút trích đặc trưng. Ngoài ra, kernel size trong \verb|MaxPooling| layer của \verb|NHITS| cần được chọn thủ công.

% dữ liệu của toi là phi chu kỳ
% Mặc dù có thể hoạt động tốt ngay cả trong bài toán dự đoán một chuỗi giá trị tương lai, các phương pháp nêu trên bắt gặp khó khăn lớn khi giải quyết các bài toán trên aperiodic time-series data, ngay cả trong bài toán dự đoán xu hướng tiếp theo của dữ liệu. Điều này là vì aperiodic time-series data thể hiện tính chu kỳ rất yếu nên các phương pháp dựa trên phân giải tần số rất khó nắm bắt được đặc trưng này. Ngoài ra, các phương pháp trên cũng có một nhược điểm tương tự các phương pháp deep learning truyền thống. Chúng không có cơ chế tổng hợp thông tin đa nguồn. Do đó, cần một hướng tiếp cận mới vừa có khả năng tổng hợp thông tin đa nguồn, vừa nắm bắt được các đặc trưng thời gian một cách hiệu quả.

Although they can perform well even in the problem of predicting a series of future value, the above methods encounter great difficulties when solving problems on aperiodic time-series data, even in the problem of predicting the next trend of the data. This is because aperiodic time-series data exhibits very weak periodicity. Thus, it is difficult for frequency-based methods to capture this feature. In addition, the above methods also have a disadvantage similar to traditional deep learning methods. They do not have a mechanism for synthesizing multi-source information. Therefore, a new approach, which can both synthesize multi-source information and effectively capture temporal features, is needed.

\subsection{Optimization-based Meta-learning}
\label{subsec:ml}

Optimization-based ML is a training method that allows model to gain experience after learning different tasks in the same task distribution. This equips the machine learning model with the ability to generalize highly and adapt quickly to new tasks after only a few training steps with limited training data. With this ability, ML is widely used in tasks that require the ability to fast adapt to new data (e.g. personalization of learning models \cite{chen2018federated, fallah2020personalized,nguyen2022meta}, domain adaptation in online learning \cite{hu2023meta, son2024meta}).

\vspace{1mm}

A basic ML algorithm is trained on multiple tasks $t$ drawn from the same task distribution $\mathcal{T}$ \cite{hospedales2021meta}. The data for task $t$ is divided into a support set $\mathcal{D}_t^{support}$ and a query set $\mathcal{D}_t^{query}$. During the learning process, two optimization steps, inner and outer optimization, are performed alternately. Inner optimization attempts to find an optimal set of parameters $\theta_t^*$ for each machine learning model on the support set of each task using the equation \ref{eq:inner_opt}.

\begin{align}
    \theta_t^* = \theta_t(\phi) = \arg\min_{\theta}{\mathcal{L}^{task}_t\left( \phi, \mathcal{D}_t^{support} \right)}
    \label{eq:inner_opt}
\end{align} In which, $\phi$ is the result of the outer optimization process, which acts as the initial value of $\theta_t$. $\mathcal{L}^{task}_t$ is the error function of the model on the support set of task $t$.

\vspace{1mm}

The algorithm then uses the optimal parameter sets $\theta_t^*$ to perform on the corresponding query set. The losses of the entire models are then aggregated to perform the outer optimization process as equation \ref{eq:outer_opt}.

\begin{align}
    \phi^* = \arg\min_{\phi}\sum_{t}{\mathcal{L}^{meta}_t\left( \theta_t(\phi), \mathcal{D}_t^{query} \right)}
    \label{eq:outer_opt}
\end{align}

By performing the above training method, the $\phi^*$ model will have a high level of generalization across different tasks, and can quickly respond to a new task after only a few training steps. In the context of aperiodic time-series data, the model can quickly learn and adapt to the new variance of data over time.

\vspace{1mm}

In the inference phase, the initial values for the model parameters are assigned $\phi^*$. The model is then adapted quickly to the support set and performed on the query set. The results on the query set are the model output.

\vspace{1mm}

One disadvantage of optimization-based ML method lies in solving equation \ref{eq:outer_opt} which requires massive overhead to compute and maintain a Hessian matrix. Even so, ML algorithms still achieve a high accuracy in handling many problem in which the quick adaptation or effective model synthesis are required.

\section{Methodology}

We propose \verb|Temporal-ML|, a ML-based method which consists of two main components that work in parallel: (1) - Temporal feature extraction; (2) - Models' parameter synthesis. The overview of the method is illustrated in Figure \ref{fig:flow}. In subsection \textit{Temporal feature extraction}, we propose to use the features extracted from \verb|BiLSTM| network. In subsection \textit{Effective synthesis of model's parameters}, \verb|MAML| is utilized to synthesize the parameters of the models.

\vspace{1mm}

Due to the contribution of \verb|BiLSTM| features, we aim to effectively extract hidden temporal features from aperiodic data. By using \verb|MAML| in the weight aggregation process, the proposed method is expected to be a reasonable and effective alternative to traditional ensemble models in effectively synthesizing external factors, minimizing the impact of variance variation, and efficiency capturing hidden long-term dependencies in the past.

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{figs/meta.png}
    \caption{The flow of meta-training and meta-testing on multi-fx dataset.}
    \label{fig:flow}
\end{figure}

\subsection{Data preparation}
\label{subsec:method_data}

\verb|Temporal-ML| bases on ML algorithms to train the model. Therefore, the data must be reorganized so that the ML algorithms can work. In case the data includes many different datasets belonging to the same field, each dataset will be considered a task. The goal of using multiple datasets of the same domain is to utilize the corelation between them as indicated in study \cite{overreactioncontrarian, mech1993portfolio}. In case the data includes a single dataset, it is necessary to divide this dataset into subsets corresponding to separate tasks. By this way, we aim to enhance the learning of information about external factor reflected in single dataset \cite{fama1970efficient}. In summary, the prepared dataset includes $n$ tasks: $\mathcal{D} = \left\{ \mathcal{D}_t \right\}_{t=1}^{n}$. The data at each task is divided into support and query sets: $\mathcal{D}_t = \left\{ \mathcal{D}_t^{support}, \mathcal{D}_t^{query} \right\}$.

\vspace{1mm}

A datapoint consists of a pair of values $(\mathbf{x}_{t-L:t}, y)$. In which, $\mathbf{x}_{t-L:t}$ includes $L$ historical values from timestamp $t$; $y\in \{0,1\}$ is the label, showing the decreasing or increasing trend of the data sample $x_{t+1}$ compared to $x_{t}$. Depending on each problem and the implementation, the elements in $\mathbf{x}_{t-L:t}$ can be a matrix or a vector. For example, consider stock data, $\mathbf{x}_{t-L:t}$ can contain $L$ vectors $\vec x_i = (\text{open, low, high, close})$ or can be a vector of close price values only.

\subsection{Temporal feature extraction}

Temporal feature extraction is performed using \verb|BiLSTM|. Specifically, since every datapoint in time-series data usually has a small number of attribute, each element in the matrix $\mathbf{x}_{t-L:t}$ (abbreviated as $\mathbf{x}$) is passed through a \verb|FullyConnected| layer (with a non-linear activation function) whose output is larger than the dimension of $\vec x_i, i\in[t-L, t]$ to obtain vector $\vec x'_i$ (equation \ref{eq:decompose}). Accordingly, the data characteristics are expressed deeper and clearer. These features are then passed through the \verb|BiLSTM| network (equation \ref{eq:feature_lstm}) to selectively extract long-term temporal dependencies ($\mathbf{h}_{BiLSTM}$). At the end, $\mathbf{h}_{BiLSTM}$ is passed through a classification head of the neural network (equation \ref{eq:clf}) to predict the label of $\mathbf{x}$.

\begin{align}
    \mathbf{x'} &= \mathbf{FullyConnected}\left( \mathbf{x} \right) \label{eq:decompose}\\
    \mathbf{h}_{BiLSTM} &= \mathbf{BiLSTM}\left( \mathbf{x'} \right) \label{eq:feature_lstm}\\
    \hat y &= \mathbf{FullyConnected}\left( \mathbf{h}_{BiLSTM} \right) \label{eq:clf}
\end{align}

\verb|BiLSTM| maintains cell-state values to selectively store temporal dependencies, as well as to mitigate the vanishing gradient problem during training. Therefore, it is well suited for solving time-series data problems. Compared to \verb|LSTM|, \verb|BiLSTM| is superior in capturing global context of a sequence as well as solving the problem of information forgetting when dealing with long sequences in \verb|LSTM|. This is because \verb|BiLSTM| can extract and synthesize information from both sides of the input sequence. In conclusion, \verb|BiLSTM| can capture the entire context with minimal information loss for long sequences. In the context of the movement prediction problem of aperiodic time-series data, \verb|BiLSTM| is expected to successfully capture continuous value variation as well as not miss important information at distant inputs.

\subsection{Efficient model synthesis}

We use \verb|MAML| to train and aggregate the weights of the models of all tasks via the \verb|CrossEntropyLoss| (equation \ref{eq:loss}).

\begin{align}
    \mathcal{L}(\mathbf{y}, \hat{\mathbf{y}}) = \mathbb{E}\left[(\mathbf{y}\log{\hat{\mathbf{y}}}) + (1-\mathbf{y})\log{(1-\hat{\mathbf{y}})}\right]
    \label{eq:loss}
\end{align}

As mentioned in section \ref{subsec:ml}, parameter optimization in the ML approach is to solve the two equations \ref{eq:inner_opt} and \ref{eq:outer_opt} using optimization methods on the support set and query set. Specifically, the optimization process includes many global steps (outer optimization), performed on all tasks participating in training. Each global step includes many local steps (inner optimization) performed on each individual task. At global step $r$, the $e$th local optimization process on the support set of task $t$ proceeds as follows:

\begin{align}
    \begin{cases}
        \theta_t^{(0)} &= \phi_{r-1} \\
        \theta_t^{(e)} &= \theta_t^{(e-1)} - \alpha \nabla_{\theta} \mathcal{L}^{task}_t\left( \theta_t^{(e-1)}, \mathcal{D}_t^{support} \right)
    \end{cases}
    \label{eq:inner_sol}
\end{align} In which, $\phi_{r-1}$ is the result of the $r-1$ outer optimization process, $\alpha$ is the inner learning rate, $\mathcal{L}_t^{task}$ evaluates the use of $\theta_t^{(e-1)}$ on $\mathcal{D}_t^{support}$.

\vspace{1mm}

Next, the outer optimization process is performed by aggregating the losses on the query set of the tasks and optimizing on it (equation \ref{eq:outer_sol}).

\begin{align}
    \begin{cases}
        \phi_0 = \text{Random Initialization}\\
        \phi_r = \phi_{r-1} - \beta \nabla_{\phi} \sum_{t=1}^n{\mathcal{L}^{meta}_t \left( \theta_t(\phi), \mathcal{D}_t^{query} \right)}
    \end{cases}
    \label{eq:outer_sol}
\end{align} Where, $\beta$ is the outer learning rate, $\theta_t^*(\phi)$ is the local optimized weight obtained from equation \ref{eq:inner_sol}, $\mathcal{L}_t^{meta}$ evaluates the use of $\theta_t(\phi)$ on $\mathcal{D}_t^{query}$.

\vspace{1mm}

Assuming the algorithm runs $E$ steps in inner optimization, the derivative quantity at equation \ref{eq:outer_sol} is rewritten as follows (the notations of dataset are removed):

\begin{align}
    &\nabla_{\phi} \sum_{t=1}^n{\mathcal{L}^{meta}_t \left( \theta_t - \alpha \nabla_{\theta} \mathcal{L}^{task}_t\left( \theta_t \right) \right)}\\
    &= \sum_{t=1}^n{ \frac{\partial \mathcal{L}^{meta}_t\left(\theta_t^{(E)}\right)}{\partial \theta_t^{(E)}} \frac{\partial \theta_t^{(E)}}{\partial \phi}}\\
    &= \sum_{t=1}^n{ \nabla_{\theta} \mathcal{L}^{meta}_t\left(\theta_t^{(E)}\right) \prod_{j=0}^{E-1} {\left[\mathbb{I} - \alpha\nabla^2_{\theta}\mathcal{L}^{task}_{t}\left(\theta_t^{(j)}\right)\right]}}
    \label{eq:outer_grad}
\end{align}

As obtained in equation \ref{eq:outer_grad}, the outer optimization process requires large overhead in terms of memory and computation to train the model. As such, the number of computational steps to find $\theta^*$ needs to be limited. In practice, methods using ML \cite{chen2018federated, fallah2020personalized, finn2017model, li2017meta, nguyen2022meta} often choose $E\in [1,5]$.

\vspace{1mm}

Under the view point of ensemble learning, the synthesis process can be regarded as an efficient way to aggregate models. Thus, information from multiple data sources is efficiently aggregated. The model can then capture the correlations between different datasets. Moreover, by dividing a dataset into multiple sequences, we ensure that the global model can learn temporal feature from entire dataset, reducing the problem of information forgetting in \verb|BiLSTM|.

\section{Numerical experiment}

\subsection{Dataset}

Foreign exchange in particular and financial indices in general are typical data types for aperiodic time-series data. Therefore, we choose this type of data to test the model. Specifically, we configure two datasets using foreign exchange data. \verb|USD/JPY| dataset consists of only data of the exchange rate between US dollar and Japanese yen. The data is sampled hourly from 2000 to 2024, including 4 attributes: \textit{open, low, high, and close price}. By carefully examining the transaction history as well as effectively aggregating data characteristics, we expect to forecast the movement of price without utilizing external information. \verb|multi-fx| dataset consists of 60 currency pairs made of 18 countries: Australia, Canada, Switzerland, Denmark, EU, United Kingdom, Hong Kong, Iceland, Japan, Norway, New Zealand, Singapore, Sweden, Turkey, United States, Mexico, China, South Africa. The data has similar attributes to \verb|USD/JPY| and is sampled daily from 2014 to 2024. By integrating analysis of multiple sources of information in the domain of FX, we expect to obtain valuable analytical insights for the indicator of interest.

\vspace{1mm}

In addition, we use two periodic datasets: Electricity Transformer Temperature (\verb|ETT-m2|) \cite{zhou2021informer} and Weather (\verb|WTH|) \cite{Kolle}. \verb|ETT-m2| dataset consists of 7 attributes, measuring parameters of a transformer in a province of China every 15 minutes from July 2016 to July 2018. \verb|WTH| dataset consists of 12 attributes, recording the weather parameters at the Weather Station of the Max Planck Biogeochemistry Institute in Jena, Germany every 10 minutes in 2020. These two datasets exhibit very strong periodicity, which frequency decomposition-based methods have been very successful in predicting. Experiments on them provide a more comprehensive insight of the proposed method's capabilities.

\subsection{Metric}

Our study uses \textit{Accuracy, Precision, Recall}, and \textit{F1-score} to evaluate the proposed method. Suppose \verb|Temporal-ML| is evaluated on a dataset $\mathcal{D}$ consisting of $a$ attributes and divided into $n$ tasks. The evaluation is performed by letting the model predict the movement of each attribute with all attributes as input, then performing two aggregation steps: (1) - Aggregation on tasks; (2) - Aggregation on attributes. This evaluation process is designed based on study \cite{challu2023nhits}.

\vspace{1mm}

\textbf{Aggregation on tasks}. Consider metric $m$ when the model predicts any attribute $k$ ($1\leq k \leq a$). After predicting attribute $k$, we obtain $n$ values: $\{m^{(k)}_1,\dots,m^{(k)}_n\}$. The process of synthesizing metrics $m$ of $n$ tasks is performed as follows:

\begin{align}
    \bar{m}^{(k)} &= \frac{1}{n}\sum_{i=1}^n{m^{(k)}_i} \label{eq:mean_task}\\
    s^{(k)}_m &= \sqrt{\frac{1}{n} \sum_{i=1}^n{\left(m^{(k)}_i - \bar{m}^{(k)}\right)^2}} \label{eq:std_task}
\end{align}

\textbf{Aggregation on attributes}. The model predicts all $a$ attributes and obtains $a$ metrics: $\left\{ \left( \bar{m}^{(k)}\pm s^{(k)}_m \right) \right\}_{k=1}^a$. The process of synthesizing metric $m$ of $a$ attributes is as follows:

\begin{align}
    \bar{m} &= \frac{1}{a}\sum_{k=1}^a{\bar{m}^{(k)}} \label{eq:mean_att}\\
    s_m &= \sqrt{\frac{1}{a} \sum_{k=1}^a{\left(s^{(k)}_m\right)^2}} \label{eq:std:att}
\end{align}

\subsection{Experimental setup}

% setup baseline
This study uses \verb|NHITS| (AAAI 2023) as the baseline model. \verb|NHITS| uses the entire attributes to predict the next trend for an attribute in a dataset. The algorithm aims to decompose the frequency bands and combine them to predict the future. For \verb|USD/JPY|, \verb|ETT-m2|, and \verb|WTH| datasets, the data in each set is divided into a training set, a validation set, and a test set with a ratio of 6:2:2, respectively. For \verb|multi-fx| dataset, since \verb|NHITS| does not have a mechanism to aggregate models across different datasets, we repeat the training process (training, validation, testing on 60\%, 20\%, 20\% of the data, respectively) for each currency pair, then aggregate metrics across currency pairs to obtain the final evaluation. In addition, a random model is also used for comparison. This model generates predicted values of 0 and 1 according to a uniform distribution. The process of synthesizing metrics of this model is similar to \verb|NHITS|.

\vspace{1mm}

% setup our method
As mentioned in subsection \ref{subsec:method_data}, to use ML, the datasets need to be structured into separate tasks. In each task, the support set accounts for 20\% of the data, the query set accounts for 80\% of the data. The statistics of tasks in datasets are presented in Table \ref{tab:stat_}. In which, the ratio of tasks used for meta training, validation, and testing is 50:25:25. The details of dividing tasks is presented in Appendix \ref{app:preprocessing}.

\vspace{1mm}

\begin{table}[h]
    \centering
    \caption{Statistics on datasets.}
    \label{tab:stat_}
    % \resizebox{\columnwidth}{!}{}
    \begin{tblr}{
        cells = {c},
        hline{1,6} = {-}{0.08em},
        hline{2} = {-}{},
        }
        Dataset                      & \#Attribute & \#Task & \#Sample    & \#Sample/Task    \\
        USD/JPY                      & 4           & 60     & 150,175     & 2,503          \\
        multi-fx                     & 4           & 120    & 154,440     & 1,287          \\
        ETT-m2                       & 7           & 48     & 69,680      & 1,452          \\
        WTH                          & 12          & 40     & 35,064      & 877            
    \end{tblr}
\end{table}

% normalization
Before data is fed to model, it is normalized by \verb|z-score|. Specifically, training data is used to compute mean and standard deviation of each attribute. Then, these values are used to normalize validation and testing data. All models are trained on training set and validated on validation set. The best model is selected based on the performance on validation set. The final performance on testing set is reported. The details of tuning process is presented in Appendix \ref{app:tuning}.

\subsection{Ablation study}
\label{subsec:ab_exp}

We evaluate each component in \verb|Temporal-ML| by removing or replacing it and compare the results to the original algorithm. Specifically, we conduct experiments on two groups of algorithms: (1) - Model with \verb|BiLSTM| replaced; (2) - Model without ML. Regarding the evaluation of these models, ML-based models are evaluated as \verb|Temporal-ML|, the remaining models are evaluated as \verb|NHITS|. The details of model architectures are presented in Appendix \ref{app:ab_model}.

\vspace{1mm}

\textbf{Model with} \verb|BiLSTM| \textbf{replaced}. In this group, we replace \verb|BiLSTM| with \verb|BiLSTM+CNN|, \verb|Transformers|. The goal of replacing \verb|BiLSTM| with \verb|BiLSTM+CNN| is to test the local feature extraction ability of the extractor. In order to do so, we concatenate the features of \verb|BiLSTM| and \verb|CNN| then performing classification. The purpose of using \verb|Transformers| is to compare the performance of attention mechanism to \verb|BiLSTM| on aperiodic time-series data.

\vspace{1mm}

\textbf{Model without ML}. To study the impact of \verb|MAML| as well as optimization-based ML algorithms on \verb|Temporal-ML|, we remove this component and only use the pure feature extraction methods introduced above: \verb|BiLSTM|, \verb|BiLSTM+CNN|, \verb|Transformers|. Through this, the ability to effectively synthesize information from multiple sources of \verb|MAML| in aperiodic time-series data will be clarified. The dataset which contains information from multiple sources is \verb|multi-fx| dataset. Thus, we only experiment on this dataset.

\section{Result \& Discussion}

\subsection{Main result}

The main results of this work are shown in Table \ref{tab:mine_nhits}. Accordingly, our proposed method outperforms \verb|NHITS| on all metrics on two aperiodic datasets. On periodic datasets, our model achieves a competitive results compared to the baseline model. Specifically, for \verb|USD/JPY| and \verb|multi-fx| dataset, \verb|Temporal-ML| improves from 12\% to 14\% on metrics compared to the \verb|NHITS|. On the other hand, for periodic datasets, the accuracy of \verb|Temporal-ML| is 1\% lower, and the remaining metrics are 4-5\% lower than \verb|NHITS| on \verb|ETT-m2| data. For strongly cyclical data such as \verb|WTH|, our algorithm outperforms \verb|NHITS| by about 1\%. In addition, \verb|Temporal-ML| also achieves 1-3\% lower dispersion on metrics than \verb|NHITS| when operating in multi-source data context.

\vspace{1mm}

\begin{table*}[H]
    \centering
    \caption{Classification results (\%) of Temporal-ML and NHITS on aperiodic (USD/JPY, multi-fx) and periodic (ETT-m2, WTH) datasets. Best results per metrics are boldfaced. (\mbox{*}): Our method.}
    \label{tab:mine_nhits}
    \begin{tabular}{c|c|cccc}
        \toprule
        \multicolumn{1}{c}{}                &                                & $accuracy$              & $precision$             & $recall$                & $F1-score$                 \\
        \hline
        \multirow{2}{*}{USD/JPY}     		& NHITS                   & $58.46$                 & $58.24$                 & $57.65$                 & $55.82$            \\
                                            & Temporal-ML\mbox{*}     & $\mathbf{70.33\pm1.69}$ & $\mathbf{70.73\pm1.82}$ & $\mathbf{70.26\pm1.93}$ & $\mathbf{69.28\pm2.50}$    \\
                                            % & Ours(\verb|BiLSTM+CNN|)  & $68.75\pm1.78$          & $69.15\pm1.55$          & $68.73\pm1.77$          & $69.14\pm2.17$             \\
        \hline
        \multirow{2}{*}{multi-fx}           & NHITS                   & $53.51\pm5.02$          & $53.90\pm7.68$          & $53.64\pm4.05$          & $50.20\pm7.35$              \\
                                            & Temporal-ML\mbox{*}     & $\mathbf{66.26\pm7.45}$ & $\mathbf{67.08\pm8.97}$ & $\mathbf{65.76\pm7.79}$ & $\mathbf{64.06\pm10.00}$   \\
                                            % & Ours(\verb|BiLSTM+CNN|)  & $66.08\pm8.19$          & $66.56\pm9.23$          & $65.41\pm8.66$          & $63.78\pm10.83$            \\
        \hline
        \multirow{2}{*}{ETT-m2}             & NHITS                   & $\mathbf{71.88}$        & $\mathbf{66.86}$        & $\mathbf{62.49}$        & $\mathbf{62.69}$   \\
                                            & Temporal-ML\mbox{*}     & $71.14\pm9.13$          & $62.21\pm7.51$          & $58.75\pm5.30$          & $57.61\pm6.74$             \\
                                            % & Ours(\verb|BiLSTM+CNN|)  & $71.40\pm9.10$          & $63.25\pm8.82$          & $60.22\pm8.42$          & $59.43\pm9.24$             \\
        \hline
        \multirow{2}{*}{WTH}		        & NHITS                   & $74.18$                 & $68.05$                 & $65.04$                 & $65.40$             \\
                                            & Temporal-ML\mbox{*}     & $\mathbf{74.97\pm2.63}$ & $\mathbf{69.13\pm3.44}$ & $\mathbf{65.98\pm3.96}$ & $\mathbf{66.00\pm3.80}$    \\
                                            % & Ours(\verb|BiLSTM+CNN|)  & $74.41\pm2.71$          & $68.17\pm4.60$          & $65.50\pm4.82$          & $65.63\pm4.47$             \\
        \bottomrule
        \end{tabular}
\end{table*}

In Table \ref{tab:mine_nhits}, the results are aggregated as the average of all features, so they are highly representative. However, these results lack detail in the evaluation. Therefore, we analyze the results of the algorithms on each attribute of the aperiodic datasets in Table \ref{tab:mine_nhits_att}. Accordingly, when \verb|NHITS| predicts the attributes \textit{hight, low, close price} on both datasets, the results are only very close to the random prediction model. This shows that prediction on aperiodic data is really difficult. On the other hand, \verb|Temporal-ML| fails to predict \textit{low price} of \verb|USD/JPY| (accuracy approaching random prediction level). All other features of both datasets achieve high convergence and are significantly higher than \verb|NHITS|. This demonstrates the superiority of the proposed algorithm on aperiodic data.

\vspace{1mm}

\begin{table*}[H]
    \centering
    \caption{Accuracy (\%) of Temporal-ML and NHITS on each attribute of USD/JPY and multi-fx dataset. Best results per metrics are boldfaced. (\mbox{*}): Our method.}
    \label{tab:mine_nhits_att}
    \begin{tabular}{c|c|cccc}
        \toprule
        \multicolumn{1}{c}{}                &                                & $Open$                  & $High$                  & $Low$                   & $Close$                     \\
        \hline
        \multirow{3}{*}{USD/JPY}            & Random                         & $49.91$                 & $49.91$                 & $50.33$                 & $50.66$                     \\
                                            & NHITS                   & $75.03$                 & $53.74$                 & $51.57$                 & $53.50$                     \\
                                            & Temporal-ML\mbox{*}     & $\mathbf{93.06\pm1.36}$ & $\mathbf{68.87\pm1.53}$ & $\mathbf{51.01\pm1.58}$ & $\mathbf{68.37\pm2.16}$     \\
        \hline
        \multirow{3}{*}{multi-fx}           & Random                         & $49.92\pm2.19$          & $49.93\pm2.22$          & $49.94\pm2.17$          & $49.85\pm2.13$              \\
                                            & NHITS                   & $58.36\pm7.13$          & $50.66\pm4.28$          & $51.17\pm3.21$          & $53.84\pm4.60$              \\
                                            & Temporal-ML\mbox{*}     & $\mathbf{78.13\pm6.33}$ & $\mathbf{62.74\pm3.30}$ & $\mathbf{58.28\pm3.86}$ & $\mathbf{65.88\pm12.49}$    \\
        \bottomrule
        \end{tabular}
\end{table*}

To clarify the capabilities of \verb|Temporal-ML|, based on the main results and the results obtained from ablation experiments (subsection \ref{subsec:ab_exp}), we analyze the proposed algorithm on two aspects: (1) - Temporal feature extraction ability; (2) - Models' parameters aggregation ability.

\subsection{Temporal feature extraction ability}

From the tables \ref{tab:mine_nhits} and \ref{tab:mine_nhits_att}, it is not difficult to see that the temporal feature extraction ability of \verb|Temporal-ML| is much better than that of \verb|NHITS|. Indeed, the feature extraction and synthesis process of \verb|NHITS| shows that this method is trying to simulate the frequency resolution process of Fourier transform. The frequency band features are the most important information for periodic data. Therefore, the predictions of \verb|NHITS| can easily achieve high accuracy on this type of data. However, it is very difficult to achieve frequency resolution on aperiodic data. Furthermore, the time-series data, although filtered through multiple stacks to extract low-to-high frequency bands, the filtering techniques used in the network are extremely simple (using only \verb|FullyConnected| with nonlinear activation functions and \verb|Pooling| layers). This is what makes \verb|NHITS| a major obstacle on \verb|multi-fx|, \verb|USD/JPY| datasets and aperiodic data in general.

\vspace{1mm}

To better understand the capabilities of \verb|BiLSTM|, we performed an ablation study on this component of \verb|Temporal-ML| and presented the results in Table \ref{tab:ab_lstm}. Accordingly, when \verb|BiLSTM| is replaced by other components (\verb|BiLSTM+CNN|, \verb|Transformers|) for temporal feature extraction, the results on all four metrics of most datasets are lower than those obtained by \verb|Temporal-ML|. Specifically, on two aperiodic datasets (\verb|USD/JPY| and \verb|multi-fx|), the algorithms using \verb|BiLSTM+CNN| and \verb|Transformers| to extract features give 2-10\% lower in prediction ability than \verb|Temporal-ML|. On periodic data, \verb|MAML(Transformer)| still gives much lower results than \verb|Temporal-ML|. However, \verb|MAML(BiLSTM+CNN)| gives results that are almost asymptotically close to the proposed algorithm. On all four datasets, the standard deviation of the metrics of the proposed algorithm is generally 1-5\% smaller than the remaining algorithms.

\vspace{1mm}

\begin{table*}[H]
    \centering
    \caption{Classification results (\%) of Temporal-ML and models with BiLSTM replaced. Best results per metrics are boldfaced. (\mbox{*}): Our method.}
    \label{tab:ab_lstm}
    \begin{tabular}{c|c|cccc} 
        \toprule
        \multicolumn{1}{c}{}      &                                          & $accuracy$               & $precision$              & $recall$                 & $F1-score$                \\ 
        \hline
        \multirow{3}{*}{USD/JPY} & Temporal-ML\mbox{*}         & $\mathbf{70.33\pm1.69}$  & $\mathbf{70.73\pm1.82}$  & $\mathbf{70.26\pm1.93}$  & $\mathbf{69.28\pm2.50}$   \\
                                        & MAML(BiLSTM+CNN)     & $68.75\pm1.78$           & $69.15\pm1.55$           & $68.73\pm1.77$           & $69.14\pm2.17$            \\
                                        & MAML(Transformers)   & $60.67\pm4.09$           & $59.74\pm6.24$           & $59.32\pm3.79$           & $53.15\pm6.19$            \\ 
        \hline
        \multirow{3}{*}{multi-fx}& Temporal-ML\mbox{*}         & $\mathbf{66.26\pm7.45}$  & $\mathbf{67.08\pm8.97}$  & $\mathbf{65.76\pm7.79}$  & $\mathbf{64.06\pm10.00}$  \\
                                        & MAML(BiLSTM+CNN)     & $66.08\pm8.19$           & $66.56\pm9.23$           & $65.41\pm8.66$           & $63.78\pm10.83$           \\
                                        & MAML(Transformers)   & $56.03\pm6.08$           & $56.32\pm6.01$           & $55.63\pm5.01$           & $52.33\pm7.08$            \\ 
        \hline
        \multirow{3}{*}{ETT-m2}  & Temporal-ML\mbox{*}         & $71.14\pm9.13$           & $62.21\pm7.51$           & $58.75\pm5.30$           & $57.61\pm6.74$            \\
                                        & MAML(BiLSTM+CNN)     & $\mathbf{71.40\pm9.10}$  & $\mathbf{63.25\pm8.82}$  & $\mathbf{60.22\pm8.42}$  & $\mathbf{59.43\pm9.24}$   \\
                                        & MAML(Transformers)   & $68.69\pm8.85$           & $58.54\pm4.38$           & $57.85\pm4.71$           & $57.07\pm5.14$            \\ 
        \hline
        \multirow{3}{*}{WTH}     & Temporal-ML\mbox{*}         & $\mathbf{74.97\pm2.63}$  & $\mathbf{69.13\pm3.44}$  & $\mathbf{65.98\pm3.96}$  & $\mathbf{66.00\pm3.80}$   \\
                                        & MAML(BiLSTM+CNN)     & $74.41\pm2.71$           & $68.17\pm4.60$           & $65.50\pm4.82$           & $65.63\pm4.47$            \\
                                        & MAML(Transformers)   & $72.28\pm3.06$           & $62.35\pm8.73$           & $60.90\pm4.97$           & $58.65\pm4.72$            \\
        \bottomrule
    \end{tabular}
\end{table*}

According to this result, it can be seen that using \verb|BiLSTM| brings high and stable results in the prediction process. Using networks such as \verb|CNN|, \verb|Transformers| not only create negative improvements but also disturbs the learning process or reduces the efficiency of the whole model. For using \verb|BiLSTM+CNN|, the learning process is disturbed because \verb|BiLSTM| itself has selected effective time dependencies during the extraction process while \verb|CNN| only extracts features but does not select. This leads to the instability of \verb|BiLSTM+CNN|'s features. If \verb|CNN| extracts good features, the network will perform well (e.g., metrics on \verb|ETT-m2|). On the contrary, the network will perform worse than using only \verb|BiLSTM| (e.g., metrics on \verb|USD/JPY|, \verb|multi-fx|, \verb|WTH|).

\vspace{1mm}

With the above analysis, we conclude that \verb|BiLSTM| is not only a suitable choice for exploring the variability of aperiodic time-series data but also a suitable algorithm to combine with ML algorithms when working on time-series data in general. By using \Verb|BiLSTM|, we have effectively extracted market fluctuations and predicted the future.

\subsection{Models' parameters aggregation ability}

In the process of synthesizing features to make the final prediction, \verb|NHITS| only uses simple addition and subtraction. This makes sense in the context of synthesizing information based on the frequency bands extracted from the input data. Indeed, data related to weather, demand for airplanes or electricity show very strong periodic characteristics because weather and human consumption behavior, although they may vary over time, still show a very clear seasonality (e.g., cold weather in winter, high demand for airplanes during holidays). Therefore, the frequency of this type of data is a good feature and can be easily extracted. After decomposing the input data into frequency bands, we can simply synthesize them to be able to predict the future. In addition, synthesizing the frequency bands by an addition can be considered very simple compared to current deep learning algorithms, which work on the hidden feature space of the data. Therefore, the synthesis process of \verb|NHITS| on aperiodic data encounters many difficulties because the input information for the process is not good and the synthesis method is ineffective. Consequently, it can be seen that aperiodic time-series data becomes a fatal weakness for \verb|NHITS| and frequency decomposition-based methods in general.

\vspace{1mm}

Dealing with the complexity of non-periodic data requires a complex structure that can efficiently aggregate features. As mentioned in the Introduction, analyzing and synthesizing information from multiple data sources in the same domain can provide useful information for predicting metrics of interest. \verb|Temporal-ML| with the ability to synthesize information based on the optimization process of ML algorithms, allows the model to capture multi-dimensional information from different data sources, which can enhance the effectiveness of predictions.

\vspace{1mm}

To evaluate the capabilities of ML algorithms, we remove this component from the proposed method and then run experiments to compare with the original results. We use ML algorithms for the purpose of synthesizing multi-source information. Therefore, comparisons are only performed on the \verb|multi-fx| dataset because this dataset contains several sub-datasets from different sources, which helps to accurately evaluate the information synthesis capability. The results are summarized in Table \ref{tab:ab_ml}. Accordingly, with the information synthesis capability, \verb|Temporal-ML| achieves 2-12\% higher in all metrics than algorithms that do not use ML in testing process. In addition, the dispersion of the metrics is also lower from 1-7\%, proving that the difference in the task prediction process is much better than the remaining algorithms.

\vspace{1mm}

\begin{table*}[H]
    \centering
    \caption{Classification results (\%) of Temporal-ML and models without MAML on multi-fx dataset. Best results per metrics are boldfaced. (\mbox{*}): Our method.}
    \label{tab:ab_ml}
    \begin{tabular}{c|cccc} 
    \toprule
                                & $accuracy$              & $precision$             & $recall$                & $F1-score$                \\ 
    \hline
    Temporal-ML\mbox{*}  & $\mathbf{66.26\pm7.45}$ & $\mathbf{67.08\pm8.97}$ & $\mathbf{65.76\pm7.79}$ & $\mathbf{64.06\pm10.00}$  \\
    BiLSTM               & $63.88\pm9.00$          & $64.18\pm13.39$         & $63.66\pm9.13$          & $60.07\pm13.09$           \\
    BiLSTM+CNN           & $62.23\pm8.63$          & $62.89\pm9.76$          & $62.20\pm8.34$          & $59.91\pm10.97$           \\
    Transformers         & $58.44\pm9.15$          & $56.71\pm15.10$         & $58.22\pm8.84$          & $52.22\pm13.12$           \\
    \bottomrule
    \end{tabular}
\end{table*}

The better performance of \verb|Temporal-ML| on \verb|multi-fx| is understandable because all the remaining methods in the \ref{tab:ab_ml} table do not have the ability to synthesize information like the proposed algorithm, but rely entirely on past data to predict the future. Therefore, not only \verb|BiLSTM|, \verb|BiLSTM+CNN| but also \verb|Transformers| cannot take advantage of the correlation between data sources.

Therefore, it can be seen that, the use of optimization-based ML algorithms in information synthesis is completely reasonable because it successfully takes advantage of the correlation between data sets and achieves high accuracy without requiring too much data as in the single-source context.

\section{Conclusion}

Problems on aperiodic time-series data pose a great challenge to current machine learning models because this data depends on both transaction history and external factors such as economic and political situations; the variance of the data changes continuously; the aperiodicity of the data is very strong. By utilizing the time constraint extraction capability of \verb|BiLSTM| and the efficient parameter synthesis capability of \verb|MAML|, we propose the \verb|Temporal-ML| algorithm with the ability to extract hidden time features as well as synthesize information from multi-source data. Based on experiments, in the process of solving the next trend prediction problem, the proposed method achieves high efficiency on two aperiodic data sets (\verb|USD/JPY| and \verb|multi-fx|) and equivalent efficiency on periodic data sets (\verb|ETT-m2| and \verb|WTH|) compared to \verb|NHITS| - a frequency decomposition based approach. In addition, by performing an ablation study, we demonstrate the superior of \verb|BiLSTM| in extracting temporal feature compared to \verb|CNN|, \verb|Transformers| and the effectiveness of \verb|MAML| in synthesizing information from multi-source data. The results of this study are expected to be a good foundation for future research on time-series data.

Our experiment only illustrates a typical case of efficient model synthesis. In the future, by using different ML algorithms such as \verb|Meta-SGD| \cite{li2017meta}, \verb|Reptile| \cite{nichol2018first}, \verb|iMAML| \cite{rajeswaran2019meta}, it is possible to generate new models with higher accuracy. On the other hand, it is possible to extend this method to solve long-horizon prediction problems. Indeed, by changing the output and error of the model, it is possible to solve these problems. However, the architecture of the sub-models needs to be re-examined to better suit the new problem.

\appendix

\section{Data preprocessing}
\label{app:preprocessing}

\subsection{Task-based preprocessing}

Since the datasets are all time-series, the training data should not reveal any information about the future. Therefore, we divide every currency pair in \verb|multi-fx| into 2 tasks, resulting in 120 tasks. The first parts of every pair are used for meta-training. The second parts of them are used for meta validating and meta testing with the ratio of 50:50. The remaining datasets are is divided into periods of time, each corresponding to a task.

\vspace{1mm}

The \verb|z-score| preprocessing is performed on training data to obtain mean and standard deviation values. These values are used to normalize the data of validation set and test set.

\subsection{Fairness in evaluation}

Let $m$ be the number of data samples in each task, the total data of training, validating, and testing is: $mn$. For \verb|NHITS|, the model is trained, validated, and tested on 60\%, 20\%, 20\% of the data, respectively:

\begin{align}
    \text{Train: } &0.6mn\\
    \text{Validation: } &0.2mn\\
    \text{Testing: } &0.2mn
\end{align}

For \verb|Temporal-ML|, the knowledge enrichment process for the model uses the entire training data and the support set of the testing/validation data. The testing/validation process is performed on the query set:

\begin{align}
    \text{Train: } &0.5mn + 2\times 20\%(0.25mn) = 0.6mn\\
    \text{Validation: } &80\% (0.25mn)  =0.2mn\\
    \text{Testing: } &80\% (0.25mn)  =0.2mn
\end{align}

Therefore, with the above data division, we achieve fairness in the number of data samples during testing.

\section{Hyperparameter tuning}
\label{app:tuning}

\subsection{Temporal-ML}

In our implementation, we use a \verb|FullyConnected| layer of 16 units with a \verb|ReLU| activation function to represent deeper and more explicit features. This feature is then passed to a \verb|BiLSTM| block. The \verb|BiLSTM| block consists of 32 hidden units, two last outputs of which are concatenated to form a final vector. This vector is then passed through a binary classification layer with a \verb|Sigmoid| activation function.

\vspace{1mm}

The fine-tuning process of ML algorithms involves many hyper-parameters such as inner batch size, outer batch size, inner training step, outer training step, etc. To make fine-tuning easy, we fix most of the parameters and only fine-tune the size of lookback window, inner and outer learning rate. Details are presented in Table \ref{tab:our_finetune}.

\begin{table}[H]
    \centering
    \caption{Search space for fine-tuning our method.}
    \label{tab:our_finetune}
    % \resizebox{\columnwidth}{!}{}
    \begin{tabular}{ll} 
        \toprule
        \multicolumn{1}{c}{Hyper-parameter} & \multicolumn{1}{c}{Seach space}   \\ 
        \hline
        Inner batch size (samples/batch)    & \{32\}                            \\
        Inner training step                 & \{3\}                             \\
        Outer batch size (tasks/batch)      & \{5\}                             \\
        Outer training step                 & \{100\}                           \\ 
        \hline
        Lookback window                     & \{10, 20, 30\}                    \\
        Inner learning rate                 & \{0.001, 0.00325, 0.0055, 0.00775, 0.01\}\\
        Outer learning rate                 & \{0.001, 0.00325, 0.0055, 0.00775, 0.01\}\\
        \bottomrule
    \end{tabular}
\end{table}

\subsection{NHITS}

We rely on study \cite{challu2023nhits} to define the search space (Table \ref{tab:nhits_finetune}) for parameter fine-tuning. For parameters not mentioned in the table, we use the default values of the \verb|NHITS| implementation in \verb|NeuralForecast| library \cite{neuralforecast}. The best results of fine-tuning process are selected and reported in this study.

\begin{table}[H]
    \centering
    \caption{Search space for fine-tuning NHITS.}
    \label{tab:nhits_finetune}
    % \resizebox{\columnwidth}{!}{}
    \begin{tabular}{ll} 
        \toprule
        \multicolumn{1}{c}{Hyper-parameter} & \multicolumn{1}{c}{Seach space}                              \\ 
        \hline
        Random seed                         & \{1\}                                                        \\
        Number of stacks                    & \{3\}                                                        \\
        Number of blocks in each stack      & \{1\}                                                        \\
        Activation function                 & \{ReLU\}                                                     \\
        Batch size                          & \{256\}                                                      \\
        Epoch                               & \{500\}                                                      \\ 
        \hline
        Lookback window                     & \{10, 20, 30\}                                                \\
        Pooling kernel                      & \{[2,2,2], [4,4,4], [8,8,8], [8,4,1], [16,8,1]\}           \\
        Stacks' coefficients                & \{[168,24,1], [24,12,1], [180,60,1],[40,20,1], [64,8,1]\}  \\
        Number of MLP layers                & \{1,2\}                                                      \\
        Learning rate                       & \{0.001, 0.00325, 0.0055, 0.00775, 0.01\}                          \\
        \bottomrule
        \end{tabular}
\end{table}

\section{Model architecture in ablation study}
\label{app:ab_model}

\verb|BiLSTM+CNN| model includes two parts: \verb|BiLSTM| and \verb|CNN|. The \verb|BiLSTM| part has the similar architecture with the one used in \verb|Temporal-ML|. The \verb|CNN| part consists of two \verb|Convolution1D| layers with \verb|ReLU| as activation function and the number of filters are 32 and 64, respectively. Each of convolution layer is followed by a \verb|MaxPooling| layer with the kernel size of 2. Next, the output of \verb|CNN| part is flatten and concatenated with the output of the \verb|BiLSTM| part. The concatenated is then passed through a classification head with \verb|Sigmoid| activation function.

\vspace{1mm}

\verb|Transformer| model is composed of two encoders. Each encoder uses three attention heads, and the key vector dimension is 256. The \verb|FeedForward| part of the encoder uses two \verb|Convolution1D| layers and one \verb|LayerNormalization|. After stacking the two encoder layers, the feature matrix is passed through a \verb|Flatten| layer and then fed into the classification head.

\printcredits

% \clearpage

%% Loading bibliography style file
%\bibliographystyle{model1-num-names}
\bibliographystyle{cas-model2-names}

% Loading bibliography database
\bibliography{cas-refs}

\end{document}

### 2025-03-03 doc/thesis/chapters/4.experiment.tex
paper + thesis

@@
 \verb|USD/JPY| dataset consists of only data of the exchange rate between US dollar and Japanese yen.
-The data is sampled hourly from 2000 to 2024, including the attributes of open, low, high, and close price.
+The data is sampled hourly from 2000 to 2024, including 4 attributes: \textit{open, low, high, and close price}.
 \textbf{By carefully examining the transaction history as well as effectively aggregating data characteristics, we expect to forecast the movement of price without utilizing external information.}
@@
 
-We evaluate each component in \verb|Temporal-ML| as well as the choice of temporal feature extraction method by removing or replacing them from the algorithm and compare the results with the original algorithm.
+We evaluate each component in \verb|Temporal-ML| by removing or replacing it from the algorithm and compare the results to the original algorithm.
 Specifically, we conduct experiments on two groups of algorithms: (1) - Model without \verb|BiLSTM|; (2) - Model without ML.

### 2025-03-03 doc/thesis/chapters/5.results.tex
paper + thesis

@@
     \centering
-    \cprotect\caption{Accuracy (\%) of \verb|Temporal-ML| and \verb|NHITS| on each attribute of \verb|USD/JPY| and \verb|multi-fx|. Best results per metrics are boldfaced. (\mbox{*}): Our method.}
+    \cprotect\caption{Accuracy (\%) of \verb|NHITS| on each attribute of \verb|USD/JPY| and \verb|multi-fx| dataset.}
     \label{tab:mine_nhits_att}
-    \resizebox{\columnwidth}{!}{\begin{tabular}{c|c|cccc}
+    % \resizebox{\columnwidth}{!}{}
+    \begin{tabular}{c|c|cccc}
         \toprule
@@
         \hline
-    \multirow{3}{*}{\Verb|USD/JPY|}     & Random                         & $49.91$                 & $49.91$                 & $50.33$                 & $50.66$                     \\
+        \multirow{2}{*}{\Verb|USD/JPY|}     & Random                         & $49.91$                 & $49.91$                 & $50.33$                 & $50.66$                     \\
                                             & \Verb|NHITS|                   & $75.03$                 & $53.74$                 & $51.57$                 & $53.50$                     \\
-                                        & \Verb|Temporal-ML|\mbox{*}     & $\mathbf{93.06\pm1.36}$ & $\mathbf{68.87\pm1.53}$ & $\mathbf{51.01\pm1.58}$ & $\mathbf{68.37\pm2.16}$     \\
+                                            % & \Verb|Temporal-ML|\mbox{*}     & $\mathbf{93.06\pm1.36}$ & $\mathbf{68.87\pm1.53}$ & $\mathbf{51.01\pm1.58}$ & $\mathbf{68.37\pm2.16}$     \\
         \hline
-    \multirow{3}{*}{\Verb|multi-fx|}    & Random                         & $49.92\pm2.19$          & $49.93\pm2.22$          & $49.94\pm2.17$          & $49.85\pm2.13$              \\
+        \multirow{2}{*}{\Verb|multi-fx|}    & Random                         & $49.92\pm2.19$          & $49.93\pm2.22$          & $49.94\pm2.17$          & $49.85\pm2.13$              \\
                                             & \Verb|NHITS|                   & $58.36\pm7.13$          & $50.66\pm4.28$          & $51.17\pm3.21$          & $53.84\pm4.60$              \\
-                                        & \Verb|Temporal-ML|\mbox{*}     & $\mathbf{78.13\pm6.33}$ & $\mathbf{62.74\pm3.30}$ & $\mathbf{58.28\pm3.86}$ & $\mathbf{65.88\pm12.49}$    \\
+                                            % & \Verb|Temporal-ML|\mbox{*}     & $\mathbf{78.13\pm6.33}$ & $\mathbf{62.74\pm3.30}$ & $\mathbf{58.28\pm3.86}$ & $\mathbf{65.88\pm12.49}$    \\
         \bottomrule
-    \end{tabular}}
+        \end{tabular}
 \end{table}
@@
     \centering
-    \cprotect\caption{Classification results (\%) of \verb|Temporal-ML| and models without \verb|BiLSTM|. Best results per metrics are boldfaced. (\mbox{*}): Our method.}
+    \cprotect\caption{Classification results (\%) of \verb|Temporal-ML| and models with \verb|BiLSTM| replaced. Best results per metrics are boldfaced. (\mbox{*}): Our method.}
     \label{tab:ab_lstm}

### 2025-03-03 doc/paper/clean.sh

@@ -1,4 +1,0 @@
-rm *.aux *.blg *.bbl *.fdb_latexmk *.lof *.lot *.out *.fls *.toc *.gz *.xml *.log
-rm Chapters/*.aux
-rm Appendix/*.aux
-rm *blx.bib

### 2025-03-03 doc/paper/els-cas-templates/clean.sh

@@ -1,0 +1,3 @@
+rm *.aux *.blg *.bbl *.fdb_latexmk *.lof *.lot *.out *.fls *.toc *.gz *.xml *.log *.abs *.bst
+rm *.cpt
+rm main-blx.bib

### 2025-03-03 doc/paper/preprint/clean.sh
commit for merging

@@ -1,0 +1,4 @@
+rm *.aux *.blg *.bbl *.fdb_latexmk *.lof *.lot *.out *.fls *.toc *.gz *.xml *.log
+rm Chapters/*.aux
+rm Appendix/*.aux
+rm *blx.bib
